### 2024-01-23 manuscript/01_intro.qmd
begin manuscript

---
bibliography: references.bib
---

# Introduction

Rhizaria are an extremely diverse super-group of single celled organisms consisting of several phyla including retaria (foraminifera and radiolaria) and cercozoa. These organisms exist in a wide range of habitats and are widely represented in plankton communities throughout the global ocean. While the taxonomy of these organisms has recently undergone several reclassifications [@biard2022], their presence in ocean ecosystems has been long known to oceanographers. Some of the earliest records of their existence are from oceanographic expeditions in the 19th century [@haekel1887]. Rhizaria are unique members of the plankton and protist community because they can reach large sizes (up to several mm in diameter) and they construct intricate mineral skeletons out of either silica, strontium, or calcium carbonate [@kimoto2015; @suzuki2015; @nakamura2015; @biard2022]. Despite their noticeable morphology and global distribution, Rhizaria were largely understudied throughout the 20th century. The bulk of modern plankton research has focused hard-bodied crustacea which are numerically dominant and easily sampled with nets and preservatives. Fragile organisms like Rhizaria, were difficult to adequately study as they can be destroyed through standard zooplankton sampling techniques. A number of studies in the late 1900s did employ alternative techniques to quantify Rhizaria including diaphragm pumps [@michaels1988] or blue-water SCUBA collections [@caron1984; @caron1995; @bijma1990]. However, the bulk of Rhizaria research was constrained to sediment traps or paleontological studies of sediment [@takahashi1983; @boltovskoy1993]. Only recently has the advent of molecular techniques and in-situ imaging tools ignited a renewed focus on Rhizaria in pelagic ecosystems [@caron2016].

The wave of new data on Rhizaria has facilitated an improved understanding of the significance in ocean ecosystem functions. Firstly, taxonomists have been able to greatly refine the understanding of evolutionary relationships amongst these diverse protists [@aurahs2009; @decelle2012; @decelle2013; @biard2015; @cavalier-smith2018; rev by @biard2022]. DNA metabarcoding studies have revealed insights into the distributional patterns [@decelle2013; @biard2017; @marsbrisbin2020; @blanco-bercial2022 @llopismonferrer2022; @sogawa2022], ecological relationships [@decelle2012; @nakamura2023], and contribution to biogeochemical fluxes [@guidi2016; @gutierrez-rodriguez2019]. Transcriptomic and proteomic approaches also have been used to quantify rhizarian contribution to community metabolism [@cohen2023]. Yet, despite the excellent taxonomic resolution provided by molecular approaches, they do not provide a truly quantitative metric for estimating Rhizarian abundance or biomass. In-situ imaging tools however, offer the ability to observe organisms in the natural state and quantify their abundance [@ohman2019a; @barth-inrev]. While there were early applicaitons of imaging tools to document Rhizaria [@dennett2002], @biard2016's report from a global imaging dataset highlighted the importance of Rhizaria to the total standing stock of marine carbon. Due to their large sizes, ability to concentrate smaller particles (ballasting), and the unique structure of their mineral skeletons, Rhizaria have the potential to massively influence ocean biogeochemical cycling. A number of studies have made large advances in estimating the contribution of Rhizaria to ocean cycling of carbon [@lampitt2009; @stukel2018; @ikenoue2019; @gutierrez-rodriguez2019], silica [@biard2018; @llopismonferrer2021], and strontium [@decelle2013]. Still, Rhizarian ecological roles are not well understood [@biard2022]. This is a major challenge as it is critical to understand the ecological role of plankton to fully incorporate them into biological oceanographic models.

The ecological role of Rhizaria in plankton communities is complicated due to the fact different taxa can exhibit every different trophic modes. As zooplankton, rhizaria are predominately heterotrophic [@biard2022], yet their feeding modes can be quite varied. Phaeodarians (family Cercozoa) are largely thought to be flux-feeders, collecting and feeding on sinking particles [@nakamura2015; @stukel2019]. Alternatively, Retaria can be either exclusively heterotrophic or mixotrophic, utilizing photosynthetic algal symbionts [@anderson2014; @decelle2015]. Mixotrophic foraminifera host a variety of endosymbiont partners [@lee2006; @decelle2015], which are thought to support early and adult life stages and contribute to total primary productivity [@kimoto2015]. Still, foraminifera are omnivorous, possibly even predominately carnivorous with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], majoritively consuming live copepods [@caron1984]. Radiolaria have several lineages all which have some taxa who are well known to host symbionts [@biard2022b]. Amongst radiolarians, arguably the most widespread are Collodaria who can be either large solitary cells or form massive colonies, up to several meters in length [@swanberg1981]. All known Collodaria species host dinoflagellate symbionts [@biard2022b] and can contribute substantially to primary productivity, particularly in oligotrophic ocean regions [@caron1995; @dennett2002]. This Collodaria-symbiont association has been suggested as a reason for their high abundances throughout the photic zone of oligotrophic environments globally [@biard2016; @biard2017]. A few Acantharean (Radiolarian order) clades host algal symbionts [@decelle2012; @biard2022b], notably with two clades forming an exclusive relationship with Phaeocystis. However, globally Acantharea are less abundant than Collodaria [@biard2022] and contribute less to total primary productivity [@michaels1995]. This may be due to the fact several clades of Acantharea are cyst-forming and strictly heterotrophs [@decelle2013; @biard2022b]. Furthermore, @marsbrisbin2020 documented apparent predation behavior in Acantharea near the surface, suggesting that there may be a large reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among Rhizarian taxa, it is reasonable to expect some form of niche partitioning. A number of studies do suggest evidence for vertical zonation between Rhizaria groups according to various trophic strategies. Taxa-specific studies of radiolarians suggest they may be restricted to the euphotic zone [@michaels1988; @boltovskoy2017]. Although some studies report Acantharea in deeper waters [@decelle2013; @gutiérrez-rodríguez2022]. Phaeodarians alternatively, are generally found in the mesopelagic where photosynthesis cannot occur but they can feed on sinking particles [@stukel2018]. In an imaging-based study of the whole Rhizaria community, @biard2020 noted clear patterns in vertical zonation which largely corresponded to different trophic roles. In the oligotrophic ocean, @blanco-bercial2022 also noted that the protist community, including Rhizaria partition along an autotroph and mixotroph to heterotroph gradient with increasing depth in the water column. Yet, few studies have made direct attempts to relate rhizaria abundances to environmental factors [@biard2020]. In part, this is due to the fact few studies have been able to sample Rhizaria in the same location over a consistent timeframe [@michaels1988; @boltovskoy1993; @michaels1995; @hull2011; @gutiérrez-rodríguez2022]. Furthermore, no studies have utilized imaging, arguably the best method for quantifying rhizaria, consistently throughout the full mesopelagic. Given this lack of information, there are many unknowns with respect to Rhizarian ecology, seasonality and phenology across different groups.

In this study, we present a comprehensive assessment of large Rhizaria measured for multiple months (greater than 1 year) using an in-situ imaging approach. With this dataset, we address two critical aims. 1) Quantification of large Rhizaria throughout the epipelagic (0-200m) and mesopelagic (200-1000m) over the course of an annual cycle. These data were collected in the Sargasso Sea, and represents the first study of its kind in an oligotrophic system. 2) We aim to test the hypothesis that Rhizaria exhibit niche partitioning according to trophic roles. This hypothesis makes several predictions, including vertical zonation, as seen in prior studies, but also that environmental variables related to trophic strategy will explain abundance patterns. Specifically, autotrophic/mixotrophic taxa will correspond to variables related to autotrophy (chl-a concentration, primary productivity, high $O_2$ ) and other rhizaria will correspond to factors which promote heterotrophy (particle concentration, flux, and low $O_2$).

### 2024-01-23 manuscript/02_methods.qmd
begin manuscript

---
bibliography: references.bib
---

# Methods

## Oceanographic Sampling

Data were collected in collaboration with the Bermuda Atlantic Time-Series Study [@michaels1996; @lomas2013] on board the R/V Atlantic Explorer. Cruises were conducted at approximately monthly intervals. Rhizaria individuals were sampled using the Underwater Vision Profiler 5 [UVP5; @picheral2010], a tool which is well established to accurately quantify large Rhizaria [@biard2016; @stukel2018; @stukel2019; @biard2020; @barth2022; @drago2022; @llopismonferrer2022; @panaïotis2023]. The UVP5 was mounted to the sampling rosette and collected data autonomously on routine casts, from which only the downcast data are utilized. The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022. Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda into the Sargasso Sea (Supplemental Figure 1; Supplemental Table 1). In general, casts extended to either 200m, 500m, or 1200m deep, with a few extended into the bathypelagic (4500m). However, Rhizaria were only typically found in large abundances throughout the epipelagic and mesopelagic zones. As such, we limit this study to results from the upper 1000m of the water column.

A variety of biotic and abiotic data were collected during each BATS cruise. Briefly, we will explain the data utilized in this study. Salinity, temperature, Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were measured at high-frequencies on each cast, attached to the same CTD rosette as the UVP. On certain casts, niskin bottles were used to collect bacterial abundance estimates (via epifluorescence microscopy) as well as measure inorganic nutrients ($NO_3$, and $Si$ as silicate/sicilic acid) at discrete depths. On each cruise, flux estimates of total mass, carbon, and nitrogen were also collected using sediment traps (in the present study we utilize flux to the mesopelagic as the flux at 200m). Also primary productivity was estimated through measuring $C^{14}$ uptake rates from in-situ incubations. Full descriptions of the BATS sampling program, methods, and data are available freely online (https://bats.bios.asu.edu/bats-data/) and reviewed in [@lomas2013].

Environmental data were processed in a variety of ways to match the format of the Rhizaria abundance estimates (See below). CTD data were collected at higher frequency than the UVP, so these data were averaged within matching bins to the UVP5 data. Data from niskin bottles were first linearly interpolated with a 1m resolution for each cast where data were available. Then, because data were collected at different frequencies over the course of a cruise, the interpolated data were averaged across casts, then averaged into matching UVP5-sized bins. Primary productivity estimates were totaled within the euphotic zone to represent a "total euphotic productivity". Finally the data from the sediment trap deployed to 200m was used to represent "flux to the mesopelagic".

## Rhizaria imaging processing and quantification

Individual vignettes of Rhizaria images were identified using the classification platform Ecotaxa [cite]. Data were pre-sorted according to a . Taxonomic classification were done based on morphology exclusively. While there are sparse taxonomic guides for in-situ images of rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020]. Using the aforementioned sources and publicly available ecotaxa projects, I constructed a guide which can be found here: (https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf). Broadly, Rhizaria were classified as Foraminifera, Radiolarians (Acantharea or Collodaria), or a variety of Phaeodarian families (Figure 1). When identification could not be confidently made between a few candidate taxa, a less specific label was used. As a result, we have data from "unidentified Rhizaria", which typically were vignettes not distinguishable between Aulacanthidae or Acantharea or "unidentified Phaeodaria", which are clearly phaeodaria but not distinguishable into a family.

The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$. However, identified rhizaria ranged from a 934$\mu m$ Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To confirm that the UVP5 was sampling adequately across all size ranges, an NBSS slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth-inrev] . However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded. The UVP5 reports the exact depth at which a particle is recorded, however to estimate abundance, observations must be binned over fixed depth intervals. In our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic [see @barth2022 for an extended discussion of UVP5 data processing). For the present study, Rhizarian abundances were estimated in 25m vertical bins, which offer a moderate sampling volume (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic) while still maintaining ecologically relevant widths. However, concentrations in a 25m bin would need to be greater than 2.428 indv. $m^{-3}$ and 3.912 indv. $m^{-3}$, in the epipealgic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth-inrev]. Because we typically observed many rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions. For quantifying and modelling Rhizaria abundances, we present integrated abundance estimates, with each cast. Due to the variable descent depths of the UVP, data are categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic (500-1000m). The average sampling volume integrated through these regions were 7.59$m^2$, 7.06$m^2$, and 11.77$m^2$, with non-detection thresholds at 0.30 indv. $m^{-2}$, 0.33 indv. $m^{-2}$, and 0.20 indv. $m^{-2}$ respectively.

## Modelling environmental controls of Rhizarian Abundance

Generalized Additive Models (GAMs) were used to assess the relationship between integrated rhizarian abudance and different environmental factors.

### 2024-01-23 manuscript/main.qmd
begin manuscript

---
title: Rhizaria in the oligotrophic exhibit clear temporal and vertical variability.
bibliography: references.bib
csl: ./deep-sea-research-part-i.csl
execute:
  echo: false
  warning: false
  message: false
author:
  - name: Alex Barth*
    affiliations:
    - name: University of South Carolina
      department: Biological Sciences
      address: 700 Sumter St 401, Columbia, SC 29208
      email: AB93@email.sc.edu
      phone: +1 (440) 668-8376
  - name: Leo Blanco-Berical
    affiliations:
    - name: Arizone State University
  - name: Rod Johnson
    affiliations: 
    - name: Bermuda Institute of Ocean Sciences
      address: Bermuda GE 01, UK
  - name: Joshua Stone
    affiliations:
    - name: University of South Carolina
      department: Biological Sciences
      address: 700 Sumter St 401, Columbia, SC 29208
format:
  pdf:
    # template-partials:
    #   - "before-body.tex"
    #   - "_titlepage.tex"
    # include-in-header:
    #   - "in-header.tex"
    number-sections: false
    linestretch: 2
    tbl-cap-location: top
    # filters: [wordcount.lua]
    keep-tex: true
---

```{r}
#| child: ./01_intro.qmd
```

### 2024-01-23 manuscript/main.tex
begin manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Rhizaria in the oligotrophic exhibit clear temporal and vertical variability.},
  pdfauthor={Alex Barth*; Leo Blanco-Berical; Rod Johnson; Joshua Stone},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Rhizaria in the oligotrophic exhibit clear temporal and vertical
variability.}
\author{Alex Barth* \and Leo Blanco-Berical \and Rod Johnson \and Joshua
Stone}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable]}{\end{tcolorbox}}\fi

\setstretch{2}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Rhizaria are an extremely diverse super-group of single celled organisms
consisting of several phyla including retaria (foraminifera and
radiolaria) and cercozoa. These organisms exist in a wide range of
habitats and are widely represented in plankton communities throughout
the global ocean. While the taxonomy of these organisms has recently
undergone several reclassifications (Wiley, 2022), their presence in
ocean ecosystems has been long known to oceanographers. Some of the
earliest records of their existence are from oceanographic expeditions
in the 19th century (Haekel, 1887). Rhizaria are unique members of the
plankton and protist community because they can reach large sizes (up to
several mm in diameter) and they construct intricate mineral skeletons
out of either silica, strontium, or calcium carbonate (Kimoto, 2015;
Nakamura and Suzuki, 2015; Suzuki and Not, 2015; Wiley, 2022). Despite
their noticeable morphology and global distribution, Rhizaria were
largely understudied throughout the 20th century. The bulk of modern
plankton research has focused hard-bodied crustacea which are
numerically dominant and easily sampled with nets and preservatives.
Fragile organisms like Rhizaria, were difficult to adequately study as
they can be destroyed through standard zooplankton sampling techniques.
A number of studies in the late 1900s did employ alternative techniques
to quantify Rhizaria including diaphragm pumps (Michaels, 1988) or
blue-water SCUBA collections (Bijma et al., 1990; Caron et al., 1995;
Caron and Be, 1984). However, the bulk of Rhizaria research was
constrained to sediment traps or paleontological studies of sediment
(Boltovskoy et al., 1993; Takahashi et al., 1983). Only recently has the
advent of molecular techniques and in-situ imaging tools ignited a
renewed focus on Rhizaria in pelagic ecosystems (Caron, 2016).

The wave of new data on Rhizaria has facilitated an improved
understanding of the significance in ocean ecosystem functions. Firstly,
taxonomists have been able to greatly refine the understanding of
evolutionary relationships amongst these diverse protists (Aurahs et
al., 2009; Biard et al., 2015; Cavalier-Smith et al., 2018; Decelle et
al., 2013, 2012; rev by Wiley, 2022). DNA metabarcoding studies have
revealed insights into the distributional patterns Llopis Monferrer et
al. (2022), ecological relationships (Decelle et al., 2012; Nakamura et
al., 2023), and contribution to biogeochemical fluxes (Guidi et al.,
2016; Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
approaches also have been used to quantify rhizarian contribution to
community metabolism (Cohen et al., 2023). Yet, despite the excellent
taxonomic resolution provided by molecular approaches, they do not
provide a truly quantitative metric for estimating Rhizarian abundance
or biomass. In-situ imaging tools however, offer the ability to observe
organisms in the natural state and quantify their abundance (Barth and
Stone, in review; Ohman, 2019). While there were early applicaitons of
imaging tools to document Rhizaria (Dennett et al., 2002), Biard et al.
(2016)'s report from a global imaging dataset highlighted the importance
of Rhizaria to the total standing stock of marine carbon. Due to their
large sizes, ability to concentrate smaller particles (ballasting), and
the unique structure of their mineral skeletons, Rhizaria have the
potential to massively influence ocean biogeochemical cycling. A number
of studies have made large advances in estimating the contribution of
Rhizaria to ocean cycling of carbon (Gutierrez-Rodriguez et al., 2019;
Ikenoue et al., 2019; Lampitt et al., 2009; Stukel et al., 2018), silica
(Biard et al., 2018; Llopis Monferrer et al., 2021), and strontium
(Decelle et al., 2013). Still, Rhizarian ecological roles are not well
understood (Wiley, 2022). This is a major challenge as it is critical to
understand the ecological role of plankton to fully incorporate them
into biological oceanographic models.

The ecological role of Rhizaria in plankton communities is complicated
due to the fact different taxa can exhibit every different trophic
modes. As zooplankton, rhizaria are predominately heterotrophic (Wiley,
2022), yet their feeding modes can be quite varied. Phaeodarians (family
Cercozoa) are largely thought to be flux-feeders, collecting and feeding
on sinking particles (Nakamura and Suzuki, 2015; Stukel et al., 2019).
Alternatively, Retaria can be either exclusively heterotrophic or
mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
Decelle et al., 2015). Mixotrophic foraminifera host a variety of
endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
thought to support early and adult life stages and contribute to total
primary productivity (Kimoto, 2015). Still, foraminifera are omnivorous,
possibly even predominately carnivorous with several studies suggesting
that they can be effective predators (Anderson and Bé, 1976; Gaskell et
al., 2019), majoritively consuming live copepods (Caron and Be, 1984).
Radiolaria have several lineages all which have some taxa who are well
known to host symbionts (Biard, 2022). Amongst radiolarians, arguably
the most widespread are Collodaria who can be either large solitary
cells or form massive colonies, up to several meters in length (Swanberg
and Anderson, 1981). All known Collodaria species host dinoflagellate
symbionts (Biard, 2022) and can contribute substantially to primary
productivity, particularly in oligotrophic ocean regions (Caron et al.,
1995; Dennett et al., 2002). This Collodaria-symbiont association has
been suggested as a reason for their high abundances throughout the
photic zone of oligotrophic environments globally (Biard et al., 2017,
2016). A few Acantharean (Radiolarian order) clades host algal symbionts
(Biard, 2022; Decelle et al., 2012), notably with two clades forming an
exclusive relationship with Phaeocystis. However, globally Acantharea
are less abundant than Collodaria (Wiley, 2022) and contribute less to
total primary productivity (Michaels et al., 1995). This may be due to
the fact several clades of Acantharea are cyst-forming and strictly
heterotrophs (Biard, 2022; Decelle et al., 2013). Furthermore, Mars
Brisbin et al. (2020) documented apparent predation behavior in
Acantharea near the surface, suggesting that there may be a large
reliance on carnivory.

Given the high abundances, yet diverse trophic strategies found among
Rhizarian taxa, it is reasonable to expect some form of niche
partitioning. A number of studies do suggest evidence for vertical
zonation between Rhizaria groups according to various trophic
strategies. Taxa-specific studies of radiolarians suggest they may be
restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
Although some studies report Acantharea in deeper waters (Decelle et
al., 2013; Gutiérrez-Rodríguez et al., 2022). Phaeodarians
alternatively, are generally found in the mesopelagic where
photosynthesis cannot occur but they can feed on sinking particles
(Stukel et al., 2018). In an imaging-based study of the whole Rhizaria
community, Biard and Ohman (2020) noted clear patterns in vertical
zonation which largely corresponded to different trophic roles. In the
oligotrophic ocean, Blanco-Bercial et al. (2022) also noted that the
protist community, including Rhizaria partition along an autotroph and
mixotroph to heterotroph gradient with increasing depth in the water
column. Yet, few studies have made direct attempts to relate rhizaria
abundances to abiotic environmental factors (Biard and Ohman, 2020). In
part, this is due to the fact few studies have been able to sample
Rhizaria in the same location over a consistent timeframe (Boltovskoy et
al., 1993; Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels
et al., 1995; Michaels, 1988). Furthermore, no studies have utilized
imaging, arguably the best method for quantifying rhizaria, consistently
throughout the full mesopelagic. Given this lack of information, there
are many unknowns with respect to Rhizarian ecology, seasonality and
phenology across different groups.

In this study, we present a comprehensive assessment of large Rhizaria
measured for multiple months (greater than 1 year) using an in-situ
imaging approach. With this dataset, we address two critical aims. 1)
Quantification of large Rhizaria throughout the epipelagic (0-200m) and
mesopelagic (200-1000m) over the course of an annual cycle. These data
were collected in the Sargasso Sea, and represents the first study of
its kind in an oligotrophic system. 2) We aim to test the hypothesis
that Rhizaria exhibit niche partitioning according to trophic roles.
This hypothesis makes several predictions, including vertical zonation,
as seen in prior studies, but also that abiotic variables related to
trophic strategy will explain abundance patterns. Specifically,
autotrophic/mixotrophic taxa will correspond to variables related to
autotrophy (chl-a concentration, primary productivity, high \(O_2\) )
and other rhizaria will correspond to factors which promote heterotrophy
(particle concentration, flux, and low \(O_2\)).

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-anderson2014}{}}%
Anderson, O.R., 2014. Living Together in the Plankton: A Survey of
Marine Protist Symbioses. Acta Protozoologica 53, 29--38.
\url{https://doi.org/10.4467/16890027AP.13.0019.1116}

\leavevmode\vadjust pre{\hypertarget{ref-anderson1976}{}}%
Anderson, O.R., Bé, A.W.H., 1976. A CYTOCHEMICAL FINE STRUCTURE STUDY OF
PHAGOTROPHY IN A PLANKTONIC FORAMINIFER, {\emph{HASTIGERINA PELAGICA}}
(d'ORBIGNY). The Biological Bulletin 151, 437--449.
\url{https://doi.org/10.2307/1540498}

\leavevmode\vadjust pre{\hypertarget{ref-aurahs2009}{}}%
Aurahs, R., Grimm, G.W., Hemleben, V., Hemleben, C., Kucera, M., 2009.
Geographical distribution of cryptic genetic types in the planktonic
foraminifer Globigerinoides ruber. Molecular Ecology 18, 1692--1706.
\url{https://doi.org/10.1111/j.1365-294X.2009.04136.x}

\leavevmode\vadjust pre{\hypertarget{ref-barth-inrev}{}}%
Barth, A., Stone, J., in review. Understanding the picture: The promis
and challenges of in-situ imagery data in the study of plankton ecology.

\leavevmode\vadjust pre{\hypertarget{ref-biard2022b}{}}%
Biard, T., 2022. Diversity and ecology of Radiolaria in modern oceans.
Environmental Microbiology 24, 2179--2200.
\url{https://doi.org/10.1111/1462-2920.16004}

\leavevmode\vadjust pre{\hypertarget{ref-biard2017}{}}%
Biard, T., Bigeard, E., Audic, S., Poulain, J., Gutierrez-Rodriguez, A.,
Pesant, S., Stemmann, L., Not, F., 2017. Biogeography and diversity of
Collodaria (Radiolaria) in the global ocean. The ISME Journal 11,
1331--1344. \url{https://doi.org/10.1038/ismej.2017.12}

\leavevmode\vadjust pre{\hypertarget{ref-biard2018}{}}%
Biard, T., Krause, J.W., Stukel, M.R., Ohman, M.D., 2018. The
Significance of Giant Phaeodarians (Rhizaria) to Biogenic Silica Export
in the California Current Ecosystem. Global Biogeochemical Cycles 32,
987--1004. \url{https://doi.org/10.1029/2018GB005877}

\leavevmode\vadjust pre{\hypertarget{ref-biard2020}{}}%
Biard, T., Ohman, M.D., 2020. Vertical niche definition of test-bearing
protists (Rhizaria) into the twilight zone revealed by in situ imaging.
Limnology and Oceanography 65, 2583--2602.
\url{https://doi.org/10.1002/lno.11472}

\leavevmode\vadjust pre{\hypertarget{ref-biard2015}{}}%
Biard, T., Pillet, L., Decelle, J., Poirier, C., Suzuki, N., Not, F.,
2015. Towards an integrative morpho-molecular classification of the
collodaria (polycystinea, radiolaria). Protist 166, 374--388.
\url{https://doi.org/10.1016/j.protis.2015.05.002}

\leavevmode\vadjust pre{\hypertarget{ref-biard2016}{}}%
Biard, T., Stemmann, L., Picheral, M., Mayot, N., Vandromme, P., Hauss,
H., Gorsky, G., Guidi, L., Kiko, R., Not, F., 2016. In situ imaging
reveals the biomass of giant protists in the global ocean. Nature 532,
504--507. \url{https://doi.org/10.1038/nature17652}

\leavevmode\vadjust pre{\hypertarget{ref-bijma1990}{}}%
Bijma, J., Erez, J., Hemleben, C., 1990.
\href{https://epic.awi.de/id/eprint/6096/1/Bij1990a.pdf}{Lunar and
semi-luan reproductive cycles in some spinose planktonic foraminifers}.
Journal of Foraminiferal Research 20, 117--127.

\leavevmode\vadjust pre{\hypertarget{ref-blanco-bercial2022}{}}%
Blanco-Bercial, L., Parsons, R., Bolaños, L.M., Johnson, R., Giovannoni,
S.J., Curry, R., 2022.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.897140}{The
protist community traces seasonality and mesoscale hydrographic features
in the oligotrophic sargasso sea}. Frontiers in Marine Science 9.

\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy2017}{}}%
Boltovskoy, D., 2017. Vertical distribution patterns of radiolaria
polycystina (protista) in the world ocean: Living ranges, isothermal
submersion and settling shells. Journal of Plankton Research 39,
330--349. \url{https://doi.org/10.1093/plankt/fbx003}

\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy1993}{}}%
Boltovskoy, D., Alder, V.A., Abelmann, A., 1993. Annual flux of
radiolaria and other shelled plankters in the eastern equatorial
atlantic at 853 m: seasonal variations and polycystine species-specific
responses. Deep Sea Research Part I: Oceanographic Research Papers 40,
1863--1895. \url{https://doi.org/10.1016/0967-0637(93)90036-3}

\leavevmode\vadjust pre{\hypertarget{ref-caron2016}{}}%
Caron, D.A., 2016.
\href{https://www.nature.com/articles/nature17892}{The rise of rhizaria
\textbar{} nature}. Nature 532, 444--445.

\leavevmode\vadjust pre{\hypertarget{ref-caron1984}{}}%
Caron, D.A., Be, A.W.H., 1984. PREDICTED AND OBSERVED FEEDING RATES OF
THE SPINOSE PLANKTONIC FORAMINIFER. BULLETIN OF MARINE SCIENCE 35.

\leavevmode\vadjust pre{\hypertarget{ref-caron1995}{}}%
Caron, D.A., Michaels, A.F., Swanberg, N.R., Howse, F.A., 1995. Primary
productivity by symbiont-bearing planktonic sarcodines (acantharia,
radiolaria, foraminifera) in surface waters near bermuda. Journal of
Plankton Research 17, 103--129.
\url{https://doi.org/10.1093/plankt/17.1.103}

\leavevmode\vadjust pre{\hypertarget{ref-cavalier-smith2018}{}}%
Cavalier-Smith, T., Chao, E.E., Lewis, R., 2018. Multigene phylogeny and
cell evolution of chromist infrakingdom Rhizaria: contrasting cell
organisation~of sister~phyla Cercozoa and Retaria. Protoplasma 255,
1517--1574. \url{https://doi.org/10.1007/s00709-018-1241-1}

\leavevmode\vadjust pre{\hypertarget{ref-cohen2023}{}}%
Cohen, N.R., Krinos, A.I., Kell, R.M., Chmiel, R.J., Moran, D.M.,
McIlvin, M.R., Lopez, P.Z., Barth, A., Stone, J., Alanis, B.A., Chan,
E.W., Breier, J.A., Jakuba, M.V., Johnson, R., Alexander, H., Saito,
M.A., 2023. Microeukaryote metabolism across the western North Atlantic
Ocean revealed through autonomous underwater profiling.
\url{https://doi.org/10.1101/2023.11.20.567900}

\leavevmode\vadjust pre{\hypertarget{ref-decelle2015}{}}%
Decelle, J., Colin, S., Foster, R.A., 2015. Photosymbiosis in Marine
Planktonic Protists, in: Ohtsuka, S., Suzaki, T., Horiguchi, T., Suzuki,
N., Not, F. (Eds.),. Springer Japan, Tokyo, pp. 465--500.
\url{https://doi.org/10.1007/978-4-431-55130-0_19}

\leavevmode\vadjust pre{\hypertarget{ref-decelle2013}{}}%
Decelle, J., Martin, P., Paborstava, K., Pond, D.W., Tarling, G., Mahé,
F., De Vargas, C., Lampitt, R., Not, F., 2013. Diversity, Ecology and
Biogeochemistry of Cyst-Forming Acantharia (Radiolaria) in the Oceans.
PLoS ONE 8, e53598. \url{https://doi.org/10.1371/journal.pone.0053598}

\leavevmode\vadjust pre{\hypertarget{ref-decelle2012}{}}%
Decelle, J., Siano, R., Probert, I., Poirier, C., Not, F., 2012.
Multiple microalgal partners in symbiosis with the acantharian
Acanthochiasma sp. (Radiolaria). Symbiosis 58, 233--244.
\url{https://doi.org/10.1007/s13199-012-0195-x}

\leavevmode\vadjust pre{\hypertarget{ref-dennett2002}{}}%
Dennett, M.R., Caron, D.A., Michaels, A.F., Gallager, S.M., Davis, C.S.,
2002. Video plankton recorder reveals high abundances of colonial
radiolaria in surface waters of the central north pacific. Journal of
Plankton Research 24, 797--805.
\url{https://doi.org/10.1093/plankt/24.8.797}

\leavevmode\vadjust pre{\hypertarget{ref-gaskell2019}{}}%
Gaskell, D.E., Ohman, M.D., Hull, P.M., 2019. Zooglider-based
measurements of planktonic foraminifera in the california current
system. Journal of Foraminiferal Research 49, 390--404.
\url{https://doi.org/10.2113/gsjfr.49.4.390}

\leavevmode\vadjust pre{\hypertarget{ref-guidi2016}{}}%
Guidi, L., Chaffron, S., Bittner, L., Eveillard, D., Larhlimi, A., Roux,
S., Darzi, Y., Audic, S., Berline, L., Brum, J.R., Coelho, L.P.,
Espinoza, J.C.I., Malviya, S., Sunagawa, S., Dimier, C., Kandels-Lewis,
S., Picheral, M., Poulain, J., Searson, S., Stemmann, L., Not, F.,
Hingamp, P., Speich, S., Follows, M., Karp-Boss, L., Boss, E., Ogata,
H., Pesant, S., Weissenbach, J., Wincker, P., Acinas, S.G., Bork, P.,
Vargas, C. de, Iudicone, D., Sullivan, M.B., Raes, J., Karsenti, E.,
Bowler, C., Gorsky, G., 2016. Plankton networks driving carbon export in
the oligotrophic ocean. Nature 532, 465--470.
\url{https://doi.org/10.1038/nature16942}

\leavevmode\vadjust pre{\hypertarget{ref-gutierrez-rodriguez2019}{}}%
Gutierrez-Rodriguez, A., Stukel, M.R., Lopes dos Santos, A., Biard, T.,
Scharek, R., Vaulot, D., Landry, M.R., Not, F., 2019. High contribution
of Rhizaria (Radiolaria) to vertical export in the California Current
Ecosystem revealed by DNA metabarcoding. The ISME Journal 13, 964--976.
\url{https://doi.org/10.1038/s41396-018-0322-7}

\leavevmode\vadjust pre{\hypertarget{ref-gutiuxe9rrez-rodruxedguez2022}{}}%
Gutiérrez-Rodríguez, A., Lopes dos Santos, A., Safi, K., Probert, I.,
Not, F., Fernández, D., Gourvil, P., Bilewitch, J., Hulston, D.,
Pinkerton, M., Nodder, S.D., 2022. Planktonic protist diversity across
contrasting subtropical and subantarctic waters of the southwest
pacific. Progress in Oceanography 206, 102809.
\url{https://doi.org/10.1016/j.pocean.2022.102809}

\leavevmode\vadjust pre{\hypertarget{ref-haekel1887}{}}%
Haekel, E., 1887.
\href{https://www.gutenberg.org/cache/epub/44527/pg44527-images.html}{Report
on the radiolaria collected by h.m.s. Challenger during the years
1873-1876, plates, by ernst haeckel}.

\leavevmode\vadjust pre{\hypertarget{ref-hull2011}{}}%
Hull, P.M., Osborn, K.J., Norris, R.D., Robison, B.H., 2011. Seasonality
and depth distribution of a mesopelagic foraminifer, Hastigerinella
digitata, in Monterey Bay, California. Limnology and Oceanography 56,
562--576. \url{https://doi.org/10.4319/lo.2011.56.2.0562}

\leavevmode\vadjust pre{\hypertarget{ref-ikenoue2019}{}}%
Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M.C., Takahashi,
K., Harada, N., Fujiki, T., 2019. Phaeodaria: An Important Carrier of
Particulate Organic Carbon in the Mesopelagic Twilight Zone of the North
Pacific Ocean. Global Biogeochemical Cycles 33, 1146--1160.
\url{https://doi.org/10.1029/2019GB006258}

\leavevmode\vadjust pre{\hypertarget{ref-kimoto2015}{}}%
Kimoto, K., 2015. Planktic foraminifera. pp. 129--178.
\url{https://doi.org/10.1007/978-4-431-55130-0_7}

\leavevmode\vadjust pre{\hypertarget{ref-lampitt2009}{}}%
Lampitt, R.S., Salter, I., Johns, D., 2009. Radiolaria: Major exporters
of organic carbon to the deep ocean. Global Biogeochemical Cycles 23.
\url{https://doi.org/10.1029/2008GB003221}

\leavevmode\vadjust pre{\hypertarget{ref-lee2006}{}}%
Lee, J.J., 2006.
\href{https://dalspace.library.dal.ca/bitstream/handle/10222/78255/VOLUME\%2042-NUMBER\%202-2006-PAGE\%2063.pdf?sequence=1}{Algal
symbiosis in larger foraminifera}. Symbiosis 42, 63--75.

\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2022}{}}%
Llopis Monferrer, N., Biard, T., Sandin, M.M., Lombard, F., Picheral,
M., Elineau, A., Guidi, L., Leynaert, A., Tréguer, P.J., Not, F., 2022.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.895995}{Siliceous
rhizaria abundances and diversity in the mediterranean sea assessed by
combined imaging and metabarcoding approaches}. Frontiers in Marine
Science 9.

\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2021}{}}%
Llopis Monferrer, N., Leynaert, A., Tréguer, P., Gutiérrez-Rodríguez,
A., Moriceau, B., Gallinari, M., Latasa, M., L'Helguen, S., Maguer,
J.-F., Safi, K., Pinkerton, M.H., Not, F., 2021. Role of small Rhizaria
and diatoms in the pelagic silica production of the Southern Ocean.
Limnology and Oceanography 66, 2187--2202.
\url{https://doi.org/10.1002/lno.11743}

\leavevmode\vadjust pre{\hypertarget{ref-marsbrisbin2020}{}}%
Mars Brisbin, M., Brunner, O.D., Grossmann, M.M., Mitarai, S., 2020.
Paired high-throughput, in situ imaging and high-throughput sequencing
illuminate acantharian abundance and vertical distribution. Limnology
and Oceanography 65, 2953--2965. \url{https://doi.org/10.1002/lno.11567}

\leavevmode\vadjust pre{\hypertarget{ref-michaels1988}{}}%
Michaels, A.F., 1988. Vertical distribution and abundance of Acantharia
and their symbionts. Marine Biology 97, 559--569.
\url{https://doi.org/10.1007/BF00391052}

\leavevmode\vadjust pre{\hypertarget{ref-michaels1995}{}}%
Michaels, A.F., Caron, D.A., Swanberg, N.R., Howse, F.A., Michaels,
C.M., 1995. Planktonic sarcodines (Acantharia, Radiolaria, Foraminifera)
in surface waters near Bermuda: abundance, biomass and vertical flux.
Journal of Plankton Research 17, 131--163.
\url{https://doi.org/10.1093/plankt/17.1.131}

\leavevmode\vadjust pre{\hypertarget{ref-nakamura2023}{}}%
Nakamura, Y., Itagaki, H., Tuji, A., Shimode, S., Yamaguchi, A., Hidaka,
K., Ogiso-Tanaka, E., 2023. DNA metabarcoding focused on
difficult-to-culture protists: An effective approach to clarify
biological interactions. Environmental Microbiology 25, 3630--3638.
\url{https://doi.org/10.1111/1462-2920.16524}

\leavevmode\vadjust pre{\hypertarget{ref-nakamura2015}{}}%
Nakamura, Y., Suzuki, N., 2015.
\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_9}{Phaeodaria:
Diverse marine cercozoans of world-wide distribution}. Springer, pp.
223--249.

\leavevmode\vadjust pre{\hypertarget{ref-ohman2019a}{}}%
Ohman, M.D., 2019. A sea of tentacles: optically discernible traits
resolved from planktonic organisms in situ. ICES Journal of Marine
Science 76, 1959--1972. \url{https://doi.org/10.1093/icesjms/fsz184}

\leavevmode\vadjust pre{\hypertarget{ref-sogawa2022}{}}%
Sogawa, S., Nakamura, Y., Nagai, S., Nishi, N., Hidaka, K., Shimizu, Y.,
Setou, T., 2022. DNA metabarcoding reveals vertical variation and hidden
diversity of alveolata and rhizaria communities in the western north
pacific. Deep Sea Research Part I: Oceanographic Research Papers 183,
103765. \url{https://doi.org/10.1016/j.dsr.2022.103765}

\leavevmode\vadjust pre{\hypertarget{ref-stukel2018}{}}%
Stukel, M.R., Biard, T., Krause, J., Ohman, M.D., 2018. Large Phaeodaria
in the twilight zone: Their role in the carbon cycle. Limnology and
Oceanography 63, 2579--2594. \url{https://doi.org/10.1002/lno.10961}

\leavevmode\vadjust pre{\hypertarget{ref-stukel2019}{}}%
Stukel, M.R., Ohman, M.D., Kelly, T.B., Biard, T., 2019.
\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00397}{The
roles of suspension-feeding and flux-feeding zooplankton as gatekeepers
of particle flux into the mesopelagic ocean in the northeast pacific}.
Frontiers in Marine Science 6.

\leavevmode\vadjust pre{\hypertarget{ref-suzuki2015}{}}%
Suzuki, N., Not, F., 2015.
\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_8}{Biology
and ecology of radiolaria}. Springer, pp. 179--222.

\leavevmode\vadjust pre{\hypertarget{ref-swanberg1981}{}}%
Swanberg, N.R., Anderson, O.R., 1981. Collozoum caudatum sp. nov.: A
giant colonial radiolarian from equatorial and Gulf Stream waters. Deep
Sea Research Part A. Oceanographic Research Papers 28, 1033--1047.
\url{https://doi.org/10.1016/0198-0149(81)90016-9}

\leavevmode\vadjust pre{\hypertarget{ref-takahashi1983}{}}%
Takahashi, K., Hurd, D.C., Honjo, S., 1983. Phaeodarian skeletons: Their
role in silica transport to the deep sea. Science 222, 616--618.
\url{https://doi.org/10.1126/science.222.4624.616}

\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
Wiley, 2022. Rhizaria. Wiley, pp. 1--11.
\url{https://doi.org/10.1002/9780470015902.a0029469}

\end{CSLReferences}

\end{document}

### 2024-01-25 manuscript/02_methods.qmd
finalize figs for draft

@@
 The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022.
-Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda into the Sargasso Sea (Supplemental Figure 1; Supplemental Table 1).
+Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda into the Sargasso Sea (Supplemental Figure 1).
 In general, casts extended to either 200m, 500m, or 1200m deep, with a few extended into the bathypelagic (4500m).
@@
 Briefly, we will explain the data utilized in this study.
+The UVP5 provided particle count data at a high-frequency from each cast.
+Particle concentration was calculated from this data for all particles 184$\mu m$ - 450$\mu m$.
+The lower size range was set by what could be reliably sampled by the UVP5's pixel resolution (\>2px) and the upper size range is representative of a potential prey field for mesozooplankton [@whitmore2021].
 Salinity, temperature, Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were measured at high-frequencies on each cast, attached to the same CTD rosette as the UVP.
@@
 Also primary productivity was estimated through measuring $C^{14}$ uptake rates from in-situ incubations.
-Full descriptions of the BATS sampling program, methods, and data are available freely online (https://bats.bios.asu.edu/bats-data/) and reviewed in [@lomas2013].
+Full descriptions of the BATS sampling program, methods, and data are available freely online [(https://bats.bios.asu.edu/bats-data/)](https://bats.bios.asu.edu/bats-data/) and reviewed in [@lomas2013].
 
@@
 
-Individual vignettes of Rhizaria images were identified using the classification platform Ecotaxa [cite].
+Individual vignettes of Rhizaria images were identified using the classification platform Ecotaxa \[cite\].
 Data were pre-sorted according to a .
@@
 While there are sparse taxonomic guides for in-situ images of rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020].
-Using the aforementioned sources and publicly available ecotaxa projects, I constructed a guide which can be found here: (https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf).
+Using the aforementioned sources and publicly available ecotaxa projects, I constructed a guide which can be found here: <https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf>.
 Broadly, Rhizaria were classified as Foraminifera, Radiolarians (Acantharea or Collodaria), or a variety of Phaeodarian families (Figure 1).
@@
 As a result, we have data from "unidentified Rhizaria", which typically were vignettes not distinguishable between Aulacanthidae or Acantharea or "unidentified Phaeodaria", which are clearly phaeodaria but not distinguishable into a family.
+
+![Example images of different Rhizaria taxa. 4mm scale bar shown in lower right. All vignettes are same scale](images/01_taxa.pdf)
 
@@
 $m^{-3}$ and 3.912 indv.
-$m^{-3}$, in the epipealgic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth-inrev].
+$m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth-inrev].
 Because we typically observed many rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions.
@@
 $m^{-2}$ respectively.
+All UVP data processing was done using the `EcotaxaTools` package in R (Barth 2023).
 
@@
 
-Generalized Additive Models (GAMs) were used to assess the relationship between integrated rhizarian abudance and different environmental factors.
+Generalized Additive Models (GAMs) were used to assess the relationship between integrated Rhizarian abundance and different environmental factors.
+GAMs offer the ability to model non-linear and non-monotonic relationships, which can be particularly useful in assessing ecological relationships [@wood2017] and have been successfully applied to Rhizarian ecology [@biard2020].
+The `mgcv` package [@wood2001] was used to construct models relating environmental parameters to each taxonomic group's integrated abundance estimates from each cast.
+To select the most parsimonious model for each analysis, a backwards step-wise approach was taken.
+First, a full model was fit using any term which may be ecologically relevant.
+Terms were fit using maximum likelihood with a double penalty approach on unnecessary smooths [@marra2011].
+The smoothness parameter was restricted (k = 6) to prevent overfitting the models.
+At each iteration of the backwards step-wise procedure, the model term with the lowest F score (least statistically significant) was removed.
+This was repeated until all model terms were statistically significant or the $R^2_{adj}$ was substantially reduced.
+Models were fit for each region; epipelagic, upper mesopelagic, and lower mesopelagic.
+In cases where observations were too sparse for a given taxonomic grouping, models were not run.

### 2024-01-25 manuscript/03_results.qmd
finalize figs for draft

---
bibliography: references.bib
---

# Results

### Environmental Variability

The BATS sampling region is southeast of Bermuda, situated in the North Sargasso Sea and the North Atlantic oligotrophic ocean gyre. Due to the sampling location, while the environmental conditions are generally low in variation and oligotrophic, there is some seasonality and considerable influence from mesoscale eddies [@mcgillicuddy1998; @lomas2013]. Variability in the water column structure was visible during the study period (Figure 2). This is best evidenced through the temperature profiles; In the late summer and early fall there was a stratified water column with high temperatures in the surface (\<75m) (Figure 2A) and slightly elevated salinity (Figure 2B). This warm, stratified period appeared more intense during the few months sampled in 2019. In 2021, we observed the stratified layer slowly dissipated into the winter months. There was a consistent oxygen minimum zone (OMZ) located at about 750m deep (Figure 2C). Although, February 2021 saw a notable downwelling event, likely due to a passing cyclonic eddy. During this time, warmer, oxic water was plunged deeper into the mesopelagic. This process was reversed in the spring months (March, April) when mixing occurred throughout the epipelagic and the surface was cooler and well mixed. Primary production was highest during the spring mixing period, evidenced both by in-situ fluorescence (Figure 2D) and productivity incubation experiments (Figure 3A). Originating near the surface, the productivity peak moved deeper throughout the spring and declined into the summer (Figure 2D). However, there was a notable, yet smaller productivity bump in the late summer and early fall (Figure 3A) which occurred deeper in the epipelagic (Figure 2D). The particle concentration (184$\mu m$ - 450$\mu m$) was closely coupled to chlorophyll-a patterns.

![Environmental profiles across time-series of study period. Y axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion (184 - 450$\mu m$). (F) Bacteria Abundance. (G) Silica. (H) Nitrate](images/02-environmental.pdf)

Overall, particle concentration was high near the surface during the 2021 spring bloom, then sank throughout the water column attenuating throughout the lower epipelagic (Figure 2E). Similarly, bacteria abundance was closely linked to overall productivity, althrough there was a more consistent layer near the top of the mesopelagic (\~250m) (Figure 2F). Concurrent with the secondary fall production peak, there was also higher particle concentration and bacterial abundance in the later summer and early fall. Interestingly, while primary productivity estimates from July-August were not that different between 2019 and 2021 (Figure 4A), chlorophyll-a florescence, particle concentration, and bacterial abundance were much higher in 2019's summer/fall (Figure 2D-F). Inorganic nutrients ($Si$ and $NO_3$) were generally well stratified, with low concentrations in the epipelagic and increasing throughout the mesopelagic. However, both nutrients did vary vertically in accordance with the 2021 February downwelling and spring mixing period (Figure 2G-H). Additionally, in the late fall of 2021, $Si$ concentrations were slightly elevated in the mid-mesopelagic (Figure 2G).

Overall mass flux to the mesopelagic was highest during the 2021 February downwelling (Figure 3B). Generally, export was similarly high during March, declining in April then increasing slightly throughout the summer and early fall. While magnitude was slightly different, this pattern was consistent with total mass, carbon and nitrogen fluxes (Figure 3B-D). Higher mass, carbon and nitrogen flux also occurred in the 2019 late summer - early fall period.

![Primary productivity (A) and flux estimates from total mass (B), carbon (C), and nitrogen (D). Values are from monthly cruises with month displayed in corresponding colors. Absent data are shown by grey bar.](images/03_prod-flux.pdf)

### Rhizaria abundance and distribution

Across all imaged mesozooplankton (\>900$\mu m$), Rhizaria comprised a considerable fraction of the total community. Considering the total abundances of the observational period, Rhizaria comprised on average, 42.6% of all mesozooplankton abundance (Supplemental Figure 3). Copepods were the second most abundant, comprising 35.5% and all living mesozooplankton were 22%. The large contribution of Rhizaria to the mesozooplankton community is most prominent in the epipelagic, where they accounted for 47% of all mesozooplankton. In the mesopelagic rhizaria were a smaller fraction, at 38% in the upper layers (200-500m) and 37% in the deeper mesopelagic (500-1000m).

![Average abundance of Rhizaria in 25m bins, across entire study period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C), Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae (G), Coelodendridae (H).](images/04_vertical.pdf)

Total average Rhizaria abundance had a bimodal distribution with respect to depth. Total abundance was highest just below the surface (0-100m), with secondary, wider peak occurring in the mid mesopelagic (Figure 4A). Although, variation in depth binned abundance was large, likely due to seasonal variability but also increased from the detection-risk described in the methods. The vertical distribution pattern and abundance varied considerably across taxonomic groups. Radiolarians were some fo the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5B). This pattern was led by Collodaria, whose colonies were abundant in the upper epipelagic and declined into the top of the mesopelagic (Figure 4C). Acantharea displayed a bimodal distribution accounting for a large portion of the total Rhizaria pattern (Figure 4B, Figure 5). Foraminifera had a similar bimodal distribution, yet their overall average densities were much lower and spread wider throughout the mesopelagic (FIgure 4E). Phaeodarian families exhibited a wide range of vertical distribution patterns. The most abundant, Aulacanthidae, also had a bimodal pattern but the density was highest in the lower mesopelagic (Figure 4D). Aulosphaeridae had low average density and was nearly homogenously distributed throughout the water column, although slightly lower int eh epipelagic (Figure 4F). Castanellidae were the only Phaeodarian who appeared to be effectively restricted to the photic zone (Figure 4G). Alternatively, Coelodendridae primarily occurred in the lower mesopelagic (Figure 4H). A few individuals from the families Tuscaroridae and Medusettidae were also observed in the mesopelagic, yet they were much rarer (data not shown).

Between the monthly cruises, Rhizaria integrated abundance varied in the epipelagic. Highest average abundance occurred in June 2021 and was lowest during the winter months (Figure 5A). The 2019 later summer - fall period also had much higher integrated abundance than similar months in 2021. While the majority of integrated abundance in the epipelagic was consistently attributable to Collodaria, Acanthrean abundance occurred sporadically and could account for a large portion of the total in some months (Figure 5B). The mesopelagic integrated abundance was much more consistent across monthly cruises, although average abundance was notably higher in 2019 (Figure 5C-F). The community composition in the mesopelagic was more diverse, mostly comprised of Phaeodarians. However, Acantharea and unidentified Rhizaria also were common members of the community (Figure 5D, 5F).

![Seasonality of Rhizarian integrated abundance for the epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower mesopelagic (500-1000m) (E-F). Left panels (A,C,E) display total integrated abundance per monthly cruise colored by month. Right panels (B, D, F) display community composition of each total abundance.](images/05_seasonality.pdf)

### Environmental Drivers of Rhizarian Abundance

For total Rhizarian integrated abundance the GAMs produced moderate fits ($R^2_{adj}$ = 0.406-0.603) (Table 1). In the epipelagic, there were several significant predictor variables including inorganic nutrients ($NO_3$ and $Si$), water quality parameters (Salinity, DO), primary production, and particle concentration (Table 1). However, the upper and lower mesopelagic were exclusively explained by particle-related variables (concentration and mass flux) (Table 1).

| **Model**                          | Term                   | edf    | F     | p-value |
|------------------------------------|------------------------|--------|-------|---------|
| **All Rhizaria Epipelagic**        | Salinity               | 2.0818 | 4.654 | \<0.001 |
| $R^2_{adj}$= 0.42                  | DO                     | 0.8272 | 0.945 | 0.0136  |
|                                    | Silica                 | 3.1206 | 15.75 | \<0.001 |
|                                    | NO3                    | 2.9970 | 6.268 | \<0.001 |
|                                    | Primary Productivity   | 1.8367 | 2.099 | \<0.001 |
|                                    | Particle Concentration | 0.9282 | 2.566 | \<0.001 |
|                                    |                        |        |       |         |
| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6847 | 0.431 | 0.073   |
| $R^2_{adj}$= 0.406                 | Avg Mass Flux          | 1.5426 | 0.923 | 0.045   |
|                                    | Particle Concentration | 3.2957 | 20.89 | \<0.001 |
|                                    |                        |        |       |         |
| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 1.6632 | 1.824 | 0.002   |
| $R^2_{adj}$= 0.603                 | Particle Concentration | 0.7694 | 0.662 | 0.027   |

: Generalized Additive Model results for integrated total Rhizarian abundance in different regions of the water column.

GAMs for individual taxa were much less consistent in their fits (Table 2). This is likely in part due to the high number of non-observations for certain taxa. Note that due to low abundances, GAMS were not constructed for Tuscaroridae or Medusettidae. Furthermore no significant terms were found for a model with Aulosphaeridae in the epipelagic nor Foraminifera in the mesopelagic.

![Partial effects of smooth terms in taxa-specific GAM models from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea (A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae (E).](images/06_epi-partials.pdf)

Epipelagic Acantharea were explained by several predictor vairables and had a good fit ($R^2_{adj}$ = 0.53, Table 2). Most notable smooths however, were mass flux and particle concentration, which had a weak positive association (Figure 6A), with July 2021 as a clear outlier where Acantharean abundances were high in the epipelagic despite lower fluxes and particle concentrations (Figure 5). Foraminifera had a good fitting GAM in the epipelagic ($R^2_{adj}$ = 0.445). There were several significant explanatory variables, althoguht the clearest pattern was observed of high temperatures associated with more Foraminifera abundance (Table 2, Figure 6B). Epipelagic Aulacanthidae similarly had several predictor variables which were significant, including both water quality parameters and particle/flux predictors (Table 2). Interestingly, Aulacanthidae had primary production as a significant predictor, yet there was not a clear association (Figure 6C). There was a fit for Collodaria in the epipelagic ($R^2_{adj}$ = 0.16), although there was a logit-like relationship where higher abundances tended to occur during higher DO conditions in the surface waters (Figure 6D). Castanellidae also had similarly poor fits in the epipelagic ($R^2_{adj}$ = 0.124) (Table 2, Figure 6E).

| Model                                | Term                   | edf   | F     | p-value |
|--------------------------------------|------------------------|-------|-------|---------|
| **Acantharea Epipelagic**            | Salinity               | 2.264 | 4.113 | \<0.001 |
| R2adj=0.53                           | O2                     | 2.579 | 6.712 | \<0.001 |
|                                      | Avg Mass Flux          | 4.810 | 22.81 | \<0.001 |
|                                      | Bacteria #/L           | 1.599 | 1.317 | 0.0117  |
|                                      | Particle Concentration | 0.853 | 1.158 | 0.0087  |
|                                      |                        |       |       |         |
| **Acantharea Upper Mesopelagic**     | Avg C Flux             | 1.296 | 0.621 | 0.0439  |
| R2adj=0.231                          | Avg N Flux             | 1.619 | 1.283 | 0.0026  |
|                                      | Particle Concentration | 0.952 | 3.886 | \<0.001 |
|                                      |                        |       |       |         |
| **Acantharea Lower Mesopelagic**     | Temperature            | 2.076 | 4.155 | \<0.001 |
| R2adj=0.509                          | Avg N Flux             | 1.494 | 1.216 | 0.0113  |
|                                      | Primary Productivity   | 0.766 | 0.648 | 0.0253  |
|                                      | Particle Concentration | 2.037 | 10.03 | \<0.001 |
| **Aulacanthidae Epipelagic**         | Salinity               | 0.792 | 0.757 | 0.0241  |
| R2adj=0.251                          | Avg N Flux             | 0.869 | 1.324 | 0.0018  |
|                                      | Primary Productivity   | 2.312 | 7.704 | \<0.001 |
|                                      | Particle Concentration | 2.008 | 2.200 | 0.0017  |
|                                      |                        |       |       |         |
| **Aulacanthidae Upper Mesopelagic**  | Particle Concentration | 2.832 | 9.802 | \<0.001 |
| R2adj = 0.158                        |                        |       |       |         |
|                                      |                        |       |       |         |
| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.622 | 1.991 | 0.002   |
| R2adj=0.298                          | Particle Concentration | 2.123 | 6.706 | \<0.001 |
|                                      |                        |       |       |         |
| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.653 | 13.06 | \<0.001 |
| R2adj=0.2                            |                        |       |       |         |
|                                      |                        |       |       |         |
| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 0.972 | 6.248 | \<0.001 |
| R2adj=.147                           |                        |       |       |         |
|                                      |                        |       |       |         |
| **Castanellidae Epipelagic**         | Avg C Flux             | 1.462 | 0.816 | 0.0421  |
| R2adj = 0.124                        | Primary Productivity   | 0.771 | 0.665 | 0.0247  |
|                                      | Particle Concentration | 3.956 | 4.623 | \<0.001 |
|                                      |                        |       |       |         |
| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.822 | 0.919 | 0.0183  |
| R2adj=.113                           | Particle Concentration | 0.970 | 6.208 | \<0.001 |
|                                      |                        |       |       |         |
| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 1.773 | 4.873 | \<0.001 |
| R2adj=.133                           |                        |       |       |         |
|                                      |                        |       |       |         |
| **Collodaria Epipelagic**            | Salinity               | 0.925 | 2.267 | \<0.001 |
| R2adj=0.16                           | O2                     | 2.015 | 2.217 | 0.002   |
|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.002   |
|                                      |                        |       |       |         |
| **Foraminifera Epipelagic**          | Temperature            | 4.414 | 8.789 | \<0.001 |
| R2adj=0.445                          | O2                     | 1.603 | 1.287 | 0.0111  |
|                                      | Avg N Flux             | 2.512 | 3.396 | \<0.001 |
|                                      | Primary Productivity   | 0.824 | 0.920 | 0.0153  |
|                                      | Particle Concentration | 0.919 | 2.257 | \<0.001 |

: Taxa-specific generalized additive models for different regions of the water column.

![Partial effects of smooth terms in taxa-specific GAM models from the upper mesopelagic (200-500m). Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/07_upmeso-partial.pdf)

In the upper mesopelagic (200-500m), abundances were generally low (Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae, Aulosphaeridae, and Coelodendridae (Table 2). All these models had generally poor fits ($R^2_{adj}$ \< 0.25). Yet, for all upper mesopelagic models, particle concentration was a significant explanatory variable (Table 2, Figure 7). Carbon flux was significant for Acantharea and nitrogen flux was significant for both Acantharea and Coelodendridae (Table 2, Figure 7A,D). The lower mesopelagic also had generally poor GAM fits for taxa specific models ($R^2_{adj}$ \<0.3), with the exception of Acantharea ($R^2_{adj}$ = 0.509). Acantharea in the lower mesopelagic was most clearly positively associated with particle concentration and nitrogen flux, as well as temperature to a slight degree (Figure 8A). For all Phaeodarians with a significant model, particle concentration was a main predictor variable (Table 2, Figure 8B-D). Aulacanthidae had the best fitting model of the Phaeodarians ($R^2_{adj}$ = 0.298), which also included mass flux as a statistically significant smooth (Figure 8B).

![Partial effects of smooth terms in taxa-specific GAM models from the lower mesopelagic (500-1000m). Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/08_lomeso-partials.pdf)

### 2024-01-25 manuscript/04_discussion.qmd
finalize figs for draft

---
bibliography: references.bib
---

# Discussion

### Overall Rhizarian abundance and patterns

In the epipelagic Rhizaria exhibited a notable seasonal pattern. Rhizarian abundances were higher in the summer months and lower during the winter. During a prior time period, @blanco-bercial2022 noted that there is considerable seasonality in the community composition of all protists. Despite the seasonality of total Rhizarian abundance, community composition was relatively consistent, with Collodaria representing the bulk of the community. It should be noted that the overall taxonomic resolution of the UVP5 is fairly low, so there may be a switching of species within the broad groups identified in this study which were not captured. Throughout the mesopelagic, month-to-month variation in 2021 was relatively low. Again this is consistent with observations from metabarcoding of the whole protist community in the same study region [@blanco-bercial2022]. This finding is not surprising as the overall seasonal variation in environmental conditions in this region were low.

Overall Rhizaria were the most commonly identified group of mesozooplankton throughout the study period. We do note that the UVP5 commonly captures Trichodesmium colonies, yet these were excluded in this comparison as they are strictly autotrophs. It should be noted that previous work has suggested that avoidance behavior with the UVP is possible, at times likely, for visual and highly mobile zooplankton [@barth2022]. Thus, the percent contribution reported here (42.7%) of Rhizaria to the total mesozooplankton community may be inflated due to under sampling of organisms such as Euphausiids and Chaetognaths which has quick escape responses. Regardless, it is worth noting that in the same region, with data collected in 2012 and 2013 using similar calculation methods, @biard2016 estimated Rhizaria only contribute 15% of the total mesozooplankton community in the upper 500m. Likely, Rhizaria display considerable interannual variability. In the present study, we noticed considerably higher Rhizarian abundance throughout the water column in 2019 compared to 2021. While this may have been driven by increased mass flux (Figure 3), more information is needed to truly understand the magnitude by which Rhizaria can vary interannually.

### Vertical Structure and Trophic Roles

In this study we present a clear pattern of vertical zonation between different Rhizaria groups. Largely, the taxonomic composition and vertical positioning were similar to Rhizaria zonation in the California Current Ecosystem [@biard2020]. It should be noted however, that the secondary abundance peak reported in the present study is lower. This is likely due to the more oligotrophic nature of the study region, were the euphotic zone penetrates deeper into the water column. Most prevalent in the epipelagic were Collodaria. These mixotrophic Radiolarians have long been reported to contribute to primary productivity in the euphotic zone [@michaels1995; @dennett2002]. Collodaria are thought to be particularly successful globally in oligotrophic regions due to their photosymbiotic relationships [@biard2016; @biard2017]. We observed the highest abundance of Collodaria during June 2021, supporting the notion they can thrive during the typically low-nutrient conditions of summer stratification. However, Collodaria also increased during the spring mixing period, suggesting that they can thrive during conditions which may typically be thought to favor autotrophs. Furthermore, while Collodaria were primarily absent from below 250m, there were a few instances of deeper colonies being observed. Global investigations of polycistine flux, suggest that deep-Collodaria in Oligotrophic regions may be a consequence of isothermal submersion [@boltovskoy2017]. Another effectively exclusively epipelagic Rhizaria was the Phaeodarian family of Castanellidae. All Phaeodarians are thought to be fully heterotrophic [@nakamura2015], nonetheless a number of studies, including this one, report Castanellidae to be typically found in the lower epipelagic [@zasko2005; @biard2018; @biard2020]. It should be considered that perhaps Castanellidae specializes in feeding sinking particles directly at the base of the epipelagic. Given it's smaller size [@nakamura2015], Castanellidae does not need a large diameter to efficiently flux feed at the typically particle rich region of the lower epipelagic. Both Castanelldiae and Collodaria had poor fitting GAMs. This is somewhat of a surprise for Collodaria who had large abundance. However, given the consistency of their abundance, it may be that this study did not capture a wide enough range of conditions for describing Collodaria's preferred niche.

The mesopelagic generally was home to known heterotrophic organisms, particularly for those which were constrained to exclusively occupy deeper waters. This is consistent with @blanco-bercial2022's observation of an auto-/mixotroph to heterotroph gradient in the protist community. The upper mesopelagic interestingly had relatively low total abundance. This low-abundance region likely reflects the dynamics of productivity and export throughout the water column. While productivity and thus sinking particles for flux feeders are high in the euphotic zone, much of this is attenuated throughout the epipelagic. So, while the base of the epipelagic may provide a rich feeding environment for Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F), the region from 200-500m might be otherwise food poor. Perhaps it is more advantageous for Rhizaria to situate deeper, in darker regions of the twilight zone. Also it should be noted that Phaeodarians utilize silica to build their opaline tests, and silica concentrations started to increase around 500m (Figure 2G). Although $Si$ was not a significant smooth for any taxa-specific model, this lack of association might be due to the overall lack of variation of $Si$ between integrated abundance of each cast. Aulosphaeridae was only found to have significant relationships, although weak fits, to particle concentration in the mesopelagic. In our study, while consistently observed, overall abundances of Aulosphaeridae were very low. In the Pacific Ocean, on California's Coast, much higher abundances of Aulosphaeridae have been reported [@zasko2005; @biard2020] and they have massive potential to impact silica export [@biard2018]. Coelodendridae were also seemingly restricted to the deeper section of the mesopelagic. This is interesting given that in the California Current, [@biard2020] found a bimodal distribution in Coelodendridae. There are several morphotypes corresponding to different taxa of Coelodendridae [@nakamura2015; @biard2020]. So it may be that only a few types of Coelodendridae were observed in this study, while the epipelagic variety was not. Alternatively, the lower epipelagic of the California Current may provide adequate habitat for Coelodendridae, which is not available in the oligotrophic Sargasso Sea.

A number of taxa were found to have a bimodal distribution, with considerable populations in both the epipelagic and mesopelagic. Aulacanthidae had a bimodal distribution, although abundances were highest in the lower mesopelagic. Foraminifera also had a bimodal distribution. Some lineages of Foraminifera are known to host photosymbionts [@kimoto2015; @biard2022], however they are also efficient predators commonly seen throughout the mesopelagic [@caron1984; @gaskell2019]. Thus it is not surprising to find their presence in both locations of the water column. Foraminifera are also known to vary their vertical distribution across their life cycle in phase with lunar cycles [@bijma1990; @kimoto2015; @gaskell2019; @biard2022]. However, the sampling scheme of the BATS program does not capture this frequency and was not investigated in the present study.

Acantharea also had a bimodal distribution, with much larger abundances than Aulacanthidae or Foraminifera. Most prior studies of Acantharea vertical distribtion found them concentrated in near surface layers of the water column [@michaels1988 @zasko2005; @biard2020]. This would support the paradigm that large Acantharean abundances may be supported by their mixotrophic abilities [@michaels1995; @suzuki2015]. While the UVP5 images cannot distinguish between mixotrophic and heterotrophic Acantharea, the GAMs constructed for Acantharean abundance found positive associations with particle concentration and mass flux, suggesting a higher reliance on heterotrophy. Recently @marsbrisbin2020 described apparent predator behavior amongst near-surface Acantharea. Thus it is likely that epipelagic Acantharea may commonly be heterotrophic. Yet, it should be noted in the Sargasso Sea, both heterotrophic and symbiotic lineages of Acantharea have been reported [@blanco-bercial2022]. Additionally, @michaels1988 noted that the majority of Acantharea (by abundance) were smaller than 160$\mu m$. While that estimate may be inflated due to inability to capture larger cells, small Acantharea were not captured in the present study. Thus, trophic strategy may shift based on sizes of Acantharea.

@decelle2013 proposed a hypothetical life cycle for cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life cycle suggests that epipelagic Acantharea are adult populations, which form cysts that sink into the mesopelagic, then reproduce and raise. Furthermore, given that horizontal transfer of symbionts between generations of Acantharea is unlikely, the newly spawned mesopelagic Acantharea are not necessarily required to rapidly return to the photic zone. This hypothesis predicts that Acanthareans in the mesopelagic would be smaller [@decelle2013]. @marsbrisbin2020 provided some support for this hypothesis, with a significant decrease in Acantharea sizes with depth. Although, the authors also observed low abundances in the mesopelagic and noted that the smaller sizes may be due to less food availability [@marsbrisbin2020]. Since food is more scarce in the mesopelagic there is is less nutritional quality, yet flux feeders would likely grow larger to increase their feeding range [@biard2020]. In the data collecting in this study, Acanthareans in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth (Figure 9, Wilcoxon rank sum text p-value \< 0.001). This provides added support for the hypothesis that cyst-forming Acantharea may utilize different sections of the water column throughout their life cycle. However to further investigate this, more work is needed with higher temporal and taxonomic resolution.

![Comparison of average sizes (ESD) amongst Rhizarian taxa which occurred throughout the water column.](images/09_size-comp.pdf)

### Conclusions and Considerations

This study provides a detailed look at Rhizarian abundance over time throughout the water column in a major oligotrophic gyre. We show that their abundances are generally related to particle concentration and flux, although lack of environmental variability may have reduced the fit of our GAMs. Considering the potential role of Rhizaria in the biological carbon pump, they may have a somewhat mixed role. In the shallower regions, where smaller Rhizaria are abundant, they may be an attenuating force on sinking particles [@stukel2019]. It should be noted that in our study, we focused on the "small" particle concentration field (\<450$\mu m$), and these particles are generally slower sinking than large particles. However, once consumed and repackaged by larger Rhizaria, they can sink quicker and contribute more to overall flux [@michaels1988]. Thus, Rhizaria may act as a ballasting mechanism. However, this is largely speculation, to truly test this, more work is needed measuring Rhizarian flux.

The vertical partitioning documented in this study do support the hypothesis that mixotrophic rhizaria will occupy shallower waters while deeper waters are dominated by heterotrophy. However the degree to which mixotrophic Rhizaria in the euphotic zone rely on heterotrophy versus symbiosis is uncertain. Collodaria were recorded as consistent and dominant members of the near surface region. These organisms have the potential to contribute considerably to the otherwise low productivity of oligotrophic regions. However, their role in food webs is not well understood. While this study represents a step-forward in our understanding of Rhizarian ecology, there are several known aspects of Rhizarian and protist ecology which are not captured in this study. Continued research on Rhizaria is much needed to better understand their ecology. Particularly extended descriptive work to capture interannual patterns. Also work defining biotic interactions, feeding rates, productivity, and life history are all rich fields of interest in Rhizaria.

### 2024-01-25 manuscript/main.qmd
finalize figs for draft

@@
   pdf:
-    # template-partials:
-    #   - "before-body.tex"
-    #   - "_titlepage.tex"
-    # include-in-header:
-    #   - "in-header.tex"
+    template-partials:
+      - "before-body.tex"
+      - "_titlepage.tex"
+    include-in-header:
+      - "in-header.tex"
     number-sections: false
@@
     tbl-cap-location: top
-    # filters: [wordcount.lua]
+    filters: [wordcount.lua]
     keep-tex: true
@@
 ```
+
+```{r}
+#| child: ./02_methods.qmd
+```
+
+```{r}
+#| child: ./03_results.qmd
+```
+
+```{r}
+#| child: ./04_discussion.qmd
+```
+
+# References:
+:::{#refs}
+:::

### 2024-01-25 manuscript/main.tex
finalize figs for draft

@@
 \documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
+]{article}
 
@@
 
-\KOMAoption{captions}{tableheading}
+\usepackage{hyphenat} % don't hyphenat titles
+\usepackage{graphicx}
+\usepackage{lineno}\linenumbers
+
+% Add any tex header commands here
 \makeatletter
@@
 \begin{document}
-\maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable]}{\end{tcolorbox}}\fi
+    \begin{titlepage}
+% This is a combination of Pandoc templating and LaTeX
+% Pandoc templating https://pandoc.org/MANUAL.html#templates
+% See the README for help
+
+\raggedleft % left align the title page
+\rule{1pt}{\textheight} % Vertical line
+\hspace{0.05\textwidth} % Whitespace between the vertical line and title page text
+% Adjust num before \textwidth to move the block left or right
+\begin{minipage}[b][\textheight][s]{0.85\textwidth}
+
+\raggedright
+% Title and subtitle
+{\large\bfseries\nohyphens{Rhizaria in the oligotrophic exhibit clear
+temporal and vertical variability.}}\\[2\baselineskip]
+
+  % Authors	
+% This hairy bit of code is just to get "and" between the last 2
+% authors. See below if you don't need that
+ {\large{Alex Barth*}}{\textsuperscript{1}}%
+%
+, 
+ {\large{Leo Blanco-Berical}}{\textsuperscript{2}}%
+%
+, 
+ {\large{Rod Johnson}}{\textsuperscript{3}}%
+%
+%
+{ and \large{Joshua Stone}}%
+{\textsuperscript{1}}%
+%
+
+% This is how to do it if you don't need the "and"
+
+  %%%%%% Affiliations
+\vspace{2\baselineskip}
+
+\hangindent=1em
+\hangafter=1
+%
+{1}.~{University of South Carolina}%
+%
+, %
+{Biological Sciences}%
+%
+%
+, %
+{700 Sumter St 401, Columbia, SC 29208}%
+%
+\par\hangindent=1em\hangafter=1%
+%
+{2}.~{Arizone State University}%
+%
+%
+\par\hangindent=1em\hangafter=1%
+%
+{3}.~{Bermuda Institute of Ocean Sciences}%
+%
+%
+, %
+{Bermuda GE 01, UK}%
+%
+
+  %%%%%% Correspondence
+\vspace{1\baselineskip}
+
+  %use \vfill instead to get the space to fill flexibly	
+%\vspace{0.25\textheight} % Whitespace between the title block and the publisher
+
+\vfill
+
+%%%%%% Cover image
+
+  % Whitespace between the title block and the tagline
+\vspace{0.1\textheight}
+
+\end{minipage}  \end{titlepage}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 the global ocean. While the taxonomy of these organisms has recently
-undergone several reclassifications (Wiley, 2022), their presence in
+undergone several reclassifications (Biard, 2022a), their presence in
 ocean ecosystems has been long known to oceanographers. Some of the
@@
 several mm in diameter) and they construct intricate mineral skeletons
-out of either silica, strontium, or calcium carbonate (Kimoto, 2015;
-Nakamura and Suzuki, 2015; Suzuki and Not, 2015; Wiley, 2022). Despite
+out of either silica, strontium, or calcium carbonate (Biard, 2022a;
+Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki and Not, 2015). Despite
 their noticeable morphology and global distribution, Rhizaria were
@@
 al., 2009; Biard et al., 2015; Cavalier-Smith et al., 2018; Decelle et
-al., 2013, 2012; rev by Wiley, 2022). DNA metabarcoding studies have
+al., 2013, 2012; rev by Biard, 2022a). DNA metabarcoding studies have
 revealed insights into the distributional patterns Llopis Monferrer et
@@
 (Decelle et al., 2013). Still, Rhizarian ecological roles are not well
-understood (Wiley, 2022). This is a major challenge as it is critical to
-understand the ecological role of plankton to fully incorporate them
+understood (Biard, 2022a). This is a major challenge as it is critical
+to understand the ecological role of plankton to fully incorporate them
 into biological oceanographic models.
@@
 due to the fact different taxa can exhibit every different trophic
-modes. As zooplankton, rhizaria are predominately heterotrophic (Wiley,
-2022), yet their feeding modes can be quite varied. Phaeodarians (family
-Cercozoa) are largely thought to be flux-feeders, collecting and feeding
-on sinking particles (Nakamura and Suzuki, 2015; Stukel et al., 2019).
-Alternatively, Retaria can be either exclusively heterotrophic or
+modes. As zooplankton, rhizaria are predominately heterotrophic (Biard,
+2022a), yet their feeding modes can be quite varied. Phaeodarians
+(family Cercozoa) are largely thought to be flux-feeders, collecting and
+feeding on sinking particles (Nakamura and Suzuki, 2015; Stukel et al.,
+2019). Alternatively, Retaria can be either exclusively heterotrophic or
 mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
@@
 Radiolaria have several lineages all which have some taxa who are well
-known to host symbionts (Biard, 2022). Amongst radiolarians, arguably
+known to host symbionts (Biard, 2022b). Amongst radiolarians, arguably
 the most widespread are Collodaria who can be either large solitary
@@
 and Anderson, 1981). All known Collodaria species host dinoflagellate
-symbionts (Biard, 2022) and can contribute substantially to primary
+symbionts (Biard, 2022b) and can contribute substantially to primary
 productivity, particularly in oligotrophic ocean regions (Caron et al.,
@@
 2016). A few Acantharean (Radiolarian order) clades host algal symbionts
-(Biard, 2022; Decelle et al., 2012), notably with two clades forming an
+(Biard, 2022b; Decelle et al., 2012), notably with two clades forming an
 exclusive relationship with Phaeocystis. However, globally Acantharea
-are less abundant than Collodaria (Wiley, 2022) and contribute less to
+are less abundant than Collodaria (Biard, 2022a) and contribute less to
 total primary productivity (Michaels et al., 1995). This may be due to
 the fact several clades of Acantharea are cyst-forming and strictly
-heterotrophs (Biard, 2022; Decelle et al., 2013). Furthermore, Mars
+heterotrophs (Biard, 2022b; Decelle et al., 2013). Furthermore, Mars
 Brisbin et al. (2020) documented apparent predation behavior in
@@
 column. Yet, few studies have made direct attempts to relate rhizaria
-abundances to abiotic environmental factors (Biard and Ohman, 2020). In
-part, this is due to the fact few studies have been able to sample
-Rhizaria in the same location over a consistent timeframe (Boltovskoy et
-al., 1993; Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels
-et al., 1995; Michaels, 1988). Furthermore, no studies have utilized
-imaging, arguably the best method for quantifying rhizaria, consistently
+abundances to environmental factors (Biard and Ohman, 2020). In part,
+this is due to the fact few studies have been able to sample Rhizaria in
+the same location over a consistent timeframe (Boltovskoy et al., 1993;
+Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels et al.,
+1995; Michaels, 1988). Furthermore, no studies have utilized imaging,
+arguably the best method for quantifying rhizaria, consistently
 throughout the full mesopelagic. Given this lack of information, there
@@
 This hypothesis makes several predictions, including vertical zonation,
-as seen in prior studies, but also that abiotic variables related to
-trophic strategy will explain abundance patterns. Specifically,
+as seen in prior studies, but also that environmental variables related
+to trophic strategy will explain abundance patterns. Specifically,
 autotrophic/mixotrophic taxa will correspond to variables related to
@@
 (particle concentration, flux, and low \(O_2\)).
+
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+\hypertarget{oceanographic-sampling}{%
+\subsection{Oceanographic Sampling}\label{oceanographic-sampling}}
+
+Data were collected in collaboration with the Bermuda Atlantic
+Time-Series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
+the R/V Atlantic Explorer. Cruises were conducted at approximately
+monthly intervals. Rhizaria individuals were sampled using the
+Underwater Vision Profiler 5 {[}UVP5; Picheral et al. (2010){]}, a tool
+which is well established to accurately quantify large Rhizaria (Barth
+and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago et
+al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023; Stukel
+et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the sampling
+rosette and collected data autonomously on routine casts, from which
+only the downcast data are utilized. The UVP5 was deployed from
+June-September 2019 then from October 2020 - January 2022. Casts were
+filtered to only include data collected in the BATS region, far offshore
+of Bermuda into the Sargasso Sea (Supplemental Figure 1). In general,
+casts extended to either 200m, 500m, or 1200m deep, with a few extended
+into the bathypelagic (4500m). However, Rhizaria were only typically
+found in large abundances throughout the epipelagic and mesopelagic
+zones. As such, we limit this study to results from the upper 1000m of
+the water column.
+
+A variety of biotic and abiotic data were collected during each BATS
+cruise. Briefly, we will explain the data utilized in this study. The
+UVP5 provided particle count data at a high-frequency from each cast.
+Particle concentration was calculated from this data for all particles
+184\(\mu m\) - 450\(\mu m\). The lower size range was set by what could
+be reliably sampled by the UVP5's pixel resolution (\textgreater2px) and
+the upper size range is representative of a potential prey field for
+mesozooplankton (Whitmore and Ohman, 2021). Salinity, temperature,
+Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were
+measured at high-frequencies on each cast, attached to the same CTD
+rosette as the UVP. On certain casts, niskin bottles were used to
+collect bacterial abundance estimates (via epifluorescence microscopy)
+as well as measure inorganic nutrients (\(NO_3\), and \(Si\) as
+silicate/sicilic acid) at discrete depths. On each cruise, flux
+estimates of total mass, carbon, and nitrogen were also collected using
+sediment traps (in the present study we utilize flux to the mesopelagic
+as the flux at 200m). Also primary productivity was estimated through
+measuring \(C^{14}\) uptake rates from in-situ incubations. Full
+descriptions of the BATS sampling program, methods, and data are
+available freely online
+\href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}
+and reviewed in (Lomas et al., 2013).
+
+Environmental data were processed in a variety of ways to match the
+format of the Rhizaria abundance estimates (See below). CTD data were
+collected at higher frequency than the UVP, so these data were averaged
+within matching bins to the UVP5 data. Data from niskin bottles were
+first linearly interpolated with a 1m resolution for each cast where
+data were available. Then, because data were collected at different
+frequencies over the course of a cruise, the interpolated data were
+averaged across casts, then averaged into matching UVP5-sized bins.
+Primary productivity estimates were totaled within the euphotic zone to
+represent a ``total euphotic productivity''. Finally the data from the
+sediment trap deployed to 200m was used to represent ``flux to the
+mesopelagic''.
+
+\hypertarget{rhizaria-imaging-processing-and-quantification}{%
+\subsection{Rhizaria imaging processing and
+quantification}\label{rhizaria-imaging-processing-and-quantification}}
+
+Individual vignettes of Rhizaria images were identified using the
+classification platform Ecotaxa {[}cite{]}. Data were pre-sorted
+according to a . Taxonomic classification were done based on morphology
+exclusively. While there are sparse taxonomic guides for in-situ images
+of rhizaria, identification largely relied on descriptions in (Nakamura
+and Suzuki, 2015; Suzuki and Not, 2015; and Biard and Ohman, 2020).
+Using the aforementioned sources and publicly available ecotaxa
+projects, I constructed a guide which can be found here:
+\url{https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf}.
+Broadly, Rhizaria were classified as Foraminifera, Radiolarians
+(Acantharea or Collodaria), or a variety of Phaeodarian families (Figure
+1). When identification could not be confidently made between a few
+candidate taxa, a less specific label was used. As a result, we have
+data from ``unidentified Rhizaria'', which typically were vignettes not
+distinguishable between Aulacanthidae or Acantharea or ``unidentified
+Phaeodaria'', which are clearly phaeodaria but not distinguishable into
+a family.
+
+\begin{figure}
+
+{\centering \includegraphics{images/01_taxa.pdf}
+
+}
+
+\caption{Example images of different Rhizaria taxa. 4mm scale bar shown
+in lower right. All vignettes are same scale}
+
+\end{figure}
+
+The UVP5 samples at \textasciitilde15Hz rate as it descends the water
+column and records the exact position of each particle larger than
+600\(\mu m\). However, identified rhizaria ranged from a 934\(\mu m\)
+Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To
+confirm that the UVP5 was sampling adequately across all size ranges, an
+NBSS slope was constructed to identify a drop-off which would indicate
+poor-sampling at the small size range (Barth and Stone, in review;
+Lombard et al., 2019) . However, it was evident from this analysis that
+all size ranges were adequately sampled across the size range
+(Supplemental Figure 2) so no data were excluded. The UVP5 reports the
+exact depth at which a particle is recorded, however to estimate
+abundance, observations must be binned over fixed depth intervals. In
+our deployments had variable descent depths and speeds with more casts
+descending to 500m than 1000m and descents quicker through the
+epipelagic than the mesopelagic {[}see Barth and Stone (2022) for an
+extended discussion of UVP5 data processing). For the present study,
+Rhizarian abundances were estimated in 25m vertical bins, which offer a
+moderate sampling volume (average 0.948\(m^3\) in the epipelagic and
+0.589\(m^3\) in the mesopelagic) while still maintaining ecologically
+relevant widths. However, concentrations in a 25m bin would need to be
+greater than 2.428 indv. \(m^{-3}\) and 3.912 indv. \(m^{-3}\), in the
+epipelagic and mesopelagic respectively, to fall below a 10\%
+non-detection risk (Barth and Stone, in review; Benfield et al., 1996).
+Because we typically observed many rhizaria taxa below these
+concentrations, we present the 25m binned data to visualize broad-scale
+average distributions. For quantifying and modelling Rhizaria
+abundances, we present integrated abundance estimates, with each cast.
+Due to the variable descent depths of the UVP, data are categorized as
+epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic
+(500-1000m). The average sampling volume integrated through these
+regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\), with
+non-detection thresholds at 0.30 indv. \(m^{-2}\), 0.33 indv.
+\(m^{-2}\), and 0.20 indv. \(m^{-2}\) respectively. All UVP data
+processing was done using the \texttt{EcotaxaTools} package in R (Barth
+2023).
+
+\hypertarget{modelling-environmental-controls-of-rhizarian-abundance}{%
+\subsection{Modelling environmental controls of Rhizarian
+Abundance}\label{modelling-environmental-controls-of-rhizarian-abundance}}
+
+Generalized Additive Models (GAMs) were used to assess the relationship
+between integrated Rhizarian abundance and different environmental
+factors. GAMs offer the ability to model non-linear and non-monotonic
+relationships, which can be particularly useful in assessing ecological
+relationships (Wood, 2017) and have been successfully applied to
+Rhizarian ecology (Biard and Ohman, 2020). The \texttt{mgcv} package
+(Wood, 2001) was used to construct models relating environmental
+parameters to each taxonomic group's integrated abundance estimates from
+each cast. To select the most parsimonious model for each analysis, a
+backwards step-wise approach was taken. First, a full model was fit
+using any term which may be ecologically relevant. Terms were fit using
+maximum likelihood with a double penalty approach on unnecessary smooths
+(Marra and Wood, 2011). The smoothness parameter was restricted (k = 6)
+to prevent overfitting the models. At each iteration of the backwards
+step-wise procedure, the model term with the lowest F score (least
+statistically significant) was removed. This was repeated until all
+model terms were statistically significant or the \(R^2_{adj}\) was
+substantially reduced. Models were fit for each region; epipelagic,
+upper mesopelagic, and lower mesopelagic. In cases where observations
+were too sparse for a given taxonomic grouping, models were not run.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{environmental-variability}{%
+\subsubsection{Environmental
+Variability}\label{environmental-variability}}
+
+The BATS sampling region is southeast of Bermuda, situated in the North
+Sargasso Sea and the North Atlantic oligotrophic ocean gyre. Due to the
+sampling location, while the environmental conditions are generally low
+in variation and oligotrophic, there is some seasonality and
+considerable influence from mesoscale eddies (Lomas et al., 2013;
+McGillicuddy et al., 1998). Variability in the water column structure
+was visible during the study period (Figure 2). This is best evidenced
+through the temperature profiles; In the late summer and early fall
+there was a stratified water column with high temperatures in the
+surface (\textless75m) (Figure 2A) and slightly elevated salinity
+(Figure 2B). This warm, stratified period appeared more intense during
+the few months sampled in 2019. In 2021, we observed the stratified
+layer slowly dissipated into the winter months. There was a consistent
+oxygen minimum zone (OMZ) located at about 750m deep (Figure 2C).
+Although, February 2021 saw a notable downwelling event, likely due to a
+passing cyclonic eddy. During this time, warmer, oxic water was plunged
+deeper into the mesopelagic. This process was reversed in the spring
+months (March, April) when mixing occurred throughout the epipelagic and
+the surface was cooler and well mixed. Primary production was highest
+during the spring mixing period, evidenced both by in-situ fluorescence
+(Figure 2D) and productivity incubation experiments (Figure 3A).
+Originating near the surface, the productivity peak moved deeper
+throughout the spring and declined into the summer (Figure 2D). However,
+there was a notable, yet smaller productivity bump in the late summer
+and early fall (Figure 3A) which occurred deeper in the epipelagic
+(Figure 2D). The particle concentration (184\(\mu m\) - 450\(\mu m\))
+was closely coupled to chlorophyll-a patterns.
+
+\begin{figure}
+
+{\centering \includegraphics{images/02-environmental.pdf}
+
+}
+
+\caption{Environmental profiles across time-series of study period. Y
+axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved
+Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion
+(184 - 450\(\mu m\)). (F) Bacteria Abundance. (G) Silica. (H) Nitrate}
+
+\end{figure}
+
+Overall, particle concentration was high near the surface during the
+2021 spring bloom, then sank throughout the water column attenuating
+throughout the lower epipelagic (Figure 2E). Similarly, bacteria
+abundance was closely linked to overall productivity, althrough there
+was a more consistent layer near the top of the mesopelagic
+(\textasciitilde250m) (Figure 2F). Concurrent with the secondary fall
+production peak, there was also higher particle concentration and
+bacterial abundance in the later summer and early fall. Interestingly,
+while primary productivity estimates from July-August were not that
+different between 2019 and 2021 (Figure 4A), chlorophyll-a florescence,
+particle concentration, and bacterial abundance were much higher in
+2019's summer/fall (Figure 2D-F). Inorganic nutrients (\(Si\) and
+\(NO_3\)) were generally well stratified, with low concentrations in the
+epipelagic and increasing throughout the mesopelagic. However, both
+nutrients did vary vertically in accordance with the 2021 February
+downwelling and spring mixing period (Figure 2G-H). Additionally, in the
+late fall of 2021, \(Si\) concentrations were slightly elevated in the
+mid-mesopelagic (Figure 2G).
+
+Overall mass flux to the mesopelagic was highest during the 2021
+February downwelling (Figure 3B). Generally, export was similarly high
+during March, declining in April then increasing slightly throughout the
+summer and early fall. While magnitude was slightly different, this
+pattern was consistent with total mass, carbon and nitrogen fluxes
+(Figure 3B-D). Higher mass, carbon and nitrogen flux also occurred in
+the 2019 late summer - early fall period.
+
+\begin{figure}
+
+{\centering \includegraphics{images/03_prod-flux.pdf}
+
+}
+
+\caption{Primary productivity (A) and flux estimates from total mass
+(B), carbon (C), and nitrogen (D). Values are from monthly cruises with
+month displayed in corresponding colors. Absent data are shown by grey
+bar.}
+
+\end{figure}
+
+\hypertarget{rhizaria-abundance-and-distribution}{%
+\subsubsection{Rhizaria abundance and
+distribution}\label{rhizaria-abundance-and-distribution}}
+
+Across all imaged mesozooplankton (\textgreater900\(\mu m\)), Rhizaria
+comprised a considerable fraction of the total community. Considering
+the total abundances of the observational period, Rhizaria comprised on
+average, 42.6\% of all mesozooplankton abundance (Supplemental Figure
+3). Copepods were the second most abundant, comprising 35.5\% and all
+living mesozooplankton were 22\%. The large contribution of Rhizaria to
+the mesozooplankton community is most prominent in the epipelagic, where
+they accounted for 47\% of all mesozooplankton. In the mesopelagic
+rhizaria were a smaller fraction, at 38\% in the upper layers (200-500m)
+and 37\% in the deeper mesopelagic (500-1000m).
+
+\begin{figure}
+
+{\centering \includegraphics{images/04_vertical.pdf}
+
+}
+
+\caption{Average abundance of Rhizaria in 25m bins, across entire study
+period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C),
+Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae
+(G), Coelodendridae (H).}
+
+\end{figure}
+
+Total average Rhizaria abundance had a bimodal distribution with respect
+to depth. Total abundance was highest just below the surface (0-100m),
+with secondary, wider peak occurring in the mid mesopelagic (Figure 4A).
+Although, variation in depth binned abundance was large, likely due to
+seasonal variability but also increased from the detection-risk
+described in the methods. The vertical distribution pattern and
+abundance varied considerably across taxonomic groups. Radiolarians were
+some fo the most abundant taxa observed, particularly in the epipelagic
+(Figure 4, Figure 5B). This pattern was led by Collodaria, whose
+colonies were abundant in the upper epipelagic and declined into the top
+of the mesopelagic (Figure 4C). Acantharea displayed a bimodal
+distribution accounting for a large portion of the total Rhizaria
+pattern (Figure 4B, Figure 5). Foraminifera had a similar bimodal
+distribution, yet their overall average densities were much lower and
+spread wider throughout the mesopelagic (FIgure 4E). Phaeodarian
+families exhibited a wide range of vertical distribution patterns. The
+most abundant, Aulacanthidae, also had a bimodal pattern but the density
+was highest in the lower mesopelagic (Figure 4D). Aulosphaeridae had low
+average density and was nearly homogenously distributed throughout the
+water column, although slightly lower int eh epipelagic (Figure 4F).
+Castanellidae were the only Phaeodarian who appeared to be effectively
+restricted to the photic zone (Figure 4G). Alternatively, Coelodendridae
+primarily occurred in the lower mesopelagic (Figure 4H). A few
+individuals from the families Tuscaroridae and Medusettidae were also
+observed in the mesopelagic, yet they were much rarer (data not shown).
+
+Between the monthly cruises, Rhizaria integrated abundance varied in the
+epipelagic. Highest average abundance occurred in June 2021 and was
+lowest during the winter months (Figure 5A). The 2019 later summer -
+fall period also had much higher integrated abundance than similar
+months in 2021. While the majority of integrated abundance in the
+epipelagic was consistently attributable to Collodaria, Acanthrean
+abundance occurred sporadically and could account for a large portion of
+the total in some months (Figure 5B). The mesopelagic integrated
+abundance was much more consistent across monthly cruises, although
+average abundance was notably higher in 2019 (Figure 5C-F). The
+community composition in the mesopelagic was more diverse, mostly
+comprised of Phaeodarians. However, Acantharea and unidentified Rhizaria
+also were common members of the community (Figure 5D, 5F).
+
+\begin{figure}
+
+{\centering \includegraphics{images/05_seasonality.pdf}
+
+}
+
+\caption{Seasonality of Rhizarian integrated abundance for the
+epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower
+mesopelagic (500-1000m) (E-F). Left panels (A,C,E) display total
+integrated abundance per monthly cruise colored by month. Right panels
+(B, D, F) display community composition of each total abundance.}
+
+\end{figure}
+
+\hypertarget{environmental-drivers-of-rhizarian-abundance}{%
+\subsubsection{Environmental Drivers of Rhizarian
+Abundance}\label{environmental-drivers-of-rhizarian-abundance}}
+
+For total Rhizarian integrated abundance the GAMs produced moderate fits
+(\(R^2_{adj}\) = 0.406-0.603) (Table 1). In the epipelagic, there were
+several significant predictor variables including inorganic nutrients
+(\(NO_3\) and \(Si\)), water quality parameters (Salinity, DO), primary
+production, and particle concentration (Table 1). However, the upper and
+lower mesopelagic were exclusively explained by particle-related
+variables (concentration and mass flux) (Table 1).
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4286}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2857}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0952}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0833}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1071}}@{}}
+\caption{Generalized Additive Model results for integrated total
+Rhizarian abundance in different regions of the water
+column.}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Model}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p-value
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Model}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p-value
+\end{minipage} \\
+\midrule()
+\endhead
+\textbf{All Rhizaria Epipelagic} & Salinity & 2.0818 & 4.654 &
+\textless0.001 \\
+\(R^2_{adj}\)= 0.42 & DO & 0.8272 & 0.945 & 0.0136 \\
+& Silica & 3.1206 & 15.75 & \textless0.001 \\
+& NO3 & 2.9970 & 6.268 & \textless0.001 \\
+& Primary Productivity & 1.8367 & 2.099 & \textless0.001 \\
+& Particle Concentration & 0.9282 & 2.566 & \textless0.001 \\
+& & & & \\
+\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6847 & 0.431 &
+0.073 \\
+\(R^2_{adj}\)= 0.406 & Avg Mass Flux & 1.5426 & 0.923 & 0.045 \\
+& Particle Concentration & 3.2957 & 20.89 & \textless0.001 \\
+& & & & \\
+\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 1.6632 & 1.824
+& 0.002 \\
+\(R^2_{adj}\)= 0.603 & Particle Concentration & 0.7694 & 0.662 &
+0.027 \\
+\bottomrule()
+\end{longtable}
+
+GAMs for individual taxa were much less consistent in their fits (Table
+2). This is likely in part due to the high number of non-observations
+for certain taxa. Note that due to low abundances, GAMS were not
+constructed for Tuscaroridae or Medusettidae. Furthermore no significant
+terms were found for a model with Aulosphaeridae in the epipelagic nor
+Foraminifera in the mesopelagic.
+
+\begin{figure}
+
+{\centering \includegraphics{images/06_epi-partials.pdf}
+
+}
+
+\caption{Partial effects of smooth terms in taxa-specific GAM models
+from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea
+(A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae
+(E).}
+
+\end{figure}
+
+Epipelagic Acantharea were explained by several predictor vairables and
+had a good fit (\(R^2_{adj}\) = 0.53, Table 2). Most notable smooths
+however, were mass flux and particle concentration, which had a weak
+positive association (Figure 6A), with July 2021 as a clear outlier
+where Acantharean abundances were high in the epipelagic despite lower
+fluxes and particle concentrations (Figure 5). Foraminifera had a good
+fitting GAM in the epipelagic (\(R^2_{adj}\) = 0.445). There were
+several significant explanatory variables, althoguht the clearest
+pattern was observed of high temperatures associated with more
+Foraminifera abundance (Table 2, Figure 6B). Epipelagic Aulacanthidae
+similarly had several predictor variables which were significant,
+including both water quality parameters and particle/flux predictors
+(Table 2). Interestingly, Aulacanthidae had primary production as a
+significant predictor, yet there was not a clear association (Figure
+6C). There was a fit for Collodaria in the epipelagic (\(R^2_{adj}\) =
+0.16), although there was a logit-like relationship where higher
+abundances tended to occur during higher DO conditions in the surface
+waters (Figure 6D). Castanellidae also had similarly poor fits in the
+epipelagic (\(R^2_{adj}\) = 0.124) (Table 2, Figure 6E).
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4471}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2824}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0824}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0824}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1059}}@{}}
+\caption{Taxa-specific generalized additive models for different regions
+of the water column.}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Model
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p-value
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Model
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p-value
+\end{minipage} \\
+\midrule()
+\endhead
+\textbf{Acantharea Epipelagic} & Salinity & 2.264 & 4.113 &
+\textless0.001 \\
+R2adj=0.53 & O2 & 2.579 & 6.712 & \textless0.001 \\
+& Avg Mass Flux & 4.810 & 22.81 & \textless0.001 \\
+& Bacteria \#/L & 1.599 & 1.317 & 0.0117 \\
+& Particle Concentration & 0.853 & 1.158 & 0.0087 \\
+& & & & \\
+\textbf{Acantharea Upper Mesopelagic} & Avg C Flux & 1.296 & 0.621 &
+0.0439 \\
+R2adj=0.231 & Avg N Flux & 1.619 & 1.283 & 0.0026 \\
+& Particle Concentration & 0.952 & 3.886 & \textless0.001 \\
+& & & & \\
+\textbf{Acantharea Lower Mesopelagic} & Temperature & 2.076 & 4.155 &
+\textless0.001 \\
+R2adj=0.509 & Avg N Flux & 1.494 & 1.216 & 0.0113 \\
+& Primary Productivity & 0.766 & 0.648 & 0.0253 \\
+& Particle Concentration & 2.037 & 10.03 & \textless0.001 \\
+\textbf{Aulacanthidae Epipelagic} & Salinity & 0.792 & 0.757 & 0.0241 \\
+R2adj=0.251 & Avg N Flux & 0.869 & 1.324 & 0.0018 \\
+& Primary Productivity & 2.312 & 7.704 & \textless0.001 \\
+& Particle Concentration & 2.008 & 2.200 & 0.0017 \\
+& & & & \\
+\textbf{Aulacanthidae Upper Mesopelagic} & Particle Concentration &
+2.832 & 9.802 & \textless0.001 \\
+R2adj = 0.158 & & & & \\
+& & & & \\
+\textbf{Aulacanthidae Lower Mesopelagic} & Avg Mass Flux & 1.622 & 1.991
+& 0.002 \\
+R2adj=0.298 & Particle Concentration & 2.123 & 6.706 & \textless0.001 \\
+& & & & \\
+\textbf{Aulosphaeridae Upper Mesopelagic} & Particle Concentration &
+2.653 & 13.06 & \textless0.001 \\
+R2adj=0.2 & & & & \\
+& & & & \\
+\textbf{Aulosphaeridae Lower Mesopelagic} & Particle Concentration &
+0.972 & 6.248 & \textless0.001 \\
+R2adj=.147 & & & & \\
+& & & & \\
+\textbf{Castanellidae Epipelagic} & Avg C Flux & 1.462 & 0.816 &
+0.0421 \\
+R2adj = 0.124 & Primary Productivity & 0.771 & 0.665 & 0.0247 \\
+& Particle Concentration & 3.956 & 4.623 & \textless0.001 \\
+& & & & \\
+\textbf{Coelodendridae Upper Mesopelagic} & Avg N Flux & 0.822 & 0.919 &
+0.0183 \\
+R2adj=.113 & Particle Concentration & 0.970 & 6.208 & \textless0.001 \\
+& & & & \\
+\textbf{Coelodendridae Lower Mesopelagic} & Particle Concentration &
+1.773 & 4.873 & \textless0.001 \\
+R2adj=.133 & & & & \\
+& & & & \\
+\textbf{Collodaria Epipelagic} & Salinity & 0.925 & 2.267 &
+\textless0.001 \\
+R2adj=0.16 & O2 & 2.015 & 2.217 & 0.002 \\
+& Bacteria \#/L & 1.843 & 2.100 & 0.002 \\
+& & & & \\
+\textbf{Foraminifera Epipelagic} & Temperature & 4.414 & 8.789 &
+\textless0.001 \\
+R2adj=0.445 & O2 & 1.603 & 1.287 & 0.0111 \\
+& Avg N Flux & 2.512 & 3.396 & \textless0.001 \\
+& Primary Productivity & 0.824 & 0.920 & 0.0153 \\
+& Particle Concentration & 0.919 & 2.257 & \textless0.001 \\
+\bottomrule()
+\end{longtable}
+
+\begin{figure}
+
+{\centering \includegraphics{images/07_upmeso-partial.pdf}
+
+}
+
+\caption{Partial effects of smooth terms in taxa-specific GAM models
+from the upper mesopelagic (200-500m). Effects are grouped by taxa;
+Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
+(D).}
+
+\end{figure}
+
+In the upper mesopelagic (200-500m), abundances were generally low
+(Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae,
+Aulosphaeridae, and Coelodendridae (Table 2). All these models had
+generally poor fits (\(R^2_{adj}\) \textless{} 0.25). Yet, for all upper
+mesopelagic models, particle concentration was a significant explanatory
+variable (Table 2, Figure 7). Carbon flux was significant for Acantharea
+and nitrogen flux was significant for both Acantharea and Coelodendridae
+(Table 2, Figure 7A,D). The lower mesopelagic also had generally poor
+GAM fits for taxa specific models (\(R^2_{adj}\) \textless0.3), with the
+exception of Acantharea (\(R^2_{adj}\) = 0.509). Acantharea in the lower
+mesopelagic was most clearly positively associated with particle
+concentration and nitrogen flux, as well as temperature to a slight
+degree (Figure 8A). For all Phaeodarians with a significant model,
+particle concentration was a main predictor variable (Table 2, Figure
+8B-D). Aulacanthidae had the best fitting model of the Phaeodarians
+(\(R^2_{adj}\) = 0.298), which also included mass flux as a
+statistically significant smooth (Figure 8B).
+
+\begin{figure}
+
+{\centering \includegraphics{images/08_lomeso-partials.pdf}
+
+}
+
+\caption{Partial effects of smooth terms in taxa-specific GAM models
+from the lower mesopelagic (500-1000m). Effects are grouped by taxa;
+Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
+(D).}
+
+\end{figure}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{overall-rhizarian-abundance-and-patterns}{%
+\subsubsection{Overall Rhizarian abundance and
+patterns}\label{overall-rhizarian-abundance-and-patterns}}
+
+In the epipelagic Rhizaria exhibited a notable seasonal pattern.
+Rhizarian abundances were higher in the summer months and lower during
+the winter. During a prior time period, Blanco-Bercial et al. (2022)
+noted that there is considerable seasonality in the community
+composition of all protists. Despite the seasonality of total Rhizarian
+abundance, community composition was relatively consistent, with
+Collodaria representing the bulk of the community. It should be noted
+that the overall taxonomic resolution of the UVP5 is fairly low, so
+there may be a switching of species within the broad groups identified
+in this study which were not captured. Throughout the mesopelagic,
+month-to-month variation in 2021 was relatively low. Again this is
+consistent with observations from metabarcoding of the whole protist
+community in the same study region (Blanco-Bercial et al., 2022). This
+finding is not surprising as the overall seasonal variation in
+environmental conditions in this region were low.
+
+Overall Rhizaria were the most commonly identified group of
+mesozooplankton throughout the study period. We do note that the UVP5
+commonly captures Trichodesmium colonies, yet these were excluded in
+this comparison as they are strictly autotrophs. It should be noted that
+previous work has suggested that avoidance behavior with the UVP is
+possible, at times likely, for visual and highly mobile zooplankton
+(Barth and Stone, 2022). Thus, the percent contribution reported here
+(42.7\%) of Rhizaria to the total mesozooplankton community may be
+inflated due to under sampling of organisms such as Euphausiids and
+Chaetognaths which has quick escape responses. Regardless, it is worth
+noting that in the same region, with data collected in 2012 and 2013
+using similar calculation methods, Biard et al. (2016) estimated
+Rhizaria only contribute 15\% of the total mesozooplankton community in
+the upper 500m. Likely, Rhizaria display considerable interannual
+variability. In the present study, we noticed considerably higher
+Rhizarian abundance throughout the water column in 2019 compared to
+2021. While this may have been driven by increased mass flux (Figure 3),
+more information is needed to truly understand the magnitude by which
+Rhizaria can vary interannually.
+
+\hypertarget{vertical-structure-and-trophic-roles}{%
+\subsubsection{Vertical Structure and Trophic
+Roles}\label{vertical-structure-and-trophic-roles}}
+
+In this study we present a clear pattern of vertical zonation between
+different Rhizaria groups. Largely, the taxonomic composition and
+vertical positioning were similar to Rhizaria zonation in the California
+Current Ecosystem (Biard and Ohman, 2020). It should be noted however,
+that the secondary abundance peak reported in the present study is
+lower. This is likely due to the more oligotrophic nature of the study
+region, were the euphotic zone penetrates deeper into the water column.
+Most prevalent in the epipelagic were Collodaria. These mixotrophic
+Radiolarians have long been reported to contribute to primary
+productivity in the euphotic zone (Dennett et al., 2002; Michaels et
+al., 1995). Collodaria are thought to be particularly successful
+globally in oligotrophic regions due to their photosymbiotic
+relationships (Biard et al., 2017, 2016). We observed the highest
+abundance of Collodaria during June 2021, supporting the notion they can
+thrive during the typically low-nutrient conditions of summer
+stratification. However, Collodaria also increased during the spring
+mixing period, suggesting that they can thrive during conditions which
+may typically be thought to favor autotrophs. Furthermore, while
+Collodaria were primarily absent from below 250m, there were a few
+instances of deeper colonies being observed. Global investigations of
+polycistine flux, suggest that deep-Collodaria in Oligotrophic regions
+may be a consequence of isothermal submersion (Boltovskoy, 2017).
+Another effectively exclusively epipelagic Rhizaria was the Phaeodarian
+family of Castanellidae. All Phaeodarians are thought to be fully
+heterotrophic (Nakamura and Suzuki, 2015), nonetheless a number of
+studies, including this one, report Castanellidae to be typically found
+in the lower epipelagic (Biard et al., 2018; Biard and Ohman, 2020;
+Zasko and Rusanov, 2005). It should be considered that perhaps
+Castanellidae specializes in feeding sinking particles directly at the
+base of the epipelagic. Given it's smaller size (Nakamura and Suzuki,
+2015), Castanellidae does not need a large diameter to efficiently flux
+feed at the typically particle rich region of the lower epipelagic. Both
+Castanelldiae and Collodaria had poor fitting GAMs. This is somewhat of
+a surprise for Collodaria who had large abundance. However, given the
+consistency of their abundance, it may be that this study did not
+capture a wide enough range of conditions for describing Collodaria's
+preferred niche.
+
+The mesopelagic generally was home to known heterotrophic organisms,
+particularly for those which were constrained to exclusively occupy
+deeper waters. This is consistent with Blanco-Bercial et al. (2022)'s
+observation of an auto-/mixotroph to heterotroph gradient in the protist
+community. The upper mesopelagic interestingly had relatively low total
+abundance. This low-abundance region likely reflects the dynamics of
+productivity and export throughout the water column. While productivity
+and thus sinking particles for flux feeders are high in the euphotic
+zone, much of this is attenuated throughout the epipelagic. So, while
+the base of the epipelagic may provide a rich feeding environment for
+Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F),
+the region from 200-500m might be otherwise food poor. Perhaps it is
+more advantageous for Rhizaria to situate deeper, in darker regions of
+the twilight zone. Also it should be noted that Phaeodarians utilize
+silica to build their opaline tests, and silica concentrations started
+to increase around 500m (Figure 2G). Although \(Si\) was not a
+significant smooth for any taxa-specific model, this lack of association
+might be due to the overall lack of variation of \(Si\) between
+integrated abundance of each cast. Aulosphaeridae was only found to have
+significant relationships, although weak fits, to particle concentration
+in the mesopelagic. In our study, while consistently observed, overall
+abundances of Aulosphaeridae were very low. In the Pacific Ocean, on
+California's Coast, much higher abundances of Aulosphaeridae have been
+reported (Biard and Ohman, 2020; Zasko and Rusanov, 2005) and they have
+massive potential to impact silica export (Biard et al., 2018).
+Coelodendridae were also seemingly restricted to the deeper section of
+the mesopelagic. This is interesting given that in the California
+Current, (Biard and Ohman, 2020) found a bimodal distribution in
+Coelodendridae. There are several morphotypes corresponding to different
+taxa of Coelodendridae (Biard and Ohman, 2020; Nakamura and Suzuki,
+2015). So it may be that only a few types of Coelodendridae were
+observed in this study, while the epipelagic variety was not.
+Alternatively, the lower epipelagic of the California Current may
+provide adequate habitat for Coelodendridae, which is not available in
+the oligotrophic Sargasso Sea.
+
+A number of taxa were found to have a bimodal distribution, with
+considerable populations in both the epipelagic and mesopelagic.
+Aulacanthidae had a bimodal distribution, although abundances were
+highest in the lower mesopelagic. Foraminifera also had a bimodal
+distribution. Some lineages of Foraminifera are known to host
+photosymbionts (Biard, 2022a; Kimoto, 2015), however they are also
+efficient predators commonly seen throughout the mesopelagic (Caron and
+Be, 1984; Gaskell et al., 2019). Thus it is not surprising to find their
+presence in both locations of the water column. Foraminifera are also
+known to vary their vertical distribution across their life cycle in
+phase with lunar cycles (Biard, 2022a; Bijma et al., 1990; Gaskell et
+al., 2019; Kimoto, 2015). However, the sampling scheme of the BATS
+program does not capture this frequency and was not investigated in the
+present study.
+
+Acantharea also had a bimodal distribution, with much larger abundances
+than Aulacanthidae or Foraminifera. Most prior studies of Acantharea
+vertical distribtion found them concentrated in near surface layers of
+the water column Zasko and Rusanov (2005). This would support the
+paradigm that large Acantharean abundances may be supported by their
+mixotrophic abilities (Michaels et al., 1995; Suzuki and Not, 2015).
+While the UVP5 images cannot distinguish between mixotrophic and
+heterotrophic Acantharea, the GAMs constructed for Acantharean abundance
+found positive associations with particle concentration and mass flux,
+suggesting a higher reliance on heterotrophy. Recently Mars Brisbin et
+al. (2020) described apparent predator behavior amongst near-surface
+Acantharea. Thus it is likely that epipelagic Acantharea may commonly be
+heterotrophic. Yet, it should be noted in the Sargasso Sea, both
+heterotrophic and symbiotic lineages of Acantharea have been reported
+(Blanco-Bercial et al., 2022). Additionally, Michaels (1988) noted that
+the majority of Acantharea (by abundance) were smaller than
+160\(\mu m\). While that estimate may be inflated due to inability to
+capture larger cells, small Acantharea were not captured in the present
+study. Thus, trophic strategy may shift based on sizes of Acantharea.
+
+Decelle et al. (2013) proposed a hypothetical life cycle for
+cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life
+cycle suggests that epipelagic Acantharea are adult populations, which
+form cysts that sink into the mesopelagic, then reproduce and raise.
+Furthermore, given that horizontal transfer of symbionts between
+generations of Acantharea is unlikely, the newly spawned mesopelagic
+Acantharea are not necessarily required to rapidly return to the photic
+zone. This hypothesis predicts that Acanthareans in the mesopelagic
+would be smaller (Decelle et al., 2013). Mars Brisbin et al. (2020)
+provided some support for this hypothesis, with a significant decrease
+in Acantharea sizes with depth. Although, the authors also observed low
+abundances in the mesopelagic and noted that the smaller sizes may be
+due to less food availability (Mars Brisbin et al., 2020). Since food is
+more scarce in the mesopelagic there is is less nutritional quality, yet
+flux feeders would likely grow larger to increase their feeding range
+(Biard and Ohman, 2020). In the data collecting in this study,
+Acanthareans in the mesopelagic were significantly smaller than the
+epipelagic, despite the other bimodal taxa (Foraminifera and
+Aulacanthidae) being significantly larger with depth (Figure 9, Wilcoxon
+rank sum text p-value \textless{} 0.001). This provides added support
+for the hypothesis that cyst-forming Acantharea may utilize different
+sections of the water column throughout their life cycle. However to
+further investigate this, more work is needed with higher temporal and
+taxonomic resolution.
+
+\begin{figure}
+
+{\centering \includegraphics{images/09_size-comp.pdf}
+
+}
+
+\caption{Comparison of average sizes (ESD) amongst Rhizarian taxa which
+occurred throughout the water column.}
+
+\end{figure}
+
+\hypertarget{conclusions-and-considerations}{%
+\subsubsection{Conclusions and
+Considerations}\label{conclusions-and-considerations}}
+
+This study provides a detailed look at Rhizarian abundance over time
+throughout the water column in a major oligotrophic gyre. We show that
+their abundances are generally related to particle concentration and
+flux, although lack of environmental variability may have reduced the
+fit of our GAMs. Considering the potential role of Rhizaria in the
+biological carbon pump, they may have a somewhat mixed role. In the
+shallower regions, where smaller Rhizaria are abundant, they may be an
+attenuating force on sinking particles (Stukel et al., 2019). It should
+be noted that in our study, we focused on the ``small'' particle
+concentration field (\textless450\(\mu m\)), and these particles are
+generally slower sinking than large particles. However, once consumed
+and repackaged by larger Rhizaria, they can sink quicker and contribute
+more to overall flux (Michaels, 1988). Thus, Rhizaria may act as a
+ballasting mechanism. However, this is largely speculation, to truly
+test this, more work is needed measuring Rhizarian flux.
+
+The vertical partitioning documented in this study do support the
+hypothesis that mixotrophic rhizaria will occupy shallower waters while
+deeper waters are dominated by heterotrophy. However the degree to which
+mixotrophic Rhizaria in the euphotic zone rely on heterotrophy versus
+symbiosis is uncertain. Collodaria were recorded as consistent and
+dominant members of the near surface region. These organisms have the
+potential to contribute considerably to the otherwise low productivity
+of oligotrophic regions. However, their role in food webs is not well
+understood. While this study represents a step-forward in our
+understanding of Rhizarian ecology, there are several known aspects of
+Rhizarian and protist ecology which are not captured in this study.
+Continued research on Rhizaria is much needed to better understand their
+ecology. Particularly extended descriptive work to capture interannual
+patterns. Also work defining biotic interactions, feeding rates,
+productivity, and life history are all rich fields of interest in
+Rhizaria.
+
+\hypertarget{references}{%
+\section{References:}\label{references}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-barth2022}{}}%
+Barth, A., Stone, J., 2022.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.898057}{Comparison
+of an in situ imaging device and net-based method to study
+mesozooplankton communities in an oligotrophic system}. Frontiers in
+Marine Science 9.
+
+\leavevmode\vadjust pre{\hypertarget{ref-benfield1996}{}}%
+Benfield, M.C., Davis, C.S., Wiebe, P.H., Gallager, S.M., Gregory Lough,
+R., Copley, N.J., 1996. Video Plankton Recorder estimates of copepod,
+pteropod and larvacean distributions from a stratified region of Georges
+Bank with comparative measurements from a MOCNESS sampler. Deep Sea
+Research Part II: Topical Studies in Oceanography 43, 1925--1945.
+\url{https://doi.org/10.1016/S0967-0645(96)00044-6}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
+Biard, T., 2022a. Rhizaria. Encyclopeida of Life Sciences, pp. 1--11.
+\url{https://doi.org/10.1002/9780470015902.a0029469}
+
 \leavevmode\vadjust pre{\hypertarget{ref-biard2022b}{}}%
-Biard, T., 2022. Diversity and ecology of Radiolaria in modern oceans.
+Biard, T., 2022b. Diversity and ecology of Radiolaria in modern oceans.
 Environmental Microbiology 24, 2179--2200.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-drago2022}{}}%
+Drago, L., Panaïotis, T., Irisson, J.-O., Babin, M., Biard, T.,
+Carlotti, F., Coppola, L., Guidi, L., Hauss, H., Karp-Boss, L., Lombard,
+F., Mcdonnell, A.M.P., Picheral, M., Rogge, A., Waite, A.M., Stemmann,
+L., Kiko, R., 2022. Global Distribution of Zooplankton Biomass Estimated
+by In Situ Imaging and Machine Learning. Frontiers in Marine Science 9.
+\url{https://doi.org/10.3389/fmars.2022.894372}
+
 \leavevmode\vadjust pre{\hypertarget{ref-gaskell2019}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lomas2013}{}}%
+Lomas, M.W., Bates, N.R., Johnson, R.J., Knap, A.H., Steinberg, D.K.,
+Carlson, C.A., 2013. Two decades and counting: 24-years of sustained
+open ocean biogeochemical measurements in the Sargasso Sea. Deep Sea
+Research Part II: Topical Studies in Oceanography 93, 16--32.
+\url{https://doi.org/10.1016/j.dsr2.2013.01.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lombard2019}{}}%
+Lombard, F., Boss, E., Waite, A.M., Vogt, M., Uitz, J., Stemmann, L.,
+Sosik, H.M., Schulz, J., Romagnan, J.-B., Picheral, M., Pearlman, J.,
+Ohman, M.D., Niehoff, B., Möller, K.O., Miloslavich, P., Lara-Lpez, A.,
+Kudela, R., Lopes, R.M., Kiko, R., Karp-Boss, L., Jaffe, J.S., Iversen,
+M.H., Irisson, J.-O., Fennel, K., Hauss, H., Guidi, L., Gorsky, G.,
+Giering, S.L.C., Gaube, P., Gallager, S., Dubelaar, G., Cowen, R.K.,
+Carlotti, F., Briseño-Avena, C., Berline, L., Benoit-Bird, K., Bax, N.,
+Batten, S., Ayata, S.D., Artigas, L.F., Appeltans, W., 2019.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00196}{Globally
+consistent quantitative observations of planktonic ecosystems}.
+Frontiers in Marine Science 6.
+
+\leavevmode\vadjust pre{\hypertarget{ref-marra2011}{}}%
+Marra, G., Wood, S.N., 2011. Practical variable selection for
+generalized additive models. Computational Statistics \& Data Analysis
+55, 2372--2387. \url{https://doi.org/10.1016/j.csda.2011.02.004}
+
 \leavevmode\vadjust pre{\hypertarget{ref-marsbrisbin2020}{}}%
@@
 and Oceanography 65, 2953--2965. \url{https://doi.org/10.1002/lno.11567}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcgillicuddy1998}{}}%
+McGillicuddy, D.J., Robinson, A.R., Siegel, D.A., Jannasch, H.W.,
+Johnson, R., Dickey, T.D., McNeil, J., Michaels, A.F., Knap, A.H., 1998.
+Influence of mesoscale eddies on new production in the Sargasso Sea.
+Nature 394, 263--266. \url{https://doi.org/10.1038/28367}
 
@@
 \url{https://doi.org/10.1093/plankt/17.1.131}
+
+\leavevmode\vadjust pre{\hypertarget{ref-michaels1996}{}}%
+Michaels, A.F., Knap, A.H., 1996. Overview of the U.S. JGOFS Bermuda
+Atlantic Time-series Study and the Hydrostation S program. Deep Sea
+Research Part II: Topical Studies in Oceanography 43, 157--198.
+\url{https://doi.org/10.1016/0967-0645(96)00004-5}
 
@@
 Science 76, 1959--1972. \url{https://doi.org/10.1093/icesjms/fsz184}
+
+\leavevmode\vadjust pre{\hypertarget{ref-panauxefotis2023}{}}%
+Panaïotis, T., Babin, M., Biard, T., Carlotti, F., Coppola, L., Guidi,
+L., Hauss, H., Karp-Boss, L., Kiko, R., Lombard, F., McDonnell, A.M.P.,
+Picheral, M., Rogge, A., Waite, A.M., Stemmann, L., Irisson, J.-O.,
+2023. Three major mesoplanktonic communities resolved by in situ imaging
+in the upper 500 m of the global ocean. Global Ecology and Biogeography
+32, 1991--2005. \url{https://doi.org/10.1111/geb.13741}
+
+\leavevmode\vadjust pre{\hypertarget{ref-picheral2010}{}}%
+Picheral, M., Guidi, L., Stemmann, L., Karl, D.M., Iddaoud, G., Gorsky,
+G., 2010. The Underwater Vision Profiler 5: An advanced instrument for
+high spatial resolution studies of particle size spectra and
+zooplankton. Limnology and Oceanography: Methods 8, 462--473.
+\url{https://doi.org/10.4319/lom.2010.8.462}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
-Wiley, 2022. Rhizaria. Wiley, pp. 1--11.
-\url{https://doi.org/10.1002/9780470015902.a0029469}
+\leavevmode\vadjust pre{\hypertarget{ref-whitmore2021}{}}%
+Whitmore, B.M., Ohman, M.D., 2021. Zooglider-measured association of
+zooplankton with the fine-scale vertical prey field. Limnology and
+Oceanography 66, 3811--3827. \url{https://doi.org/10.1002/lno.11920}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wood2017}{}}%
+Wood, S.N., 2017. Generalized additive models: an introduction with R,
+Second edition. ed, Texts in statistical science. CRC Press/Taylor \&
+Francis Group, Boca Raton London New York.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wood2001}{}}%
+Wood, S.N., 2001. mgcv: GAMs and Generalized Ridge Regression for R 1.
+
+\leavevmode\vadjust pre{\hypertarget{ref-zasko2005}{}}%
+Zasko, D.N., Rusanov, I.I., 2005. Vertical Distribution of Radiolarians
+and Their Role in Epipelagic Communities of the East Pacific Rise and
+the Gulf of California. Biology Bulletin 32, 279--287.
+\url{https://doi.org/10.1007/s10525-005-0103-5}
 

### 2024-01-25 R/00_data-import.R

@@ -141,27 +141,6 @@ coords <- zoops$meta |>
   select(profileid, latitude, longitude, sampledate,
          ctd_origfilename, tod) |> 
   left_join(ctd_coords, by = "ctd_origfilename")
-# 
-# # |- All cast map ---------------------
-# library("rnaturalearth")
-# library("rnaturalearthdata")
-# 
-# world <- ne_countries(scale = "medium", returnclass = "sf")
-# 
-# all_cast_map <- ggplot()+
-#   geom_sf(data = world, fill = 'green')+
-#   geom_point(aes(x = coords$lon, coords$lat,
-#                  shape = coords$tod, col = coords$tod))+
-#   coord_sf(xlim = c(-65,-62), ylim = c(30,33))+
-#   geom_segment(aes(x = -65.1, xend = -62.9,
-#                    y = 31.8, yend = 32.75),
-#                size = 2, col = 'red')+
-#   theme_bw()+
-#   labs(x = "",y = "", shape = "")+
-#   scale_color_manual(values = c('beige','black'))+
-#   theme(panel.background = element_rect(fill = 'cornflowerblue'),
-#         axis.text.x = element_text(angle = 45, hjust = c(1,1)),
-#         legend.position = 'bottom')
 
 # |- Removing casts outside range ------------------------
 drop_casts <- which(

### 2024-01-25 R/s01_cast-map.R

@@ -1,3 +1,37 @@
 ###
 # Cast Map
 ###
+
+ctd_data <- readRDS('./data/00_ctd-data.rds')
+zoops <- readRDS("./data/00_zoop-uvp.rds")
+
+# need to check ctd vs uvp for correct coords
+ctd_coords <- ctd_data |> 
+  list_to_tib('cruise_id') |> 
+  select(ctd_origfilename, lat, lon, cruise_id, datetime) |> 
+  distinct()
+
+coords <- zoops$meta |> 
+  select(profileid, latitude, longitude, sampledate,
+         ctd_origfilename, tod) |> 
+  left_join(ctd_coords, by = "ctd_origfilename")
+
+
+# |- All cast map ---------------------
+library("rnaturalearth")
+library("rnaturalearthdata")
+
+world <- ne_countries(scale = "medium", returnclass = "sf")
+
+all_cast_map <- ggplot() +
+  geom_sf(data = world, fill = 'green')+
+  geom_point(aes(x = coords$lon, y =coords$lat))+
+  coord_sf(xlim = c(-65,-62), ylim = c(30,33))+
+  theme_bw()+
+  labs(x = "",y = "", shape = "")+
+  theme(panel.background = element_rect(fill = 'cornflowerblue'),
+        axis.text.x = element_text(angle = 45, hjust = c(1,1)),
+        legend.position = 'bottom')
+
+
+windows()

### 2024-01-29 manuscript/01_intro.qmd
render with line numbers

@@
 
-Rhizaria are an extremely diverse super-group of single celled organisms consisting of several phyla including retaria (foraminifera and radiolaria) and cercozoa.
+Rhizaria are an extremely diverse super-group of single-celled organisms consisting of several phyla including Retaria (foraminifera and radiolaria) and Cercozoa.
 These organisms exist in a wide range of habitats and are widely represented in plankton communities throughout the global ocean.
@@
 Despite their noticeable morphology and global distribution, Rhizaria were largely understudied throughout the 20th century.
-The bulk of modern plankton research has focused hard-bodied crustacea which are numerically dominant and easily sampled with nets and preservatives.
-Fragile organisms like Rhizaria, were difficult to adequately study as they can be destroyed through standard zooplankton sampling techniques.
+The bulk of modern plankton research has focused on hard-bodied crustacea which are numerically dominant and easily sampled with nets and preservatives.
+Fragile organisms like Rhizaria, were difficult to adequately study as they can be destroyed through standard zooplankton sampling techniques and preserve poorly.
 A number of studies in the late 1900s did employ alternative techniques to quantify Rhizaria including diaphragm pumps [@michaels1988] or blue-water SCUBA collections [@caron1984; @caron1995; @bijma1990].
@@
 While there were early applicaitons of imaging tools to document Rhizaria [@dennett2002], @biard2016's report from a global imaging dataset highlighted the importance of Rhizaria to the total standing stock of marine carbon.
-Due to their large sizes, ability to concentrate smaller particles (ballasting), and the unique structure of their mineral skeletons, Rhizaria have the potential to massively influence ocean biogeochemical cycling.
+Due to their large sizes, ability to concentrate smaller particles and the unique structure of their mineral skeletons, Rhizaria have the potential to massively influence ocean biogeochemical cycling.
 A number of studies have made large advances in estimating the contribution of Rhizaria to ocean cycling of carbon [@lampitt2009; @stukel2018; @ikenoue2019; @gutierrez-rodriguez2019], silica [@biard2018; @llopismonferrer2021], and strontium [@decelle2013].
@@
 Mixotrophic foraminifera host a variety of endosymbiont partners [@lee2006; @decelle2015], which are thought to support early and adult life stages and contribute to total primary productivity [@kimoto2015].
-Still, foraminifera are omnivorous, possibly even predominately carnivorous with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], majoritively consuming live copepods [@caron1984].
+Still, foraminifera are omnivorous, possibly even predominately carnivorous with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], mainly consuming live copepods [@caron1984].
 Radiolaria have several lineages all which have some taxa who are well known to host symbionts [@biard2022b].
@@
 In an imaging-based study of the whole Rhizaria community, @biard2020 noted clear patterns in vertical zonation which largely corresponded to different trophic roles.
-In the oligotrophic ocean, @blanco-bercial2022 also noted that the protist community, including Rhizaria partition along an autotroph and mixotroph to heterotroph gradient with increasing depth in the water column.
+In the oligotrophic ocean, @blanco-bercial2022 also noted that the protist community, including Rhizaria, partition along an autotroph and mixotroph to heterotroph gradient with increasing depth in the water column.
 Yet, few studies have made direct attempts to relate rhizaria abundances to environmental factors [@biard2020].

### 2024-01-29 manuscript/02_methods.qmd
render with line numbers

@@
 The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022.
-Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda into the Sargasso Sea (Supplemental Figure 1).
+Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda into the Sargasso Sea (approximately 31.0$^oN$-32.5$^oN$, 64.25$^oW$-63$^oW$; Supplemental Figure 1).
 In general, casts extended to either 200m, 500m, or 1200m deep, with a few extended into the bathypelagic (4500m).
@@
 
-Individual vignettes of Rhizaria images were identified using the classification platform Ecotaxa \[cite\].
-Data were pre-sorted according to a .
+Individual vignettes of Rhizaria images were identified using the classification platform Ecotaxa [@picheral].
+Data were pre-sorted utilizing a random-forest classifier and pre-trained learning set.
 Taxonomic classification were done based on morphology exclusively.
 While there are sparse taxonomic guides for in-situ images of rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020].
-Using the aforementioned sources and publicly available ecotaxa projects, I constructed a guide which can be found here: <https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf>.
+Using the aforementioned sources and publicly available ecotaxa projects, I constructed a guide which can be found here: [https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf](#0){.uri}.
 Broadly, Rhizaria were classified as Foraminifera, Radiolarians (Acantharea or Collodaria), or a variety of Phaeodarian families (Figure 1).
@@
 
-![Example images of different Rhizaria taxa. 4mm scale bar shown in lower right. All vignettes are same scale](images/01_taxa.pdf)
-
 The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$.
 However, identified rhizaria ranged from a 934$\mu m$ Aulacanthidae cell to a Collodarian colony over 10mm in diameter.
-To confirm that the UVP5 was sampling adequately across all size ranges, an NBSS slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth-inrev] .
+To confirm that the UVP5 was sampling adequately across all size ranges, an NBSS slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth-inrev].
 However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded.
 The UVP5 reports the exact depth at which a particle is recorded, however to estimate abundance, observations must be binned over fixed depth intervals.
-In our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic [see @barth2022 for an extended discussion of UVP5 data processing).
-For the present study, Rhizarian abundances were estimated in 25m vertical bins, which offer a moderate sampling volume (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic) while still maintaining ecologically relevant widths.
+Our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic (see @barth2022 for an extended discussion of UVP5 data processing).
+For the present study, Rhizarian abundances were estimated in 25m vertical bins, which offer a moderate sampling volume per bin (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic) while still maintaining ecologically relevant widths.
 However, concentrations in a 25m bin would need to be greater than 2.428 indv.
 $m^{-3}$ and 3.912 indv.
-$m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth-inrev].
+$m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk \[@benfield1996; @barth-inrev\].
 Because we typically observed many rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions.
@@
 In cases where observations were too sparse for a given taxonomic grouping, models were not run.
+All code and full models are available in code, as well as intermediate data products at <https://github.com/TheAlexBarth/RhizariaSeasonality>.

### 2024-01-29 manuscript/03_results.qmd
render with line numbers

@@
 
-![Environmental profiles across time-series of study period.
-Y axis shows depth in meters.
-(A) Temperature.
-(B) Salinity.
-(C) Dissolved Oxygen.
-(D) In-situ chlorophyll fluorescence.
-(E) Particle concenctrion (184 - 450$\mu m$).
-(F) Bacteria Abundance.
-(G) Silica.
-(H) Nitrate](images/02-environmental.pdf)
-
-Overall, particle concentration was high near the surface during the 2021 spring bloom, then sank throughout the water column attenuating throughout the lower epipelagic (Figure 2E).
-Similarly, bacteria abundance was closely linked to overall productivity, althrough there was a more consistent layer near the top of the mesopelagic (\~250m) (Figure 2F).
+Overall, particle concentration was high near the surface during the 2021 spring bloom, then moved deeper throughout the water column attenuating throughout the lower epipelagic (Figure 2E).
+Similarly, bacteria abundance was closely linked to overall productivity, althrough there was a more consistent moderate-abundance layer near the top of the mesopelagic (\~250m) (Figure 2F).
 Concurrent with the secondary fall production peak, there was also higher particle concentration and bacterial abundance in the later summer and early fall.
@@
 
-![Primary productivity (A) and flux estimates from total mass (B), carbon (C), and nitrogen (D).
-Values are from monthly cruises with month displayed in corresponding colors.
-Absent data are shown by grey bar.](images/03_prod-flux.pdf)
-
 ### Rhizaria abundance and distribution
@@
 Considering the total abundances of the observational period, Rhizaria comprised on average, 42.6% of all mesozooplankton abundance (Supplemental Figure 3).
-Copepods were the second most abundant, comprising 35.5% and all living mesozooplankton were 22%.
+Copepods were the second most abundant, comprising 35.5% and all other living mesozooplankton were 22%.
 The large contribution of Rhizaria to the mesozooplankton community is most prominent in the epipelagic, where they accounted for 47% of all mesozooplankton.
@@
 
-![Average abundance of Rhizaria in 25m bins, across entire study period.
-Shown are total Rhizaria (A), Acantharea (B), Collodaria (C), Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae (G), Coelodendridae (H).](images/04_vertical.pdf)
-
 Total average Rhizaria abundance had a bimodal distribution with respect to depth.
 Total abundance was highest just below the surface (0-100m), with secondary, wider peak occurring in the mid mesopelagic (Figure 4A).
-Although, variation in depth binned abundance was large, likely due to seasonal variability but also increased from the detection-risk described in the methods.
+Variation in depth binned abundance was large, likely due to seasonal variability but also increased from the detection-risk described in the methods.
 The vertical distribution pattern and abundance varied considerably across taxonomic groups.
-Radiolarians were some fo the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5B).
+Radiolarians were some of the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5B).
 This pattern was led by Collodaria, whose colonies were abundant in the upper epipelagic and declined into the top of the mesopelagic (Figure 4C).
 Acantharea displayed a bimodal distribution accounting for a large portion of the total Rhizaria pattern (Figure 4B, Figure 5).
-Foraminifera had a similar bimodal distribution, yet their overall average densities were much lower and spread wider throughout the mesopelagic (FIgure 4E).
+Foraminifera had a similar bimodal distribution, yet their overall average densities were much lower and spread wider throughout the mesopelagic (Figure 4E).
 Phaeodarian families exhibited a wide range of vertical distribution patterns.
 The most abundant, Aulacanthidae, also had a bimodal pattern but the density was highest in the lower mesopelagic (Figure 4D).
-Aulosphaeridae had low average density and was nearly homogenously distributed throughout the water column, although slightly lower int eh epipelagic (Figure 4F).
+Aulosphaeridae had low average density and was nearly homogeneously distributed throughout the water column, although slightly lower in the epipelagic (Figure 4F).
 Castanellidae were the only Phaeodarian who appeared to be effectively restricted to the photic zone (Figure 4G).
@@
 
-![Seasonality of Rhizarian integrated abundance for the epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower mesopelagic (500-1000m) (E-F).
-Left panels (A,C,E) display total integrated abundance per monthly cruise colored by month.
-Right panels (B, D, F) display community composition of each total abundance.](images/05_seasonality.pdf)
+### Body size throughout the water column.
+
+Very few taxa had consistent distributions throughout the water column.
+Only Acanthrea, Foraminifera, Aulacanthidae, and Aulosphaeridae were consistently abundant in the epipelagic and mesopelagic.
+To investigate if the populations or morphologies shifted throughout the water column, we compared the sizes (Equivalent Spherical Diameters, ESD) between mesopelagic and epipelagic groups for each taxa.
+All groups were significantly different on average (Wilcox Rank Sum p-value <0.001).
+Acantharea were smaller, on average in the mesopelagic while all other taxa tended to be larger (Figure 6).
 
@@
 
-| **Model**                          | Term                   | edf    | F     | p-value |
-|------------------------------------|------------------------|--------|-------|---------|
-| **All Rhizaria Epipelagic**        | Salinity               | 2.0818 | 4.654 | \<0.001 |
-| $R^2_{adj}$= 0.42                  | DO                     | 0.8272 | 0.945 | 0.0136  |
-|                                    | Silica                 | 3.1206 | 15.75 | \<0.001 |
-|                                    | NO3                    | 2.9970 | 6.268 | \<0.001 |
-|                                    | Primary Productivity   | 1.8367 | 2.099 | \<0.001 |
-|                                    | Particle Concentration | 0.9282 | 2.566 | \<0.001 |
-|                                    |                        |        |       |         |
-| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6847 | 0.431 | 0.073   |
-| $R^2_{adj}$= 0.406                 | Avg Mass Flux          | 1.5426 | 0.923 | 0.045   |
-|                                    | Particle Concentration | 3.2957 | 20.89 | \<0.001 |
-|                                    |                        |        |       |         |
-| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 1.6632 | 1.824 | 0.002   |
-| $R^2_{adj}$= 0.603                 | Particle Concentration | 0.7694 | 0.662 | 0.027   |
-
-: Generalized Additive Model results for integrated total Rhizarian abundance in different regions of the water column.
-
 GAMs for individual taxa were much less consistent in their fits (Table 2).
 This is likely in part due to the high number of non-observations for certain taxa.
-Note that due to low abundances, GAMS were not constructed for Tuscaroridae or Medusettidae.
+Note that due to low abundances, GAMs were not constructed for Tuscaroridae or Medusettidae.
 Furthermore no significant terms were found for a model with Aulosphaeridae in the epipelagic nor Foraminifera in the mesopelagic.
 
-![Partial effects of smooth terms in taxa-specific GAM models from the epipelagic (0-200m).
-Effects are grouped by taxa; Acantharea (A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae (E).](images/06_epi-partials.pdf)
-
-Epipelagic Acantharea were explained by several predictor vairables and had a good fit ($R^2_{adj}$ = 0.53, Table 2).
-Most notable smooths however, were mass flux and particle concentration, which had a weak positive association (Figure 6A), with July 2021 as a clear outlier where Acantharean abundances were high in the epipelagic despite lower fluxes and particle concentrations (Figure 5).
+Epipelagic Acantharea were explained by several predictor variables and had a good fit ($R^2_{adj}$ = 0.53, Table 2).
+Most notable smooths were mass flux and particle concentration, which had a weak positive association (Figure 7A), with July 2021 as a clear outlier where Acantharean abundances were high in the epipelagic despite lower fluxes and particle concentrations (Figure 5).
 Foraminifera had a good fitting GAM in the epipelagic ($R^2_{adj}$ = 0.445).
-There were several significant explanatory variables, althoguht the clearest pattern was observed of high temperatures associated with more Foraminifera abundance (Table 2, Figure 6B).
+There were several significant explanatory variables, although the clearest pattern was observed of high temperatures associated with more Foraminifera abundance (Table 2, Figure 7B).
 Epipelagic Aulacanthidae similarly had several predictor variables which were significant, including both water quality parameters and particle/flux predictors (Table 2).
-Interestingly, Aulacanthidae had primary production as a significant predictor, yet there was not a clear association (Figure 6C).
-There was a fit for Collodaria in the epipelagic ($R^2_{adj}$ = 0.16), although there was a logit-like relationship where higher abundances tended to occur during higher DO conditions in the surface waters (Figure 6D).
-Castanellidae also had similarly poor fits in the epipelagic ($R^2_{adj}$ = 0.124) (Table 2, Figure 6E).
-
-| Model                                | Term                   | edf   | F     | p-value |
-|--------------------------------------|------------------------|-------|-------|---------|
-| **Acantharea Epipelagic**            | Salinity               | 2.264 | 4.113 | \<0.001 |
-| R2adj=0.53                           | O2                     | 2.579 | 6.712 | \<0.001 |
-|                                      | Avg Mass Flux          | 4.810 | 22.81 | \<0.001 |
-|                                      | Bacteria #/L           | 1.599 | 1.317 | 0.0117  |
-|                                      | Particle Concentration | 0.853 | 1.158 | 0.0087  |
-|                                      |                        |       |       |         |
-| **Acantharea Upper Mesopelagic**     | Avg C Flux             | 1.296 | 0.621 | 0.0439  |
-| R2adj=0.231                          | Avg N Flux             | 1.619 | 1.283 | 0.0026  |
-|                                      | Particle Concentration | 0.952 | 3.886 | \<0.001 |
-|                                      |                        |       |       |         |
-| **Acantharea Lower Mesopelagic**     | Temperature            | 2.076 | 4.155 | \<0.001 |
-| R2adj=0.509                          | Avg N Flux             | 1.494 | 1.216 | 0.0113  |
-|                                      | Primary Productivity   | 0.766 | 0.648 | 0.0253  |
-|                                      | Particle Concentration | 2.037 | 10.03 | \<0.001 |
-| **Aulacanthidae Epipelagic**         | Salinity               | 0.792 | 0.757 | 0.0241  |
-| R2adj=0.251                          | Avg N Flux             | 0.869 | 1.324 | 0.0018  |
-|                                      | Primary Productivity   | 2.312 | 7.704 | \<0.001 |
-|                                      | Particle Concentration | 2.008 | 2.200 | 0.0017  |
-|                                      |                        |       |       |         |
-| **Aulacanthidae Upper Mesopelagic**  | Particle Concentration | 2.832 | 9.802 | \<0.001 |
-| R2adj = 0.158                        |                        |       |       |         |
-|                                      |                        |       |       |         |
-| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.622 | 1.991 | 0.002   |
-| R2adj=0.298                          | Particle Concentration | 2.123 | 6.706 | \<0.001 |
-|                                      |                        |       |       |         |
-| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.653 | 13.06 | \<0.001 |
-| R2adj=0.2                            |                        |       |       |         |
-|                                      |                        |       |       |         |
-| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 0.972 | 6.248 | \<0.001 |
-| R2adj=.147                           |                        |       |       |         |
-|                                      |                        |       |       |         |
-| **Castanellidae Epipelagic**         | Avg C Flux             | 1.462 | 0.816 | 0.0421  |
-| R2adj = 0.124                        | Primary Productivity   | 0.771 | 0.665 | 0.0247  |
-|                                      | Particle Concentration | 3.956 | 4.623 | \<0.001 |
-|                                      |                        |       |       |         |
-| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.822 | 0.919 | 0.0183  |
-| R2adj=.113                           | Particle Concentration | 0.970 | 6.208 | \<0.001 |
-|                                      |                        |       |       |         |
-| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 1.773 | 4.873 | \<0.001 |
-| R2adj=.133                           |                        |       |       |         |
-|                                      |                        |       |       |         |
-| **Collodaria Epipelagic**            | Salinity               | 0.925 | 2.267 | \<0.001 |
-| R2adj=0.16                           | O2                     | 2.015 | 2.217 | 0.002   |
-|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.002   |
-|                                      |                        |       |       |         |
-| **Foraminifera Epipelagic**          | Temperature            | 4.414 | 8.789 | \<0.001 |
-| R2adj=0.445                          | O2                     | 1.603 | 1.287 | 0.0111  |
-|                                      | Avg N Flux             | 2.512 | 3.396 | \<0.001 |
-|                                      | Primary Productivity   | 0.824 | 0.920 | 0.0153  |
-|                                      | Particle Concentration | 0.919 | 2.257 | \<0.001 |
-
-: Taxa-specific generalized additive models for different regions of the water column.
-
-![Partial effects of smooth terms in taxa-specific GAM models from the upper mesopelagic (200-500m).
-Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/07_upmeso-partial.pdf)
+Interestingly, Aulacanthidae had primary production as a significant predictor, yet there was not a clear association (Figure 7C).
+There was a fit for Collodaria in the epipelagic ($R^2_{adj}$ = 0.16), although there was a logit-like relationship where higher abundances tended to occur during higher DO conditions in the surface waters (Figure 7D).
+Castanellidae also had similarly poor fits in the epipelagic ($R^2_{adj}$ = 0.124) (Table 2, Figure 7E).
 
@@
 All these models had generally poor fits ($R^2_{adj}$ \< 0.25).
-Yet, for all upper mesopelagic models, particle concentration was a significant explanatory variable (Table 2, Figure 7).
-Carbon flux was significant for Acantharea and nitrogen flux was significant for both Acantharea and Coelodendridae (Table 2, Figure 7A,D).
+Yet, for all upper mesopelagic models, particle concentration was a significant explanatory variable (Table 2, Figure 8).
+Carbon flux was significant for Acantharea and nitrogen flux was significant for both Acantharea and Coelodendridae (Table 2, Figure 8A,D).
 The lower mesopelagic also had generally poor GAM fits for taxa specific models ($R^2_{adj}$ \<0.3), with the exception of Acantharea ($R^2_{adj}$ = 0.509).
-Acantharea in the lower mesopelagic was most clearly positively associated with particle concentration and nitrogen flux, as well as temperature to a slight degree (Figure 8A).
-For all Phaeodarians with a significant model, particle concentration was a main predictor variable (Table 2, Figure 8B-D).
-Aulacanthidae had the best fitting model of the Phaeodarians ($R^2_{adj}$ = 0.298), which also included mass flux as a statistically significant smooth (Figure 8B).
-
-![Partial effects of smooth terms in taxa-specific GAM models from the lower mesopelagic (500-1000m).
-Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/08_lomeso-partials.pdf)
+Acantharea in the lower mesopelagic was most clearly positively associated with particle concentration and nitrogen flux, as well as temperature to a slight degree (Figure 9A).
+For all Phaeodarians with a significant model, particle concentration was a main predictor variable (Table 2, Figure 9B-D).
+Aulacanthidae had the best fitting model of the Phaeodarians ($R^2_{adj}$ = 0.298), which also included mass flux as a statistically significant smooth (Figure 9B).

### 2024-01-29 manuscript/04_discussion.qmd
render with line numbers

@@
 It should be noted that previous work has suggested that avoidance behavior with the UVP is possible, at times likely, for visual and highly mobile zooplankton [@barth2022].
-Thus, the percent contribution reported here (42.7%) of Rhizaria to the total mesozooplankton community may be inflated due to under sampling of organisms such as Euphausiids and Chaetognaths which has quick escape responses.
+Thus, the percent contribution reported here (42.7%) of Rhizaria to the total mesozooplankton community may be inflated due to under sampling of organisms such as Euphausiids and Chaetognaths which have quick escape responses.
 Regardless, it is worth noting that in the same region, with data collected in 2012 and 2013 using similar calculation methods, @biard2016 estimated Rhizaria only contribute 15% of the total mesozooplankton community in the upper 500m.
@@
 In the present study, we noticed considerably higher Rhizarian abundance throughout the water column in 2019 compared to 2021.
-While this may have been driven by increased mass flux (Figure 3), more information is needed to truly understand the magnitude by which Rhizaria can vary interannually.
+While this may have been driven by increased mass flux, more information is needed to truly understand the magnitude by which Rhizaria can vary interannually.
+
+### Relationship to environmental parameters
+
+In general, the fit of most GAMs were moderate to poor.
+One possible reason for the poor fits may have been that for some taxa, conditions were not variable enough to capture a range of conditions at which they may exist.
+For instance, Collodaria were the most abundant taxa observed, yet the fit of their GAM was particularly poor.
+In studies which covered a wider range of parameters, Collodaria has been shown to strongly vary with changes in parameters such as temperature, chlorophyll-a, mixing, and water clarity [@biard2017; @biard2020].
+Alternatively, Acantharea had relatively good fitting GAMs.
+These taxa also had some of the largest variation from month to month on cruises.
+Thus, it may be that in the oligotrophic, the relatively stable conditions can support certain taxa while others are more sporadic.
+It should also be noted that due to the challenge of adequately sampling enough volume to overcome low-detection issues, GAMs were run on integrated data.
+However, variation with environmental parameters throughout the water column are likely, just not captured in the modelling aspect of this study.
+One consistent parameter which had significant positive associations was particle concentration.
+This observation is not surprising as most rhizaria likely to some extent engage in flux feeding.
 
@@
 Furthermore, while Collodaria were primarily absent from below 250m, there were a few instances of deeper colonies being observed.
-Global investigations of polycistine flux, suggest that deep-Collodaria in Oligotrophic regions may be a consequence of isothermal submersion [@boltovskoy2017].
+Global investigations of polycystine flux, suggest that deep-Collodaria in Oligotrophic regions may be a consequence of isothermal submersion [@boltovskoy2017].
 Another effectively exclusively epipelagic Rhizaria was the Phaeodarian family of Castanellidae.
 All Phaeodarians are thought to be fully heterotrophic [@nakamura2015], nonetheless a number of studies, including this one, report Castanellidae to be typically found in the lower epipelagic [@zasko2005; @biard2018; @biard2020].
-It should be considered that perhaps Castanellidae specializes in feeding sinking particles directly at the base of the epipelagic.
+It should be considered that perhaps Castanellidae specializes in feeding on sinking particles directly at the base of the epipelagic.
 Given it's smaller size [@nakamura2015], Castanellidae does not need a large diameter to efficiently flux feed at the typically particle rich region of the lower epipelagic.
@@
 Acantharea also had a bimodal distribution, with much larger abundances than Aulacanthidae or Foraminifera.
-Most prior studies of Acantharea vertical distribtion found them concentrated in near surface layers of the water column [@michaels1988 @zasko2005; @biard2020].
+Most prior studies of Acantharea vertical distribution found them concentrated in near surface layers of the water column [@michaels1988 @zasko2005; @biard2020].
 This would support the paradigm that large Acantharean abundances may be supported by their mixotrophic abilities [@michaels1995; @suzuki2015].
@@
 @decelle2013 proposed a hypothetical life cycle for cyst-bearing (strictly heterotrophic) Acantharea.
-This hypothesized life cycle suggests that epipelagic Acantharea are adult populations, which form cysts that sink into the mesopelagic, then reproduce and raise.
-Furthermore, given that horizontal transfer of symbionts between generations of Acantharea is unlikely, the newly spawned mesopelagic Acantharea are not necessarily required to rapidly return to the photic zone.
+This hypothesized life cycle suggests that epipelagic Acantharea are adult populations, which form cysts that sink into the mesopelagic, then reproduce and rise.
+Furthermore, given that horizontal transfer of symbionts between generations of Acantharea is unlikely due to their spawning behavior, the newly spawned mesopelagic Acantharea are not necessarily required to rapidly return to the photic zone [@decelle2012; @decelle2013].
 This hypothesis predicts that Acanthareans in the mesopelagic would be smaller [@decelle2013].
 @marsbrisbin2020 provided some support for this hypothesis, with a significant decrease in Acantharea sizes with depth.
-Although, the authors also observed low abundances in the mesopelagic and noted that the smaller sizes may be due to less food availability [@marsbrisbin2020].
-Since food is more scarce in the mesopelagic there is is less nutritional quality, yet flux feeders would likely grow larger to increase their feeding range [@biard2020].
-In the data collecting in this study, Acanthareans in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth (Figure 9, Wilcoxon rank sum text p-value \< 0.001).
+Although the authors also observed low abundances in the mesopelagic and noted that the smaller sizes may be due to lower food availability [@marsbrisbin2020].
+Since food is more scarce in the mesopelagi, nutritional quality lower, yet flux feeders would likely grow larger to increase their feeding range [@biard2020].
+In the data collecting in this study, Acanthareans in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth.
 This provides added support for the hypothesis that cyst-forming Acantharea may utilize different sections of the water column throughout their life cycle.
 However to further investigate this, more work is needed with higher temporal and taxonomic resolution.
-
-![Comparison of average sizes (ESD) amongst Rhizarian taxa which occurred throughout the water column.](images/09_size-comp.pdf)
 
@@
 However, once consumed and repackaged by larger Rhizaria, they can sink quicker and contribute more to overall flux [@michaels1988].
-Thus, Rhizaria may act as a ballasting mechanism.
+Thus, Rhizaria may act as an aggregation mechanism.
 However, this is largely speculation, to truly test this, more work is needed measuring Rhizarian flux.
@@
 However, their role in food webs is not well understood.
-While this study represents a step-forward in our understanding of Rhizarian ecology, there are several known aspects of Rhizarian and protist ecology which are not captured in this study.
-Continued research on Rhizaria is much needed to better understand their ecology.
+While this study represents a step-forward in our understanding of Rhizarian ecology, continued research on Rhizaria is much needed to better understand their ecology.
 Particularly extended descriptive work to capture interannual patterns.

### 2024-01-29 manuscript/main.qmd
render with line numbers

@@
 ---
-title: Rhizaria in the oligotrophic exhibit clear temporal and vertical variability.
+title: Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.
 bibliography: references.bib
@@
       phone: +1 (440) 668-8376
-  - name: Leo Blanco-Berical
+  - name: Leocadio Blanco-Berical
     affiliations:
     - name: Arizone State University
+      department: School of Ocean Futures
+      address: Bermuda Institute of Ocean Sciences 17 Biological Station, St Georges, GE 01
   - name: Rod Johnson
     affiliations: 
-    - name: Bermuda Institute of Ocean Sciences
-      address: Bermuda GE 01, UK
+    - name: Arizone State University
+      department: School of Ocean Futures
+      address: Bermuda Institute of Ocean Sciences 17 Biological Station, St Georges, GE 01
   - name: Joshua Stone
@@
     keep-tex: true
+abstract: |
+  Recently studies have shown that Rhizaria, a super-group of marine protists have a large role in pelagic ecosystems.
+They are unique in that they construct mineral tests out of silica, calcium carbonate, or strontium sulfate.
+As a consequence, Rhizaria can have large impacts on the ocean’s cycling of carbon and other elements.
+However, less is known about Rhizarian ecology or their role in the pelagic food-web.
+Some taxa, like certain Radiolarians are mixotrophic, hosting algal symbionts.
+While other taxa are flux-feeders or even predatory carnivores.
+Some prior research has suggested that Rhizaria will partition vertically in the water column, likely due to different trophic strategies.
+However, very few studies have investigated their populations over extended periods of time.
+In this study, we present data investigating Rhizarian abundance and vertical distribution from over a year of monthly cruises into the Sargasso Sea.
+This study represents the first quantification of Rhizaria throughout the mesopelagic zone in an oligotrophic system for an extended period of time.
+We use this data to investigate the hypothesis that Rhizaria taxonomic groups will partition due to trophic mode.
+We also investigate how their abundance varies in accordance with a variety of environmental parameters.
+Rhizaria abundance was quantified using an Underwater Vision Profiler (UVP5), an in-situ imaging device.
+Ultimately, we show that different Rhizaria taxa will have unique vertical distribution patterns.
+Models relating their abundance to environmental parameters have mixed results, yet particle concentration is a common predictive variable, supporting the importance of heterotrophy amongst many taxa.
 ---
@@
 # References:
-:::{#refs}
+
+::: {#refs}
 :::
+
+\newpage
+
+| **Model**                          | Term                   | edf    | F     | p-value |
+|-------------------|--------------|--------------|--------------|--------------|
+| **All Rhizaria Epipelagic**        | Salinity               | 2.0818 | 4.654 | \<0.001 |
+| $R^2_{adj}$= 0.42                  | DO                     | 0.8272 | 0.945 | 0.0136  |
+|                                    | Silica                 | 3.1206 | 15.75 | \<0.001 |
+|                                    | NO3                    | 2.9970 | 6.268 | \<0.001 |
+|                                    | Primary Productivity   | 1.8367 | 2.099 | \<0.001 |
+|                                    | Particle Concentration | 0.9282 | 2.566 | \<0.001 |
+|                                    |                        |        |       |         |
+| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6847 | 0.431 | 0.073   |
+| $R^2_{adj}$= 0.406                 | Avg Mass Flux          | 1.5426 | 0.923 | 0.045   |
+|                                    | Particle Concentration | 3.2957 | 20.89 | \<0.001 |
+|                                    |                        |        |       |         |
+| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 1.6632 | 1.824 | 0.002   |
+| $R^2_{adj}$= 0.603                 | Particle Concentration | 0.7694 | 0.662 | 0.027   |
+
+: Generalized Additive Model results for integrated total Rhizarian abundance in different regions of the water column.
+
+\newpage
+
+| Model                                | Term                   | edf   | F     | p-value |
+|-------------------|--------------|--------------|--------------|--------------|
+| **Acantharea Epipelagic**            | Salinity               | 2.264 | 4.113 | \<0.001 |
+| R2adj=0.53                           | O2                     | 2.579 | 6.712 | \<0.001 |
+|                                      | Avg Mass Flux          | 4.810 | 22.81 | \<0.001 |
+|                                      | Bacteria #/L           | 1.599 | 1.317 | 0.0117  |
+|                                      | Particle Concentration | 0.853 | 1.158 | 0.0087  |
+|                                      |                        |       |       |         |
+| **Acantharea Upper Mesopelagic**     | Avg C Flux             | 1.296 | 0.621 | 0.0439  |
+| R2adj=0.231                          | Avg N Flux             | 1.619 | 1.283 | 0.0026  |
+|                                      | Particle Concentration | 0.952 | 3.886 | \<0.001 |
+|                                      |                        |       |       |         |
+| **Acantharea Lower Mesopelagic**     | Temperature            | 2.076 | 4.155 | \<0.001 |
+| R2adj=0.509                          | Avg N Flux             | 1.494 | 1.216 | 0.0113  |
+|                                      | Primary Productivity   | 0.766 | 0.648 | 0.0253  |
+|                                      | Particle Concentration | 2.037 | 10.03 | \<0.001 |
+| **Aulacanthidae Epipelagic**         | Salinity               | 0.792 | 0.757 | 0.0241  |
+| R2adj=0.251                          | Avg N Flux             | 0.869 | 1.324 | 0.0018  |
+|                                      | Primary Productivity   | 2.312 | 7.704 | \<0.001 |
+|                                      | Particle Concentration | 2.008 | 2.200 | 0.0017  |
+|                                      |                        |       |       |         |
+| **Aulacanthidae Upper Mesopelagic**  | Particle Concentration | 2.832 | 9.802 | \<0.001 |
+| R2adj = 0.158                        |                        |       |       |         |
+|                                      |                        |       |       |         |
+| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.622 | 1.991 | 0.002   |
+| R2adj=0.298                          | Particle Concentration | 2.123 | 6.706 | \<0.001 |
+|                                      |                        |       |       |         |
+| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.653 | 13.06 | \<0.001 |
+| R2adj=0.2                            |                        |       |       |         |
+|                                      |                        |       |       |         |
+| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 0.972 | 6.248 | \<0.001 |
+| R2adj=.147                           |                        |       |       |         |
+|                                      |                        |       |       |         |
+| **Castanellidae Epipelagic**         | Avg C Flux             | 1.462 | 0.816 | 0.0421  |
+| R2adj = 0.124                        | Primary Productivity   | 0.771 | 0.665 | 0.0247  |
+|                                      | Particle Concentration | 3.956 | 4.623 | \<0.001 |
+|                                      |                        |       |       |         |
+| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.822 | 0.919 | 0.0183  |
+| R2adj=.113                           | Particle Concentration | 0.970 | 6.208 | \<0.001 |
+|                                      |                        |       |       |         |
+| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 1.773 | 4.873 | \<0.001 |
+| R2adj=.133                           |                        |       |       |         |
+|                                      |                        |       |       |         |
+| **Collodaria Epipelagic**            | Salinity               | 0.925 | 2.267 | \<0.001 |
+| R2adj=0.16                           | O2                     | 2.015 | 2.217 | 0.002   |
+|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.002   |
+|                                      |                        |       |       |         |
+| **Foraminifera Epipelagic**          | Temperature            | 4.414 | 8.789 | \<0.001 |
+| R2adj=0.445                          | O2                     | 1.603 | 1.287 | 0.0111  |
+|                                      | Avg N Flux             | 2.512 | 3.396 | \<0.001 |
+|                                      | Primary Productivity   | 0.824 | 0.920 | 0.0153  |
+|                                      | Particle Concentration | 0.919 | 2.257 | \<0.001 |
+: Taxa-specific generalized additive models for different regions of the water column.
+
+\newpage
+
+![Example images of different Rhizaria taxa. 4mm scale bar shown in lower right. All vignettes are same scale](images/01_taxa.pdf)
+
+\newpage
+
+![Environmental profiles across time-series of study period.
+Y axis shows depth in meters.
+(A) Temperature.
+(B) Salinity.
+(C) Dissolved Oxygen.
+(D) In-situ chlorophyll fluorescence.
+(E) Particle concenctrion (184 - 450$\mu m$).
+(F) Bacteria Abundance.
+(G) Silica.
+(H) Nitrate](images/02-environmental.pdf)
+
+\newpage
+
+![Primary productivity (A) and flux estimates from total mass (B), carbon (C), and nitrogen (D).
+Values are from monthly cruises with month displayed in corresponding colors.
+Absent data are shown by grey bar.](images/03_prod-flux.pdf)
+
+\newpage
+
+![Average abundance of Rhizaria in 25m bins, across entire study period.
+Shown are total Rhizaria (A), Acantharea (B), Collodaria (C), Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae (G), Coelodendridae (H).](images/04_vertical.pdf)
+
+\newpage
+
+![Seasonality of Rhizarian integrated abundance for the epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower mesopelagic (500-1000m) (E-F).
+Left panels (A,C,E) display total integrated abundance per monthly cruise colored by month.
+Right panels (B, D, F) display community composition of each total abundance.](images/05_seasonality.pdf)
+
+\newpage
+
+![Comparison of average sizes (ESD) amongst Rhizarian taxa which occurred throughout the water column.](images/09_size-comp.pdf)
+
+\newpage
+
+![Partial effects of smooth terms in taxa-specific GAM models from the epipelagic (0-200m).
+Effects are grouped by taxa; Acantharea (A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae (E).](images/06_epi-partials.pdf)
+
+\newpage
+
+![Partial effects of smooth terms in taxa-specific GAM models from the upper mesopelagic (200-500m).
+Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/07_upmeso-partial.pdf)
+
+\newpage
+
+![Partial effects of smooth terms in taxa-specific GAM models from the lower mesopelagic (500-1000m).
+Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/08_lomeso-partials.pdf)
+
+\newpage
+
+end.

### 2024-01-29 manuscript/main.tex
render with line numbers

@@
 \hypersetup{
-  pdftitle={Rhizaria in the oligotrophic exhibit clear temporal and vertical variability.},
-  pdfauthor={Alex Barth*; Leo Blanco-Berical; Rod Johnson; Joshua Stone},
+  pdftitle={Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.},
+  pdfauthor={Alex Barth*; Leocadio Blanco-Berical; Rod Johnson; Joshua Stone},
   colorlinks=true,
@@
 
-\title{Rhizaria in the oligotrophic exhibit clear temporal and vertical
-variability.}
-\author{Alex Barth* \and Leo Blanco-Berical \and Rod Johnson \and Joshua
-Stone}
+\title{Rhizaria in the oligotrophic ocean exhibit clear temporal and
+vertical variability.}
+\author{Alex Barth* \and Leocadio Blanco-Berical \and Rod
+Johnson \and Joshua Stone}
 \date{}
@@
 % Title and subtitle
-{\large\bfseries\nohyphens{Rhizaria in the oligotrophic exhibit clear
-temporal and vertical variability.}}\\[2\baselineskip]
+{\large\bfseries\nohyphens{Rhizaria in the oligotrophic ocean exhibit
+clear temporal and vertical variability.}}\\[2\baselineskip]
 
@@
 , 
- {\large{Leo Blanco-Berical}}{\textsuperscript{2}}%
+ {\large{Leocadio Blanco-Berical}}{\textsuperscript{2}}%
 %
 , 
- {\large{Rod Johnson}}{\textsuperscript{3}}%
+ {\large{Rod Johnson}}{\textsuperscript{2}}%
 %
@@
 %
-%
-\par\hangindent=1em\hangafter=1%
-%
-{3}.~{Bermuda Institute of Ocean Sciences}%
+, %
+{School of Ocean Futures}%
 %
@@
 , %
-{Bermuda GE 01, UK}%
+{Bermuda Institute of Ocean Sciences 17 Biological Station, St Georges,
+GE 01}%
 %
@@
 
+%%%%%%%%% Abstract
+\begin{abstract}
+Recently studies have shown that Rhizaria, a super-group of marine
+protists have a large role in pelagic ecosystems. They are unique in
+that they construct mineral tests out of silica, calcium carbonate, or
+strontium sulfate. As a consequence, Rhizaria can have large impacts on
+the ocean's cycling of carbon and other elements. However, less is known
+about Rhizarian ecology or their role in the pelagic food-web. Some
+taxa, like certain Radiolarians are mixotrophic, hosting algal
+symbionts. While other taxa are flux-feeders or even predatory
+carnivores. Some prior research has suggested that Rhizaria will
+partition vertically in the water column, likely due to different
+trophic strategies. However, very few studies have investigated their
+populations over extended periods of time. In this study, we present
+data investigating Rhizarian abundance and vertical distribution from
+over a year of monthly cruises into the Sargasso Sea. This study
+represents the first quantification of Rhizaria throughout the
+mesopelagic zone in an oligotrophic system for an extended period of
+time. We use this data to investigate the hypothesis that Rhizaria
+taxonomic groups will partition due to trophic mode. We also investigate
+how their abundance varies in accordance with a variety of environmental
+parameters. Rhizaria abundance was quantified using an Underwater Vision
+Profiler (UVP5), an in-situ imaging device. Ultimately, we show that
+different Rhizaria taxa will have unique vertical distribution patterns.
+Models relating their abundance to environmental parameters have mixed
+results, yet particle concentration is a common predictive variable,
+supporting the importance of heterotrophy amongst many taxa.    
+\end{abstract}
+
 \vfill
@@
 \end{minipage}  \end{titlepage}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-Rhizaria are an extremely diverse super-group of single celled organisms
-consisting of several phyla including retaria (foraminifera and
-radiolaria) and cercozoa. These organisms exist in a wide range of
+Rhizaria are an extremely diverse super-group of single-celled organisms
+consisting of several phyla including Retaria (foraminifera and
+radiolaria) and Cercozoa. These organisms exist in a wide range of
 habitats and are widely represented in plankton communities throughout
@@
 largely understudied throughout the 20th century. The bulk of modern
-plankton research has focused hard-bodied crustacea which are
+plankton research has focused on hard-bodied crustacea which are
 numerically dominant and easily sampled with nets and preservatives.
 Fragile organisms like Rhizaria, were difficult to adequately study as
-they can be destroyed through standard zooplankton sampling techniques.
-A number of studies in the late 1900s did employ alternative techniques
-to quantify Rhizaria including diaphragm pumps (Michaels, 1988) or
-blue-water SCUBA collections (Bijma et al., 1990; Caron et al., 1995;
-Caron and Be, 1984). However, the bulk of Rhizaria research was
-constrained to sediment traps or paleontological studies of sediment
-(Boltovskoy et al., 1993; Takahashi et al., 1983). Only recently has the
-advent of molecular techniques and in-situ imaging tools ignited a
-renewed focus on Rhizaria in pelagic ecosystems (Caron, 2016).
+they can be destroyed through standard zooplankton sampling techniques
+and preserve poorly. A number of studies in the late 1900s did employ
+alternative techniques to quantify Rhizaria including diaphragm pumps
+(Michaels, 1988) or blue-water SCUBA collections (Bijma et al., 1990;
+Caron et al., 1995; Caron and Be, 1984). However, the bulk of Rhizaria
+research was constrained to sediment traps or paleontological studies of
+sediment (Boltovskoy et al., 1993; Takahashi et al., 1983). Only
+recently has the advent of molecular techniques and in-situ imaging
+tools ignited a renewed focus on Rhizaria in pelagic ecosystems (Caron,
+2016).
 
@@
 of Rhizaria to the total standing stock of marine carbon. Due to their
-large sizes, ability to concentrate smaller particles (ballasting), and
-the unique structure of their mineral skeletons, Rhizaria have the
-potential to massively influence ocean biogeochemical cycling. A number
-of studies have made large advances in estimating the contribution of
-Rhizaria to ocean cycling of carbon (Gutierrez-Rodriguez et al., 2019;
-Ikenoue et al., 2019; Lampitt et al., 2009; Stukel et al., 2018), silica
-(Biard et al., 2018; Llopis Monferrer et al., 2021), and strontium
-(Decelle et al., 2013). Still, Rhizarian ecological roles are not well
-understood (Biard, 2022a). This is a major challenge as it is critical
-to understand the ecological role of plankton to fully incorporate them
+large sizes, ability to concentrate smaller particles and the unique
+structure of their mineral skeletons, Rhizaria have the potential to
+massively influence ocean biogeochemical cycling. A number of studies
+have made large advances in estimating the contribution of Rhizaria to
+ocean cycling of carbon (Gutierrez-Rodriguez et al., 2019; Ikenoue et
+al., 2019; Lampitt et al., 2009; Stukel et al., 2018), silica (Biard et
+al., 2018; Llopis Monferrer et al., 2021), and strontium (Decelle et
+al., 2013). Still, Rhizarian ecological roles are not well understood
+(Biard, 2022a). This is a major challenge as it is critical to
+understand the ecological role of plankton to fully incorporate them
 into biological oceanographic models.
@@
 that they can be effective predators (Anderson and Bé, 1976; Gaskell et
-al., 2019), majoritively consuming live copepods (Caron and Be, 1984).
+al., 2019), mainly consuming live copepods (Caron and Be, 1984).
 Radiolaria have several lineages all which have some taxa who are well
@@
 oligotrophic ocean, Blanco-Bercial et al. (2022) also noted that the
-protist community, including Rhizaria partition along an autotroph and
+protist community, including Rhizaria, partition along an autotroph and
 mixotroph to heterotroph gradient with increasing depth in the water
@@
 filtered to only include data collected in the BATS region, far offshore
-of Bermuda into the Sargasso Sea (Supplemental Figure 1). In general,
-casts extended to either 200m, 500m, or 1200m deep, with a few extended
-into the bathypelagic (4500m). However, Rhizaria were only typically
-found in large abundances throughout the epipelagic and mesopelagic
-zones. As such, we limit this study to results from the upper 1000m of
-the water column.
+of Bermuda into the Sargasso Sea (approximately 31.0\(^oN\)-32.5\(^oN\),
+64.25\(^oW\)-63\(^oW\); Supplemental Figure 1). In general, casts
+extended to either 200m, 500m, or 1200m deep, with a few extended into
+the bathypelagic (4500m). However, Rhizaria were only typically found in
+large abundances throughout the epipelagic and mesopelagic zones. As
+such, we limit this study to results from the upper 1000m of the water
+column.
 
@@
 Individual vignettes of Rhizaria images were identified using the
-classification platform Ecotaxa {[}cite{]}. Data were pre-sorted
-according to a . Taxonomic classification were done based on morphology
-exclusively. While there are sparse taxonomic guides for in-situ images
-of rhizaria, identification largely relied on descriptions in (Nakamura
-and Suzuki, 2015; Suzuki and Not, 2015; and Biard and Ohman, 2020).
-Using the aforementioned sources and publicly available ecotaxa
-projects, I constructed a guide which can be found here:
-\url{https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf}.
+classification platform Ecotaxa (Picheral et al., n.d.). Data were
+pre-sorted utilizing a random-forest classifier and pre-trained learning
+set. Taxonomic classification were done based on morphology exclusively.
+While there are sparse taxonomic guides for in-situ images of rhizaria,
+identification largely relied on descriptions in (Nakamura and Suzuki,
+2015; Suzuki and Not, 2015; and Biard and Ohman, 2020). Using the
+aforementioned sources and publicly available ecotaxa projects, I
+constructed a guide which can be found here:
+\protect\hyperlink{0}{https://thealexbarth.github.io/media//Project\_Items/Oligotrophic\_Community/ecotaxa\_UVP-guide-stone-lab.pdf}.
 Broadly, Rhizaria were classified as Foraminifera, Radiolarians
@@
 
-\begin{figure}
-
-{\centering \includegraphics{images/01_taxa.pdf}
-
-}
-
-\caption{Example images of different Rhizaria taxa. 4mm scale bar shown
-in lower right. All vignettes are same scale}
-
-\end{figure}
-
 The UVP5 samples at \textasciitilde15Hz rate as it descends the water
@@
 poor-sampling at the small size range (Barth and Stone, in review;
-Lombard et al., 2019) . However, it was evident from this analysis that
+Lombard et al., 2019). However, it was evident from this analysis that
 all size ranges were adequately sampled across the size range
@@
 exact depth at which a particle is recorded, however to estimate
-abundance, observations must be binned over fixed depth intervals. In
-our deployments had variable descent depths and speeds with more casts
+abundance, observations must be binned over fixed depth intervals. Our
+deployments had variable descent depths and speeds with more casts
 descending to 500m than 1000m and descents quicker through the
-epipelagic than the mesopelagic {[}see Barth and Stone (2022) for an
+epipelagic than the mesopelagic (see Barth and Stone (2022) for an
 extended discussion of UVP5 data processing). For the present study,
 Rhizarian abundances were estimated in 25m vertical bins, which offer a
-moderate sampling volume (average 0.948\(m^3\) in the epipelagic and
-0.589\(m^3\) in the mesopelagic) while still maintaining ecologically
-relevant widths. However, concentrations in a 25m bin would need to be
-greater than 2.428 indv. \(m^{-3}\) and 3.912 indv. \(m^{-3}\), in the
-epipelagic and mesopelagic respectively, to fall below a 10\%
-non-detection risk (Barth and Stone, in review; Benfield et al., 1996).
-Because we typically observed many rhizaria taxa below these
-concentrations, we present the 25m binned data to visualize broad-scale
-average distributions. For quantifying and modelling Rhizaria
-abundances, we present integrated abundance estimates, with each cast.
-Due to the variable descent depths of the UVP, data are categorized as
-epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic
-(500-1000m). The average sampling volume integrated through these
-regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\), with
-non-detection thresholds at 0.30 indv. \(m^{-2}\), 0.33 indv.
+moderate sampling volume per bin (average 0.948\(m^3\) in the epipelagic
+and 0.589\(m^3\) in the mesopelagic) while still maintaining
+ecologically relevant widths. However, concentrations in a 25m bin would
+need to be greater than 2.428 indv. \(m^{-3}\) and 3.912 indv.
+\(m^{-3}\), in the epipelagic and mesopelagic respectively, to fall
+below a 10\% non-detection risk {[}Benfield et al. (1996); Barth and
+Stone (in review){]}. Because we typically observed many rhizaria taxa
+below these concentrations, we present the 25m binned data to visualize
+broad-scale average distributions. For quantifying and modelling
+Rhizaria abundances, we present integrated abundance estimates, with
+each cast. Due to the variable descent depths of the UVP, data are
+categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and
+lower mesopelagic (500-1000m). The average sampling volume integrated
+through these regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\),
+with non-detection thresholds at 0.30 indv. \(m^{-2}\), 0.33 indv.
 \(m^{-2}\), and 0.20 indv. \(m^{-2}\) respectively. All UVP data
@@
 upper mesopelagic, and lower mesopelagic. In cases where observations
-were too sparse for a given taxonomic grouping, models were not run.
+were too sparse for a given taxonomic grouping, models were not run. All
+code and full models are available in code, as well as intermediate data
+products at \url{https://github.com/TheAlexBarth/RhizariaSeasonality}.
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{images/02-environmental.pdf}
-
-}
-
-\caption{Environmental profiles across time-series of study period. Y
-axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved
-Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion
-(184 - 450\(\mu m\)). (F) Bacteria Abundance. (G) Silica. (H) Nitrate}
-
-\end{figure}
-
 Overall, particle concentration was high near the surface during the
-2021 spring bloom, then sank throughout the water column attenuating
-throughout the lower epipelagic (Figure 2E). Similarly, bacteria
-abundance was closely linked to overall productivity, althrough there
-was a more consistent layer near the top of the mesopelagic
-(\textasciitilde250m) (Figure 2F). Concurrent with the secondary fall
-production peak, there was also higher particle concentration and
-bacterial abundance in the later summer and early fall. Interestingly,
-while primary productivity estimates from July-August were not that
-different between 2019 and 2021 (Figure 4A), chlorophyll-a florescence,
-particle concentration, and bacterial abundance were much higher in
-2019's summer/fall (Figure 2D-F). Inorganic nutrients (\(Si\) and
-\(NO_3\)) were generally well stratified, with low concentrations in the
-epipelagic and increasing throughout the mesopelagic. However, both
-nutrients did vary vertically in accordance with the 2021 February
-downwelling and spring mixing period (Figure 2G-H). Additionally, in the
-late fall of 2021, \(Si\) concentrations were slightly elevated in the
-mid-mesopelagic (Figure 2G).
+2021 spring bloom, then moved deeper throughout the water column
+attenuating throughout the lower epipelagic (Figure 2E). Similarly,
+bacteria abundance was closely linked to overall productivity, althrough
+there was a more consistent moderate-abundance layer near the top of the
+mesopelagic (\textasciitilde250m) (Figure 2F). Concurrent with the
+secondary fall production peak, there was also higher particle
+concentration and bacterial abundance in the later summer and early
+fall. Interestingly, while primary productivity estimates from
+July-August were not that different between 2019 and 2021 (Figure 4A),
+chlorophyll-a florescence, particle concentration, and bacterial
+abundance were much higher in 2019's summer/fall (Figure 2D-F).
+Inorganic nutrients (\(Si\) and \(NO_3\)) were generally well
+stratified, with low concentrations in the epipelagic and increasing
+throughout the mesopelagic. However, both nutrients did vary vertically
+in accordance with the 2021 February downwelling and spring mixing
+period (Figure 2G-H). Additionally, in the late fall of 2021, \(Si\)
+concentrations were slightly elevated in the mid-mesopelagic (Figure
+2G).
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{images/03_prod-flux.pdf}
-
-}
-
-\caption{Primary productivity (A) and flux estimates from total mass
-(B), carbon (C), and nitrogen (D). Values are from monthly cruises with
-month displayed in corresponding colors. Absent data are shown by grey
-bar.}
-
-\end{figure}
-
 \hypertarget{rhizaria-abundance-and-distribution}{%
@@
 3). Copepods were the second most abundant, comprising 35.5\% and all
-living mesozooplankton were 22\%. The large contribution of Rhizaria to
-the mesozooplankton community is most prominent in the epipelagic, where
-they accounted for 47\% of all mesozooplankton. In the mesopelagic
-rhizaria were a smaller fraction, at 38\% in the upper layers (200-500m)
-and 37\% in the deeper mesopelagic (500-1000m).
-
-\begin{figure}
-
-{\centering \includegraphics{images/04_vertical.pdf}
-
-}
-
-\caption{Average abundance of Rhizaria in 25m bins, across entire study
-period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C),
-Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae
-(G), Coelodendridae (H).}
-
-\end{figure}
+other living mesozooplankton were 22\%. The large contribution of
+Rhizaria to the mesozooplankton community is most prominent in the
+epipelagic, where they accounted for 47\% of all mesozooplankton. In the
+mesopelagic rhizaria were a smaller fraction, at 38\% in the upper
+layers (200-500m) and 37\% in the deeper mesopelagic (500-1000m).
 
@@
 with secondary, wider peak occurring in the mid mesopelagic (Figure 4A).
-Although, variation in depth binned abundance was large, likely due to
-seasonal variability but also increased from the detection-risk
-described in the methods. The vertical distribution pattern and
-abundance varied considerably across taxonomic groups. Radiolarians were
-some fo the most abundant taxa observed, particularly in the epipelagic
-(Figure 4, Figure 5B). This pattern was led by Collodaria, whose
-colonies were abundant in the upper epipelagic and declined into the top
-of the mesopelagic (Figure 4C). Acantharea displayed a bimodal
-distribution accounting for a large portion of the total Rhizaria
-pattern (Figure 4B, Figure 5). Foraminifera had a similar bimodal
-distribution, yet their overall average densities were much lower and
-spread wider throughout the mesopelagic (FIgure 4E). Phaeodarian
-families exhibited a wide range of vertical distribution patterns. The
-most abundant, Aulacanthidae, also had a bimodal pattern but the density
-was highest in the lower mesopelagic (Figure 4D). Aulosphaeridae had low
-average density and was nearly homogenously distributed throughout the
-water column, although slightly lower int eh epipelagic (Figure 4F).
-Castanellidae were the only Phaeodarian who appeared to be effectively
-restricted to the photic zone (Figure 4G). Alternatively, Coelodendridae
-primarily occurred in the lower mesopelagic (Figure 4H). A few
-individuals from the families Tuscaroridae and Medusettidae were also
-observed in the mesopelagic, yet they were much rarer (data not shown).
+Variation in depth binned abundance was large, likely due to seasonal
+variability but also increased from the detection-risk described in the
+methods. The vertical distribution pattern and abundance varied
+considerably across taxonomic groups. Radiolarians were some of the most
+abundant taxa observed, particularly in the epipelagic (Figure 4, Figure
+5B). This pattern was led by Collodaria, whose colonies were abundant in
+the upper epipelagic and declined into the top of the mesopelagic
+(Figure 4C). Acantharea displayed a bimodal distribution accounting for
+a large portion of the total Rhizaria pattern (Figure 4B, Figure 5).
+Foraminifera had a similar bimodal distribution, yet their overall
+average densities were much lower and spread wider throughout the
+mesopelagic (Figure 4E). Phaeodarian families exhibited a wide range of
+vertical distribution patterns. The most abundant, Aulacanthidae, also
+had a bimodal pattern but the density was highest in the lower
+mesopelagic (Figure 4D). Aulosphaeridae had low average density and was
+nearly homogeneously distributed throughout the water column, although
+slightly lower in the epipelagic (Figure 4F). Castanellidae were the
+only Phaeodarian who appeared to be effectively restricted to the photic
+zone (Figure 4G). Alternatively, Coelodendridae primarily occurred in
+the lower mesopelagic (Figure 4H). A few individuals from the families
+Tuscaroridae and Medusettidae were also observed in the mesopelagic, yet
+they were much rarer (data not shown).
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{images/05_seasonality.pdf}
-
-}
-
-\caption{Seasonality of Rhizarian integrated abundance for the
-epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower
-mesopelagic (500-1000m) (E-F). Left panels (A,C,E) display total
-integrated abundance per monthly cruise colored by month. Right panels
-(B, D, F) display community composition of each total abundance.}
-
-\end{figure}
+\hypertarget{body-size-throughout-the-water-column.}{%
+\subsubsection{Body size throughout the water
+column.}\label{body-size-throughout-the-water-column.}}
+
+Very few taxa had consistent distributions throughout the water column.
+Only Acanthrea, Foraminifera, Aulacanthidae, and Aulosphaeridae were
+consistently abundant in the epipelagic and mesopelagic. To investigate
+if the populations or morphologies shifted throughout the water column,
+we compared the sizes (Equivalent Spherical Diameters, ESD) between
+mesopelagic and epipelagic groups for each taxa. All groups were
+significantly different on average (Wilcox Rank Sum p-value
+\textless0.001). Acantharea were smaller, on average in the mesopelagic
+while all other taxa tended to be larger (Figure 6).
 
@@
 
+GAMs for individual taxa were much less consistent in their fits (Table
+2). This is likely in part due to the high number of non-observations
+for certain taxa. Note that due to low abundances, GAMs were not
+constructed for Tuscaroridae or Medusettidae. Furthermore no significant
+terms were found for a model with Aulosphaeridae in the epipelagic nor
+Foraminifera in the mesopelagic.
+
+Epipelagic Acantharea were explained by several predictor variables and
+had a good fit (\(R^2_{adj}\) = 0.53, Table 2). Most notable smooths
+were mass flux and particle concentration, which had a weak positive
+association (Figure 7A), with July 2021 as a clear outlier where
+Acantharean abundances were high in the epipelagic despite lower fluxes
+and particle concentrations (Figure 5). Foraminifera had a good fitting
+GAM in the epipelagic (\(R^2_{adj}\) = 0.445). There were several
+significant explanatory variables, although the clearest pattern was
+observed of high temperatures associated with more Foraminifera
+abundance (Table 2, Figure 7B). Epipelagic Aulacanthidae similarly had
+several predictor variables which were significant, including both water
+quality parameters and particle/flux predictors (Table 2).
+Interestingly, Aulacanthidae had primary production as a significant
+predictor, yet there was not a clear association (Figure 7C). There was
+a fit for Collodaria in the epipelagic (\(R^2_{adj}\) = 0.16), although
+there was a logit-like relationship where higher abundances tended to
+occur during higher DO conditions in the surface waters (Figure 7D).
+Castanellidae also had similarly poor fits in the epipelagic
+(\(R^2_{adj}\) = 0.124) (Table 2, Figure 7E).
+
+In the upper mesopelagic (200-500m), abundances were generally low
+(Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae,
+Aulosphaeridae, and Coelodendridae (Table 2). All these models had
+generally poor fits (\(R^2_{adj}\) \textless{} 0.25). Yet, for all upper
+mesopelagic models, particle concentration was a significant explanatory
+variable (Table 2, Figure 8). Carbon flux was significant for Acantharea
+and nitrogen flux was significant for both Acantharea and Coelodendridae
+(Table 2, Figure 8A,D). The lower mesopelagic also had generally poor
+GAM fits for taxa specific models (\(R^2_{adj}\) \textless0.3), with the
+exception of Acantharea (\(R^2_{adj}\) = 0.509). Acantharea in the lower
+mesopelagic was most clearly positively associated with particle
+concentration and nitrogen flux, as well as temperature to a slight
+degree (Figure 9A). For all Phaeodarians with a significant model,
+particle concentration was a main predictor variable (Table 2, Figure
+9B-D). Aulacanthidae had the best fitting model of the Phaeodarians
+(\(R^2_{adj}\) = 0.298), which also included mass flux as a
+statistically significant smooth (Figure 9B).
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{overall-rhizarian-abundance-and-patterns}{%
+\subsubsection{Overall Rhizarian abundance and
+patterns}\label{overall-rhizarian-abundance-and-patterns}}
+
+In the epipelagic Rhizaria exhibited a notable seasonal pattern.
+Rhizarian abundances were higher in the summer months and lower during
+the winter. During a prior time period, Blanco-Bercial et al. (2022)
+noted that there is considerable seasonality in the community
+composition of all protists. Despite the seasonality of total Rhizarian
+abundance, community composition was relatively consistent, with
+Collodaria representing the bulk of the community. It should be noted
+that the overall taxonomic resolution of the UVP5 is fairly low, so
+there may be a switching of species within the broad groups identified
+in this study which were not captured. Throughout the mesopelagic,
+month-to-month variation in 2021 was relatively low. Again this is
+consistent with observations from metabarcoding of the whole protist
+community in the same study region (Blanco-Bercial et al., 2022). This
+finding is not surprising as the overall seasonal variation in
+environmental conditions in this region were low.
+
+Overall Rhizaria were the most commonly identified group of
+mesozooplankton throughout the study period. We do note that the UVP5
+commonly captures Trichodesmium colonies, yet these were excluded in
+this comparison as they are strictly autotrophs. It should be noted that
+previous work has suggested that avoidance behavior with the UVP is
+possible, at times likely, for visual and highly mobile zooplankton
+(Barth and Stone, 2022). Thus, the percent contribution reported here
+(42.7\%) of Rhizaria to the total mesozooplankton community may be
+inflated due to under sampling of organisms such as Euphausiids and
+Chaetognaths which have quick escape responses. Regardless, it is worth
+noting that in the same region, with data collected in 2012 and 2013
+using similar calculation methods, Biard et al. (2016) estimated
+Rhizaria only contribute 15\% of the total mesozooplankton community in
+the upper 500m. Likely, Rhizaria display considerable interannual
+variability. In the present study, we noticed considerably higher
+Rhizarian abundance throughout the water column in 2019 compared to
+2021. While this may have been driven by increased mass flux, more
+information is needed to truly understand the magnitude by which
+Rhizaria can vary interannually.
+
+\hypertarget{relationship-to-environmental-parameters}{%
+\subsubsection{Relationship to environmental
+parameters}\label{relationship-to-environmental-parameters}}
+
+In general, the fit of most GAMs were moderate to poor. One possible
+reason for the poor fits may have been that for some taxa, conditions
+were not variable enough to capture a range of conditions at which they
+may exist. For instance, Collodaria were the most abundant taxa
+observed, yet the fit of their GAM was particularly poor. In studies
+which covered a wider range of parameters, Collodaria has been shown to
+strongly vary with changes in parameters such as temperature,
+chlorophyll-a, mixing, and water clarity (Biard et al., 2017; Biard and
+Ohman, 2020). Alternatively, Acantharea had relatively good fitting
+GAMs. These taxa also had some of the largest variation from month to
+month on cruises. Thus, it may be that in the oligotrophic, the
+relatively stable conditions can support certain taxa while others are
+more sporadic. It should also be noted that due to the challenge of
+adequately sampling enough volume to overcome low-detection issues, GAMs
+were run on integrated data. However, variation with environmental
+parameters throughout the water column are likely, just not captured in
+the modelling aspect of this study. One consistent parameter which had
+significant positive associations was particle concentration. This
+observation is not surprising as most rhizaria likely to some extent
+engage in flux feeding.
+
+\hypertarget{vertical-structure-and-trophic-roles}{%
+\subsubsection{Vertical Structure and Trophic
+Roles}\label{vertical-structure-and-trophic-roles}}
+
+In this study we present a clear pattern of vertical zonation between
+different Rhizaria groups. Largely, the taxonomic composition and
+vertical positioning were similar to Rhizaria zonation in the California
+Current Ecosystem (Biard and Ohman, 2020). It should be noted however,
+that the secondary abundance peak reported in the present study is
+lower. This is likely due to the more oligotrophic nature of the study
+region, were the euphotic zone penetrates deeper into the water column.
+Most prevalent in the epipelagic were Collodaria. These mixotrophic
+Radiolarians have long been reported to contribute to primary
+productivity in the euphotic zone (Dennett et al., 2002; Michaels et
+al., 1995). Collodaria are thought to be particularly successful
+globally in oligotrophic regions due to their photosymbiotic
+relationships (Biard et al., 2017, 2016). We observed the highest
+abundance of Collodaria during June 2021, supporting the notion they can
+thrive during the typically low-nutrient conditions of summer
+stratification. However, Collodaria also increased during the spring
+mixing period, suggesting that they can thrive during conditions which
+may typically be thought to favor autotrophs. Furthermore, while
+Collodaria were primarily absent from below 250m, there were a few
+instances of deeper colonies being observed. Global investigations of
+polycystine flux, suggest that deep-Collodaria in Oligotrophic regions
+may be a consequence of isothermal submersion (Boltovskoy, 2017).
+Another effectively exclusively epipelagic Rhizaria was the Phaeodarian
+family of Castanellidae. All Phaeodarians are thought to be fully
+heterotrophic (Nakamura and Suzuki, 2015), nonetheless a number of
+studies, including this one, report Castanellidae to be typically found
+in the lower epipelagic (Biard et al., 2018; Biard and Ohman, 2020;
+Zasko and Rusanov, 2005). It should be considered that perhaps
+Castanellidae specializes in feeding on sinking particles directly at
+the base of the epipelagic. Given it's smaller size (Nakamura and
+Suzuki, 2015), Castanellidae does not need a large diameter to
+efficiently flux feed at the typically particle rich region of the lower
+epipelagic. Both Castanelldiae and Collodaria had poor fitting GAMs.
+This is somewhat of a surprise for Collodaria who had large abundance.
+However, given the consistency of their abundance, it may be that this
+study did not capture a wide enough range of conditions for describing
+Collodaria's preferred niche.
+
+The mesopelagic generally was home to known heterotrophic organisms,
+particularly for those which were constrained to exclusively occupy
+deeper waters. This is consistent with Blanco-Bercial et al. (2022)'s
+observation of an auto-/mixotroph to heterotroph gradient in the protist
+community. The upper mesopelagic interestingly had relatively low total
+abundance. This low-abundance region likely reflects the dynamics of
+productivity and export throughout the water column. While productivity
+and thus sinking particles for flux feeders are high in the euphotic
+zone, much of this is attenuated throughout the epipelagic. So, while
+the base of the epipelagic may provide a rich feeding environment for
+Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F),
+the region from 200-500m might be otherwise food poor. Perhaps it is
+more advantageous for Rhizaria to situate deeper, in darker regions of
+the twilight zone. Also it should be noted that Phaeodarians utilize
+silica to build their opaline tests, and silica concentrations started
+to increase around 500m (Figure 2G). Although \(Si\) was not a
+significant smooth for any taxa-specific model, this lack of association
+might be due to the overall lack of variation of \(Si\) between
+integrated abundance of each cast. Aulosphaeridae was only found to have
+significant relationships, although weak fits, to particle concentration
+in the mesopelagic. In our study, while consistently observed, overall
+abundances of Aulosphaeridae were very low. In the Pacific Ocean, on
+California's Coast, much higher abundances of Aulosphaeridae have been
+reported (Biard and Ohman, 2020; Zasko and Rusanov, 2005) and they have
+massive potential to impact silica export (Biard et al., 2018).
+Coelodendridae were also seemingly restricted to the deeper section of
+the mesopelagic. This is interesting given that in the California
+Current, (Biard and Ohman, 2020) found a bimodal distribution in
+Coelodendridae. There are several morphotypes corresponding to different
+taxa of Coelodendridae (Biard and Ohman, 2020; Nakamura and Suzuki,
+2015). So it may be that only a few types of Coelodendridae were
+observed in this study, while the epipelagic variety was not.
+Alternatively, the lower epipelagic of the California Current may
+provide adequate habitat for Coelodendridae, which is not available in
+the oligotrophic Sargasso Sea.
+
+A number of taxa were found to have a bimodal distribution, with
+considerable populations in both the epipelagic and mesopelagic.
+Aulacanthidae had a bimodal distribution, although abundances were
+highest in the lower mesopelagic. Foraminifera also had a bimodal
+distribution. Some lineages of Foraminifera are known to host
+photosymbionts (Biard, 2022a; Kimoto, 2015), however they are also
+efficient predators commonly seen throughout the mesopelagic (Caron and
+Be, 1984; Gaskell et al., 2019). Thus it is not surprising to find their
+presence in both locations of the water column. Foraminifera are also
+known to vary their vertical distribution across their life cycle in
+phase with lunar cycles (Biard, 2022a; Bijma et al., 1990; Gaskell et
+al., 2019; Kimoto, 2015). However, the sampling scheme of the BATS
+program does not capture this frequency and was not investigated in the
+present study.
+
+Acantharea also had a bimodal distribution, with much larger abundances
+than Aulacanthidae or Foraminifera. Most prior studies of Acantharea
+vertical distribution found them concentrated in near surface layers of
+the water column Zasko and Rusanov (2005). This would support the
+paradigm that large Acantharean abundances may be supported by their
+mixotrophic abilities (Michaels et al., 1995; Suzuki and Not, 2015).
+While the UVP5 images cannot distinguish between mixotrophic and
+heterotrophic Acantharea, the GAMs constructed for Acantharean abundance
+found positive associations with particle concentration and mass flux,
+suggesting a higher reliance on heterotrophy. Recently Mars Brisbin et
+al. (2020) described apparent predator behavior amongst near-surface
+Acantharea. Thus it is likely that epipelagic Acantharea may commonly be
+heterotrophic. Yet, it should be noted in the Sargasso Sea, both
+heterotrophic and symbiotic lineages of Acantharea have been reported
+(Blanco-Bercial et al., 2022). Additionally, Michaels (1988) noted that
+the majority of Acantharea (by abundance) were smaller than
+160\(\mu m\). While that estimate may be inflated due to inability to
+capture larger cells, small Acantharea were not captured in the present
+study. Thus, trophic strategy may shift based on sizes of Acantharea.
+
+Decelle et al. (2013) proposed a hypothetical life cycle for
+cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life
+cycle suggests that epipelagic Acantharea are adult populations, which
+form cysts that sink into the mesopelagic, then reproduce and rise.
+Furthermore, given that horizontal transfer of symbionts between
+generations of Acantharea is unlikely due to their spawning behavior,
+the newly spawned mesopelagic Acantharea are not necessarily required to
+rapidly return to the photic zone (Decelle et al., 2013, 2012). This
+hypothesis predicts that Acanthareans in the mesopelagic would be
+smaller (Decelle et al., 2013). Mars Brisbin et al. (2020) provided some
+support for this hypothesis, with a significant decrease in Acantharea
+sizes with depth. Although the authors also observed low abundances in
+the mesopelagic and noted that the smaller sizes may be due to lower
+food availability (Mars Brisbin et al., 2020). Since food is more scarce
+in the mesopelagi, nutritional quality lower, yet flux feeders would
+likely grow larger to increase their feeding range (Biard and Ohman,
+2020). In the data collecting in this study, Acanthareans in the
+mesopelagic were significantly smaller than the epipelagic, despite the
+other bimodal taxa (Foraminifera and Aulacanthidae) being significantly
+larger with depth. This provides added support for the hypothesis that
+cyst-forming Acantharea may utilize different sections of the water
+column throughout their life cycle. However to further investigate this,
+more work is needed with higher temporal and taxonomic resolution.
+
+\hypertarget{conclusions-and-considerations}{%
+\subsubsection{Conclusions and
+Considerations}\label{conclusions-and-considerations}}
+
+This study provides a detailed look at Rhizarian abundance over time
+throughout the water column in a major oligotrophic gyre. We show that
+their abundances are generally related to particle concentration and
+flux, although lack of environmental variability may have reduced the
+fit of our GAMs. Considering the potential role of Rhizaria in the
+biological carbon pump, they may have a somewhat mixed role. In the
+shallower regions, where smaller Rhizaria are abundant, they may be an
+attenuating force on sinking particles (Stukel et al., 2019). It should
+be noted that in our study, we focused on the ``small'' particle
+concentration field (\textless450\(\mu m\)), and these particles are
+generally slower sinking than large particles. However, once consumed
+and repackaged by larger Rhizaria, they can sink quicker and contribute
+more to overall flux (Michaels, 1988). Thus, Rhizaria may act as an
+aggregation mechanism. However, this is largely speculation, to truly
+test this, more work is needed measuring Rhizarian flux.
+
+The vertical partitioning documented in this study do support the
+hypothesis that mixotrophic rhizaria will occupy shallower waters while
+deeper waters are dominated by heterotrophy. However the degree to which
+mixotrophic Rhizaria in the euphotic zone rely on heterotrophy versus
+symbiosis is uncertain. Collodaria were recorded as consistent and
+dominant members of the near surface region. These organisms have the
+potential to contribute considerably to the otherwise low productivity
+of oligotrophic regions. However, their role in food webs is not well
+understood. While this study represents a step-forward in our
+understanding of Rhizarian ecology, continued research on Rhizaria is
+much needed to better understand their ecology. Particularly extended
+descriptive work to capture interannual patterns. Also work defining
+biotic interactions, feeding rates, productivity, and life history are
+all rich fields of interest in Rhizaria.
+
+\hypertarget{references}{%
+\section{References:}\label{references}}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-anderson2014}{}}%
+Anderson, O.R., 2014. Living Together in the Plankton: A Survey of
+Marine Protist Symbioses. Acta Protozoologica 53, 29--38.
+\url{https://doi.org/10.4467/16890027AP.13.0019.1116}
+
+\leavevmode\vadjust pre{\hypertarget{ref-anderson1976}{}}%
+Anderson, O.R., Bé, A.W.H., 1976. A CYTOCHEMICAL FINE STRUCTURE STUDY OF
+PHAGOTROPHY IN A PLANKTONIC FORAMINIFER, {\emph{HASTIGERINA PELAGICA}}
+(d'ORBIGNY). The Biological Bulletin 151, 437--449.
+\url{https://doi.org/10.2307/1540498}
+
+\leavevmode\vadjust pre{\hypertarget{ref-aurahs2009}{}}%
+Aurahs, R., Grimm, G.W., Hemleben, V., Hemleben, C., Kucera, M., 2009.
+Geographical distribution of cryptic genetic types in the planktonic
+foraminifer Globigerinoides ruber. Molecular Ecology 18, 1692--1706.
+\url{https://doi.org/10.1111/j.1365-294X.2009.04136.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-barth-inrev}{}}%
+Barth, A., Stone, J., in review. Understanding the picture: The promis
+and challenges of in-situ imagery data in the study of plankton ecology.
+
+\leavevmode\vadjust pre{\hypertarget{ref-barth2022}{}}%
+Barth, A., Stone, J., 2022.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.898057}{Comparison
+of an in situ imaging device and net-based method to study
+mesozooplankton communities in an oligotrophic system}. Frontiers in
+Marine Science 9.
+
+\leavevmode\vadjust pre{\hypertarget{ref-benfield1996}{}}%
+Benfield, M.C., Davis, C.S., Wiebe, P.H., Gallager, S.M., Gregory Lough,
+R., Copley, N.J., 1996. Video Plankton Recorder estimates of copepod,
+pteropod and larvacean distributions from a stratified region of Georges
+Bank with comparative measurements from a MOCNESS sampler. Deep Sea
+Research Part II: Topical Studies in Oceanography 43, 1925--1945.
+\url{https://doi.org/10.1016/S0967-0645(96)00044-6}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
+Biard, T., 2022a. Rhizaria. Encyclopeida of Life Sciences, pp. 1--11.
+\url{https://doi.org/10.1002/9780470015902.a0029469}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2022b}{}}%
+Biard, T., 2022b. Diversity and ecology of Radiolaria in modern oceans.
+Environmental Microbiology 24, 2179--2200.
+\url{https://doi.org/10.1111/1462-2920.16004}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2017}{}}%
+Biard, T., Bigeard, E., Audic, S., Poulain, J., Gutierrez-Rodriguez, A.,
+Pesant, S., Stemmann, L., Not, F., 2017. Biogeography and diversity of
+Collodaria (Radiolaria) in the global ocean. The ISME Journal 11,
+1331--1344. \url{https://doi.org/10.1038/ismej.2017.12}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2018}{}}%
+Biard, T., Krause, J.W., Stukel, M.R., Ohman, M.D., 2018. The
+Significance of Giant Phaeodarians (Rhizaria) to Biogenic Silica Export
+in the California Current Ecosystem. Global Biogeochemical Cycles 32,
+987--1004. \url{https://doi.org/10.1029/2018GB005877}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2020}{}}%
+Biard, T., Ohman, M.D., 2020. Vertical niche definition of test-bearing
+protists (Rhizaria) into the twilight zone revealed by in situ imaging.
+Limnology and Oceanography 65, 2583--2602.
+\url{https://doi.org/10.1002/lno.11472}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2015}{}}%
+Biard, T., Pillet, L., Decelle, J., Poirier, C., Suzuki, N., Not, F.,
+2015. Towards an integrative morpho-molecular classification of the
+collodaria (polycystinea, radiolaria). Protist 166, 374--388.
+\url{https://doi.org/10.1016/j.protis.2015.05.002}
+
+\leavevmode\vadjust pre{\hypertarget{ref-biard2016}{}}%
+Biard, T., Stemmann, L., Picheral, M., Mayot, N., Vandromme, P., Hauss,
+H., Gorsky, G., Guidi, L., Kiko, R., Not, F., 2016. In situ imaging
+reveals the biomass of giant protists in the global ocean. Nature 532,
+504--507. \url{https://doi.org/10.1038/nature17652}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bijma1990}{}}%
+Bijma, J., Erez, J., Hemleben, C., 1990.
+\href{https://epic.awi.de/id/eprint/6096/1/Bij1990a.pdf}{Lunar and
+semi-luan reproductive cycles in some spinose planktonic foraminifers}.
+Journal of Foraminiferal Research 20, 117--127.
+
+\leavevmode\vadjust pre{\hypertarget{ref-blanco-bercial2022}{}}%
+Blanco-Bercial, L., Parsons, R., Bolaños, L.M., Johnson, R., Giovannoni,
+S.J., Curry, R., 2022.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.897140}{The
+protist community traces seasonality and mesoscale hydrographic features
+in the oligotrophic sargasso sea}. Frontiers in Marine Science 9.
+
+\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy2017}{}}%
+Boltovskoy, D., 2017. Vertical distribution patterns of radiolaria
+polycystina (protista) in the world ocean: Living ranges, isothermal
+submersion and settling shells. Journal of Plankton Research 39,
+330--349. \url{https://doi.org/10.1093/plankt/fbx003}
+
+\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy1993}{}}%
+Boltovskoy, D., Alder, V.A., Abelmann, A., 1993. Annual flux of
+radiolaria and other shelled plankters in the eastern equatorial
+atlantic at 853 m: seasonal variations and polycystine species-specific
+responses. Deep Sea Research Part I: Oceanographic Research Papers 40,
+1863--1895. \url{https://doi.org/10.1016/0967-0637(93)90036-3}
+
+\leavevmode\vadjust pre{\hypertarget{ref-caron2016}{}}%
+Caron, D.A., 2016.
+\href{https://www.nature.com/articles/nature17892}{The rise of rhizaria
+\textbar{} nature}. Nature 532, 444--445.
+
+\leavevmode\vadjust pre{\hypertarget{ref-caron1984}{}}%
+Caron, D.A., Be, A.W.H., 1984. PREDICTED AND OBSERVED FEEDING RATES OF
+THE SPINOSE PLANKTONIC FORAMINIFER. BULLETIN OF MARINE SCIENCE 35.
+
+\leavevmode\vadjust pre{\hypertarget{ref-caron1995}{}}%
+Caron, D.A., Michaels, A.F., Swanberg, N.R., Howse, F.A., 1995. Primary
+productivity by symbiont-bearing planktonic sarcodines (acantharia,
+radiolaria, foraminifera) in surface waters near bermuda. Journal of
+Plankton Research 17, 103--129.
+\url{https://doi.org/10.1093/plankt/17.1.103}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cavalier-smith2018}{}}%
+Cavalier-Smith, T., Chao, E.E., Lewis, R., 2018. Multigene phylogeny and
+cell evolution of chromist infrakingdom Rhizaria: contrasting cell
+organisation~of sister~phyla Cercozoa and Retaria. Protoplasma 255,
+1517--1574. \url{https://doi.org/10.1007/s00709-018-1241-1}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cohen2023}{}}%
+Cohen, N.R., Krinos, A.I., Kell, R.M., Chmiel, R.J., Moran, D.M.,
+McIlvin, M.R., Lopez, P.Z., Barth, A., Stone, J., Alanis, B.A., Chan,
+E.W., Breier, J.A., Jakuba, M.V., Johnson, R., Alexander, H., Saito,
+M.A., 2023. Microeukaryote metabolism across the western North Atlantic
+Ocean revealed through autonomous underwater profiling.
+\url{https://doi.org/10.1101/2023.11.20.567900}
+
+\leavevmode\vadjust pre{\hypertarget{ref-decelle2015}{}}%
+Decelle, J., Colin, S., Foster, R.A., 2015. Photosymbiosis in Marine
+Planktonic Protists, in: Ohtsuka, S., Suzaki, T., Horiguchi, T., Suzuki,
+N., Not, F. (Eds.),. Springer Japan, Tokyo, pp. 465--500.
+\url{https://doi.org/10.1007/978-4-431-55130-0_19}
+
+\leavevmode\vadjust pre{\hypertarget{ref-decelle2013}{}}%
+Decelle, J., Martin, P., Paborstava, K., Pond, D.W., Tarling, G., Mahé,
+F., De Vargas, C., Lampitt, R., Not, F., 2013. Diversity, Ecology and
+Biogeochemistry of Cyst-Forming Acantharia (Radiolaria) in the Oceans.
+PLoS ONE 8, e53598. \url{https://doi.org/10.1371/journal.pone.0053598}
+
+\leavevmode\vadjust pre{\hypertarget{ref-decelle2012}{}}%
+Decelle, J., Siano, R., Probert, I., Poirier, C., Not, F., 2012.
+Multiple microalgal partners in symbiosis with the acantharian
+Acanthochiasma sp. (Radiolaria). Symbiosis 58, 233--244.
+\url{https://doi.org/10.1007/s13199-012-0195-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dennett2002}{}}%
+Dennett, M.R., Caron, D.A., Michaels, A.F., Gallager, S.M., Davis, C.S.,
+2002. Video plankton recorder reveals high abundances of colonial
+radiolaria in surface waters of the central north pacific. Journal of
+Plankton Research 24, 797--805.
+\url{https://doi.org/10.1093/plankt/24.8.797}
+
+\leavevmode\vadjust pre{\hypertarget{ref-drago2022}{}}%
+Drago, L., Panaïotis, T., Irisson, J.-O., Babin, M., Biard, T.,
+Carlotti, F., Coppola, L., Guidi, L., Hauss, H., Karp-Boss, L., Lombard,
+F., Mcdonnell, A.M.P., Picheral, M., Rogge, A., Waite, A.M., Stemmann,
+L., Kiko, R., 2022. Global Distribution of Zooplankton Biomass Estimated
+by In Situ Imaging and Machine Learning. Frontiers in Marine Science 9.
+\url{https://doi.org/10.3389/fmars.2022.894372}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gaskell2019}{}}%
+Gaskell, D.E., Ohman, M.D., Hull, P.M., 2019. Zooglider-based
+measurements of planktonic foraminifera in the california current
+system. Journal of Foraminiferal Research 49, 390--404.
+\url{https://doi.org/10.2113/gsjfr.49.4.390}
+
+\leavevmode\vadjust pre{\hypertarget{ref-guidi2016}{}}%
+Guidi, L., Chaffron, S., Bittner, L., Eveillard, D., Larhlimi, A., Roux,
+S., Darzi, Y., Audic, S., Berline, L., Brum, J.R., Coelho, L.P.,
+Espinoza, J.C.I., Malviya, S., Sunagawa, S., Dimier, C., Kandels-Lewis,
+S., Picheral, M., Poulain, J., Searson, S., Stemmann, L., Not, F.,
+Hingamp, P., Speich, S., Follows, M., Karp-Boss, L., Boss, E., Ogata,
+H., Pesant, S., Weissenbach, J., Wincker, P., Acinas, S.G., Bork, P.,
+Vargas, C. de, Iudicone, D., Sullivan, M.B., Raes, J., Karsenti, E.,
+Bowler, C., Gorsky, G., 2016. Plankton networks driving carbon export in
+the oligotrophic ocean. Nature 532, 465--470.
+\url{https://doi.org/10.1038/nature16942}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gutierrez-rodriguez2019}{}}%
+Gutierrez-Rodriguez, A., Stukel, M.R., Lopes dos Santos, A., Biard, T.,
+Scharek, R., Vaulot, D., Landry, M.R., Not, F., 2019. High contribution
+of Rhizaria (Radiolaria) to vertical export in the California Current
+Ecosystem revealed by DNA metabarcoding. The ISME Journal 13, 964--976.
+\url{https://doi.org/10.1038/s41396-018-0322-7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gutiuxe9rrez-rodruxedguez2022}{}}%
+Gutiérrez-Rodríguez, A., Lopes dos Santos, A., Safi, K., Probert, I.,
+Not, F., Fernández, D., Gourvil, P., Bilewitch, J., Hulston, D.,
+Pinkerton, M., Nodder, S.D., 2022. Planktonic protist diversity across
+contrasting subtropical and subantarctic waters of the southwest
+pacific. Progress in Oceanography 206, 102809.
+\url{https://doi.org/10.1016/j.pocean.2022.102809}
+
+\leavevmode\vadjust pre{\hypertarget{ref-haekel1887}{}}%
+Haekel, E., 1887.
+\href{https://www.gutenberg.org/cache/epub/44527/pg44527-images.html}{Report
+on the radiolaria collected by h.m.s. Challenger during the years
+1873-1876, plates, by ernst haeckel}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hull2011}{}}%
+Hull, P.M., Osborn, K.J., Norris, R.D., Robison, B.H., 2011. Seasonality
+and depth distribution of a mesopelagic foraminifer, Hastigerinella
+digitata, in Monterey Bay, California. Limnology and Oceanography 56,
+562--576. \url{https://doi.org/10.4319/lo.2011.56.2.0562}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ikenoue2019}{}}%
+Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M.C., Takahashi,
+K., Harada, N., Fujiki, T., 2019. Phaeodaria: An Important Carrier of
+Particulate Organic Carbon in the Mesopelagic Twilight Zone of the North
+Pacific Ocean. Global Biogeochemical Cycles 33, 1146--1160.
+\url{https://doi.org/10.1029/2019GB006258}
+
+\leavevmode\vadjust pre{\hypertarget{ref-kimoto2015}{}}%
+Kimoto, K., 2015. Planktic foraminifera. pp. 129--178.
+\url{https://doi.org/10.1007/978-4-431-55130-0_7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lampitt2009}{}}%
+Lampitt, R.S., Salter, I., Johns, D., 2009. Radiolaria: Major exporters
+of organic carbon to the deep ocean. Global Biogeochemical Cycles 23.
+\url{https://doi.org/10.1029/2008GB003221}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lee2006}{}}%
+Lee, J.J., 2006.
+\href{https://dalspace.library.dal.ca/bitstream/handle/10222/78255/VOLUME\%2042-NUMBER\%202-2006-PAGE\%2063.pdf?sequence=1}{Algal
+symbiosis in larger foraminifera}. Symbiosis 42, 63--75.
+
+\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2022}{}}%
+Llopis Monferrer, N., Biard, T., Sandin, M.M., Lombard, F., Picheral,
+M., Elineau, A., Guidi, L., Leynaert, A., Tréguer, P.J., Not, F., 2022.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.895995}{Siliceous
+rhizaria abundances and diversity in the mediterranean sea assessed by
+combined imaging and metabarcoding approaches}. Frontiers in Marine
+Science 9.
+
+\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2021}{}}%
+Llopis Monferrer, N., Leynaert, A., Tréguer, P., Gutiérrez-Rodríguez,
+A., Moriceau, B., Gallinari, M., Latasa, M., L'Helguen, S., Maguer,
+J.-F., Safi, K., Pinkerton, M.H., Not, F., 2021. Role of small Rhizaria
+and diatoms in the pelagic silica production of the Southern Ocean.
+Limnology and Oceanography 66, 2187--2202.
+\url{https://doi.org/10.1002/lno.11743}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lomas2013}{}}%
+Lomas, M.W., Bates, N.R., Johnson, R.J., Knap, A.H., Steinberg, D.K.,
+Carlson, C.A., 2013. Two decades and counting: 24-years of sustained
+open ocean biogeochemical measurements in the Sargasso Sea. Deep Sea
+Research Part II: Topical Studies in Oceanography 93, 16--32.
+\url{https://doi.org/10.1016/j.dsr2.2013.01.008}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lombard2019}{}}%
+Lombard, F., Boss, E., Waite, A.M., Vogt, M., Uitz, J., Stemmann, L.,
+Sosik, H.M., Schulz, J., Romagnan, J.-B., Picheral, M., Pearlman, J.,
+Ohman, M.D., Niehoff, B., Möller, K.O., Miloslavich, P., Lara-Lpez, A.,
+Kudela, R., Lopes, R.M., Kiko, R., Karp-Boss, L., Jaffe, J.S., Iversen,
+M.H., Irisson, J.-O., Fennel, K., Hauss, H., Guidi, L., Gorsky, G.,
+Giering, S.L.C., Gaube, P., Gallager, S., Dubelaar, G., Cowen, R.K.,
+Carlotti, F., Briseño-Avena, C., Berline, L., Benoit-Bird, K., Bax, N.,
+Batten, S., Ayata, S.D., Artigas, L.F., Appeltans, W., 2019.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00196}{Globally
+consistent quantitative observations of planktonic ecosystems}.
+Frontiers in Marine Science 6.
+
+\leavevmode\vadjust pre{\hypertarget{ref-marra2011}{}}%
+Marra, G., Wood, S.N., 2011. Practical variable selection for
+generalized additive models. Computational Statistics \& Data Analysis
+55, 2372--2387. \url{https://doi.org/10.1016/j.csda.2011.02.004}
+
+\leavevmode\vadjust pre{\hypertarget{ref-marsbrisbin2020}{}}%
+Mars Brisbin, M., Brunner, O.D., Grossmann, M.M., Mitarai, S., 2020.
+Paired high-throughput, in situ imaging and high-throughput sequencing
+illuminate acantharian abundance and vertical distribution. Limnology
+and Oceanography 65, 2953--2965. \url{https://doi.org/10.1002/lno.11567}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcgillicuddy1998}{}}%
+McGillicuddy, D.J., Robinson, A.R., Siegel, D.A., Jannasch, H.W.,
+Johnson, R., Dickey, T.D., McNeil, J., Michaels, A.F., Knap, A.H., 1998.
+Influence of mesoscale eddies on new production in the Sargasso Sea.
+Nature 394, 263--266. \url{https://doi.org/10.1038/28367}
+
+\leavevmode\vadjust pre{\hypertarget{ref-michaels1988}{}}%
+Michaels, A.F., 1988. Vertical distribution and abundance of Acantharia
+and their symbionts. Marine Biology 97, 559--569.
+\url{https://doi.org/10.1007/BF00391052}
+
+\leavevmode\vadjust pre{\hypertarget{ref-michaels1995}{}}%
+Michaels, A.F., Caron, D.A., Swanberg, N.R., Howse, F.A., Michaels,
+C.M., 1995. Planktonic sarcodines (Acantharia, Radiolaria, Foraminifera)
+in surface waters near Bermuda: abundance, biomass and vertical flux.
+Journal of Plankton Research 17, 131--163.
+\url{https://doi.org/10.1093/plankt/17.1.131}
+
+\leavevmode\vadjust pre{\hypertarget{ref-michaels1996}{}}%
+Michaels, A.F., Knap, A.H., 1996. Overview of the U.S. JGOFS Bermuda
+Atlantic Time-series Study and the Hydrostation S program. Deep Sea
+Research Part II: Topical Studies in Oceanography 43, 157--198.
+\url{https://doi.org/10.1016/0967-0645(96)00004-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nakamura2023}{}}%
+Nakamura, Y., Itagaki, H., Tuji, A., Shimode, S., Yamaguchi, A., Hidaka,
+K., Ogiso-Tanaka, E., 2023. DNA metabarcoding focused on
+difficult-to-culture protists: An effective approach to clarify
+biological interactions. Environmental Microbiology 25, 3630--3638.
+\url{https://doi.org/10.1111/1462-2920.16524}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nakamura2015}{}}%
+Nakamura, Y., Suzuki, N., 2015.
+\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_9}{Phaeodaria:
+Diverse marine cercozoans of world-wide distribution}. Springer, pp.
+223--249.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ohman2019a}{}}%
+Ohman, M.D., 2019. A sea of tentacles: optically discernible traits
+resolved from planktonic organisms in situ. ICES Journal of Marine
+Science 76, 1959--1972. \url{https://doi.org/10.1093/icesjms/fsz184}
+
+\leavevmode\vadjust pre{\hypertarget{ref-panauxefotis2023}{}}%
+Panaïotis, T., Babin, M., Biard, T., Carlotti, F., Coppola, L., Guidi,
+L., Hauss, H., Karp-Boss, L., Kiko, R., Lombard, F., McDonnell, A.M.P.,
+Picheral, M., Rogge, A., Waite, A.M., Stemmann, L., Irisson, J.-O.,
+2023. Three major mesoplanktonic communities resolved by in situ imaging
+in the upper 500 m of the global ocean. Global Ecology and Biogeography
+32, 1991--2005. \url{https://doi.org/10.1111/geb.13741}
+
+\leavevmode\vadjust pre{\hypertarget{ref-picheral}{}}%
+Picheral, M., Colin, S.P., Irisson, J.-O., n.d.
+\href{https://ecotaxa.obs-vlfr.fr/}{Ecotaxa}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-picheral2010}{}}%
+Picheral, M., Guidi, L., Stemmann, L., Karl, D.M., Iddaoud, G., Gorsky,
+G., 2010. The Underwater Vision Profiler 5: An advanced instrument for
+high spatial resolution studies of particle size spectra and
+zooplankton. Limnology and Oceanography: Methods 8, 462--473.
+\url{https://doi.org/10.4319/lom.2010.8.462}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sogawa2022}{}}%
+Sogawa, S., Nakamura, Y., Nagai, S., Nishi, N., Hidaka, K., Shimizu, Y.,
+Setou, T., 2022. DNA metabarcoding reveals vertical variation and hidden
+diversity of alveolata and rhizaria communities in the western north
+pacific. Deep Sea Research Part I: Oceanographic Research Papers 183,
+103765. \url{https://doi.org/10.1016/j.dsr.2022.103765}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stukel2018}{}}%
+Stukel, M.R., Biard, T., Krause, J., Ohman, M.D., 2018. Large Phaeodaria
+in the twilight zone: Their role in the carbon cycle. Limnology and
+Oceanography 63, 2579--2594. \url{https://doi.org/10.1002/lno.10961}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stukel2019}{}}%
+Stukel, M.R., Ohman, M.D., Kelly, T.B., Biard, T., 2019.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00397}{The
+roles of suspension-feeding and flux-feeding zooplankton as gatekeepers
+of particle flux into the mesopelagic ocean in the northeast pacific}.
+Frontiers in Marine Science 6.
+
+\leavevmode\vadjust pre{\hypertarget{ref-suzuki2015}{}}%
+Suzuki, N., Not, F., 2015.
+\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_8}{Biology
+and ecology of radiolaria}. Springer, pp. 179--222.
+
+\leavevmode\vadjust pre{\hypertarget{ref-swanberg1981}{}}%
+Swanberg, N.R., Anderson, O.R., 1981. Collozoum caudatum sp. nov.: A
+giant colonial radiolarian from equatorial and Gulf Stream waters. Deep
+Sea Research Part A. Oceanographic Research Papers 28, 1033--1047.
+\url{https://doi.org/10.1016/0198-0149(81)90016-9}
+
+\leavevmode\vadjust pre{\hypertarget{ref-takahashi1983}{}}%
+Takahashi, K., Hurd, D.C., Honjo, S., 1983. Phaeodarian skeletons: Their
+role in silica transport to the deep sea. Science 222, 616--618.
+\url{https://doi.org/10.1126/science.222.4624.616}
+
+\leavevmode\vadjust pre{\hypertarget{ref-whitmore2021}{}}%
+Whitmore, B.M., Ohman, M.D., 2021. Zooglider-measured association of
+zooplankton with the fine-scale vertical prey field. Limnology and
+Oceanography 66, 3811--3827. \url{https://doi.org/10.1002/lno.11920}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wood2017}{}}%
+Wood, S.N., 2017. Generalized additive models: an introduction with R,
+Second edition. ed, Texts in statistical science. CRC Press/Taylor \&
+Francis Group, Boca Raton London New York.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wood2001}{}}%
+Wood, S.N., 2001. mgcv: GAMs and Generalized Ridge Regression for R 1.
+
+\leavevmode\vadjust pre{\hypertarget{ref-zasko2005}{}}%
+Zasko, D.N., Rusanov, I.I., 2005. Vertical Distribution of Radiolarians
+and Their Role in Epipelagic Communities of the East Pacific Rise and
+the Gulf of California. Biology Bulletin 32, 279--287.
+\url{https://doi.org/10.1007/s10525-005-0103-5}
+
+\end{CSLReferences}
+
+\newpage
+
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4286}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2857}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0952}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0833}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1071}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2533}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}@{}}
 \caption{Generalized Additive Model results for integrated total
@@
 
-GAMs for individual taxa were much less consistent in their fits (Table
-2). This is likely in part due to the high number of non-observations
-for certain taxa. Note that due to low abundances, GAMS were not
-constructed for Tuscaroridae or Medusettidae. Furthermore no significant
-terms were found for a model with Aulosphaeridae in the epipelagic nor
-Foraminifera in the mesopelagic.
-
-\begin{figure}
-
-{\centering \includegraphics{images/06_epi-partials.pdf}
-
-}
-
-\caption{Partial effects of smooth terms in taxa-specific GAM models
-from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea
-(A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae
-(E).}
-
-\end{figure}
-
-Epipelagic Acantharea were explained by several predictor vairables and
-had a good fit (\(R^2_{adj}\) = 0.53, Table 2). Most notable smooths
-however, were mass flux and particle concentration, which had a weak
-positive association (Figure 6A), with July 2021 as a clear outlier
-where Acantharean abundances were high in the epipelagic despite lower
-fluxes and particle concentrations (Figure 5). Foraminifera had a good
-fitting GAM in the epipelagic (\(R^2_{adj}\) = 0.445). There were
-several significant explanatory variables, althoguht the clearest
-pattern was observed of high temperatures associated with more
-Foraminifera abundance (Table 2, Figure 6B). Epipelagic Aulacanthidae
-similarly had several predictor variables which were significant,
-including both water quality parameters and particle/flux predictors
-(Table 2). Interestingly, Aulacanthidae had primary production as a
-significant predictor, yet there was not a clear association (Figure
-6C). There was a fit for Collodaria in the epipelagic (\(R^2_{adj}\) =
-0.16), although there was a logit-like relationship where higher
-abundances tended to occur during higher DO conditions in the surface
-waters (Figure 6D). Castanellidae also had similarly poor fits in the
-epipelagic (\(R^2_{adj}\) = 0.124) (Table 2, Figure 6E).
+\newpage
 
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4471}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2824}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0824}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0824}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1059}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2533}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}@{}}
 \caption{Taxa-specific generalized additive models for different regions
@@
 
+\newpage
+
+\begin{figure}
+
+{\centering \includegraphics{images/01_taxa.pdf}
+
+}
+
+\caption{Example images of different Rhizaria taxa. 4mm scale bar shown
+in lower right. All vignettes are same scale}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}
+
+{\centering \includegraphics{images/02-environmental.pdf}
+
+}
+
+\caption{Environmental profiles across time-series of study period. Y
+axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved
+Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion
+(184 - 450\(\mu m\)). (F) Bacteria Abundance. (G) Silica. (H) Nitrate}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}
+
+{\centering \includegraphics{images/03_prod-flux.pdf}
+
+}
+
+\caption{Primary productivity (A) and flux estimates from total mass
+(B), carbon (C), and nitrogen (D). Values are from monthly cruises with
+month displayed in corresponding colors. Absent data are shown by grey
+bar.}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}
+
+{\centering \includegraphics{images/04_vertical.pdf}
+
+}
+
+\caption{Average abundance of Rhizaria in 25m bins, across entire study
+period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C),
+Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae
+(G), Coelodendridae (H).}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}
+
+{\centering \includegraphics{images/05_seasonality.pdf}
+
+}
+
+\caption{Seasonality of Rhizarian integrated abundance for the
+epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower
+mesopelagic (500-1000m) (E-F). Left panels (A,C,E) display total
+integrated abundance per monthly cruise colored by month. Right panels
+(B, D, F) display community composition of each total abundance.}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}
+
+{\centering \includegraphics{images/09_size-comp.pdf}
+
+}
+
+\caption{Comparison of average sizes (ESD) amongst Rhizarian taxa which
+occurred throughout the water column.}
+
+\end{figure}
+
+\newpage
+
+\begin{figure}
+
+{\centering \includegraphics{images/06_epi-partials.pdf}
+
+}
+
+\caption{Partial effects of smooth terms in taxa-specific GAM models
+from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea
+(A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae
+(E).}
+
+\end{figure}
+
+\newpage
+
 \begin{figure}
@@
 
-In the upper mesopelagic (200-500m), abundances were generally low
-(Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae,
-Aulosphaeridae, and Coelodendridae (Table 2). All these models had
-generally poor fits (\(R^2_{adj}\) \textless{} 0.25). Yet, for all upper
-mesopelagic models, particle concentration was a significant explanatory
-variable (Table 2, Figure 7). Carbon flux was significant for Acantharea
-and nitrogen flux was significant for both Acantharea and Coelodendridae
-(Table 2, Figure 7A,D). The lower mesopelagic also had generally poor
-GAM fits for taxa specific models (\(R^2_{adj}\) \textless0.3), with the
-exception of Acantharea (\(R^2_{adj}\) = 0.509). Acantharea in the lower
-mesopelagic was most clearly positively associated with particle
-concentration and nitrogen flux, as well as temperature to a slight
-degree (Figure 8A). For all Phaeodarians with a significant model,
-particle concentration was a main predictor variable (Table 2, Figure
-8B-D). Aulacanthidae had the best fitting model of the Phaeodarians
-(\(R^2_{adj}\) = 0.298), which also included mass flux as a
-statistically significant smooth (Figure 8B).
+\newpage
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\hypertarget{overall-rhizarian-abundance-and-patterns}{%
-\subsubsection{Overall Rhizarian abundance and
-patterns}\label{overall-rhizarian-abundance-and-patterns}}
-
-In the epipelagic Rhizaria exhibited a notable seasonal pattern.
-Rhizarian abundances were higher in the summer months and lower during
-the winter. During a prior time period, Blanco-Bercial et al. (2022)
-noted that there is considerable seasonality in the community
-composition of all protists. Despite the seasonality of total Rhizarian
-abundance, community composition was relatively consistent, with
-Collodaria representing the bulk of the community. It should be noted
-that the overall taxonomic resolution of the UVP5 is fairly low, so
-there may be a switching of species within the broad groups identified
-in this study which were not captured. Throughout the mesopelagic,
-month-to-month variation in 2021 was relatively low. Again this is
-consistent with observations from metabarcoding of the whole protist
-community in the same study region (Blanco-Bercial et al., 2022). This
-finding is not surprising as the overall seasonal variation in
-environmental conditions in this region were low.
-
-Overall Rhizaria were the most commonly identified group of
-mesozooplankton throughout the study period. We do note that the UVP5
-commonly captures Trichodesmium colonies, yet these were excluded in
-this comparison as they are strictly autotrophs. It should be noted that
-previous work has suggested that avoidance behavior with the UVP is
-possible, at times likely, for visual and highly mobile zooplankton
-(Barth and Stone, 2022). Thus, the percent contribution reported here
-(42.7\%) of Rhizaria to the total mesozooplankton community may be
-inflated due to under sampling of organisms such as Euphausiids and
-Chaetognaths which has quick escape responses. Regardless, it is worth
-noting that in the same region, with data collected in 2012 and 2013
-using similar calculation methods, Biard et al. (2016) estimated
-Rhizaria only contribute 15\% of the total mesozooplankton community in
-the upper 500m. Likely, Rhizaria display considerable interannual
-variability. In the present study, we noticed considerably higher
-Rhizarian abundance throughout the water column in 2019 compared to
-2021. While this may have been driven by increased mass flux (Figure 3),
-more information is needed to truly understand the magnitude by which
-Rhizaria can vary interannually.
-
-\hypertarget{vertical-structure-and-trophic-roles}{%
-\subsubsection{Vertical Structure and Trophic
-Roles}\label{vertical-structure-and-trophic-roles}}
-
-In this study we present a clear pattern of vertical zonation between
-different Rhizaria groups. Largely, the taxonomic composition and
-vertical positioning were similar to Rhizaria zonation in the California
-Current Ecosystem (Biard and Ohman, 2020). It should be noted however,
-that the secondary abundance peak reported in the present study is
-lower. This is likely due to the more oligotrophic nature of the study
-region, were the euphotic zone penetrates deeper into the water column.
-Most prevalent in the epipelagic were Collodaria. These mixotrophic
-Radiolarians have long been reported to contribute to primary
-productivity in the euphotic zone (Dennett et al., 2002; Michaels et
-al., 1995). Collodaria are thought to be particularly successful
-globally in oligotrophic regions due to their photosymbiotic
-relationships (Biard et al., 2017, 2016). We observed the highest
-abundance of Collodaria during June 2021, supporting the notion they can
-thrive during the typically low-nutrient conditions of summer
-stratification. However, Collodaria also increased during the spring
-mixing period, suggesting that they can thrive during conditions which
-may typically be thought to favor autotrophs. Furthermore, while
-Collodaria were primarily absent from below 250m, there were a few
-instances of deeper colonies being observed. Global investigations of
-polycistine flux, suggest that deep-Collodaria in Oligotrophic regions
-may be a consequence of isothermal submersion (Boltovskoy, 2017).
-Another effectively exclusively epipelagic Rhizaria was the Phaeodarian
-family of Castanellidae. All Phaeodarians are thought to be fully
-heterotrophic (Nakamura and Suzuki, 2015), nonetheless a number of
-studies, including this one, report Castanellidae to be typically found
-in the lower epipelagic (Biard et al., 2018; Biard and Ohman, 2020;
-Zasko and Rusanov, 2005). It should be considered that perhaps
-Castanellidae specializes in feeding sinking particles directly at the
-base of the epipelagic. Given it's smaller size (Nakamura and Suzuki,
-2015), Castanellidae does not need a large diameter to efficiently flux
-feed at the typically particle rich region of the lower epipelagic. Both
-Castanelldiae and Collodaria had poor fitting GAMs. This is somewhat of
-a surprise for Collodaria who had large abundance. However, given the
-consistency of their abundance, it may be that this study did not
-capture a wide enough range of conditions for describing Collodaria's
-preferred niche.
-
-The mesopelagic generally was home to known heterotrophic organisms,
-particularly for those which were constrained to exclusively occupy
-deeper waters. This is consistent with Blanco-Bercial et al. (2022)'s
-observation of an auto-/mixotroph to heterotroph gradient in the protist
-community. The upper mesopelagic interestingly had relatively low total
-abundance. This low-abundance region likely reflects the dynamics of
-productivity and export throughout the water column. While productivity
-and thus sinking particles for flux feeders are high in the euphotic
-zone, much of this is attenuated throughout the epipelagic. So, while
-the base of the epipelagic may provide a rich feeding environment for
-Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F),
-the region from 200-500m might be otherwise food poor. Perhaps it is
-more advantageous for Rhizaria to situate deeper, in darker regions of
-the twilight zone. Also it should be noted that Phaeodarians utilize
-silica to build their opaline tests, and silica concentrations started
-to increase around 500m (Figure 2G). Although \(Si\) was not a
-significant smooth for any taxa-specific model, this lack of association
-might be due to the overall lack of variation of \(Si\) between
-integrated abundance of each cast. Aulosphaeridae was only found to have
-significant relationships, although weak fits, to particle concentration
-in the mesopelagic. In our study, while consistently observed, overall
-abundances of Aulosphaeridae were very low. In the Pacific Ocean, on
-California's Coast, much higher abundances of Aulosphaeridae have been
-reported (Biard and Ohman, 2020; Zasko and Rusanov, 2005) and they have
-massive potential to impact silica export (Biard et al., 2018).
-Coelodendridae were also seemingly restricted to the deeper section of
-the mesopelagic. This is interesting given that in the California
-Current, (Biard and Ohman, 2020) found a bimodal distribution in
-Coelodendridae. There are several morphotypes corresponding to different
-taxa of Coelodendridae (Biard and Ohman, 2020; Nakamura and Suzuki,
-2015). So it may be that only a few types of Coelodendridae were
-observed in this study, while the epipelagic variety was not.
-Alternatively, the lower epipelagic of the California Current may
-provide adequate habitat for Coelodendridae, which is not available in
-the oligotrophic Sargasso Sea.
-
-A number of taxa were found to have a bimodal distribution, with
-considerable populations in both the epipelagic and mesopelagic.
-Aulacanthidae had a bimodal distribution, although abundances were
-highest in the lower mesopelagic. Foraminifera also had a bimodal
-distribution. Some lineages of Foraminifera are known to host
-photosymbionts (Biard, 2022a; Kimoto, 2015), however they are also
-efficient predators commonly seen throughout the mesopelagic (Caron and
-Be, 1984; Gaskell et al., 2019). Thus it is not surprising to find their
-presence in both locations of the water column. Foraminifera are also
-known to vary their vertical distribution across their life cycle in
-phase with lunar cycles (Biard, 2022a; Bijma et al., 1990; Gaskell et
-al., 2019; Kimoto, 2015). However, the sampling scheme of the BATS
-program does not capture this frequency and was not investigated in the
-present study.
-
-Acantharea also had a bimodal distribution, with much larger abundances
-than Aulacanthidae or Foraminifera. Most prior studies of Acantharea
-vertical distribtion found them concentrated in near surface layers of
-the water column Zasko and Rusanov (2005). This would support the
-paradigm that large Acantharean abundances may be supported by their
-mixotrophic abilities (Michaels et al., 1995; Suzuki and Not, 2015).
-While the UVP5 images cannot distinguish between mixotrophic and
-heterotrophic Acantharea, the GAMs constructed for Acantharean abundance
-found positive associations with particle concentration and mass flux,
-suggesting a higher reliance on heterotrophy. Recently Mars Brisbin et
-al. (2020) described apparent predator behavior amongst near-surface
-Acantharea. Thus it is likely that epipelagic Acantharea may commonly be
-heterotrophic. Yet, it should be noted in the Sargasso Sea, both
-heterotrophic and symbiotic lineages of Acantharea have been reported
-(Blanco-Bercial et al., 2022). Additionally, Michaels (1988) noted that
-the majority of Acantharea (by abundance) were smaller than
-160\(\mu m\). While that estimate may be inflated due to inability to
-capture larger cells, small Acantharea were not captured in the present
-study. Thus, trophic strategy may shift based on sizes of Acantharea.
-
-Decelle et al. (2013) proposed a hypothetical life cycle for
-cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life
-cycle suggests that epipelagic Acantharea are adult populations, which
-form cysts that sink into the mesopelagic, then reproduce and raise.
-Furthermore, given that horizontal transfer of symbionts between
-generations of Acantharea is unlikely, the newly spawned mesopelagic
-Acantharea are not necessarily required to rapidly return to the photic
-zone. This hypothesis predicts that Acanthareans in the mesopelagic
-would be smaller (Decelle et al., 2013). Mars Brisbin et al. (2020)
-provided some support for this hypothesis, with a significant decrease
-in Acantharea sizes with depth. Although, the authors also observed low
-abundances in the mesopelagic and noted that the smaller sizes may be
-due to less food availability (Mars Brisbin et al., 2020). Since food is
-more scarce in the mesopelagic there is is less nutritional quality, yet
-flux feeders would likely grow larger to increase their feeding range
-(Biard and Ohman, 2020). In the data collecting in this study,
-Acanthareans in the mesopelagic were significantly smaller than the
-epipelagic, despite the other bimodal taxa (Foraminifera and
-Aulacanthidae) being significantly larger with depth (Figure 9, Wilcoxon
-rank sum text p-value \textless{} 0.001). This provides added support
-for the hypothesis that cyst-forming Acantharea may utilize different
-sections of the water column throughout their life cycle. However to
-further investigate this, more work is needed with higher temporal and
-taxonomic resolution.
-
-\begin{figure}
-
-{\centering \includegraphics{images/09_size-comp.pdf}
-
-}
-
-\caption{Comparison of average sizes (ESD) amongst Rhizarian taxa which
-occurred throughout the water column.}
-
-\end{figure}
-
-\hypertarget{conclusions-and-considerations}{%
-\subsubsection{Conclusions and
-Considerations}\label{conclusions-and-considerations}}
-
-This study provides a detailed look at Rhizarian abundance over time
-throughout the water column in a major oligotrophic gyre. We show that
-their abundances are generally related to particle concentration and
-flux, although lack of environmental variability may have reduced the
-fit of our GAMs. Considering the potential role of Rhizaria in the
-biological carbon pump, they may have a somewhat mixed role. In the
-shallower regions, where smaller Rhizaria are abundant, they may be an
-attenuating force on sinking particles (Stukel et al., 2019). It should
-be noted that in our study, we focused on the ``small'' particle
-concentration field (\textless450\(\mu m\)), and these particles are
-generally slower sinking than large particles. However, once consumed
-and repackaged by larger Rhizaria, they can sink quicker and contribute
-more to overall flux (Michaels, 1988). Thus, Rhizaria may act as a
-ballasting mechanism. However, this is largely speculation, to truly
-test this, more work is needed measuring Rhizarian flux.
-
-The vertical partitioning documented in this study do support the
-hypothesis that mixotrophic rhizaria will occupy shallower waters while
-deeper waters are dominated by heterotrophy. However the degree to which
-mixotrophic Rhizaria in the euphotic zone rely on heterotrophy versus
-symbiosis is uncertain. Collodaria were recorded as consistent and
-dominant members of the near surface region. These organisms have the
-potential to contribute considerably to the otherwise low productivity
-of oligotrophic regions. However, their role in food webs is not well
-understood. While this study represents a step-forward in our
-understanding of Rhizarian ecology, there are several known aspects of
-Rhizarian and protist ecology which are not captured in this study.
-Continued research on Rhizaria is much needed to better understand their
-ecology. Particularly extended descriptive work to capture interannual
-patterns. Also work defining biotic interactions, feeding rates,
-productivity, and life history are all rich fields of interest in
-Rhizaria.
-
-\hypertarget{references}{%
-\section{References:}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-anderson2014}{}}%
-Anderson, O.R., 2014. Living Together in the Plankton: A Survey of
-Marine Protist Symbioses. Acta Protozoologica 53, 29--38.
-\url{https://doi.org/10.4467/16890027AP.13.0019.1116}
-
-\leavevmode\vadjust pre{\hypertarget{ref-anderson1976}{}}%
-Anderson, O.R., Bé, A.W.H., 1976. A CYTOCHEMICAL FINE STRUCTURE STUDY OF
-PHAGOTROPHY IN A PLANKTONIC FORAMINIFER, {\emph{HASTIGERINA PELAGICA}}
-(d'ORBIGNY). The Biological Bulletin 151, 437--449.
-\url{https://doi.org/10.2307/1540498}
-
-\leavevmode\vadjust pre{\hypertarget{ref-aurahs2009}{}}%
-Aurahs, R., Grimm, G.W., Hemleben, V., Hemleben, C., Kucera, M., 2009.
-Geographical distribution of cryptic genetic types in the planktonic
-foraminifer Globigerinoides ruber. Molecular Ecology 18, 1692--1706.
-\url{https://doi.org/10.1111/j.1365-294X.2009.04136.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-barth-inrev}{}}%
-Barth, A., Stone, J., in review. Understanding the picture: The promis
-and challenges of in-situ imagery data in the study of plankton ecology.
-
-\leavevmode\vadjust pre{\hypertarget{ref-barth2022}{}}%
-Barth, A., Stone, J., 2022.
-\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.898057}{Comparison
-of an in situ imaging device and net-based method to study
-mesozooplankton communities in an oligotrophic system}. Frontiers in
-Marine Science 9.
-
-\leavevmode\vadjust pre{\hypertarget{ref-benfield1996}{}}%
-Benfield, M.C., Davis, C.S., Wiebe, P.H., Gallager, S.M., Gregory Lough,
-R., Copley, N.J., 1996. Video Plankton Recorder estimates of copepod,
-pteropod and larvacean distributions from a stratified region of Georges
-Bank with comparative measurements from a MOCNESS sampler. Deep Sea
-Research Part II: Topical Studies in Oceanography 43, 1925--1945.
-\url{https://doi.org/10.1016/S0967-0645(96)00044-6}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
-Biard, T., 2022a. Rhizaria. Encyclopeida of Life Sciences, pp. 1--11.
-\url{https://doi.org/10.1002/9780470015902.a0029469}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2022b}{}}%
-Biard, T., 2022b. Diversity and ecology of Radiolaria in modern oceans.
-Environmental Microbiology 24, 2179--2200.
-\url{https://doi.org/10.1111/1462-2920.16004}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2017}{}}%
-Biard, T., Bigeard, E., Audic, S., Poulain, J., Gutierrez-Rodriguez, A.,
-Pesant, S., Stemmann, L., Not, F., 2017. Biogeography and diversity of
-Collodaria (Radiolaria) in the global ocean. The ISME Journal 11,
-1331--1344. \url{https://doi.org/10.1038/ismej.2017.12}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2018}{}}%
-Biard, T., Krause, J.W., Stukel, M.R., Ohman, M.D., 2018. The
-Significance of Giant Phaeodarians (Rhizaria) to Biogenic Silica Export
-in the California Current Ecosystem. Global Biogeochemical Cycles 32,
-987--1004. \url{https://doi.org/10.1029/2018GB005877}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2020}{}}%
-Biard, T., Ohman, M.D., 2020. Vertical niche definition of test-bearing
-protists (Rhizaria) into the twilight zone revealed by in situ imaging.
-Limnology and Oceanography 65, 2583--2602.
-\url{https://doi.org/10.1002/lno.11472}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2015}{}}%
-Biard, T., Pillet, L., Decelle, J., Poirier, C., Suzuki, N., Not, F.,
-2015. Towards an integrative morpho-molecular classification of the
-collodaria (polycystinea, radiolaria). Protist 166, 374--388.
-\url{https://doi.org/10.1016/j.protis.2015.05.002}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2016}{}}%
-Biard, T., Stemmann, L., Picheral, M., Mayot, N., Vandromme, P., Hauss,
-H., Gorsky, G., Guidi, L., Kiko, R., Not, F., 2016. In situ imaging
-reveals the biomass of giant protists in the global ocean. Nature 532,
-504--507. \url{https://doi.org/10.1038/nature17652}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bijma1990}{}}%
-Bijma, J., Erez, J., Hemleben, C., 1990.
-\href{https://epic.awi.de/id/eprint/6096/1/Bij1990a.pdf}{Lunar and
-semi-luan reproductive cycles in some spinose planktonic foraminifers}.
-Journal of Foraminiferal Research 20, 117--127.
-
-\leavevmode\vadjust pre{\hypertarget{ref-blanco-bercial2022}{}}%
-Blanco-Bercial, L., Parsons, R., Bolaños, L.M., Johnson, R., Giovannoni,
-S.J., Curry, R., 2022.
-\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.897140}{The
-protist community traces seasonality and mesoscale hydrographic features
-in the oligotrophic sargasso sea}. Frontiers in Marine Science 9.
-
-\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy2017}{}}%
-Boltovskoy, D., 2017. Vertical distribution patterns of radiolaria
-polycystina (protista) in the world ocean: Living ranges, isothermal
-submersion and settling shells. Journal of Plankton Research 39,
-330--349. \url{https://doi.org/10.1093/plankt/fbx003}
-
-\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy1993}{}}%
-Boltovskoy, D., Alder, V.A., Abelmann, A., 1993. Annual flux of
-radiolaria and other shelled plankters in the eastern equatorial
-atlantic at 853 m: seasonal variations and polycystine species-specific
-responses. Deep Sea Research Part I: Oceanographic Research Papers 40,
-1863--1895. \url{https://doi.org/10.1016/0967-0637(93)90036-3}
-
-\leavevmode\vadjust pre{\hypertarget{ref-caron2016}{}}%
-Caron, D.A., 2016.
-\href{https://www.nature.com/articles/nature17892}{The rise of rhizaria
-\textbar{} nature}. Nature 532, 444--445.
-
-\leavevmode\vadjust pre{\hypertarget{ref-caron1984}{}}%
-Caron, D.A., Be, A.W.H., 1984. PREDICTED AND OBSERVED FEEDING RATES OF
-THE SPINOSE PLANKTONIC FORAMINIFER. BULLETIN OF MARINE SCIENCE 35.
-
-\leavevmode\vadjust pre{\hypertarget{ref-caron1995}{}}%
-Caron, D.A., Michaels, A.F., Swanberg, N.R., Howse, F.A., 1995. Primary
-productivity by symbiont-bearing planktonic sarcodines (acantharia,
-radiolaria, foraminifera) in surface waters near bermuda. Journal of
-Plankton Research 17, 103--129.
-\url{https://doi.org/10.1093/plankt/17.1.103}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cavalier-smith2018}{}}%
-Cavalier-Smith, T., Chao, E.E., Lewis, R., 2018. Multigene phylogeny and
-cell evolution of chromist infrakingdom Rhizaria: contrasting cell
-organisation~of sister~phyla Cercozoa and Retaria. Protoplasma 255,
-1517--1574. \url{https://doi.org/10.1007/s00709-018-1241-1}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cohen2023}{}}%
-Cohen, N.R., Krinos, A.I., Kell, R.M., Chmiel, R.J., Moran, D.M.,
-McIlvin, M.R., Lopez, P.Z., Barth, A., Stone, J., Alanis, B.A., Chan,
-E.W., Breier, J.A., Jakuba, M.V., Johnson, R., Alexander, H., Saito,
-M.A., 2023. Microeukaryote metabolism across the western North Atlantic
-Ocean revealed through autonomous underwater profiling.
-\url{https://doi.org/10.1101/2023.11.20.567900}
-
-\leavevmode\vadjust pre{\hypertarget{ref-decelle2015}{}}%
-Decelle, J., Colin, S., Foster, R.A., 2015. Photosymbiosis in Marine
-Planktonic Protists, in: Ohtsuka, S., Suzaki, T., Horiguchi, T., Suzuki,
-N., Not, F. (Eds.),. Springer Japan, Tokyo, pp. 465--500.
-\url{https://doi.org/10.1007/978-4-431-55130-0_19}
-
-\leavevmode\vadjust pre{\hypertarget{ref-decelle2013}{}}%
-Decelle, J., Martin, P., Paborstava, K., Pond, D.W., Tarling, G., Mahé,
-F., De Vargas, C., Lampitt, R., Not, F., 2013. Diversity, Ecology and
-Biogeochemistry of Cyst-Forming Acantharia (Radiolaria) in the Oceans.
-PLoS ONE 8, e53598. \url{https://doi.org/10.1371/journal.pone.0053598}
-
-\leavevmode\vadjust pre{\hypertarget{ref-decelle2012}{}}%
-Decelle, J., Siano, R., Probert, I., Poirier, C., Not, F., 2012.
-Multiple microalgal partners in symbiosis with the acantharian
-Acanthochiasma sp. (Radiolaria). Symbiosis 58, 233--244.
-\url{https://doi.org/10.1007/s13199-012-0195-x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dennett2002}{}}%
-Dennett, M.R., Caron, D.A., Michaels, A.F., Gallager, S.M., Davis, C.S.,
-2002. Video plankton recorder reveals high abundances of colonial
-radiolaria in surface waters of the central north pacific. Journal of
-Plankton Research 24, 797--805.
-\url{https://doi.org/10.1093/plankt/24.8.797}
-
-\leavevmode\vadjust pre{\hypertarget{ref-drago2022}{}}%
-Drago, L., Panaïotis, T., Irisson, J.-O., Babin, M., Biard, T.,
-Carlotti, F., Coppola, L., Guidi, L., Hauss, H., Karp-Boss, L., Lombard,
-F., Mcdonnell, A.M.P., Picheral, M., Rogge, A., Waite, A.M., Stemmann,
-L., Kiko, R., 2022. Global Distribution of Zooplankton Biomass Estimated
-by In Situ Imaging and Machine Learning. Frontiers in Marine Science 9.
-\url{https://doi.org/10.3389/fmars.2022.894372}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gaskell2019}{}}%
-Gaskell, D.E., Ohman, M.D., Hull, P.M., 2019. Zooglider-based
-measurements of planktonic foraminifera in the california current
-system. Journal of Foraminiferal Research 49, 390--404.
-\url{https://doi.org/10.2113/gsjfr.49.4.390}
-
-\leavevmode\vadjust pre{\hypertarget{ref-guidi2016}{}}%
-Guidi, L., Chaffron, S., Bittner, L., Eveillard, D., Larhlimi, A., Roux,
-S., Darzi, Y., Audic, S., Berline, L., Brum, J.R., Coelho, L.P.,
-Espinoza, J.C.I., Malviya, S., Sunagawa, S., Dimier, C., Kandels-Lewis,
-S., Picheral, M., Poulain, J., Searson, S., Stemmann, L., Not, F.,
-Hingamp, P., Speich, S., Follows, M., Karp-Boss, L., Boss, E., Ogata,
-H., Pesant, S., Weissenbach, J., Wincker, P., Acinas, S.G., Bork, P.,
-Vargas, C. de, Iudicone, D., Sullivan, M.B., Raes, J., Karsenti, E.,
-Bowler, C., Gorsky, G., 2016. Plankton networks driving carbon export in
-the oligotrophic ocean. Nature 532, 465--470.
-\url{https://doi.org/10.1038/nature16942}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gutierrez-rodriguez2019}{}}%
-Gutierrez-Rodriguez, A., Stukel, M.R., Lopes dos Santos, A., Biard, T.,
-Scharek, R., Vaulot, D., Landry, M.R., Not, F., 2019. High contribution
-of Rhizaria (Radiolaria) to vertical export in the California Current
-Ecosystem revealed by DNA metabarcoding. The ISME Journal 13, 964--976.
-\url{https://doi.org/10.1038/s41396-018-0322-7}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gutiuxe9rrez-rodruxedguez2022}{}}%
-Gutiérrez-Rodríguez, A., Lopes dos Santos, A., Safi, K., Probert, I.,
-Not, F., Fernández, D., Gourvil, P., Bilewitch, J., Hulston, D.,
-Pinkerton, M., Nodder, S.D., 2022. Planktonic protist diversity across
-contrasting subtropical and subantarctic waters of the southwest
-pacific. Progress in Oceanography 206, 102809.
-\url{https://doi.org/10.1016/j.pocean.2022.102809}
-
-\leavevmode\vadjust pre{\hypertarget{ref-haekel1887}{}}%
-Haekel, E., 1887.
-\href{https://www.gutenberg.org/cache/epub/44527/pg44527-images.html}{Report
-on the radiolaria collected by h.m.s. Challenger during the years
-1873-1876, plates, by ernst haeckel}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-hull2011}{}}%
-Hull, P.M., Osborn, K.J., Norris, R.D., Robison, B.H., 2011. Seasonality
-and depth distribution of a mesopelagic foraminifer, Hastigerinella
-digitata, in Monterey Bay, California. Limnology and Oceanography 56,
-562--576. \url{https://doi.org/10.4319/lo.2011.56.2.0562}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ikenoue2019}{}}%
-Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M.C., Takahashi,
-K., Harada, N., Fujiki, T., 2019. Phaeodaria: An Important Carrier of
-Particulate Organic Carbon in the Mesopelagic Twilight Zone of the North
-Pacific Ocean. Global Biogeochemical Cycles 33, 1146--1160.
-\url{https://doi.org/10.1029/2019GB006258}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kimoto2015}{}}%
-Kimoto, K., 2015. Planktic foraminifera. pp. 129--178.
-\url{https://doi.org/10.1007/978-4-431-55130-0_7}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lampitt2009}{}}%
-Lampitt, R.S., Salter, I., Johns, D., 2009. Radiolaria: Major exporters
-of organic carbon to the deep ocean. Global Biogeochemical Cycles 23.
-\url{https://doi.org/10.1029/2008GB003221}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lee2006}{}}%
-Lee, J.J., 2006.
-\href{https://dalspace.library.dal.ca/bitstream/handle/10222/78255/VOLUME\%2042-NUMBER\%202-2006-PAGE\%2063.pdf?sequence=1}{Algal
-symbiosis in larger foraminifera}. Symbiosis 42, 63--75.
-
-\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2022}{}}%
-Llopis Monferrer, N., Biard, T., Sandin, M.M., Lombard, F., Picheral,
-M., Elineau, A., Guidi, L., Leynaert, A., Tréguer, P.J., Not, F., 2022.
-\href{https://www.frontiersin.org/articles/10.3389/fmars.2022.895995}{Siliceous
-rhizaria abundances and diversity in the mediterranean sea assessed by
-combined imaging and metabarcoding approaches}. Frontiers in Marine
-Science 9.
-
-\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2021}{}}%
-Llopis Monferrer, N., Leynaert, A., Tréguer, P., Gutiérrez-Rodríguez,
-A., Moriceau, B., Gallinari, M., Latasa, M., L'Helguen, S., Maguer,
-J.-F., Safi, K., Pinkerton, M.H., Not, F., 2021. Role of small Rhizaria
-and diatoms in the pelagic silica production of the Southern Ocean.
-Limnology and Oceanography 66, 2187--2202.
-\url{https://doi.org/10.1002/lno.11743}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lomas2013}{}}%
-Lomas, M.W., Bates, N.R., Johnson, R.J., Knap, A.H., Steinberg, D.K.,
-Carlson, C.A., 2013. Two decades and counting: 24-years of sustained
-open ocean biogeochemical measurements in the Sargasso Sea. Deep Sea
-Research Part II: Topical Studies in Oceanography 93, 16--32.
-\url{https://doi.org/10.1016/j.dsr2.2013.01.008}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lombard2019}{}}%
-Lombard, F., Boss, E., Waite, A.M., Vogt, M., Uitz, J., Stemmann, L.,
-Sosik, H.M., Schulz, J., Romagnan, J.-B., Picheral, M., Pearlman, J.,
-Ohman, M.D., Niehoff, B., Möller, K.O., Miloslavich, P., Lara-Lpez, A.,
-Kudela, R., Lopes, R.M., Kiko, R., Karp-Boss, L., Jaffe, J.S., Iversen,
-M.H., Irisson, J.-O., Fennel, K., Hauss, H., Guidi, L., Gorsky, G.,
-Giering, S.L.C., Gaube, P., Gallager, S., Dubelaar, G., Cowen, R.K.,
-Carlotti, F., Briseño-Avena, C., Berline, L., Benoit-Bird, K., Bax, N.,
-Batten, S., Ayata, S.D., Artigas, L.F., Appeltans, W., 2019.
-\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00196}{Globally
-consistent quantitative observations of planktonic ecosystems}.
-Frontiers in Marine Science 6.
-
-\leavevmode\vadjust pre{\hypertarget{ref-marra2011}{}}%
-Marra, G., Wood, S.N., 2011. Practical variable selection for
-generalized additive models. Computational Statistics \& Data Analysis
-55, 2372--2387. \url{https://doi.org/10.1016/j.csda.2011.02.004}
-
-\leavevmode\vadjust pre{\hypertarget{ref-marsbrisbin2020}{}}%
-Mars Brisbin, M., Brunner, O.D., Grossmann, M.M., Mitarai, S., 2020.
-Paired high-throughput, in situ imaging and high-throughput sequencing
-illuminate acantharian abundance and vertical distribution. Limnology
-and Oceanography 65, 2953--2965. \url{https://doi.org/10.1002/lno.11567}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcgillicuddy1998}{}}%
-McGillicuddy, D.J., Robinson, A.R., Siegel, D.A., Jannasch, H.W.,
-Johnson, R., Dickey, T.D., McNeil, J., Michaels, A.F., Knap, A.H., 1998.
-Influence of mesoscale eddies on new production in the Sargasso Sea.
-Nature 394, 263--266. \url{https://doi.org/10.1038/28367}
-
-\leavevmode\vadjust pre{\hypertarget{ref-michaels1988}{}}%
-Michaels, A.F., 1988. Vertical distribution and abundance of Acantharia
-and their symbionts. Marine Biology 97, 559--569.
-\url{https://doi.org/10.1007/BF00391052}
-
-\leavevmode\vadjust pre{\hypertarget{ref-michaels1995}{}}%
-Michaels, A.F., Caron, D.A., Swanberg, N.R., Howse, F.A., Michaels,
-C.M., 1995. Planktonic sarcodines (Acantharia, Radiolaria, Foraminifera)
-in surface waters near Bermuda: abundance, biomass and vertical flux.
-Journal of Plankton Research 17, 131--163.
-\url{https://doi.org/10.1093/plankt/17.1.131}
-
-\leavevmode\vadjust pre{\hypertarget{ref-michaels1996}{}}%
-Michaels, A.F., Knap, A.H., 1996. Overview of the U.S. JGOFS Bermuda
-Atlantic Time-series Study and the Hydrostation S program. Deep Sea
-Research Part II: Topical Studies in Oceanography 43, 157--198.
-\url{https://doi.org/10.1016/0967-0645(96)00004-5}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nakamura2023}{}}%
-Nakamura, Y., Itagaki, H., Tuji, A., Shimode, S., Yamaguchi, A., Hidaka,
-K., Ogiso-Tanaka, E., 2023. DNA metabarcoding focused on
-difficult-to-culture protists: An effective approach to clarify
-biological interactions. Environmental Microbiology 25, 3630--3638.
-\url{https://doi.org/10.1111/1462-2920.16524}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nakamura2015}{}}%
-Nakamura, Y., Suzuki, N., 2015.
-\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_9}{Phaeodaria:
-Diverse marine cercozoans of world-wide distribution}. Springer, pp.
-223--249.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ohman2019a}{}}%
-Ohman, M.D., 2019. A sea of tentacles: optically discernible traits
-resolved from planktonic organisms in situ. ICES Journal of Marine
-Science 76, 1959--1972. \url{https://doi.org/10.1093/icesjms/fsz184}
-
-\leavevmode\vadjust pre{\hypertarget{ref-panauxefotis2023}{}}%
-Panaïotis, T., Babin, M., Biard, T., Carlotti, F., Coppola, L., Guidi,
-L., Hauss, H., Karp-Boss, L., Kiko, R., Lombard, F., McDonnell, A.M.P.,
-Picheral, M., Rogge, A., Waite, A.M., Stemmann, L., Irisson, J.-O.,
-2023. Three major mesoplanktonic communities resolved by in situ imaging
-in the upper 500 m of the global ocean. Global Ecology and Biogeography
-32, 1991--2005. \url{https://doi.org/10.1111/geb.13741}
-
-\leavevmode\vadjust pre{\hypertarget{ref-picheral2010}{}}%
-Picheral, M., Guidi, L., Stemmann, L., Karl, D.M., Iddaoud, G., Gorsky,
-G., 2010. The Underwater Vision Profiler 5: An advanced instrument for
-high spatial resolution studies of particle size spectra and
-zooplankton. Limnology and Oceanography: Methods 8, 462--473.
-\url{https://doi.org/10.4319/lom.2010.8.462}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sogawa2022}{}}%
-Sogawa, S., Nakamura, Y., Nagai, S., Nishi, N., Hidaka, K., Shimizu, Y.,
-Setou, T., 2022. DNA metabarcoding reveals vertical variation and hidden
-diversity of alveolata and rhizaria communities in the western north
-pacific. Deep Sea Research Part I: Oceanographic Research Papers 183,
-103765. \url{https://doi.org/10.1016/j.dsr.2022.103765}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stukel2018}{}}%
-Stukel, M.R., Biard, T., Krause, J., Ohman, M.D., 2018. Large Phaeodaria
-in the twilight zone: Their role in the carbon cycle. Limnology and
-Oceanography 63, 2579--2594. \url{https://doi.org/10.1002/lno.10961}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stukel2019}{}}%
-Stukel, M.R., Ohman, M.D., Kelly, T.B., Biard, T., 2019.
-\href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00397}{The
-roles of suspension-feeding and flux-feeding zooplankton as gatekeepers
-of particle flux into the mesopelagic ocean in the northeast pacific}.
-Frontiers in Marine Science 6.
-
-\leavevmode\vadjust pre{\hypertarget{ref-suzuki2015}{}}%
-Suzuki, N., Not, F., 2015.
-\href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_8}{Biology
-and ecology of radiolaria}. Springer, pp. 179--222.
-
-\leavevmode\vadjust pre{\hypertarget{ref-swanberg1981}{}}%
-Swanberg, N.R., Anderson, O.R., 1981. Collozoum caudatum sp. nov.: A
-giant colonial radiolarian from equatorial and Gulf Stream waters. Deep
-Sea Research Part A. Oceanographic Research Papers 28, 1033--1047.
-\url{https://doi.org/10.1016/0198-0149(81)90016-9}
-
-\leavevmode\vadjust pre{\hypertarget{ref-takahashi1983}{}}%
-Takahashi, K., Hurd, D.C., Honjo, S., 1983. Phaeodarian skeletons: Their
-role in silica transport to the deep sea. Science 222, 616--618.
-\url{https://doi.org/10.1126/science.222.4624.616}
-
-\leavevmode\vadjust pre{\hypertarget{ref-whitmore2021}{}}%
-Whitmore, B.M., Ohman, M.D., 2021. Zooglider-measured association of
-zooplankton with the fine-scale vertical prey field. Limnology and
-Oceanography 66, 3811--3827. \url{https://doi.org/10.1002/lno.11920}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wood2017}{}}%
-Wood, S.N., 2017. Generalized additive models: an introduction with R,
-Second edition. ed, Texts in statistical science. CRC Press/Taylor \&
-Francis Group, Boca Raton London New York.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wood2001}{}}%
-Wood, S.N., 2001. mgcv: GAMs and Generalized Ridge Regression for R 1.
-
-\leavevmode\vadjust pre{\hypertarget{ref-zasko2005}{}}%
-Zasko, D.N., Rusanov, I.I., 2005. Vertical Distribution of Radiolarians
-and Their Role in Epipelagic Communities of the East Pacific Rise and
-the Gulf of California. Biology Bulletin 32, 279--287.
-\url{https://doi.org/10.1007/s10525-005-0103-5}
-
-\end{CSLReferences}
+\newpage
+
+end.
 

### 2024-01-29 R/s01_cast-map.R

@@ -1,6 +1,9 @@
 ###
 # Cast Map
 ###
+library(EcotaxaTools)
+library(ggplot2)
+library(dplyr)
 
 ctd_data <- readRDS('./data/00_ctd-data.rds')
 zoops <- readRDS("./data/00_zoop-uvp.rds")
@@ -33,5 +36,6 @@ all_cast_map <- ggplot() +
         axis.text.x = element_text(angle = 45, hjust = c(1,1)),
         legend.position = 'bottom')
 
+ggsave('./output/Supplement/s01_cast-map.pdf', all_cast_map)
 
 windows()

### 2024-05-02 manuscript/01_intro.qmd
update and rerender

@@
 The bulk of modern plankton research has focused on hard-bodied crustacea which are numerically dominant and easily sampled with nets and preservatives.
-Fragile organisms like Rhizaria, were difficult to adequately study as they can be destroyed through standard zooplankton sampling techniques and preserve poorly.
+Fragile organisms like Rhizaria were difficult to adequately study as they can be destroyed through standard zooplankton sampling techniques and preserve poorly.
 A number of studies in the late 1900s did employ alternative techniques to quantify Rhizaria including diaphragm pumps [@michaels1988] or blue-water SCUBA collections [@caron1984; @caron1995; @bijma1990].
@@
 Firstly, taxonomists have been able to greatly refine the understanding of evolutionary relationships amongst these diverse protists [@aurahs2009; @decelle2012; @decelle2013; @biard2015; @cavalier-smith2018; rev by @biard2022].
-DNA metabarcoding studies have revealed insights into the distributional patterns [@decelle2013; @biard2017; @marsbrisbin2020; @blanco-bercial2022 @llopismonferrer2022; @sogawa2022], ecological relationships [@decelle2012; @nakamura2023], and contribution to biogeochemical fluxes [@guidi2016; @gutierrez-rodriguez2019].
-Transcriptomic and proteomic approaches also have been used to quantify rhizarian contribution to community metabolism [@cohen2023].
-Yet, despite the excellent taxonomic resolution provided by molecular approaches, they do not provide a truly quantitative metric for estimating Rhizarian abundance or biomass.
-In-situ imaging tools however, offer the ability to observe organisms in the natural state and quantify their abundance [@ohman2019a; @barth-inrev].
-While there were early applicaitons of imaging tools to document Rhizaria [@dennett2002], @biard2016's report from a global imaging dataset highlighted the importance of Rhizaria to the total standing stock of marine carbon.
+DNA metabarcoding studies have revealed insights into the distributional patterns [@decelle2013; @biard2017; @marsbrisbin2020; @blanco-bercial2022; @llopismonferrer2022; @sogawa2022], ecological relationships [@decelle2012; @nakamura2023], and contribution to biogeochemical fluxes [@guidi2016; @gutierrez-rodriguez2019].
+Transcriptomic and proteomic approaches also have been used to quantify Rhizaria contribution to community metabolism [@cohen2023].
+Yet, despite the excellent taxonomic resolution provided by molecular approaches, they do not provide a truly quantitative metric for estimating Rhizaria abundance or biomass.
+In-situ imaging tools however, offer the ability to observe organisms in the natural state and quantify their abundance [@ohman2019a; @barth-inpress].
+@biard2016 utilized in-situ imaging at a global scale to suggest Rhizaria were substantial contributors to the ocean carbon standing stock.
+While more recent calculations suggest lower carbon contribution [@laget2024], Rhizaria nonetheless have substantial influences on biogeochemical cycling.
 Due to their large sizes, ability to concentrate smaller particles and the unique structure of their mineral skeletons, Rhizaria have the potential to massively influence ocean biogeochemical cycling.
 A number of studies have made large advances in estimating the contribution of Rhizaria to ocean cycling of carbon [@lampitt2009; @stukel2018; @ikenoue2019; @gutierrez-rodriguez2019], silica [@biard2018; @llopismonferrer2021], and strontium [@decelle2013].
-Still, Rhizarian ecological roles are not well understood [@biard2022].
+Still, Rhizaria ecological roles are not well understood [@biard2022].
 This is a major challenge as it is critical to understand the ecological role of plankton to fully incorporate them into biological oceanographic models.
@@
 As zooplankton, rhizaria are predominately heterotrophic [@biard2022], yet their feeding modes can be quite varied.
-Phaeodarians (family Cercozoa) are largely thought to be flux-feeders, collecting and feeding on sinking particles [@nakamura2015; @stukel2019].
+Phaeodarias (family Cercozoa) are largely thought to be flux-feeders, collecting and feeding on sinking particles [@nakamura2015; @stukel2019].
 Alternatively, Retaria can be either exclusively heterotrophic or mixotrophic, utilizing photosynthetic algal symbionts [@anderson2014; @decelle2015].
-Mixotrophic foraminifera host a variety of endosymbiont partners [@lee2006; @decelle2015], which are thought to support early and adult life stages and contribute to total primary productivity [@kimoto2015].
-Still, foraminifera are omnivorous, possibly even predominately carnivorous with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], mainly consuming live copepods [@caron1984].
-Radiolaria have several lineages all which have some taxa who are well known to host symbionts [@biard2022b].
-Amongst radiolarians, arguably the most widespread are Collodaria who can be either large solitary cells or form massive colonies, up to several meters in length [@swanberg1981].
+Mixotrophic foraminifera host a variety of endosymbiont partners [@lee2006; @decelle2015], which are thought to support early and adult life stages and significantly contribute to total primary productivity [@kimoto2015].
+Still, foraminifera are omnivorous, possibly even predominately carnivorous, with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], mainly consuming live copepods [@caron1984].
+Radiolaria have several lineages, all of which have some taxa well known to host symbionts [@biard2022b].
+Amongst Radiolarias, arguably the most widespread are Collodaria who can be either large solitary cells or form massive colonies, up to several meters in length [@swanberg1981].
 All known Collodaria species host dinoflagellate symbionts [@biard2022b] and can contribute substantially to primary productivity, particularly in oligotrophic ocean regions [@caron1995; @dennett2002].
 This Collodaria-symbiont association has been suggested as a reason for their high abundances throughout the photic zone of oligotrophic environments globally [@biard2016; @biard2017].
-A few Acantharean (Radiolarian order) clades host algal symbionts [@decelle2012; @biard2022b], notably with two clades forming an exclusive relationship with Phaeocystis.
-However, globally Acantharea are less abundant than Collodaria [@biard2022] and contribute less to total primary productivity [@michaels1995].
+A few Acantharea (Radiolaria order) clades host algal symbionts [@decelle2012; @biard2022b], notably with two clades forming an exclusive relationship with *Phaeocystis*.
+However, globally, Acantharea are less abundant than Collodaria [@biard2022] and contribute less to total primary productivity [@michaels1995].
 This may be due to the fact several clades of Acantharea are cyst-forming and strictly heterotrophs [@decelle2013; @biard2022b].
-Furthermore, @marsbrisbin2020 documented apparent predation behavior in Acantharea near the surface, suggesting that there may be a large reliance on carnivory.
+Furthermore, @marsbrisbin2020 documented apparent predation behavior in Acantharea near the surface, suggesting that there may be a larger reliance on carnivory.
 
-Given the high abundances, yet diverse trophic strategies found among Rhizarian taxa, it is reasonable to expect some form of niche partitioning.
+Given the high abundances, yet diverse trophic strategies found among Rhizaria taxa, it is reasonable to expect some form of niche partitioning.
 A number of studies do suggest evidence for vertical zonation between Rhizaria groups according to various trophic strategies.
-Taxa-specific studies of radiolarians suggest they may be restricted to the euphotic zone [@michaels1988; @boltovskoy2017].
-Although some studies report Acantharea in deeper waters [@decelle2013; @gutiérrez-rodríguez2022].
-Phaeodarians alternatively, are generally found in the mesopelagic where photosynthesis cannot occur but they can feed on sinking particles [@stukel2018].
+Taxa-specific studies of Radiolarias suggest they may be restricted to the euphotic zone [@michaels1988; @boltovskoy2017].
+Although some studies report Acantharea in deeper waters [@decelle2013; @gutiérrez-rodríguez2022], Phaeodarias, alternatively, are generally found in the mesopelagic where photosynthesis cannot occur, but they can feed on sinking particles [@stukel2018].
 In an imaging-based study of the whole Rhizaria community, @biard2020 noted clear patterns in vertical zonation which largely corresponded to different trophic roles.
@@
 Furthermore, no studies have utilized imaging, arguably the best method for quantifying rhizaria, consistently throughout the full mesopelagic.
-Given this lack of information, there are many unknowns with respect to Rhizarian ecology, seasonality and phenology across different groups.
+Given this lack of information, there are many unknowns with respect to Rhizaria ecology, seasonality and phenology across different groups.
 
-In this study, we present a comprehensive assessment of large Rhizaria measured for multiple months (greater than 1 year) using an in-situ imaging approach.
+In this study, we present a comprehensive assessment of large Rhizaria measured for over a year from regularly occurring cruises at monthly intervals.
+We utilized an in-situ imaging approach to facilitate abundance calculations.
 With this dataset, we address two critical aims.
 1) Quantification of large Rhizaria throughout the epipelagic (0-200m) and mesopelagic (200-1000m) over the course of an annual cycle.
-These data were collected in the Sargasso Sea, and represents the first study of its kind in an oligotrophic system.
-2) We aim to test the hypothesis that Rhizaria exhibit niche partitioning according to trophic roles.
+These data were collected in the Sargasso Sea, and represents the first study of its kind in an oligotrophic system; and 2) We aim to test the hypothesis that Rhizaria exhibit niche partitioning according to trophic roles.
 This hypothesis makes several predictions, including vertical zonation, as seen in prior studies, but also that environmental variables related to trophic strategy will explain abundance patterns.
-Specifically, autotrophic/mixotrophic taxa will correspond to variables related to autotrophy (chl-a concentration, primary productivity, high $O_2$ ) and other rhizaria will correspond to factors which promote heterotrophy (particle concentration, flux, and low $O_2$).
+Specifically, autotrophic/mixotrophic taxa will correspond to variables related to autotrophy (chl-a concentration, primary productivity, local DO maxima) and other rhizaria will correspond to factors which promote heterotrophy (particle concentration, flux, and local DO minima).

### 2024-05-02 manuscript/02_methods.qmd
update and rerender

@@
 
-Data were collected in collaboration with the Bermuda Atlantic Time-Series Study [@michaels1996; @lomas2013] on board the R/V Atlantic Explorer.
+Data were collected in collaboration with the Bermuda Atlantic Time-series Study [@michaels1996; @lomas2013] on board the R/V Atlantic Explorer.
 Cruises were conducted at approximately monthly intervals.
@@
 The UVP5 was mounted to the sampling rosette and collected data autonomously on routine casts, from which only the downcast data are utilized.
-The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022.
-Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda into the Sargasso Sea (approximately 31.0$^oN$-32.5$^oN$, 64.25$^oW$-63$^oW$; Supplemental Figure 1).
+The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022, during which time the BATS region was sampled for 3-5 days at monthly intervals.
+Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda in the Sargasso Sea (approximately 31.0$^oN$-32.5$^oN$, 64.25$^oW$-63$^oW$; Supplemental Figure 1).
 In general, casts extended to either 200m, 500m, or 1200m deep, with a few extended into the bathypelagic (4500m).
@@
 Particle concentration was calculated from this data for all particles 184$\mu m$ - 450$\mu m$.
-The lower size range was set by what could be reliably sampled by the UVP5's pixel resolution (\>2px) and the upper size range is representative of a potential prey field for mesozooplankton [@whitmore2021].
-Salinity, temperature, Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were measured at high-frequencies on each cast, attached to the same CTD rosette as the UVP.
-On certain casts, niskin bottles were used to collect bacterial abundance estimates (via epifluorescence microscopy) as well as measure inorganic nutrients ($NO_3$, and $Si$ as silicate/sicilic acid) at discrete depths.
-On each cruise, flux estimates of total mass, carbon, and nitrogen were also collected using sediment traps (in the present study we utilize flux to the mesopelagic as the flux at 200m).
-Also primary productivity was estimated through measuring $C^{14}$ uptake rates from in-situ incubations.
-Full descriptions of the BATS sampling program, methods, and data are available freely online [(https://bats.bios.asu.edu/bats-data/)](https://bats.bios.asu.edu/bats-data/) and reviewed in [@lomas2013].
+The lower size range was set by what could be reliably sampled by the UVP5's pixel resolution (\>2px; 0.092mm per pixel) and the upper size range is representative of a potential prey field for mesozooplankton [@whitmore2021].
+For each UVP cast supporting continuous profiles of the CTD parameters salinity, temperature, and auxiliary CTD channels; Dissolved Oxygen (DO), in-situ chlorophyll fluorescence were measured at 24Hz using the BATS CTD package.
+On select casts, Niskin bottles were used to collect bacterial abundance estimates (via epifluorescence microscopy) as well as measure inorganic nutrients ($NO_3$, and $Si$ as silicate/sicilic acid) at discrete depths.
+On each cruise, flux estimates of total mass, organic carbon, and nitrogen were also collected using sediment traps; in the present study we utilized flux to the mesopelagic as the flux at 200m.
+Also primary productivity was estimated through measuring $^{14}C$ uptake rates from in-situ incubations.
+For full descriptions of the BATS sampling program and methods, see @knap1997 and @lomas2013 for a review.
+Additionally, data can be viewed online [(https://bats.bios.asu.edu/bats-data/)](https://bats.bios.asu.edu/bats-data/).
 
-Environmental data were processed in a variety of ways to match the format of the Rhizaria abundance estimates (See below).
-CTD data were collected at higher frequency than the UVP, so these data were averaged within matching bins to the UVP5 data.
-Data from niskin bottles were first linearly interpolated with a 1m resolution for each cast where data were available.
-Then, because data were collected at different frequencies over the course of a cruise, the interpolated data were averaged across casts, then averaged into matching UVP5-sized bins.
+Environmental data were processed in a variety of ways to match the format of the Rhizaria abundance estimates (see below).
+CTD data were collected at higher frequency than the UVP (24Hz vs 15Hz respectively), so these data were averaged within matching bins to the UVP5 data.
+Data from Niskin bottles were first linearly interpolated in depth at 1m resolution then time averaged over the cruise, then subsequently averaged into matching UVP5-sized bins.
 Primary productivity estimates were totaled within the euphotic zone to represent a "total euphotic productivity".
-Finally the data from the sediment trap deployed to 200m was used to represent "flux to the mesopelagic".
 
@@
 While there are sparse taxonomic guides for in-situ images of rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020].
-Using the aforementioned sources and publicly available ecotaxa projects, I constructed a guide which can be found here: [https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf](#0){.uri}.
-Broadly, Rhizaria were classified as Foraminifera, Radiolarians (Acantharea or Collodaria), or a variety of Phaeodarian families (Figure 1).
+Using the aforementioned sources and publicly available ecotaxa projects, we constructed a guide accessible at: <https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf>.
+Broadly, Rhizaria were classified as Foraminifera, Radiolarias (Acantharea or Collodaria), or as a variety of Phaeodaria families (Figure 1).
 When identification could not be confidently made between a few candidate taxa, a less specific label was used.
-As a result, we have data from "unidentified Rhizaria", which typically were vignettes not distinguishable between Aulacanthidae or Acantharea or "unidentified Phaeodaria", which are clearly phaeodaria but not distinguishable into a family.
+As a result, we have data from "unidentified Rhizaria", which typically were vignettes not distinguishable between Aulacanthidae or Acantharea or "unidentified Phaeodaria", which are clearly Phaeodaria but not distinguishable into a family.
+
+![Example images of different Rhizaria taxa. 4mm scale bar shown in lower right. All vignettes are same scale](images/01_taxa.pdf)
 
@@
 However, identified rhizaria ranged from a 934$\mu m$ Aulacanthidae cell to a Collodarian colony over 10mm in diameter.
-To confirm that the UVP5 was sampling adequately across all size ranges, an NBSS slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth-inrev].
+To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth-inpress].
 However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded.
@@
 Our deployments had variable descent depths and speeds with more casts descending to 500m than 1000m and descents quicker through the epipelagic than the mesopelagic (see @barth2022 for an extended discussion of UVP5 data processing).
-For the present study, Rhizarian abundances were estimated in 25m vertical bins, which offer a moderate sampling volume per bin (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic) while still maintaining ecologically relevant widths.
-However, concentrations in a 25m bin would need to be greater than 2.428 indv.
-$m^{-3}$ and 3.912 indv.
-$m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk \[@benfield1996; @barth-inrev\].
+For the present study, Rhizaria abundances were estimated in 25m vertical bins, which offer a moderate sampling volume per bin (average 0.948$m^3$ in the epipelagic and 0.589$m^3$ in the mesopelagic; Supplemental Table 1) while still maintaining ecologically relevant widths.
+However, concentrations in a 25m bin would need to be greater than 2.428 ind.
+$m^{-3}$ and 3.912 ind.
+$m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth-inpress].
 Because we typically observed many rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions.
@@
 Due to the variable descent depths of the UVP, data are categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic (500-1000m).
-The average sampling volume integrated through these regions were 7.59$m^2$, 7.06$m^2$, and 11.77$m^2$, with non-detection thresholds at 0.30 indv.
-$m^{-2}$, 0.33 indv.
-$m^{-2}$, and 0.20 indv.
+The average sampling volume integrated through these regions were 7.59$m^2$, 7.06$m^2$, and 11.77$m^2$, with non-detection thresholds at 0.30 ind.
+$m^{-2}$, 0.33 ind.
+$m^{-2}$, and 0.20 ind.
 $m^{-2}$ respectively.
@@
 
-## Modelling environmental controls of Rhizarian Abundance
+## Modelling environmental controls of Rhizaria Abundance
 
-Generalized Additive Models (GAMs) were used to assess the relationship between integrated Rhizarian abundance and different environmental factors.
-GAMs offer the ability to model non-linear and non-monotonic relationships, which can be particularly useful in assessing ecological relationships [@wood2017] and have been successfully applied to Rhizarian ecology [@biard2020].
+Generalized Additive Models (GAMs) were used to assess the relationship between integrated Rhizaria abundance and different environmental factors.
+GAMs offer the ability to model non-linear and non-monotonic relationships, which can be particularly useful in assessing ecological relationships [@wood2017] and have been successfully applied to Rhizaria ecology [@biard2020].
 The `mgcv` package [@wood2001] was used to construct models relating environmental parameters to each taxonomic group's integrated abundance estimates from each cast.

### 2024-05-02 manuscript/03_results.qmd
update and rerender

@@
 
-The BATS sampling region is southeast of Bermuda, situated in the North Sargasso Sea and the North Atlantic oligotrophic ocean gyre.
-Due to the sampling location, while the environmental conditions are generally low in variation and oligotrophic, there is some seasonality and considerable influence from mesoscale eddies [@mcgillicuddy1998; @lomas2013].
+The BATS sampling region is southeast of Bermuda, situated in the oligotrophic North Atlantic Subtropical Gyre.
+Due to the sampling location, while the environmental conditions are generally low in variation and oligotrophic, there is considerable influence from deep winter-mixing and summer stratification as well as secondary influences from mesoscale eddies which result in spatiotemporal heterogeniety [@mcgillicuddy1998; @lomas2013].
 Variability in the water column structure was visible during the study period (Figure 2).
@@
 This warm, stratified period appeared more intense during the few months sampled in 2019.
-In 2021, we observed the stratified layer slowly dissipated into the winter months.
-There was a consistent oxygen minimum zone (OMZ) located at about 750m deep (Figure 2C).
-Although, February 2021 saw a notable downwelling event, likely due to a passing cyclonic eddy.
-During this time, warmer, oxic water was plunged deeper into the mesopelagic.
-This process was reversed in the spring months (March, April) when mixing occurred throughout the epipelagic and the surface was cooler and well mixed.
+In 2021, we observed the stratified layer slowly dissipated into the winter months following mixed layer entrainment.
+There was a consistent oxygen minimum zone (OMZ) located at about 800m deep (Figure 2C).
+February 2021 saw a notable downwelling event, likely due to a passing anti-cyclonic eddy which impacted the local region during early 2021.
+During this phaes, warmer, oxic water was significantly depressed deeper into the mesopelagic.
+This process was reversed in the spring months (March, April) primary due to the interaction of convective mixing and the passing of a strong cyclonic eddy resulting in a deep cold mixed layer.
 Primary production was highest during the spring mixing period, evidenced both by in-situ fluorescence (Figure 2D) and productivity incubation experiments (Figure 3A).
@@
 
+![Environmental profiles across time-series of study period.
+Y axis shows depth in meters.
+(A) Temperature.
+(B) Salinity.
+(C) Dissolved Oxygen.
+(D) In-situ chlorophyll fluorescence.
+(E) Particle concenctrion (184 - 450$\mu m$).
+(F) Bacteria Abundance.
+(G) Silica.
+(H) Nitrate](images/02-environmental.pdf)
+
 Overall, particle concentration was high near the surface during the 2021 spring bloom, then moved deeper throughout the water column attenuating throughout the lower epipelagic (Figure 2E).
-Similarly, bacteria abundance was closely linked to overall productivity, althrough there was a more consistent moderate-abundance layer near the top of the mesopelagic (\~250m) (Figure 2F).
+Similarly, heterotrophic bacteria abundance was closely linked to overall productivity, althrough there was a more consistent moderate-abundance layer near the top of the mesopelagic (\~250m) (Figure 2F).
 Concurrent with the secondary fall production peak, there was also higher particle concentration and bacterial abundance in the later summer and early fall.
-Interestingly, while primary productivity estimates from July-August were not that different between 2019 and 2021 (Figure 4A), chlorophyll-a florescence, particle concentration, and bacterial abundance were much higher in 2019's summer/fall (Figure 2D-F).
+Interestingly, while primary productivity estimates from July-August were not that different between 2019 and 2021 (Figure 3A), chlorophyll-a florescence, particle concentration, and bacterial abundance were much higher in 2019's summer/fall (Figure 2D-F).
 Inorganic nutrients ($Si$ and $NO_3$) were generally well stratified, with low concentrations in the epipelagic and increasing throughout the mesopelagic.
@@
 Additionally, in the late fall of 2021, $Si$ concentrations were slightly elevated in the mid-mesopelagic (Figure 2G).
+
+![Primary productivity (A) and flux estimates from total mass (B), carbon (C), and nitrogen (D).
+Values are from monthly cruises with month displayed in corresponding colors.
+Absent data are shown by grey bar.](images/03_prod-flux.pdf)
 
@@
 The large contribution of Rhizaria to the mesozooplankton community is most prominent in the epipelagic, where they accounted for 47% of all mesozooplankton.
-In the mesopelagic rhizaria were a smaller fraction, at 38% in the upper layers (200-500m) and 37% in the deeper mesopelagic (500-1000m).
+In the mesopelagic rhizaria were a smaller (but still prominent fraction), at 38% in the upper layers (200-500m) and 37% in the deeper mesopelagic (500-1000m).
+
+![Average abundance of Rhizaria in 25m bins, across entire study period.
+Shown are total Rhizaria (A), Acantharea (B), Collodaria (C), Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae (G), Coelodendridae (H).](images/04_vertical.pdf)
 
@@
 The vertical distribution pattern and abundance varied considerably across taxonomic groups.
-Radiolarians were some of the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5B).
+Radiolarias were some of the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5B).
 This pattern was led by Collodaria, whose colonies were abundant in the upper epipelagic and declined into the top of the mesopelagic (Figure 4C).
@@
 Foraminifera had a similar bimodal distribution, yet their overall average densities were much lower and spread wider throughout the mesopelagic (Figure 4E).
-Phaeodarian families exhibited a wide range of vertical distribution patterns.
+Phaeodaria families exhibited a wide range of vertical distribution patterns.
 The most abundant, Aulacanthidae, also had a bimodal pattern but the density was highest in the lower mesopelagic (Figure 4D).
 Aulosphaeridae had low average density and was nearly homogeneously distributed throughout the water column, although slightly lower in the epipelagic (Figure 4F).
-Castanellidae were the only Phaeodarian who appeared to be effectively restricted to the photic zone (Figure 4G).
+Castanellidae were the only Phaeodaria who appeared to be effectively restricted to the photic zone (Figure 4G).
 Alternatively, Coelodendridae primarily occurred in the lower mesopelagic (Figure 4H).
 A few individuals from the families Tuscaroridae and Medusettidae were also observed in the mesopelagic, yet they were much rarer (data not shown).
+
+![Seasonality of Rhizaria integrated abundance for the epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower mesopelagic (500-1000m) (E-F).
+Left panels (A,C,E) display total integrated abundance per monthly cruise colored by month.
+Right panels (B, D, F) display community composition of each total abundance.](images/05_seasonality.pdf)
 
@@
 The 2019 later summer - fall period also had much higher integrated abundance than similar months in 2021.
-While the majority of integrated abundance in the epipelagic was consistently attributable to Collodaria, Acanthrean abundance occurred sporadically and could account for a large portion of the total in some months (Figure 5B).
+While the majority of integrated abundance in the epipelagic was consistently attributable to Collodaria, Acanthrea abundance occurred sporadically and could account for a large portion of the total in some months (Figure 5B).
 The mesopelagic integrated abundance was much more consistent across monthly cruises, although average abundance was notably higher in 2019 (Figure 5C-F).
-The community composition in the mesopelagic was more diverse, mostly comprised of Phaeodarians.
+The community composition in the mesopelagic was more diverse, mostly comprised of Phaeodarias.
 However, Acantharea and unidentified Rhizaria also were common members of the community (Figure 5D, 5F).
@@
 To investigate if the populations or morphologies shifted throughout the water column, we compared the sizes (Equivalent Spherical Diameters, ESD) between mesopelagic and epipelagic groups for each taxa.
-All groups were significantly different on average (Wilcox Rank Sum p-value <0.001).
+All groups were significantly different on average (Wilcox Rank Sum p-value \<0.001).
 Acantharea were smaller, on average in the mesopelagic while all other taxa tended to be larger (Figure 6).
 
-### Environmental Drivers of Rhizarian Abundance
+![Comparison of average sizes (ESD) amongst Rhizaria taxa which occurred throughout the water column.](images/09_size-comp.pdf)
 
-For total Rhizarian integrated abundance the GAMs produced moderate fits ($R^2_{adj}$ = 0.406-0.603) (Table 1).
+### Environmental Drivers of Rhizaria Abundance
+
+For total Rhizaria integrated abundance the GAMs produced moderate fits ($R^2_{adj}$ = 0.406-0.603) (Table 1).
 In the epipelagic, there were several significant predictor variables including inorganic nutrients ($NO_3$ and $Si$), water quality parameters (Salinity, DO), primary production, and particle concentration (Table 1).
 However, the upper and lower mesopelagic were exclusively explained by particle-related variables (concentration and mass flux) (Table 1).
+
+| **Model**                          | Term                   | edf    | F     | p       |
+|--------------------|---------------|-------------|-------------|-------------|
+| **All Rhizaria Epipelagic**        | Salinity               | 2.0818 | 4.654 | \<0.001 |
+| $R^2_{adj}$= 0.42                  | DO                     | 0.8272 | 0.945 | 0.0136  |
+|                                    | Silica                 | 3.1206 | 15.75 | \<0.001 |
+|                                    | NO3                    | 2.9970 | 6.268 | \<0.001 |
+|                                    | Primary Productivity   | 1.8367 | 2.099 | \<0.001 |
+|                                    | Particle Concentration | 0.9282 | 2.566 | \<0.001 |
+| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6847 | 0.431 | 0.073   |
+| $R^2_{adj}$= 0.406                 | Avg Mass Flux          | 1.5426 | 0.923 | 0.045   |
+|                                    | Particle Concentration | 3.2957 | 20.89 | \<0.001 |
+| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 1.6632 | 1.824 | 0.002   |
+| $R^2_{adj}$= 0.603                 | Particle Concentration | 0.7694 | 0.662 | 0.027   |
+
+: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column.
 
@@
 Note that due to low abundances, GAMs were not constructed for Tuscaroridae or Medusettidae.
-Furthermore no significant terms were found for a model with Aulosphaeridae in the epipelagic nor Foraminifera in the mesopelagic.
+Furthermore no, significant terms were found for a model with Aulosphaeridae in the epipelagic nor Foraminifera in the mesopelagic.
+
+![Partial effects of smooth terms in taxa-specific GAM models from the epipelagic (0-200m).
+Effects are grouped by taxa; Acantharea (A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae (E).](images/06_epi-partials.pdf)
 
 Epipelagic Acantharea were explained by several predictor variables and had a good fit ($R^2_{adj}$ = 0.53, Table 2).
-Most notable smooths were mass flux and particle concentration, which had a weak positive association (Figure 7A), with July 2021 as a clear outlier where Acantharean abundances were high in the epipelagic despite lower fluxes and particle concentrations (Figure 5).
+Most notable smooths were mass flux and particle concentration, which had a weak positive association (Figure 7A), with July 2021 as a clear outlier where Acantharea abundances were high in the epipelagic despite lower fluxes and particle concentrations (Figure 5).
 Foraminifera had a good fitting GAM in the epipelagic ($R^2_{adj}$ = 0.445).
@@
 
+| Model                                | Term                   | edf   | F     | p       |
+|---------------|---------------|---------------|---------------|---------------|
+| **Acantharea Epipelagic**            | Salinity               | 2.264 | 4.113 | \<0.001 |
+| R2adj=0.53                           | O2                     | 2.579 | 6.712 | \<0.001 |
+|                                      | Avg Mass Flux          | 4.810 | 22.81 | \<0.001 |
+|                                      | Bacteria #/L           | 1.599 | 1.317 | 0.0117  |
+|                                      | Particle Concentration | 0.853 | 1.158 | 0.0087  |
+| **Acantharea Upper Mesopelagic**     | Avg C Flux             | 1.296 | 0.621 | 0.0439  |
+| R2adj=0.231                          | Avg N Flux             | 1.619 | 1.283 | 0.0026  |
+|                                      | Particle Concentration | 0.952 | 3.886 | \<0.001 |
+| **Acantharea Lower Mesopelagic**     | Temperature            | 2.076 | 4.155 | \<0.001 |
+| R2adj=0.509                          | Avg N Flux             | 1.494 | 1.216 | 0.0113  |
+|                                      | Primary Productivity   | 0.766 | 0.648 | 0.0253  |
+|                                      | Particle Concentration | 2.037 | 10.03 | \<0.001 |
+| **Aulacanthidae Epipelagic**         | Salinity               | 0.792 | 0.757 | 0.0241  |
+| R2adj=0.251                          | Avg N Flux             | 0.869 | 1.324 | 0.0018  |
+|                                      | Primary Productivity   | 2.312 | 7.704 | \<0.001 |
+|                                      | Particle Concentration | 2.008 | 2.200 | 0.0017  |
+| **Aulacanthidae Upper Mesopelagic**  | Particle Concentration | 2.832 | 9.802 | \<0.001 |
+| R2adj = 0.158                        |                        |       |       |         |
+| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.622 | 1.991 | 0.002   |
+| R2adj=0.298                          | Particle Concentration | 2.123 | 6.706 | \<0.001 |
+| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.653 | 13.06 | \<0.001 |
+| R2adj=0.2                            |                        |       |       |         |
+| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 0.972 | 6.248 | \<0.001 |
+| R2adj=.147                           |                        |       |       |         |
+| **Castanellidae Epipelagic**         | Avg C Flux             | 1.462 | 0.816 | 0.0421  |
+| R2adj = 0.124                        | Primary Productivity   | 0.771 | 0.665 | 0.0247  |
+|                                      | Particle Concentration | 3.956 | 4.623 | \<0.001 |
+| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.822 | 0.919 | 0.0183  |
+| R2adj=.113                           | Particle Concentration | 0.970 | 6.208 | \<0.001 |
+| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 1.773 | 4.873 | \<0.001 |
+| R2adj=.133                           |                        |       |       |         |
+| **Collodaria Epipelagic**            | Salinity               | 0.925 | 2.267 | \<0.001 |
+| R2adj=0.16                           | O2                     | 2.015 | 2.217 | 0.002   |
+|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.002   |
+| **Foraminifera Epipelagic**          | Temperature            | 4.414 | 8.789 | \<0.001 |
+| R2adj=0.445                          | O2                     | 1.603 | 1.287 | 0.0111  |
+|                                      | Avg N Flux             | 2.512 | 3.396 | \<0.001 |
+|                                      | Primary Productivity   | 0.824 | 0.920 | 0.0153  |
+|                                      | Particle Concentration | 0.919 | 2.257 | \<0.001 |
+
+: Taxa-specific generalized additive models for different regions of the water column.
+
+![Partial effects of smooth terms in taxa-specific GAM models from the upper mesopelagic (200-500m).
+Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/07_upmeso-partial.pdf)
+
 In the upper mesopelagic (200-500m), abundances were generally low (Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae, Aulosphaeridae, and Coelodendridae (Table 2).
@@
 Acantharea in the lower mesopelagic was most clearly positively associated with particle concentration and nitrogen flux, as well as temperature to a slight degree (Figure 9A).
-For all Phaeodarians with a significant model, particle concentration was a main predictor variable (Table 2, Figure 9B-D).
-Aulacanthidae had the best fitting model of the Phaeodarians ($R^2_{adj}$ = 0.298), which also included mass flux as a statistically significant smooth (Figure 9B).
+For all Phaeodarias with a significant model, particle concentration was a main predictor variable (Table 2, Figure 9B-D).
+Aulacanthidae had the best fitting model of the Phaeodarias ($R^2_{adj}$ = 0.298), which also included mass flux as a statistically significant smooth (Figure 9B).
+
+![Partial effects of smooth terms in taxa-specific GAM models from the lower mesopelagic (500-1000m).
+Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/08_lomeso-partials.pdf)

### 2024-05-02 manuscript/04_discussion.qmd
update and rerender

@@
 
-### Overall Rhizarian abundance and patterns
+### Overall Rhizaria abundance and patterns
 
 In the epipelagic Rhizaria exhibited a notable seasonal pattern.
-Rhizarian abundances were higher in the summer months and lower during the winter.
+Rhizaria abundances were higher in the summer months and lower during the winter.
 During a prior time period, @blanco-bercial2022 noted that there is considerable seasonality in the community composition of all protists.
-Despite the seasonality of total Rhizarian abundance, community composition was relatively consistent, with Collodaria representing the bulk of the community.
+Despite the seasonality of total Rhizaria abundance, community composition was relatively consistent, with Collodaria representing the bulk of the community.
 It should be noted that the overall taxonomic resolution of the UVP5 is fairly low, so there may be a switching of species within the broad groups identified in this study which were not captured.
 Throughout the mesopelagic, month-to-month variation in 2021 was relatively low.
-Again this is consistent with observations from metabarcoding of the whole protist community in the same study region [@blanco-bercial2022].
+Again, this is consistent with observations from metabarcoding of the whole protist community in the same study region [@blanco-bercial2022].
 This finding is not surprising as the overall seasonal variation in environmental conditions in this region were low.
 
-Overall Rhizaria were the most commonly identified group of mesozooplankton throughout the study period.
-We do note that the UVP5 commonly captures Trichodesmium colonies, yet these were excluded in this comparison as they are strictly autotrophs.
+Overall Rhizaria were the most commonly identified group of mesoplankton throughout the study period.
+We do note that the UVP5 commonly captures *Trichodesmium* colonies, yet these were excluded in this comparison as they are strictly autotrophs.
 It should be noted that previous work has suggested that avoidance behavior with the UVP is possible, at times likely, for visual and highly mobile zooplankton [@barth2022].
@@
 Likely, Rhizaria display considerable interannual variability.
-In the present study, we noticed considerably higher Rhizarian abundance throughout the water column in 2019 compared to 2021.
+In the present study, we noticed considerably higher Rhizaria abundance throughout the water column in 2019 compared to 2021.
 While this may have been driven by increased mass flux, more information is needed to truly understand the magnitude by which Rhizaria can vary interannually.
@@
 Global investigations of polycystine flux, suggest that deep-Collodaria in Oligotrophic regions may be a consequence of isothermal submersion [@boltovskoy2017].
-Another effectively exclusively epipelagic Rhizaria was the Phaeodarian family of Castanellidae.
-All Phaeodarians are thought to be fully heterotrophic [@nakamura2015], nonetheless a number of studies, including this one, report Castanellidae to be typically found in the lower epipelagic [@zasko2005; @biard2018; @biard2020].
+Alternatively, surface waters at BATS often mix into the mode water during the seasonal mixing, so Collodaria in the deeper waters may be a result of diapyncal mixing.
+Another effectively exclusively epipelagic Rhizaria was the Phaeodaria family of Castanellidae.
+All Phaeodaria are thought to be fully heterotrophic [@nakamura2015], nonetheless a number of studies, including this one, report Castanellidae to be typically found in the lower epipelagic [@zasko2005; @biard2018; @biard2020].
 It should be considered that perhaps Castanellidae specializes in feeding on sinking particles directly at the base of the epipelagic.
@@
 Perhaps it is more advantageous for Rhizaria to situate deeper, in darker regions of the twilight zone.
-Also it should be noted that Phaeodarians utilize silica to build their opaline tests, and silica concentrations started to increase around 500m (Figure 2G).
+Also it should be noted that Phaeodaria utilize silica to build their opaline tests, and silica concentrations started to increase around 500m (Figure 2G).
 Although $Si$ was not a significant smooth for any taxa-specific model, this lack of association might be due to the overall lack of variation of $Si$ between integrated abundance of each cast.
@@
 Most prior studies of Acantharea vertical distribution found them concentrated in near surface layers of the water column [@michaels1988 @zasko2005; @biard2020].
-This would support the paradigm that large Acantharean abundances may be supported by their mixotrophic abilities [@michaels1995; @suzuki2015].
-While the UVP5 images cannot distinguish between mixotrophic and heterotrophic Acantharea, the GAMs constructed for Acantharean abundance found positive associations with particle concentration and mass flux, suggesting a higher reliance on heterotrophy.
+This would support the paradigm that large Acantharea abundances may be supported by their mixotrophic abilities [@michaels1995; @suzuki2015].
+While the UVP5 images cannot distinguish between mixotrophic and heterotrophic Acantharea, the GAMs constructed for Acantharea abundance found positive associations with particle concentration and mass flux, suggesting a higher reliance on heterotrophy.
 Recently @marsbrisbin2020 described apparent predator behavior amongst near-surface Acantharea.
@@
 Furthermore, given that horizontal transfer of symbionts between generations of Acantharea is unlikely due to their spawning behavior, the newly spawned mesopelagic Acantharea are not necessarily required to rapidly return to the photic zone [@decelle2012; @decelle2013].
-This hypothesis predicts that Acanthareans in the mesopelagic would be smaller [@decelle2013].
+This hypothesis predicts that Acanthareas in the mesopelagic would be smaller [@decelle2013].
 @marsbrisbin2020 provided some support for this hypothesis, with a significant decrease in Acantharea sizes with depth.
@@
 Since food is more scarce in the mesopelagi, nutritional quality lower, yet flux feeders would likely grow larger to increase their feeding range [@biard2020].
-In the data collecting in this study, Acanthareans in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth.
+In the data collecting in this study, Acanthareas in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth.
 This provides added support for the hypothesis that cyst-forming Acantharea may utilize different sections of the water column throughout their life cycle.
@@
 
-This study provides a detailed look at Rhizarian abundance over time throughout the water column in a major oligotrophic gyre.
+This study provides a detailed look at Rhizaria abundance over time throughout the water column in a major oligotrophic gyre.
 We show that their abundances are generally related to particle concentration and flux, although lack of environmental variability may have reduced the fit of our GAMs.
@@
 Thus, Rhizaria may act as an aggregation mechanism.
-However, this is largely speculation, to truly test this, more work is needed measuring Rhizarian flux.
+However, this is largely speculation, to truly test this, more work is needed measuring Rhizaria flux.
 
@@
 However, their role in food webs is not well understood.
-While this study represents a step-forward in our understanding of Rhizarian ecology, continued research on Rhizaria is much needed to better understand their ecology.
+While this study represents a step-forward in our understanding of Rhizaria ecology, continued research on Rhizaria is much needed to better understand their ecology.
 Particularly extended descriptive work to capture interannual patterns.

### 2024-05-02 manuscript/main.qmd
update and rerender

--- 2024-01-25 manuscript/main.qmd
@@
       address: 700 Sumter St 401, Columbia, SC 29208
-      email: AB93@email.sc.edu
-      phone: +1 (440) 668-8376
-  - name: Leo Blanco-Berical
+    - name: University of Texas at Austin
+      department: Marine Science Institute
+      address: 750 Channel Drive, Port Aransas, Texas, USA
+      email: alexander.barth1@austin.utexas.edu
+  - name: Leocadio Blanco-Berical
     affiliations:
     - name: Arizone State University
+      department: School of Ocean Futures
+      address: Bermuda Institute of Ocean Sciences 17 Biological Station, St Georges, GE 01
   - name: Rod Johnson
     affiliations: 
-    - name: Bermuda Institute of Ocean Sciences
-      address: Bermuda GE 01, UK
+    - name: Arizone State University
+      department: School of Ocean Futures
+      address: Bermuda Institute of Ocean Sciences 17 Biological Station, St Georges, GE 01
   - name: Joshua Stone
@@
     keep-tex: true
+    tbl-colwidths: [43, 33.75, 7.75, 7.75, 7.75]
+    geometry:
+      - left=1in
+      - textwidth=6.5in
+      - right=1in
+
+abstract: |
+  Recently studies have shown that Rhizaria, a super-group of marine protists, have a large role in pelagic ecosystems.
+They are unique in that they construct mineral tests out of silica, calcium carbonate, or strontium sulfate.
+As a consequence, Rhizaria can have large impacts on the ocean’s cycling of carbon and other elements.
+However, less is known about Rhizaria ecology or their role in the pelagic food-web.
+Some taxa, like certain Radiolarians, are mixotrophic, hosting algal symbionts.
+While other taxa are flux-feeders or even predatory carnivores.
+Some prior research has suggested that Rhizaria will partition vertically in the water column, likely due to different trophic strategies.
+However, very few studies have investigated their populations over extended periods of time.
+In this study, we present data investigating Rhizaria abundance and vertical distribution from over a year of monthly cruises in the Sargasso Sea.
+This study represents the first quantification of Rhizaria throughout the mesopelagic zone in an oligotrophic system for an extended period of time.
+We use this data to investigate the hypothesis that Rhizaria taxonomic groups will partition due to trophic mode.
+We also investigate how their abundance varies in accordance with environmental parameters.
+Rhizaria abundance was quantified using an Underwater Vision Profiler (UVP5), an in-situ imaging device.
+Ultimately, we show that different Rhizaria taxa will have unique vertical distribution patterns.
+Models relating their abundance to environmental parameters have mixed results, yet particle concentration is a common predictive variable, supporting the importance of heterotrophy amongst many taxa.
 ---
@@
 # References:
-:::{#refs}
+
+::: {#refs}
 :::

### 2024-05-02 manuscript/main.tex
update and rerender

@@
 \usepackage{xcolor}
+\usepackage[left=1in,textwidth=6.5in,right=1in]{geometry}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 % authors. See below if you don't need that
- {\large{Alex Barth*}}{\textsuperscript{1}}%
+ {\large{Alex
+Barth*}}{\textsuperscript{1}}\textsuperscript{,}{\textsuperscript{2}}%
 %
 , 
- {\large{Leocadio Blanco-Berical}}{\textsuperscript{2}}%
+ {\large{Leocadio Blanco-Berical}}{\textsuperscript{3}}%
 %
 , 
- {\large{Rod Johnson}}{\textsuperscript{2}}%
+ {\large{Rod Johnson}}{\textsuperscript{3}}%
 %
@@
 %
-{2}.~{Arizone State University}%
+{2}.~{University of Texas at Austin}%
+%
+, %
+{Marine Science Institute}%
+%
+%
+, %
+{750 Channel Drive, Port Aransas, Texas, USA}%
+%
+\par\hangindent=1em\hangafter=1%
+%
+{3}.~{Arizone State University}%
 %
@@
 Recently studies have shown that Rhizaria, a super-group of marine
-protists have a large role in pelagic ecosystems. They are unique in
+protists, have a large role in pelagic ecosystems. They are unique in
 that they construct mineral tests out of silica, calcium carbonate, or
@@
 the ocean's cycling of carbon and other elements. However, less is known
-about Rhizarian ecology or their role in the pelagic food-web. Some
-taxa, like certain Radiolarians are mixotrophic, hosting algal
-symbionts. While other taxa are flux-feeders or even predatory
-carnivores. Some prior research has suggested that Rhizaria will
-partition vertically in the water column, likely due to different
-trophic strategies. However, very few studies have investigated their
-populations over extended periods of time. In this study, we present
-data investigating Rhizarian abundance and vertical distribution from
-over a year of monthly cruises into the Sargasso Sea. This study
-represents the first quantification of Rhizaria throughout the
-mesopelagic zone in an oligotrophic system for an extended period of
-time. We use this data to investigate the hypothesis that Rhizaria
-taxonomic groups will partition due to trophic mode. We also investigate
-how their abundance varies in accordance with a variety of environmental
-parameters. Rhizaria abundance was quantified using an Underwater Vision
-Profiler (UVP5), an in-situ imaging device. Ultimately, we show that
-different Rhizaria taxa will have unique vertical distribution patterns.
-Models relating their abundance to environmental parameters have mixed
-results, yet particle concentration is a common predictive variable,
-supporting the importance of heterotrophy amongst many taxa.    
+about Rhizaria ecology or their role in the pelagic food-web. Some taxa,
+like certain Radiolarians, are mixotrophic, hosting algal symbionts.
+While other taxa are flux-feeders or even predatory carnivores. Some
+prior research has suggested that Rhizaria will partition vertically in
+the water column, likely due to different trophic strategies. However,
+very few studies have investigated their populations over extended
+periods of time. In this study, we present data investigating Rhizaria
+abundance and vertical distribution from over a year of monthly cruises
+in the Sargasso Sea. This study represents the first quantification of
+Rhizaria throughout the mesopelagic zone in an oligotrophic system for
+an extended period of time. We use this data to investigate the
+hypothesis that Rhizaria taxonomic groups will partition due to trophic
+mode. We also investigate how their abundance varies in accordance with
+environmental parameters. Rhizaria abundance was quantified using an
+Underwater Vision Profiler (UVP5), an in-situ imaging device.
+Ultimately, we show that different Rhizaria taxa will have unique
+vertical distribution patterns. Models relating their abundance to
+environmental parameters have mixed results, yet particle concentration
+is a common predictive variable, supporting the importance of
+heterotrophy amongst many taxa.    
 \end{abstract}
@@
 \end{minipage}  \end{titlepage}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 numerically dominant and easily sampled with nets and preservatives.
-Fragile organisms like Rhizaria, were difficult to adequately study as
+Fragile organisms like Rhizaria were difficult to adequately study as
 they can be destroyed through standard zooplankton sampling techniques
@@
 al., 2013, 2012; rev by Biard, 2022a). DNA metabarcoding studies have
-revealed insights into the distributional patterns Llopis Monferrer et
-al. (2022), ecological relationships (Decelle et al., 2012; Nakamura et
-al., 2023), and contribution to biogeochemical fluxes (Guidi et al.,
-2016; Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
-approaches also have been used to quantify rhizarian contribution to
+revealed insights into the distributional patterns (Biard et al., 2017;
+Blanco-Bercial et al., 2022; Decelle et al., 2013; Llopis Monferrer et
+al., 2022; Mars Brisbin et al., 2020; Sogawa et al., 2022), ecological
+relationships (Decelle et al., 2012; Nakamura et al., 2023), and
+contribution to biogeochemical fluxes (Guidi et al., 2016;
+Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
+approaches also have been used to quantify Rhizaria contribution to
 community metabolism (Cohen et al., 2023). Yet, despite the excellent
 taxonomic resolution provided by molecular approaches, they do not
-provide a truly quantitative metric for estimating Rhizarian abundance
-or biomass. In-situ imaging tools however, offer the ability to observe
+provide a truly quantitative metric for estimating Rhizaria abundance or
+biomass. In-situ imaging tools however, offer the ability to observe
 organisms in the natural state and quantify their abundance (Barth and
-Stone, in review; Ohman, 2019). While there were early applicaitons of
-imaging tools to document Rhizaria (Dennett et al., 2002), Biard et al.
-(2016)'s report from a global imaging dataset highlighted the importance
-of Rhizaria to the total standing stock of marine carbon. Due to their
-large sizes, ability to concentrate smaller particles and the unique
-structure of their mineral skeletons, Rhizaria have the potential to
-massively influence ocean biogeochemical cycling. A number of studies
-have made large advances in estimating the contribution of Rhizaria to
-ocean cycling of carbon (Gutierrez-Rodriguez et al., 2019; Ikenoue et
-al., 2019; Lampitt et al., 2009; Stukel et al., 2018), silica (Biard et
-al., 2018; Llopis Monferrer et al., 2021), and strontium (Decelle et
-al., 2013). Still, Rhizarian ecological roles are not well understood
-(Biard, 2022a). This is a major challenge as it is critical to
-understand the ecological role of plankton to fully incorporate them
-into biological oceanographic models.
+Stone, In press; Ohman, 2019). Biard et al. (2016) utilized in-situ
+imaging at a global scale to suggest Rhizaria were substantial
+contributors to the ocean carbon standing stock. While more recent
+calculations suggest lower carbon contribution (Laget et al., 2024),
+Rhizaria nonetheless have substantial influences on biogeochemical
+cycling. Due to their large sizes, ability to concentrate smaller
+particles and the unique structure of their mineral skeletons, Rhizaria
+have the potential to massively influence ocean biogeochemical cycling.
+A number of studies have made large advances in estimating the
+contribution of Rhizaria to ocean cycling of carbon (Gutierrez-Rodriguez
+et al., 2019; Ikenoue et al., 2019; Lampitt et al., 2009; Stukel et al.,
+2018), silica (Biard et al., 2018; Llopis Monferrer et al., 2021), and
+strontium (Decelle et al., 2013). Still, Rhizaria ecological roles are
+not well understood (Biard, 2022a). This is a major challenge as it is
+critical to understand the ecological role of plankton to fully
+incorporate them into biological oceanographic models.
 
@@
 modes. As zooplankton, rhizaria are predominately heterotrophic (Biard,
-2022a), yet their feeding modes can be quite varied. Phaeodarians
-(family Cercozoa) are largely thought to be flux-feeders, collecting and
-feeding on sinking particles (Nakamura and Suzuki, 2015; Stukel et al.,
-2019). Alternatively, Retaria can be either exclusively heterotrophic or
+2022a), yet their feeding modes can be quite varied. Phaeodarias (family
+Cercozoa) are largely thought to be flux-feeders, collecting and feeding
+on sinking particles (Nakamura and Suzuki, 2015; Stukel et al., 2019).
+Alternatively, Retaria can be either exclusively heterotrophic or
 mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
@@
 endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
-thought to support early and adult life stages and contribute to total
-primary productivity (Kimoto, 2015). Still, foraminifera are omnivorous,
-possibly even predominately carnivorous with several studies suggesting
-that they can be effective predators (Anderson and Bé, 1976; Gaskell et
-al., 2019), mainly consuming live copepods (Caron and Be, 1984).
-Radiolaria have several lineages all which have some taxa who are well
-known to host symbionts (Biard, 2022b). Amongst radiolarians, arguably
-the most widespread are Collodaria who can be either large solitary
-cells or form massive colonies, up to several meters in length (Swanberg
-and Anderson, 1981). All known Collodaria species host dinoflagellate
-symbionts (Biard, 2022b) and can contribute substantially to primary
-productivity, particularly in oligotrophic ocean regions (Caron et al.,
-1995; Dennett et al., 2002). This Collodaria-symbiont association has
-been suggested as a reason for their high abundances throughout the
-photic zone of oligotrophic environments globally (Biard et al., 2017,
-2016). A few Acantharean (Radiolarian order) clades host algal symbionts
-(Biard, 2022b; Decelle et al., 2012), notably with two clades forming an
-exclusive relationship with Phaeocystis. However, globally Acantharea
-are less abundant than Collodaria (Biard, 2022a) and contribute less to
-total primary productivity (Michaels et al., 1995). This may be due to
-the fact several clades of Acantharea are cyst-forming and strictly
-heterotrophs (Biard, 2022b; Decelle et al., 2013). Furthermore, Mars
-Brisbin et al. (2020) documented apparent predation behavior in
-Acantharea near the surface, suggesting that there may be a large
-reliance on carnivory.
+thought to support early and adult life stages and significantly
+contribute to total primary productivity (Kimoto, 2015). Still,
+foraminifera are omnivorous, possibly even predominately carnivorous,
+with several studies suggesting that they can be effective predators
+(Anderson and Bé, 1976; Gaskell et al., 2019), mainly consuming live
+copepods (Caron and Be, 1984). Radiolaria have several lineages, all of
+which have some taxa well known to host symbionts (Biard, 2022b).
+Amongst Radiolarias, arguably the most widespread are Collodaria who can
+be either large solitary cells or form massive colonies, up to several
+meters in length (Swanberg and Anderson, 1981). All known Collodaria
+species host dinoflagellate symbionts (Biard, 2022b) and can contribute
+substantially to primary productivity, particularly in oligotrophic
+ocean regions (Caron et al., 1995; Dennett et al., 2002). This
+Collodaria-symbiont association has been suggested as a reason for their
+high abundances throughout the photic zone of oligotrophic environments
+globally (Biard et al., 2017, 2016). A few Acantharea (Radiolaria order)
+clades host algal symbionts (Biard, 2022b; Decelle et al., 2012),
+notably with two clades forming an exclusive relationship with
+\emph{Phaeocystis}. However, globally, Acantharea are less abundant than
+Collodaria (Biard, 2022a) and contribute less to total primary
+productivity (Michaels et al., 1995). This may be due to the fact
+several clades of Acantharea are cyst-forming and strictly heterotrophs
+(Biard, 2022b; Decelle et al., 2013). Furthermore, Mars Brisbin et al.
+(2020) documented apparent predation behavior in Acantharea near the
+surface, suggesting that there may be a larger reliance on carnivory.
 
 Given the high abundances, yet diverse trophic strategies found among
-Rhizarian taxa, it is reasonable to expect some form of niche
+Rhizaria taxa, it is reasonable to expect some form of niche
 partitioning. A number of studies do suggest evidence for vertical
 zonation between Rhizaria groups according to various trophic
-strategies. Taxa-specific studies of radiolarians suggest they may be
+strategies. Taxa-specific studies of Radiolarias suggest they may be
 restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
 Although some studies report Acantharea in deeper waters (Decelle et
-al., 2013; Gutiérrez-Rodríguez et al., 2022). Phaeodarians
+al., 2013; Gutiérrez-Rodríguez et al., 2022), Phaeodarias,
 alternatively, are generally found in the mesopelagic where
-photosynthesis cannot occur but they can feed on sinking particles
+photosynthesis cannot occur, but they can feed on sinking particles
 (Stukel et al., 2018). In an imaging-based study of the whole Rhizaria
@@
 throughout the full mesopelagic. Given this lack of information, there
-are many unknowns with respect to Rhizarian ecology, seasonality and
+are many unknowns with respect to Rhizaria ecology, seasonality and
 phenology across different groups.
@@
 In this study, we present a comprehensive assessment of large Rhizaria
-measured for multiple months (greater than 1 year) using an in-situ
-imaging approach. With this dataset, we address two critical aims. 1)
-Quantification of large Rhizaria throughout the epipelagic (0-200m) and
-mesopelagic (200-1000m) over the course of an annual cycle. These data
-were collected in the Sargasso Sea, and represents the first study of
-its kind in an oligotrophic system. 2) We aim to test the hypothesis
-that Rhizaria exhibit niche partitioning according to trophic roles.
-This hypothesis makes several predictions, including vertical zonation,
-as seen in prior studies, but also that environmental variables related
-to trophic strategy will explain abundance patterns. Specifically,
-autotrophic/mixotrophic taxa will correspond to variables related to
-autotrophy (chl-a concentration, primary productivity, high \(O_2\) )
-and other rhizaria will correspond to factors which promote heterotrophy
-(particle concentration, flux, and low \(O_2\)).
+measured for over a year from regularly occurring cruises at monthly
+intervals. We utilized an in-situ imaging approach to facilitate
+abundance calculations. With this dataset, we address two critical aims.
+1) Quantification of large Rhizaria throughout the epipelagic (0-200m)
+and mesopelagic (200-1000m) over the course of an annual cycle. These
+data were collected in the Sargasso Sea, and represents the first study
+of its kind in an oligotrophic system; and 2) We aim to test the
+hypothesis that Rhizaria exhibit niche partitioning according to trophic
+roles. This hypothesis makes several predictions, including vertical
+zonation, as seen in prior studies, but also that environmental
+variables related to trophic strategy will explain abundance patterns.
+Specifically, autotrophic/mixotrophic taxa will correspond to variables
+related to autotrophy (chl-a concentration, primary productivity, local
+DO maxima) and other rhizaria will correspond to factors which promote
+heterotrophy (particle concentration, flux, and local DO minima).
 
@@
 Data were collected in collaboration with the Bermuda Atlantic
-Time-Series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
+Time-series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
 the R/V Atlantic Explorer. Cruises were conducted at approximately
@@
 only the downcast data are utilized. The UVP5 was deployed from
-June-September 2019 then from October 2020 - January 2022. Casts were
-filtered to only include data collected in the BATS region, far offshore
-of Bermuda into the Sargasso Sea (approximately 31.0\(^oN\)-32.5\(^oN\),
-64.25\(^oW\)-63\(^oW\); Supplemental Figure 1). In general, casts
-extended to either 200m, 500m, or 1200m deep, with a few extended into
-the bathypelagic (4500m). However, Rhizaria were only typically found in
-large abundances throughout the epipelagic and mesopelagic zones. As
-such, we limit this study to results from the upper 1000m of the water
-column.
+June-September 2019 then from October 2020 - January 2022, during which
+time the BATS region was sampled for 3-5 days at monthly intervals.
+Casts were filtered to only include data collected in the BATS region,
+far offshore of Bermuda in the Sargasso Sea (approximately
+31.0\(^oN\)-32.5\(^oN\), 64.25\(^oW\)-63\(^oW\); Supplemental Figure 1).
+In general, casts extended to either 200m, 500m, or 1200m deep, with a
+few extended into the bathypelagic (4500m). However, Rhizaria were only
+typically found in large abundances throughout the epipelagic and
+mesopelagic zones. As such, we limit this study to results from the
+upper 1000m of the water column.
 
@@
 184\(\mu m\) - 450\(\mu m\). The lower size range was set by what could
-be reliably sampled by the UVP5's pixel resolution (\textgreater2px) and
-the upper size range is representative of a potential prey field for
-mesozooplankton (Whitmore and Ohman, 2021). Salinity, temperature,
-Dissolved Oxygen (DO), and in-situ chlorophyll fluorescence were
-measured at high-frequencies on each cast, attached to the same CTD
-rosette as the UVP. On certain casts, niskin bottles were used to
-collect bacterial abundance estimates (via epifluorescence microscopy)
-as well as measure inorganic nutrients (\(NO_3\), and \(Si\) as
-silicate/sicilic acid) at discrete depths. On each cruise, flux
-estimates of total mass, carbon, and nitrogen were also collected using
-sediment traps (in the present study we utilize flux to the mesopelagic
-as the flux at 200m). Also primary productivity was estimated through
-measuring \(C^{14}\) uptake rates from in-situ incubations. Full
-descriptions of the BATS sampling program, methods, and data are
-available freely online
-\href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}
-and reviewed in (Lomas et al., 2013).
+be reliably sampled by the UVP5's pixel resolution (\textgreater2px;
+0.092mm per pixel) and the upper size range is representative of a
+potential prey field for mesozooplankton (Whitmore and Ohman, 2021). For
+each UVP cast supporting continuous profiles of the CTD parameters
+salinity, temperature, and auxiliary CTD channels; Dissolved Oxygen
+(DO), in-situ chlorophyll fluorescence were measured at 24Hz using the
+BATS CTD package. On select casts, Niskin bottles were used to collect
+bacterial abundance estimates (via epifluorescence microscopy) as well
+as measure inorganic nutrients (\(NO_3\), and \(Si\) as silicate/sicilic
+acid) at discrete depths. On each cruise, flux estimates of total mass,
+organic carbon, and nitrogen were also collected using sediment traps;
+in the present study we utilized flux to the mesopelagic as the flux at
+200m. Also primary productivity was estimated through measuring
+\(^{14}C\) uptake rates from in-situ incubations. For full descriptions
+of the BATS sampling program and methods, see Knap et al. (1997) and
+Lomas et al. (2013) for a review. Additionally, data can be viewed
+online
+\href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}.
 
 Environmental data were processed in a variety of ways to match the
-format of the Rhizaria abundance estimates (See below). CTD data were
-collected at higher frequency than the UVP, so these data were averaged
-within matching bins to the UVP5 data. Data from niskin bottles were
-first linearly interpolated with a 1m resolution for each cast where
-data were available. Then, because data were collected at different
-frequencies over the course of a cruise, the interpolated data were
-averaged across casts, then averaged into matching UVP5-sized bins.
-Primary productivity estimates were totaled within the euphotic zone to
-represent a ``total euphotic productivity''. Finally the data from the
-sediment trap deployed to 200m was used to represent ``flux to the
-mesopelagic''.
+format of the Rhizaria abundance estimates (see below). CTD data were
+collected at higher frequency than the UVP (24Hz vs 15Hz respectively),
+so these data were averaged within matching bins to the UVP5 data. Data
+from Niskin bottles were first linearly interpolated in depth at 1m
+resolution then time averaged over the cruise, then subsequently
+averaged into matching UVP5-sized bins. Primary productivity estimates
+were totaled within the euphotic zone to represent a ``total euphotic
+productivity''.
 
@@
 2015; Suzuki and Not, 2015; and Biard and Ohman, 2020). Using the
-aforementioned sources and publicly available ecotaxa projects, I
-constructed a guide which can be found here:
-\protect\hyperlink{0}{https://thealexbarth.github.io/media//Project\_Items/Oligotrophic\_Community/ecotaxa\_UVP-guide-stone-lab.pdf}.
-Broadly, Rhizaria were classified as Foraminifera, Radiolarians
-(Acantharea or Collodaria), or a variety of Phaeodarian families (Figure
-1). When identification could not be confidently made between a few
-candidate taxa, a less specific label was used. As a result, we have
+aforementioned sources and publicly available ecotaxa projects, we
+constructed a guide accessible at:
+\url{https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf}.
+Broadly, Rhizaria were classified as Foraminifera, Radiolarias
+(Acantharea or Collodaria), or as a variety of Phaeodaria families
+(Figure 1). When identification could not be confidently made between a
+few candidate taxa, a less specific label was used. As a result, we have
 data from ``unidentified Rhizaria'', which typically were vignettes not
 distinguishable between Aulacanthidae or Acantharea or ``unidentified
-Phaeodaria'', which are clearly phaeodaria but not distinguishable into
+Phaeodaria'', which are clearly Phaeodaria but not distinguishable into
 a family.
+
+\begin{figure}
+
+{\centering \includegraphics{images/01_taxa.pdf}
+
+}
+
+\caption{Example images of different Rhizaria taxa. 4mm scale bar shown
+in lower right. All vignettes are same scale}
+
+\end{figure}
 
@@
 confirm that the UVP5 was sampling adequately across all size ranges, an
-NBSS slope was constructed to identify a drop-off which would indicate
-poor-sampling at the small size range (Barth and Stone, in review;
-Lombard et al., 2019). However, it was evident from this analysis that
-all size ranges were adequately sampled across the size range
-(Supplemental Figure 2) so no data were excluded. The UVP5 reports the
-exact depth at which a particle is recorded, however to estimate
-abundance, observations must be binned over fixed depth intervals. Our
-deployments had variable descent depths and speeds with more casts
-descending to 500m than 1000m and descents quicker through the
-epipelagic than the mesopelagic (see Barth and Stone (2022) for an
-extended discussion of UVP5 data processing). For the present study,
-Rhizarian abundances were estimated in 25m vertical bins, which offer a
-moderate sampling volume per bin (average 0.948\(m^3\) in the epipelagic
-and 0.589\(m^3\) in the mesopelagic) while still maintaining
-ecologically relevant widths. However, concentrations in a 25m bin would
-need to be greater than 2.428 indv. \(m^{-3}\) and 3.912 indv.
-\(m^{-3}\), in the epipelagic and mesopelagic respectively, to fall
-below a 10\% non-detection risk {[}Benfield et al. (1996); Barth and
-Stone (in review){]}. Because we typically observed many rhizaria taxa
-below these concentrations, we present the 25m binned data to visualize
-broad-scale average distributions. For quantifying and modelling
-Rhizaria abundances, we present integrated abundance estimates, with
-each cast. Due to the variable descent depths of the UVP, data are
-categorized as epipelagic (0-200m), upper mesopelagic (200-500m), and
-lower mesopelagic (500-1000m). The average sampling volume integrated
-through these regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\),
-with non-detection thresholds at 0.30 indv. \(m^{-2}\), 0.33 indv.
-\(m^{-2}\), and 0.20 indv. \(m^{-2}\) respectively. All UVP data
-processing was done using the \texttt{EcotaxaTools} package in R (Barth
-2023).
-
-\hypertarget{modelling-environmental-controls-of-rhizarian-abundance}{%
-\subsection{Modelling environmental controls of Rhizarian
-Abundance}\label{modelling-environmental-controls-of-rhizarian-abundance}}
+normalized biomass size spectrum (NBSS) slope was constructed to
+identify a drop-off which would indicate poor-sampling at the small size
+range (Barth and Stone, In press; Lombard et al., 2019). However, it was
+evident from this analysis that all size ranges were adequately sampled
+across the size range (Supplemental Figure 2) so no data were excluded.
+The UVP5 reports the exact depth at which a particle is recorded,
+however to estimate abundance, observations must be binned over fixed
+depth intervals. Our deployments had variable descent depths and speeds
+with more casts descending to 500m than 1000m and descents quicker
+through the epipelagic than the mesopelagic (see Barth and Stone (2022)
+for an extended discussion of UVP5 data processing). For the present
+study, Rhizaria abundances were estimated in 25m vertical bins, which
+offer a moderate sampling volume per bin (average 0.948\(m^3\) in the
+epipelagic and 0.589\(m^3\) in the mesopelagic; Supplemental Table 1)
+while still maintaining ecologically relevant widths. However,
+concentrations in a 25m bin would need to be greater than 2.428 ind.
+\(m^{-3}\) and 3.912 ind. \(m^{-3}\), in the epipelagic and mesopelagic
+respectively, to fall below a 10\% non-detection risk (Barth and Stone,
+In press; Benfield et al., 1996). Because we typically observed many
+rhizaria taxa below these concentrations, we present the 25m binned data
+to visualize broad-scale average distributions. For quantifying and
+modelling Rhizaria abundances, we present integrated abundance
+estimates, with each cast. Due to the variable descent depths of the
+UVP, data are categorized as epipelagic (0-200m), upper mesopelagic
+(200-500m), and lower mesopelagic (500-1000m). The average sampling
+volume integrated through these regions were 7.59\(m^2\), 7.06\(m^2\),
+and 11.77\(m^2\), with non-detection thresholds at 0.30 ind. \(m^{-2}\),
+0.33 ind. \(m^{-2}\), and 0.20 ind. \(m^{-2}\) respectively. All UVP
+data processing was done using the \texttt{EcotaxaTools} package in R
+(Barth 2023).
+
+\hypertarget{modelling-environmental-controls-of-rhizaria-abundance}{%
+\subsection{Modelling environmental controls of Rhizaria
+Abundance}\label{modelling-environmental-controls-of-rhizaria-abundance}}
 
 Generalized Additive Models (GAMs) were used to assess the relationship
-between integrated Rhizarian abundance and different environmental
+between integrated Rhizaria abundance and different environmental
 factors. GAMs offer the ability to model non-linear and non-monotonic
@@
 relationships (Wood, 2017) and have been successfully applied to
-Rhizarian ecology (Biard and Ohman, 2020). The \texttt{mgcv} package
+Rhizaria ecology (Biard and Ohman, 2020). The \texttt{mgcv} package
 (Wood, 2001) was used to construct models relating environmental
@@
 
-The BATS sampling region is southeast of Bermuda, situated in the North
-Sargasso Sea and the North Atlantic oligotrophic ocean gyre. Due to the
-sampling location, while the environmental conditions are generally low
-in variation and oligotrophic, there is some seasonality and
-considerable influence from mesoscale eddies (Lomas et al., 2013;
-McGillicuddy et al., 1998). Variability in the water column structure
-was visible during the study period (Figure 2). This is best evidenced
-through the temperature profiles; In the late summer and early fall
-there was a stratified water column with high temperatures in the
-surface (\textless75m) (Figure 2A) and slightly elevated salinity
-(Figure 2B). This warm, stratified period appeared more intense during
-the few months sampled in 2019. In 2021, we observed the stratified
-layer slowly dissipated into the winter months. There was a consistent
-oxygen minimum zone (OMZ) located at about 750m deep (Figure 2C).
-Although, February 2021 saw a notable downwelling event, likely due to a
-passing cyclonic eddy. During this time, warmer, oxic water was plunged
-deeper into the mesopelagic. This process was reversed in the spring
-months (March, April) when mixing occurred throughout the epipelagic and
-the surface was cooler and well mixed. Primary production was highest
-during the spring mixing period, evidenced both by in-situ fluorescence
-(Figure 2D) and productivity incubation experiments (Figure 3A).
-Originating near the surface, the productivity peak moved deeper
-throughout the spring and declined into the summer (Figure 2D). However,
-there was a notable, yet smaller productivity bump in the late summer
-and early fall (Figure 3A) which occurred deeper in the epipelagic
-(Figure 2D). The particle concentration (184\(\mu m\) - 450\(\mu m\))
-was closely coupled to chlorophyll-a patterns.
+The BATS sampling region is southeast of Bermuda, situated in the
+oligotrophic North Atlantic Subtropical Gyre. Due to the sampling
+location, while the environmental conditions are generally low in
+variation and oligotrophic, there is considerable influence from deep
+winter-mixing and summer stratification as well as secondary influences
+from mesoscale eddies which result in spatiotemporal heterogeniety
+(Lomas et al., 2013; McGillicuddy et al., 1998). Variability in the
+water column structure was visible during the study period (Figure 2).
+This is best evidenced through the temperature profiles; In the late
+summer and early fall there was a stratified water column with high
+temperatures in the surface (\textless75m) (Figure 2A) and slightly
+elevated salinity (Figure 2B). This warm, stratified period appeared
+more intense during the few months sampled in 2019. In 2021, we observed
+the stratified layer slowly dissipated into the winter months following
+mixed layer entrainment. There was a consistent oxygen minimum zone
+(OMZ) located at about 800m deep (Figure 2C). February 2021 saw a
+notable downwelling event, likely due to a passing anti-cyclonic eddy
+which impacted the local region during early 2021. During this phaes,
+warmer, oxic water was significantly depressed deeper into the
+mesopelagic. This process was reversed in the spring months (March,
+April) primary due to the interaction of convective mixing and the
+passing of a strong cyclonic eddy resulting in a deep cold mixed layer.
+Primary production was highest during the spring mixing period,
+evidenced both by in-situ fluorescence (Figure 2D) and productivity
+incubation experiments (Figure 3A). Originating near the surface, the
+productivity peak moved deeper throughout the spring and declined into
+the summer (Figure 2D). However, there was a notable, yet smaller
+productivity bump in the late summer and early fall (Figure 3A) which
+occurred deeper in the epipelagic (Figure 2D). The particle
+concentration (184\(\mu m\) - 450\(\mu m\)) was closely coupled to
+chlorophyll-a patterns.
+
+\begin{figure}
+
+{\centering \includegraphics{images/02-environmental.pdf}
+
+}
+
+\caption{Environmental profiles across time-series of study period. Y
+axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved
+Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion
+(184 - 450\(\mu m\)). (F) Bacteria Abundance. (G) Silica. (H) Nitrate}
+
+\end{figure}
 
@@
 attenuating throughout the lower epipelagic (Figure 2E). Similarly,
-bacteria abundance was closely linked to overall productivity, althrough
-there was a more consistent moderate-abundance layer near the top of the
-mesopelagic (\textasciitilde250m) (Figure 2F). Concurrent with the
-secondary fall production peak, there was also higher particle
-concentration and bacterial abundance in the later summer and early
-fall. Interestingly, while primary productivity estimates from
-July-August were not that different between 2019 and 2021 (Figure 4A),
-chlorophyll-a florescence, particle concentration, and bacterial
-abundance were much higher in 2019's summer/fall (Figure 2D-F).
-Inorganic nutrients (\(Si\) and \(NO_3\)) were generally well
+heterotrophic bacteria abundance was closely linked to overall
+productivity, althrough there was a more consistent moderate-abundance
+layer near the top of the mesopelagic (\textasciitilde250m) (Figure 2F).
+Concurrent with the secondary fall production peak, there was also
+higher particle concentration and bacterial abundance in the later
+summer and early fall. Interestingly, while primary productivity
+estimates from July-August were not that different between 2019 and 2021
+(Figure 3A), chlorophyll-a florescence, particle concentration, and
+bacterial abundance were much higher in 2019's summer/fall (Figure
+2D-F). Inorganic nutrients (\(Si\) and \(NO_3\)) were generally well
 stratified, with low concentrations in the epipelagic and increasing
@@
 2G).
+
+\begin{figure}
+
+{\centering \includegraphics{images/03_prod-flux.pdf}
+
+}
+
+\caption{Primary productivity (A) and flux estimates from total mass
+(B), carbon (C), and nitrogen (D). Values are from monthly cruises with
+month displayed in corresponding colors. Absent data are shown by grey
+bar.}
+
+\end{figure}
 
@@
 epipelagic, where they accounted for 47\% of all mesozooplankton. In the
-mesopelagic rhizaria were a smaller fraction, at 38\% in the upper
-layers (200-500m) and 37\% in the deeper mesopelagic (500-1000m).
+mesopelagic rhizaria were a smaller (but still prominent fraction), at
+38\% in the upper layers (200-500m) and 37\% in the deeper mesopelagic
+(500-1000m).
+
+\begin{figure}
+
+{\centering \includegraphics{images/04_vertical.pdf}
+
+}
+
+\caption{Average abundance of Rhizaria in 25m bins, across entire study
+period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C),
+Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae
+(G), Coelodendridae (H).}
+
+\end{figure}
 
@@
 methods. The vertical distribution pattern and abundance varied
-considerably across taxonomic groups. Radiolarians were some of the most
+considerably across taxonomic groups. Radiolarias were some of the most
 abundant taxa observed, particularly in the epipelagic (Figure 4, Figure
@@
 average densities were much lower and spread wider throughout the
-mesopelagic (Figure 4E). Phaeodarian families exhibited a wide range of
+mesopelagic (Figure 4E). Phaeodaria families exhibited a wide range of
 vertical distribution patterns. The most abundant, Aulacanthidae, also
@@
 slightly lower in the epipelagic (Figure 4F). Castanellidae were the
-only Phaeodarian who appeared to be effectively restricted to the photic
+only Phaeodaria who appeared to be effectively restricted to the photic
 zone (Figure 4G). Alternatively, Coelodendridae primarily occurred in
@@
 they were much rarer (data not shown).
+
+\begin{figure}
+
+{\centering \includegraphics{images/05_seasonality.pdf}
+
+}
+
+\caption{Seasonality of Rhizaria integrated abundance for the epipelagic
+(0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower mesopelagic
+(500-1000m) (E-F). Left panels (A,C,E) display total integrated
+abundance per monthly cruise colored by month. Right panels (B, D, F)
+display community composition of each total abundance.}
+
+\end{figure}
 
@@
 months in 2021. While the majority of integrated abundance in the
-epipelagic was consistently attributable to Collodaria, Acanthrean
+epipelagic was consistently attributable to Collodaria, Acanthrea
 abundance occurred sporadically and could account for a large portion of
@@
 community composition in the mesopelagic was more diverse, mostly
-comprised of Phaeodarians. However, Acantharea and unidentified Rhizaria
+comprised of Phaeodarias. However, Acantharea and unidentified Rhizaria
 also were common members of the community (Figure 5D, 5F).
@@
 
-\hypertarget{environmental-drivers-of-rhizarian-abundance}{%
-\subsubsection{Environmental Drivers of Rhizarian
-Abundance}\label{environmental-drivers-of-rhizarian-abundance}}
-
-For total Rhizarian integrated abundance the GAMs produced moderate fits
+\begin{figure}
+
+{\centering \includegraphics{images/09_size-comp.pdf}
+
+}
+
+\caption{Comparison of average sizes (ESD) amongst Rhizaria taxa which
+occurred throughout the water column.}
+
+\end{figure}
+
+\hypertarget{environmental-drivers-of-rhizaria-abundance}{%
+\subsubsection{Environmental Drivers of Rhizaria
+Abundance}\label{environmental-drivers-of-rhizaria-abundance}}
+
+For total Rhizaria integrated abundance the GAMs produced moderate fits
 (\(R^2_{adj}\) = 0.406-0.603) (Table 1). In the epipelagic, there were
@@
 
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4300}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3400}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}@{}}
+\caption{Generalized Additive Model results for integrated total
+Rhizaria abundance in different regions of the water
+column.}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Model}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Model}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p
+\end{minipage} \\
+\midrule()
+\endhead
+\textbf{All Rhizaria Epipelagic} & Salinity & 2.0818 & 4.654 &
+\textless0.001 \\
+\(R^2_{adj}\)= 0.42 & DO & 0.8272 & 0.945 & 0.0136 \\
+& Silica & 3.1206 & 15.75 & \textless0.001 \\
+& NO3 & 2.9970 & 6.268 & \textless0.001 \\
+& Primary Productivity & 1.8367 & 2.099 & \textless0.001 \\
+& Particle Concentration & 0.9282 & 2.566 & \textless0.001 \\
+\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6847 & 0.431 &
+0.073 \\
+\(R^2_{adj}\)= 0.406 & Avg Mass Flux & 1.5426 & 0.923 & 0.045 \\
+& Particle Concentration & 3.2957 & 20.89 & \textless0.001 \\
+\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 1.6632 & 1.824
+& 0.002 \\
+\(R^2_{adj}\)= 0.603 & Particle Concentration & 0.7694 & 0.662 &
+0.027 \\
+\bottomrule()
+\end{longtable}
+
 GAMs for individual taxa were much less consistent in their fits (Table
@@
 for certain taxa. Note that due to low abundances, GAMs were not
-constructed for Tuscaroridae or Medusettidae. Furthermore no significant
-terms were found for a model with Aulosphaeridae in the epipelagic nor
-Foraminifera in the mesopelagic.
+constructed for Tuscaroridae or Medusettidae. Furthermore no,
+significant terms were found for a model with Aulosphaeridae in the
+epipelagic nor Foraminifera in the mesopelagic.
+
+\begin{figure}
+
+{\centering \includegraphics{images/06_epi-partials.pdf}
+
+}
+
+\caption{Partial effects of smooth terms in taxa-specific GAM models
+from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea
+(A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae
+(E).}
+
+\end{figure}
 
@@
 association (Figure 7A), with July 2021 as a clear outlier where
-Acantharean abundances were high in the epipelagic despite lower fluxes
+Acantharea abundances were high in the epipelagic despite lower fluxes
 and particle concentrations (Figure 5). Foraminifera had a good fitting
@@
 
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4300}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3400}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0800}}@{}}
+\caption{Taxa-specific generalized additive models for different regions
+of the water column.}\tabularnewline
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Model
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p
+\end{minipage} \\
+\midrule()
+\endfirsthead
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedright
+Model
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Term
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+edf
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+F
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+p
+\end{minipage} \\
+\midrule()
+\endhead
+\textbf{Acantharea Epipelagic} & Salinity & 2.264 & 4.113 &
+\textless0.001 \\
+R2adj=0.53 & O2 & 2.579 & 6.712 & \textless0.001 \\
+& Avg Mass Flux & 4.810 & 22.81 & \textless0.001 \\
+& Bacteria \#/L & 1.599 & 1.317 & 0.0117 \\
+& Particle Concentration & 0.853 & 1.158 & 0.0087 \\
+\textbf{Acantharea Upper Mesopelagic} & Avg C Flux & 1.296 & 0.621 &
+0.0439 \\
+R2adj=0.231 & Avg N Flux & 1.619 & 1.283 & 0.0026 \\
+& Particle Concentration & 0.952 & 3.886 & \textless0.001 \\
+\textbf{Acantharea Lower Mesopelagic} & Temperature & 2.076 & 4.155 &
+\textless0.001 \\
+R2adj=0.509 & Avg N Flux & 1.494 & 1.216 & 0.0113 \\
+& Primary Productivity & 0.766 & 0.648 & 0.0253 \\
+& Particle Concentration & 2.037 & 10.03 & \textless0.001 \\
+\textbf{Aulacanthidae Epipelagic} & Salinity & 0.792 & 0.757 & 0.0241 \\
+R2adj=0.251 & Avg N Flux & 0.869 & 1.324 & 0.0018 \\
+& Primary Productivity & 2.312 & 7.704 & \textless0.001 \\
+& Particle Concentration & 2.008 & 2.200 & 0.0017 \\
+\textbf{Aulacanthidae Upper Mesopelagic} & Particle Concentration &
+2.832 & 9.802 & \textless0.001 \\
+R2adj = 0.158 & & & & \\
+\textbf{Aulacanthidae Lower Mesopelagic} & Avg Mass Flux & 1.622 & 1.991
+& 0.002 \\
+R2adj=0.298 & Particle Concentration & 2.123 & 6.706 & \textless0.001 \\
+\textbf{Aulosphaeridae Upper Mesopelagic} & Particle Concentration &
+2.653 & 13.06 & \textless0.001 \\
+R2adj=0.2 & & & & \\
+\textbf{Aulosphaeridae Lower Mesopelagic} & Particle Concentration &
+0.972 & 6.248 & \textless0.001 \\
+R2adj=.147 & & & & \\
+\textbf{Castanellidae Epipelagic} & Avg C Flux & 1.462 & 0.816 &
+0.0421 \\
+R2adj = 0.124 & Primary Productivity & 0.771 & 0.665 & 0.0247 \\
+& Particle Concentration & 3.956 & 4.623 & \textless0.001 \\
+\textbf{Coelodendridae Upper Mesopelagic} & Avg N Flux & 0.822 & 0.919 &
+0.0183 \\
+R2adj=.113 & Particle Concentration & 0.970 & 6.208 & \textless0.001 \\
+\textbf{Coelodendridae Lower Mesopelagic} & Particle Concentration &
+1.773 & 4.873 & \textless0.001 \\
+R2adj=.133 & & & & \\
+\textbf{Collodaria Epipelagic} & Salinity & 0.925 & 2.267 &
+\textless0.001 \\
+R2adj=0.16 & O2 & 2.015 & 2.217 & 0.002 \\
+& Bacteria \#/L & 1.843 & 2.100 & 0.002 \\
+\textbf{Foraminifera Epipelagic} & Temperature & 4.414 & 8.789 &
+\textless0.001 \\
+R2adj=0.445 & O2 & 1.603 & 1.287 & 0.0111 \\
+& Avg N Flux & 2.512 & 3.396 & \textless0.001 \\
+& Primary Productivity & 0.824 & 0.920 & 0.0153 \\
+& Particle Concentration & 0.919 & 2.257 & \textless0.001 \\
+\bottomrule()
+\end{longtable}
+
+\begin{figure}
+
+{\centering \includegraphics{images/07_upmeso-partial.pdf}
+
+}
+
+\caption{Partial effects of smooth terms in taxa-specific GAM models
+from the upper mesopelagic (200-500m). Effects are grouped by taxa;
+Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
+(D).}
+
+\end{figure}
+
 In the upper mesopelagic (200-500m), abundances were generally low
@@
 concentration and nitrogen flux, as well as temperature to a slight
-degree (Figure 9A). For all Phaeodarians with a significant model,
+degree (Figure 9A). For all Phaeodarias with a significant model,
 particle concentration was a main predictor variable (Table 2, Figure
-9B-D). Aulacanthidae had the best fitting model of the Phaeodarians
+9B-D). Aulacanthidae had the best fitting model of the Phaeodarias
 (\(R^2_{adj}\) = 0.298), which also included mass flux as a
@@
 
+\begin{figure}
+
+{\centering \includegraphics{images/08_lomeso-partials.pdf}
+
+}
+
+\caption{Partial effects of smooth terms in taxa-specific GAM models
+from the lower mesopelagic (500-1000m). Effects are grouped by taxa;
+Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
+(D).}
+
+\end{figure}
+
 \hypertarget{discussion}{%
@@
 
-\hypertarget{overall-rhizarian-abundance-and-patterns}{%
-\subsubsection{Overall Rhizarian abundance and
-patterns}\label{overall-rhizarian-abundance-and-patterns}}
+\hypertarget{overall-rhizaria-abundance-and-patterns}{%
+\subsubsection{Overall Rhizaria abundance and
+patterns}\label{overall-rhizaria-abundance-and-patterns}}
 
 In the epipelagic Rhizaria exhibited a notable seasonal pattern.
-Rhizarian abundances were higher in the summer months and lower during
+Rhizaria abundances were higher in the summer months and lower during
 the winter. During a prior time period, Blanco-Bercial et al. (2022)
 noted that there is considerable seasonality in the community
-composition of all protists. Despite the seasonality of total Rhizarian
+composition of all protists. Despite the seasonality of total Rhizaria
 abundance, community composition was relatively consistent, with
@@
 in this study which were not captured. Throughout the mesopelagic,
-month-to-month variation in 2021 was relatively low. Again this is
+month-to-month variation in 2021 was relatively low. Again, this is
 consistent with observations from metabarcoding of the whole protist
@@
 
-Overall Rhizaria were the most commonly identified group of
-mesozooplankton throughout the study period. We do note that the UVP5
-commonly captures Trichodesmium colonies, yet these were excluded in
-this comparison as they are strictly autotrophs. It should be noted that
+Overall Rhizaria were the most commonly identified group of mesoplankton
+throughout the study period. We do note that the UVP5 commonly captures
+\emph{Trichodesmium} colonies, yet these were excluded in this
+comparison as they are strictly autotrophs. It should be noted that
 previous work has suggested that avoidance behavior with the UVP is
@@
 variability. In the present study, we noticed considerably higher
-Rhizarian abundance throughout the water column in 2019 compared to
-2021. While this may have been driven by increased mass flux, more
-information is needed to truly understand the magnitude by which
-Rhizaria can vary interannually.
+Rhizaria abundance throughout the water column in 2019 compared to 2021.
+While this may have been driven by increased mass flux, more information
+is needed to truly understand the magnitude by which Rhizaria can vary
+interannually.
 
@@
 may be a consequence of isothermal submersion (Boltovskoy, 2017).
-Another effectively exclusively epipelagic Rhizaria was the Phaeodarian
-family of Castanellidae. All Phaeodarians are thought to be fully
-heterotrophic (Nakamura and Suzuki, 2015), nonetheless a number of
-studies, including this one, report Castanellidae to be typically found
-in the lower epipelagic (Biard et al., 2018; Biard and Ohman, 2020;
-Zasko and Rusanov, 2005). It should be considered that perhaps
-Castanellidae specializes in feeding on sinking particles directly at
-the base of the epipelagic. Given it's smaller size (Nakamura and
-Suzuki, 2015), Castanellidae does not need a large diameter to
-efficiently flux feed at the typically particle rich region of the lower
-epipelagic. Both Castanelldiae and Collodaria had poor fitting GAMs.
-This is somewhat of a surprise for Collodaria who had large abundance.
-However, given the consistency of their abundance, it may be that this
-study did not capture a wide enough range of conditions for describing
-Collodaria's preferred niche.
+Alternatively, surface waters at BATS often mix into the mode water
+during the seasonal mixing, so Collodaria in the deeper waters may be a
+result of diapyncal mixing. Another effectively exclusively epipelagic
+Rhizaria was the Phaeodaria family of Castanellidae. All Phaeodaria are
+thought to be fully heterotrophic (Nakamura and Suzuki, 2015),
+nonetheless a number of studies, including this one, report
+Castanellidae to be typically found in the lower epipelagic (Biard et
+al., 2018; Biard and Ohman, 2020; Zasko and Rusanov, 2005). It should be
+considered that perhaps Castanellidae specializes in feeding on sinking
+particles directly at the base of the epipelagic. Given it's smaller
+size (Nakamura and Suzuki, 2015), Castanellidae does not need a large
+diameter to efficiently flux feed at the typically particle rich region
+of the lower epipelagic. Both Castanelldiae and Collodaria had poor
+fitting GAMs. This is somewhat of a surprise for Collodaria who had
+large abundance. However, given the consistency of their abundance, it
+may be that this study did not capture a wide enough range of conditions
+for describing Collodaria's preferred niche.
 
@@
 more advantageous for Rhizaria to situate deeper, in darker regions of
-the twilight zone. Also it should be noted that Phaeodarians utilize
+the twilight zone. Also it should be noted that Phaeodaria utilize
 silica to build their opaline tests, and silica concentrations started
@@
 the water column Zasko and Rusanov (2005). This would support the
-paradigm that large Acantharean abundances may be supported by their
+paradigm that large Acantharea abundances may be supported by their
 mixotrophic abilities (Michaels et al., 1995; Suzuki and Not, 2015).
 While the UVP5 images cannot distinguish between mixotrophic and
-heterotrophic Acantharea, the GAMs constructed for Acantharean abundance
+heterotrophic Acantharea, the GAMs constructed for Acantharea abundance
 found positive associations with particle concentration and mass flux,
@@
 rapidly return to the photic zone (Decelle et al., 2013, 2012). This
-hypothesis predicts that Acanthareans in the mesopelagic would be
-smaller (Decelle et al., 2013). Mars Brisbin et al. (2020) provided some
-support for this hypothesis, with a significant decrease in Acantharea
-sizes with depth. Although the authors also observed low abundances in
-the mesopelagic and noted that the smaller sizes may be due to lower
-food availability (Mars Brisbin et al., 2020). Since food is more scarce
-in the mesopelagi, nutritional quality lower, yet flux feeders would
-likely grow larger to increase their feeding range (Biard and Ohman,
-2020). In the data collecting in this study, Acanthareans in the
-mesopelagic were significantly smaller than the epipelagic, despite the
-other bimodal taxa (Foraminifera and Aulacanthidae) being significantly
-larger with depth. This provides added support for the hypothesis that
-cyst-forming Acantharea may utilize different sections of the water
-column throughout their life cycle. However to further investigate this,
-more work is needed with higher temporal and taxonomic resolution.
+hypothesis predicts that Acanthareas in the mesopelagic would be smaller
+(Decelle et al., 2013). Mars Brisbin et al. (2020) provided some support
+for this hypothesis, with a significant decrease in Acantharea sizes
+with depth. Although the authors also observed low abundances in the
+mesopelagic and noted that the smaller sizes may be due to lower food
+availability (Mars Brisbin et al., 2020). Since food is more scarce in
+the mesopelagi, nutritional quality lower, yet flux feeders would likely
+grow larger to increase their feeding range (Biard and Ohman, 2020). In
+the data collecting in this study, Acanthareas in the mesopelagic were
+significantly smaller than the epipelagic, despite the other bimodal
+taxa (Foraminifera and Aulacanthidae) being significantly larger with
+depth. This provides added support for the hypothesis that cyst-forming
+Acantharea may utilize different sections of the water column throughout
+their life cycle. However to further investigate this, more work is
+needed with higher temporal and taxonomic resolution.
 
@@
 
-This study provides a detailed look at Rhizarian abundance over time
+This study provides a detailed look at Rhizaria abundance over time
 throughout the water column in a major oligotrophic gyre. We show that
@@
 aggregation mechanism. However, this is largely speculation, to truly
-test this, more work is needed measuring Rhizarian flux.
+test this, more work is needed measuring Rhizaria flux.
 
@@
 understood. While this study represents a step-forward in our
-understanding of Rhizarian ecology, continued research on Rhizaria is
+understanding of Rhizaria ecology, continued research on Rhizaria is
 much needed to better understand their ecology. Particularly extended
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-barth-inrev}{}}%
-Barth, A., Stone, J., in review. Understanding the picture: The promis
+\leavevmode\vadjust pre{\hypertarget{ref-barth-inpress}{}}%
+Barth, A., Stone, J., In press. Understanding the picture: The promis
 and challenges of in-situ imagery data in the study of plankton ecology.
+Journal of Plankton Research: Horizons.
 
@@
 \url{https://doi.org/10.1007/978-4-431-55130-0_7}
+
+\leavevmode\vadjust pre{\hypertarget{ref-knap1997}{}}%
+Knap, A.H., Michaels, A.F., Steinberg, D.K., Bahr, F., Bates, N.R.,
+Bell, S., Countway, P., Close, A.R., Doyle, A.P., Dow, R.L., Howse,
+F.A., Gundersen, K., Johnson, R.J., Kelly, R., Little, R., Orcutt, K.,
+Parsons, R., Rathburn, C., Sanderson, M., Stone, S., 1997.
+\href{https://eprints.soton.ac.uk/361194/}{BATS methods manual, version
+4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-laget2024}{}}%
+Laget, M., Drago, L., Panaïotis, T., Kiko, R., Stemmann, L., Rogge, A.,
+Llopis-Monferrer, N., Leynaert, A., Irisson, J.-O., Biard, T., 2024.
+Global census of the significance of giant mesopelagic protists to the
+marine carbon and silicon cycles. Nature Communications 15, 3341.
+\url{https://doi.org/10.1038/s41467-024-47651-4}
 
@@
 
-\newpage
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}@{}}
-\caption{Generalized Additive Model results for integrated total
-Rhizarian abundance in different regions of the water
-column.}\tabularnewline
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Model}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Term
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-edf
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-p-value
-\end{minipage} \\
-\midrule()
-\endfirsthead
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Model}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Term
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-edf
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-p-value
-\end{minipage} \\
-\midrule()
-\endhead
-\textbf{All Rhizaria Epipelagic} & Salinity & 2.0818 & 4.654 &
-\textless0.001 \\
-\(R^2_{adj}\)= 0.42 & DO & 0.8272 & 0.945 & 0.0136 \\
-& Silica & 3.1206 & 15.75 & \textless0.001 \\
-& NO3 & 2.9970 & 6.268 & \textless0.001 \\
-& Primary Productivity & 1.8367 & 2.099 & \textless0.001 \\
-& Particle Concentration & 0.9282 & 2.566 & \textless0.001 \\
-& & & & \\
-\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6847 & 0.431 &
-0.073 \\
-\(R^2_{adj}\)= 0.406 & Avg Mass Flux & 1.5426 & 0.923 & 0.045 \\
-& Particle Concentration & 3.2957 & 20.89 & \textless0.001 \\
-& & & & \\
-\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 1.6632 & 1.824
-& 0.002 \\
-\(R^2_{adj}\)= 0.603 & Particle Concentration & 0.7694 & 0.662 &
-0.027 \\
-\bottomrule()
-\end{longtable}
-
-\newpage
-
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2533}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1867}}@{}}
-\caption{Taxa-specific generalized additive models for different regions
-of the water column.}\tabularnewline
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Model
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Term
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-edf
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-p-value
-\end{minipage} \\
-\midrule()
-\endfirsthead
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Model
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Term
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-edf
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-F
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-p-value
-\end{minipage} \\
-\midrule()
-\endhead
-\textbf{Acantharea Epipelagic} & Salinity & 2.264 & 4.113 &
-\textless0.001 \\
-R2adj=0.53 & O2 & 2.579 & 6.712 & \textless0.001 \\
-& Avg Mass Flux & 4.810 & 22.81 & \textless0.001 \\
-& Bacteria \#/L & 1.599 & 1.317 & 0.0117 \\
-& Particle Concentration & 0.853 & 1.158 & 0.0087 \\
-& & & & \\
-\textbf{Acantharea Upper Mesopelagic} & Avg C Flux & 1.296 & 0.621 &
-0.0439 \\
-R2adj=0.231 & Avg N Flux & 1.619 & 1.283 & 0.0026 \\
-& Particle Concentration & 0.952 & 3.886 & \textless0.001 \\
-& & & & \\
-\textbf{Acantharea Lower Mesopelagic} & Temperature & 2.076 & 4.155 &
-\textless0.001 \\
-R2adj=0.509 & Avg N Flux & 1.494 & 1.216 & 0.0113 \\
-& Primary Productivity & 0.766 & 0.648 & 0.0253 \\
-& Particle Concentration & 2.037 & 10.03 & \textless0.001 \\
-\textbf{Aulacanthidae Epipelagic} & Salinity & 0.792 & 0.757 & 0.0241 \\
-R2adj=0.251 & Avg N Flux & 0.869 & 1.324 & 0.0018 \\
-& Primary Productivity & 2.312 & 7.704 & \textless0.001 \\
-& Particle Concentration & 2.008 & 2.200 & 0.0017 \\
-& & & & \\
-\textbf{Aulacanthidae Upper Mesopelagic} & Particle Concentration &
-2.832 & 9.802 & \textless0.001 \\
-R2adj = 0.158 & & & & \\
-& & & & \\
-\textbf{Aulacanthidae Lower Mesopelagic} & Avg Mass Flux & 1.622 & 1.991
-& 0.002 \\
-R2adj=0.298 & Particle Concentration & 2.123 & 6.706 & \textless0.001 \\
-& & & & \\
-\textbf{Aulosphaeridae Upper Mesopelagic} & Particle Concentration &
-2.653 & 13.06 & \textless0.001 \\
-R2adj=0.2 & & & & \\
-& & & & \\
-\textbf{Aulosphaeridae Lower Mesopelagic} & Particle Concentration &
-0.972 & 6.248 & \textless0.001 \\
-R2adj=.147 & & & & \\
-& & & & \\
-\textbf{Castanellidae Epipelagic} & Avg C Flux & 1.462 & 0.816 &
-0.0421 \\
-R2adj = 0.124 & Primary Productivity & 0.771 & 0.665 & 0.0247 \\
-& Particle Concentration & 3.956 & 4.623 & \textless0.001 \\
-& & & & \\
-\textbf{Coelodendridae Upper Mesopelagic} & Avg N Flux & 0.822 & 0.919 &
-0.0183 \\
-R2adj=.113 & Particle Concentration & 0.970 & 6.208 & \textless0.001 \\
-& & & & \\
-\textbf{Coelodendridae Lower Mesopelagic} & Particle Concentration &
-1.773 & 4.873 & \textless0.001 \\
-R2adj=.133 & & & & \\
-& & & & \\
-\textbf{Collodaria Epipelagic} & Salinity & 0.925 & 2.267 &
-\textless0.001 \\
-R2adj=0.16 & O2 & 2.015 & 2.217 & 0.002 \\
-& Bacteria \#/L & 1.843 & 2.100 & 0.002 \\
-& & & & \\
-\textbf{Foraminifera Epipelagic} & Temperature & 4.414 & 8.789 &
-\textless0.001 \\
-R2adj=0.445 & O2 & 1.603 & 1.287 & 0.0111 \\
-& Avg N Flux & 2.512 & 3.396 & \textless0.001 \\
-& Primary Productivity & 0.824 & 0.920 & 0.0153 \\
-& Particle Concentration & 0.919 & 2.257 & \textless0.001 \\
-\bottomrule()
-\end{longtable}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/01_taxa.pdf}
-
-}
-
-\caption{Example images of different Rhizaria taxa. 4mm scale bar shown
-in lower right. All vignettes are same scale}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/02-environmental.pdf}
-
-}
-
-\caption{Environmental profiles across time-series of study period. Y
-axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved
-Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion
-(184 - 450\(\mu m\)). (F) Bacteria Abundance. (G) Silica. (H) Nitrate}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/03_prod-flux.pdf}
-
-}
-
-\caption{Primary productivity (A) and flux estimates from total mass
-(B), carbon (C), and nitrogen (D). Values are from monthly cruises with
-month displayed in corresponding colors. Absent data are shown by grey
-bar.}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/04_vertical.pdf}
-
-}
-
-\caption{Average abundance of Rhizaria in 25m bins, across entire study
-period. Shown are total Rhizaria (A), Acantharea (B), Collodaria (C),
-Aulacanthidae (D), Foraminifera (E), Aulosphaeridae (F), Castanellidae
-(G), Coelodendridae (H).}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/05_seasonality.pdf}
-
-}
-
-\caption{Seasonality of Rhizarian integrated abundance for the
-epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower
-mesopelagic (500-1000m) (E-F). Left panels (A,C,E) display total
-integrated abundance per monthly cruise colored by month. Right panels
-(B, D, F) display community composition of each total abundance.}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/09_size-comp.pdf}
-
-}
-
-\caption{Comparison of average sizes (ESD) amongst Rhizarian taxa which
-occurred throughout the water column.}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/06_epi-partials.pdf}
-
-}
-
-\caption{Partial effects of smooth terms in taxa-specific GAM models
-from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea
-(A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae
-(E).}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/07_upmeso-partial.pdf}
-
-}
-
-\caption{Partial effects of smooth terms in taxa-specific GAM models
-from the upper mesopelagic (200-500m). Effects are grouped by taxa;
-Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
-(D).}
-
-\end{figure}
-
-\newpage
-
-\begin{figure}
-
-{\centering \includegraphics{images/08_lomeso-partials.pdf}
-
-}
-
-\caption{Partial effects of smooth terms in taxa-specific GAM models
-from the lower mesopelagic (500-1000m). Effects are grouped by taxa;
-Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
-(D).}
-
-\end{figure}
-
-\newpage
-
-end.
-
 \end{document}

### 2024-05-02 R/01_size-vol-checks.R
add to supplement

@@ -139,6 +139,29 @@ meso_avg * ((300)/25)
 non_detect(0.1, epi_avg * (200/25))
 non_detect(0.1, meso_avg * ((1000-200) /25))
 
+###
+# Table for supplement of sampling volumes #######################
+###
+
+meta_basics <- data.frame(
+  profileid = uvp_data$meta$profileid,
+  cruise_id = uvp_data$meta$cruise_id,
+  year_mo = as.Date(
+    paste(year(uvp_data$meta$sampledate),month(uvp_data$meta$sampledate), '01', sep = '-'),
+    format = '%Y-%m-%d'
+  )
+)
+meta_basics$month = month(meta_basics$year_mo, label = T)
+
+vol_by_cast |> 
+  left_join(meta_basics, by = 'profileid') |> 
+  group_by(zone, month, cruise_id, year(year_mo)) |> 
+  summarize(n_casts = length(unique(profileid)),
+            mean_vol = mean(vol_sampled),
+            sd_vol = sd(vol_sampled),
+            tot_vol = sum(vol_sampled)) |> 
+  write.csv('./output/Supplement/sT01_vol-by-cast.csv', row.names = F)
+
 
 ###
 # Calculate Binned Abundances #############

### 2024-05-02 R/s01_cast-map.R

@@ -3,6 +3,8 @@
 ###
 library(EcotaxaTools)
 library(ggplot2)
+library(ggOceanMaps)
+library(ggspatial)
 library(dplyr)
 
 ctd_data <- readRDS('./data/00_ctd-data.rds')
@@ -26,16 +28,24 @@ library("rnaturalearthdata")
 
 world <- ne_countries(scale = "medium", returnclass = "sf")
 
-all_cast_map <- ggplot() +
-  geom_sf(data = world, fill = 'green')+
+all_cast_map <- basemap(limits = c(-65,-62,30,33), bathymetry = T, bathy.alpha = 0.5) +  
   geom_point(aes(x = coords$lon, y =coords$lat))+
-  coord_sf(xlim = c(-65,-62), ylim = c(30,33))+
+  theme_bw()+
+  labs(x = "",y = "", shape = "")
+
+
+atlantic_ocean <- ggplot() +
+  geom_sf(data = world, fill = '#7b7b7b') +
+  geom_rect(aes(xmin = -65, xmax = -62,
+                ymin = 30, ymax = 33),
+            fill = 'grey', color = 'black', alpha = 0.1) +
+  coord_sf(xlim = c(-79,-5), ylim = c(23.5,48))+
   theme_bw()+
   labs(x = "",y = "", shape = "")+
-  theme(panel.background = element_rect(fill = 'cornflowerblue'),
+  theme(panel.background = element_rect(fill = 'lightgrey'),
         axis.text.x = element_text(angle = 45, hjust = c(1,1)),
         legend.position = 'bottom')
 
+ggsave('./output/Supplement/s01_background-map.pdf', atlantic_ocean)
 ggsave('./output/Supplement/s01_cast-map.pdf', all_cast_map)
 
-windows()

### 2024-06-12 R/01_par-conc.R
Make changes to par_conc for two size classes

@@ -10,13 +10,27 @@ library(dplyr)
 uvp_data <- readRDS('./data/00_zoop-uvp.rds') 
 
 par_conc <- uvp_data |> 
-  uvp_par_conc(min_esd = 0.184, max_esd = .450)
+  uvp_par_conc(bin_limits = c(0.184, 0.450, 0.900))
+
+# rename particles sizes
+renamer <- function(x) {
+  switch(x,
+         '(0.184,0.45]' = 'small',
+         '(0.45,0.9]' = 'large')
+}
+
+# loop through and apply
+for(cast in names(par_conc)) {
+  par_conc[[cast]][['esd_bin']] <- par_conc[[cast]][['esd_bin']]|> 
+    sapply(renamer)
+}
+
 
 db_cutter <- function(pdf) {
   pdf$db <- cut(pdf$depth, seq(0,1000,25))
   
   pdf <- pdf |> 
-    group_by(db) |> 
+    group_by(db, esd_bin) |> 
     summarize(par_conc = mean(par_conc))
   
   return(pdf)

### 2024-06-12 R/02_vertical-distributions.R

@@ -12,7 +12,7 @@ library(EcotaxaTools)
 # Data prep #################
 ###
 
-# |- Load in data -------
+## |- Load in data -------
 rhiz <- readRDS('./data/01_rhiz-densities.rds')
 meta <- readRDS('./data/00_zoop-uvp.rds')$meta
 
@@ -20,7 +20,7 @@ meta <- readRDS('./data/00_zoop-uvp.rds')$meta
 meta$month <- month(meta$sampledate)
 meta$year <- year(meta$sampledate)
 
-# |- make a all-rhizaria dataframe --------------
+## |- make a all-rhizaria dataframe --------------
 
 sum_by_cast <- function(cast_df) {
   out_df <- cast_df |> 
@@ -49,7 +49,7 @@ rhiz_avg_group <- rhiz |>
   summarize(mean = mean(conc_m3),
             sd = sd(conc_m3))
 
-# |- Save for later plots ---------------
+## |- Save for later plots ---------------
 saveRDS(
   list(
     all = total_avg_rhiz,
@@ -64,7 +64,7 @@ saveRDS(
 # Integrate for Seasonality ##########
 ####
 
-# |- Set up--------------------------
+## |- Set up--------------------------
 
 # write a function to make integrations based on depths and reject non-matches
 
@@ -128,8 +128,8 @@ cruise_id_convert <- function(rhiz_df) {
   return(rhiz_df)
 }
 
-# |- Total Calcs -----------------------------------
-# |-|- Epipelagic --------------------
+## |- Total Calcs -----------------------------------
+## |-|- Epipelagic --------------------
 
 all_epi <- total_rhiz |> 
   lapply(integrate_range_total, 0, 200) |> 
@@ -154,7 +154,7 @@ all_epi_sum <- all_epi_sum |>
   cruise_id_convert()
 
 
-# |-|- Upper Meso --------------------
+## |-|- Upper Meso --------------------
 
 all_upmeso <- total_rhiz |> 
   lapply(integrate_range_total, 200, 500) |> 
@@ -179,7 +179,7 @@ all_upmeso_sum <- all_upmeso_sum |>
   cruise_id_convert()
 
 
-# |-|- Lower Meso --------------------
+## |-|- Lower Meso --------------------
 
 all_lomeso <- total_rhiz |> 
   lapply(integrate_range_total, 500, 1000) |> 
[…]

### 2024-06-12 R/03_environmental-data-mgmt.R

@@ -12,21 +12,21 @@ library(lubridate)
 # Data Prep ############
 ##
 
-# |- Original UVP data -------------
+## |- Original UVP data -------------
 meta <- readRDS('./data/00_zoop-uvp.rds')$meta
 
-# |- Particle data --------------------
+## |- Particle data --------------------
 par_conc <- readRDS('./data/01_par-conc.rds')
 
 par_df <- par_conc |> 
   list_to_tib('profileid')
 
-# |- Rhizaria Data -------------------------------------------------
+## |- Rhizaria Data -------------------------------------------------
 
 # bin-specific data
 binned_rhiz <- readRDS('./data/01_rhiz-densities.rds')
 
-# |-|- make a all-rhizaria dataframe --------------
+## |-|- make a all-rhizaria dataframe --------------
 
 sum_by_cast <- function(cast_df) {
   out_df <- cast_df |> 
@@ -43,7 +43,7 @@ intg_total_rhiz <- readRDS('./data/02_integrated-all-rhiz.rds')
 intg_taxa <- readRDS('./data/02_integrated-taxa.rds')
 
 
-# |- Environmental Data --------------------------------------
+## |- Environmental Data --------------------------------------
 
 # readin
 bottle_data <- readRDS('./data/00_bottles.rds')
@@ -51,7 +51,7 @@ ctd_data <- readRDS('./data/00_ctd-data.rds')
 flux_data <- readRDS('./data/00_flux.rds')
 prod_data <- readRDS('./data/00_prod.rds')
 
-# |-|- CTD Data ----------------------------
+## |-|- CTD Data ----------------------------
 
 ctd_binner <- function(ctd_df) {
   
@@ -89,12 +89,12 @@ ctd_bins <- ctd_data |>
 ctd_df <- ctd_bins |> 
   do.call(what = rbind,)
 
-# |-|-|- Correct the CTD Data ---------------
+## |-|-|- Correct the CTD Data ---------------
 ctd_df$o2[which(ctd_df$o2 < 0)] <- 0
 ctd_df$sal[which(ctd_df$sal < 0)] <- 0
 
 
-# |-|- Bottle Data ------------------------
+## |-|- Bottle Data ------------------------
 
 bottle_data$cruise_id <- bottle_data$ctd_origfilename |> 
   as.character() |> 
@@ -114,7 +114,7 @@ drop_idx <- bottle_data[,c(18:26)] |>
 
 bottle_data <- bottle_data[-which(drop_idx == 0),]
 
[…]

### 2024-06-12 R/04_gam-intg.R

@@ -9,6 +9,7 @@ library(EcotaxaTools)
 library(dplyr)
 library(tidyr)
 library(mgcViz)
+source('./R/utils.R')
 
 ##
 # Load in Data #########
@@ -27,14 +28,19 @@ for(taxa in names(taxa_rhiz$epi)) {
 # Total Rhizaria ##########
 ###
 
-# |- Epi pelagic -----------------------
+## |- Epi pelagic -----------------------
 
 # full model 
+full_epi_formula <-   intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) + s(RFU, k = 6) +
+  s(Si, k = 6) + s(Bact_enumb, k = 6) + s(NO3, k = 6) +
+  s(avg_mass_flux_200, k = 6) + s(avg_fbc_200, k = 6) + s(avg_fbn_200, k = 6) +
+  s(pp, k = 6) + s(small, k = 6) + s(large, k = 6)
+
 full_epi_total <- gam(
   intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) + s(RFU, k = 6) +
     s(Si, k = 6) + s(Bact_enumb, k = 6) + s(NO3, k = 6) +
     s(avg_mass_flux_200, k = 6) + s(avg_fbc_200, k = 6) + s(avg_fbn_200, k = 6) +
-    s(pp, k = 6) + s(par_conc, k = 6),
+    s(pp, k = 6) + s(small, k = 6) + s(large, k = 6),
   data = tot_rhiz$epi, method = 'ML', select = T
 )
 full_epi_total |>  summary()
@@ -50,7 +56,7 @@ red_epi_total |>  summary()
 AIC(red_epi_total, full_epi_total)
 BIC(red_epi_total, full_epi_total)
 
-# |- UpMeso ----------------------------------
+## |- UpMeso ----------------------------------
 
 full_upmeso_tot <- gam(
   intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) +
@@ -73,7 +79,7 @@ red_upmeso_tot |> summary()
 AIC(full_upmeso_tot, red_upmeso_tot)
 BIC(full_upmeso_tot, red_upmeso_tot)
 
-# |- LoMeso -----------------------------
+## |- LoMeso -----------------------------
 
 
 full_lomeso_tot <- gam(
@@ -99,7 +105,7 @@ red_lomeso_tot |> summary()
 # Acantharea #########
 ##
[…]

### 2024-06-12 R/plot_0103_environmental-interps.R

@@ -12,7 +12,7 @@ library(tidyr)
 library(scales)
 source('./R/utils.R')
 
-# |- Interpolating Function -----------------------
+## |- Interpolating Function -----------------------
 
 surf_interp <- function(df,
                         value,
@@ -50,26 +50,35 @@ par_df$Date <- as.Date(par_df$sampledate)
 
 # for casts occuring on the same date, mean
 par_df <- par_df |> 
-  group_by(Date, db) |> 
+  group_by(Date, db, esd_bin) |> 
   summarize(par_conc = mean(par_conc)) |> 
   ungroup() |> 
   bin_format()
 
 par_df$Depth = par_df$mp
 
-# |- Interpolate -----------------
+## |- Interpolate -----------------
 
-par_interp <- surf_interp(par_df, 'par_conc')
+## |-|- NEEDS REVISION FOR SMALL & LARGE --------------------------
+
+## |-|- Small -----------------------------------
+par_df <- par_df |> 
+  pivot_wider(names_from = 'esd_bin',
+              values_from = 'par_conc')
+
+
+small_par_interp <- surf_interp(par_df, 'small')
+small_par_interp$small[small_par_interp$small > 1] = 1
 
 par_df_early <- par_df[par_df$Date <= '2019-12-31',]
 par_df_late <- par_df[par_df$Date > '2019-12-31',]
 
-par_early_interp <- par_interp[par_interp$Date <= '2019-09-30',]
-par_late_interp <- par_interp[par_interp$Date >= '2020-10-24',]
+small_par_early_interp <- small_par_interp[small_par_interp$Date <= '2019-09-30',]
+small_par_late_interp <- small_par_interp[small_par_interp$Date >= '2020-10-24',]
 
-par_early_plot <- ggplot() +
-  geom_tile(data = par_early_interp,
-            aes(x = Date, y = Depth, fill = par_conc)) +
+small_par_early_plot <- ggplot() +
+  geom_tile(data = small_par_early_interp,
+            aes(x = Date, y = Depth, fill = small)) +
   geom_point(data = par_df_early,
              aes(x = as.POSIXct(Date), y = Depth),
              size = 0.1, color = 'lightgrey', alpha = 0.75) +
@@ -81,27 +90,76 @@ par_early_plot <- ggplot() +
         axis.title = element_blank(),
         axis.text = element_text(size = 2))
 
-par_late_plot <- ggplot() +
-  geom_tile(data = par_late_interp,
[…]

### 2024-06-12 R/utils.R

@@ -3,6 +3,63 @@
 ##
 
 ###
+# Backstepping #########
+###
+# Function to perform backward stepwise selection on GAM model
+backward_stepwise_gam <- function(formula, data, 
+                                  alpha = 0.1,
+                                  method = "ML", select = TRUE) {
+  # Initial model
+  model <- gam(formula, data = data, method = method, select = select)
+  summary_model <- summary(model)
+  old_rsq <- summary_model$r.sq
+  
+  counter = 0
+  # Loop until no non-significant terms remain
+  while (TRUE) {
+    # Get p-values of the terms
+    pvalues <- summary_model$s.table[, "p-value"]
+    
+    # Check if any terms are non-significant
+    if (all(pvalues < alpha)) {
+      break
+    }
+    
+    # Update the formula by removing the least significant term
+    terms <- attr(terms(formula), "term.labels")
+    terms <- terms[-which.max(pvalues)]
+    formula <- as.formula(paste("intg ~", paste(terms, collapse = " + ")))
+    
+    # Refit the model with the updated formula
+    new_mod <- gam(formula, data = data, method = method, select = select)
+    new_summary_mod <- summary(new_mod)
+    
+    # check to see if rsq got much worse
+    if(new_summary_mod$r.sq < (summary_model$r.sq - 0.1)) {
+      break
+    } else if(length(terms) < 1){
+      print('No significant terms')
+      break
+    } else {
+      model = new_mod
+      summary_model = new_summary_mod
+      counter = counter + 1
+      cat('\n')
+      cat('_________________________________________')
+      cat('\n')
+      cat(paste0('Model update number: ', counter))
+      cat('\n')
+      print(summary_model)
+    }
+  }
+  
+  return(model)
+}
+
+
+
+
+###
 # Plot colors #######
 ###
 

### 2024-06-13 R/01_par-conc.R
update particles

@@ -10,7 +10,7 @@ library(dplyr)
 uvp_data <- readRDS('./data/00_zoop-uvp.rds') 
 
 par_conc <- uvp_data |> 
-  uvp_par_conc(min_esd = 0.184, max_esd = .450)
+  uvp_par_conc(min_esd = 0.184, max_esd = 0.9)
 
 db_cutter <- function(pdf) {
   pdf$db <- cut(pdf$depth, seq(0,1000,25))

### 2024-06-13 R/04_gam-intg.R

@@ -9,6 +9,7 @@ library(EcotaxaTools)
 library(dplyr)
 library(tidyr)
 library(mgcViz)
+source('./R/utils.R')
 
 ##
 # Load in Data #########
@@ -27,53 +28,46 @@ for(taxa in names(taxa_rhiz$epi)) {
 # Total Rhizaria ##########
 ###
 
-# |- Epi pelagic -----------------------
+## |- Epi pelagic -----------------------
 
 # full model 
-full_epi_total <- gam(
-  intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) + s(RFU, k = 6) +
-    s(Si, k = 6) + s(Bact_enumb, k = 6) + s(NO3, k = 6) +
-    s(avg_mass_flux_200, k = 6) + s(avg_fbc_200, k = 6) + s(avg_fbn_200, k = 6) +
-    s(pp, k = 6) + s(par_conc, k = 6),
+
+full_epi_form <-   intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) + s(RFU, k = 6) +
+  s(Si, k = 6) + s(Bact_enumb, k = 6) + s(NO3, k = 6) +
+  s(avg_mass_flux_200, k = 6) + s(avg_fbc_200, k = 6) + s(avg_fbn_200, k = 6) +
+  s(pp, k = 6) + s(par_conc, k = 6)
+
+
+full_epi_total <- gam(full_epi_form,
   data = tot_rhiz$epi, method = 'ML', select = T
 )
 full_epi_total |>  summary()
+full_epi_total |> concurvity()
 
 # reduced model
-red_epi_total <- gam(
-  intg ~ s(sal, k = 6) + s(o2, k = 6) +
-    s(Si, k = 6) + s(NO3, k = 6) +
-    s(pp, k = 6) + s(par_conc, k = 6),
-  data = tot_rhiz$epi, method = 'ML', select = T
-)
+red_epi_total <- backward_stepwise_gam(full_epi_form, tot_rhiz$epi)
 red_epi_total |>  summary()
-AIC(red_epi_total, full_epi_total)
-BIC(red_epi_total, full_epi_total)
 
-# |- UpMeso ----------------------------------
+## |- UpMeso ----------------------------------
 
-full_upmeso_tot <- gam(
[…]

### 2024-06-13 R/TEMP_CONCURVEFUNC.R
add model checking auto functions

@@ -1,0 +1,94 @@
+library(mgcv)
+
+# Function to fit GAM and check for concurvity
+fit_gam_check_concurvity <- function(formula, data, k = 6, method = "ML", 
+                                     select = TRUE, threshold = 0.8) {
+  while (TRUE) {
+    # Fit the model
+    model <- gam(formula, data = data, method = method, select = select)
+    
+    # Check concurvity
+    conc <- concurvity(model, full = FALSE)$estimate
+    #format the matrix
+    conc <- conc[2:nrow(conc),]
+    conc <- conc[,2:ncol(conc)]
+    conc[lower.tri(conc, diag = T)] <- NA
+    lower.tri(conc)
+    
+    # Find pairs with concurvity above the threshold
+    # If no high concurvity pairs found, break the loop
+    if (all(conc < threshold)) {
+      break
+    } else {
+      max_conc <- max(conc, na.rm = T)
+      max_idx <- which(conc == max_conc, arr.ind = T)
+      
+      if(nrow(max_idx) > 1) {
+        cat(paste0('\n', 'There are more than one max concurvity scores',
+                   '\n', 'They are for terms: ', rownames(max_idx), '\n',
+                   'I didn\'t plan for this, fix funciton in utils or self evaluate'))
+      } else {
+        term1 <- rownames(conc)[max_idx[1]] |> add_ks(k)
+        term2 <- colnames(conc)[max_idx[2]]
+        
+        ### STOP HERE TO FIX issue with adding ks to the model terms
+        
+        
+        
+      }
+    }
+    # Identify terms in the pair with the highest concurvity
+    term1 <- rownames(high_conc_pairs)[1]
+    term2 <- colnames(high_conc_pairs)[1]
+    
+    # Create formulas with each term removed
+    terms <- attr(terms(formula), "term.labels")
+    formula1 <- as.formula(paste("intg ~", paste(setdiff(terms, term1), collapse = " + ")))
+    formula2 <- as.formula(paste("intg ~", paste(setdiff(terms, term2), collapse = " + ")))
+    
+    # Fit models with each term removed
+    model1 <- gam(formula1, data = data, method = method, select = select)
+    model2 <- gam(formula2, data = data, method = method, select = select)
+    
+    # Compare models (e.g., using AIC)
+    aic1 <- AIC(model1)
[…]

### 2024-06-13 R/utils.R

@@ -1,6 +1,59 @@
 ##
 # Utilities ##
 ##
+# Function to perform backward stepwise selection on GAM model
+backward_stepwise_gam <- function(formula, data, 
+                                  alpha = 0.1,
+                                  method = "ML", select = TRUE) {
+  # Initial model
+  model <- gam(formula, data = data, method = method, select = select)
+  summary_model <- summary(model)
+  old_rsq <- summary_model$r.sq
+  
+  counter = 0
+  # Loop until no non-significant terms remain
+  while (TRUE) {
+    # Get p-values of the terms
+    pvalues <- summary_model$s.table[, "p-value"]
+    
+    # Check if any terms are non-significant
+    if (all(pvalues < alpha)) {
+      break
+    }
+    
+    # Update the formula by removing the least significant term
+    terms <- attr(terms(formula), "term.labels")
+    terms <- terms[-which.max(pvalues)]
+    formula <- as.formula(paste("intg ~", paste(terms, collapse = " + ")))
+    
+    # Refit the model with the updated formula
+    new_mod <- gam(formula, data = data, method = method, select = select)
+    new_summary_mod <- summary(new_mod)
+    
+    # check to see if rsq got much worse
+    if(new_summary_mod$r.sq < (summary_model$r.sq - 0.05)) {
+      break
+    } else if(length(terms) < 1){
+      print('No significant terms')
+      break
+    } else {
+      model = new_mod
+      summary_model = new_summary_mod
+      counter = counter + 1
+      cat('\n')
+      cat('_________________________________________')
+      cat('\n')
+      cat(paste0('Model update number: ', counter))
+      cat('\n')
+      print(summary_model)
+    }
+  }
+  
+  return(model)
+}
+
+
+
 
 ###
 # Plot colors #######

### 2024-06-13 R/04_gam-intg.R
generate models with autoselecting tools

@@ -3,7 +3,7 @@
 ###
 
 rm(list = ls())
-
+set.seed(3)
 library(mgcv)
 library(EcotaxaTools)
 library(dplyr)
@@ -41,12 +41,11 @@ full_epi_form <-   intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) + s(RFU,
 full_epi_total <- gam(full_epi_form,
   data = tot_rhiz$epi, method = 'ML', select = T
 )
-full_epi_total |>  summary()
-full_epi_total |> concurvity()
-
+full_epi_total |> summary()
 # reduced model
-red_epi_total <- backward_stepwise_gam(full_epi_form, tot_rhiz$epi)
-red_epi_total |>  summary()
+red_epi_total <- backward_stepwise_gam(full_epi_form, tot_rhiz$epi,
+                                       k = 6, method = 'ML', select = T,
+                                       conc_threshold = 0.8)
 
 ## |- UpMeso ----------------------------------
 
@@ -55,7 +54,6 @@ full_upmeso_form <-   intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) +
   s(avg_mass_flux_200, k = 6) + s(avg_fbc_200, k = 6) + s(avg_fbn_200, k = 6) +
   s(par_conc, k = 6)
 
-
 full_upmeso_tot <- gam(full_upmeso_form,
   data = tot_rhiz$upmeso, method = 'ML', select = T 
 )
@@ -63,30 +61,21 @@ full_upmeso_tot |> summary()
 
 
 red_upmeso_tot <- backward_stepwise_gam(full_upmeso_form, tot_rhiz$upmeso)
-red_upmeso_tot |> summary()
-AIC(full_upmeso_tot, red_upmeso_tot)
-BIC(full_upmeso_tot, red_upmeso_tot)
 
 ## |- LoMeso -----------------------------
 
+full_lomeso_form <-   intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) +
+  s(Si, k = 6) +
+  s(avg_mass_flux_200, k = 6) + s(avg_fbc_200, k = 6) + s(avg_fbn_200, k = 6) +
+  s(par_conc, k = 6)
 
-full_lomeso_tot <- gam(
-  intg ~ s(temp, k = 6) + s(sal, k = 6) + s(o2, k = 6) +
[…]

### 2024-06-13 R/utils.R

@@ -1,12 +1,35 @@
 ##
 # Utilities ##
 ##
-# Function to perform backward stepwise selection on GAM model
+
+
+
+#' Function to perform backward stepwise selection on GAM model
+#' 
+#' This function performs backward stepwise selection on a GAM model by
+#' iteratively removing the least significant term until all terms are
+#' significant. The function also checks for concurvity in the model and
+#' refits the model if concurvity is detected.
+#' 
+#' @param formula a formula object specifying the model
+#' @param data a data frame containing the data
+#' @param alpha the significance level for removing terms
+#' @param k the number of knots to use in the model
+#' @param method the method to use for fitting the model
+#' @param select whether to use the select argument in the gam function
+#' @param conc_threshold the threshold for concurvity
 backward_stepwise_gam <- function(formula, data, 
-                                  alpha = 0.1,
-                                  method = "ML", select = TRUE) {
-  # Initial model
-  model <- gam(formula, data = data, method = method, select = select)
+                                  alpha = 0.1, k = 6,
+                                  method = "ML", select = TRUE,
+                                  conc_threshold = 0.8) {
+  
+  # Initial model passed to concurvity check:
+  model <- fit_gam_check_concurvity(formula, data, 
+                                    k = k, 
+                                    method = method, 
+                                    select = select,
+                                    threshold = conc_threshold)
+  
   summary_model <- summary(model)
   old_rsq <- summary_model$r.sq
   
@@ -18,20 +41,28 @@ backward_stepwise_gam <- function(formula, data,
     
     # Check if any terms are non-significant
     if (all(pvalues < alpha)) {
+      cat(paste0('\n', 'Breaking due to alpha threshold'))
       break
     }
     
     # Update the formula by removing the least significant term
+    formula <- formula(model)
     terms <- attr(terms(formula), "term.labels")
     terms <- terms[-which.max(pvalues)]
     formula <- as.formula(paste("intg ~", paste(terms, collapse = " + ")))
     
     # Refit the model with the updated formula
-    new_mod <- gam(formula, data = data, method = method, select = select)
+    # this is the proposal model
+    # need to concurvity check the new model!
+    new_mod <- fit_gam_check_concurvity(formula, data, 
[…]

### 2024-06-13 R/TEMP_CONCURVEFUNC.R
delete temp function file

@@ -1,94 +1,0 @@
-library(mgcv)
-
-# Function to fit GAM and check for concurvity
-fit_gam_check_concurvity <- function(formula, data, k = 6, method = "ML", 
-                                     select = TRUE, threshold = 0.8) {
-  while (TRUE) {
-    # Fit the model
-    model <- gam(formula, data = data, method = method, select = select)
-    
-    # Check concurvity
-    conc <- concurvity(model, full = FALSE)$estimate
-    #format the matrix
-    conc <- conc[2:nrow(conc),]
-    conc <- conc[,2:ncol(conc)]
-    conc[lower.tri(conc, diag = T)] <- NA
-    lower.tri(conc)
-    
-    # Find pairs with concurvity above the threshold
-    # If no high concurvity pairs found, break the loop
-    if (all(conc < threshold)) {
-      break
-    } else {
-      max_conc <- max(conc, na.rm = T)
-      max_idx <- which(conc == max_conc, arr.ind = T)
-      
-      if(nrow(max_idx) > 1) {
-        cat(paste0('\n', 'There are more than one max concurvity scores',
-                   '\n', 'They are for terms: ', rownames(max_idx), '\n',
-                   'I didn\'t plan for this, fix funciton in utils or self evaluate'))
-      } else {
-        term1 <- rownames(conc)[max_idx[1]] |> add_ks(k)
-        term2 <- colnames(conc)[max_idx[2]]
-        
-        ### STOP HERE TO FIX issue with adding ks to the model terms
-        
-        
-        
-      }
-    }
-    # Identify terms in the pair with the highest concurvity
-    term1 <- rownames(high_conc_pairs)[1]
-    term2 <- colnames(high_conc_pairs)[1]
-    
-    # Create formulas with each term removed
-    terms <- attr(terms(formula), "term.labels")
-    formula1 <- as.formula(paste("intg ~", paste(setdiff(terms, term1), collapse = " + ")))
-    formula2 <- as.formula(paste("intg ~", paste(setdiff(terms, term2), collapse = " + ")))
-    
-    # Fit models with each term removed
-    model1 <- gam(formula1, data = data, method = method, select = select)
-    model2 <- gam(formula2, data = data, method = method, select = select)
-    
-    # Compare models (e.g., using AIC)
-    aic1 <- AIC(model1)
[…]

### 2024-06-14 manuscript/03_results.qmd
create supplemental smooth plots

@@
 |--------------------|---------------|-------------|-------------|-------------|
-| **All Rhizaria Epipelagic**        | Salinity               | 2.0818 | 4.654 | \<0.001 |
-| $R^2_{adj}$= 0.42                  | DO                     | 0.8272 | 0.945 | 0.0136  |
-|                                    | Silica                 | 3.1206 | 15.75 | \<0.001 |
-|                                    | NO3                    | 2.9970 | 6.268 | \<0.001 |
-|                                    | Primary Productivity   | 1.8367 | 2.099 | \<0.001 |
-|                                    | Particle Concentration | 0.9282 | 2.566 | \<0.001 |
-| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6847 | 0.431 | 0.073   |
-| $R^2_{adj}$= 0.406                 | Avg Mass Flux          | 1.5426 | 0.923 | 0.045   |
-|                                    | Particle Concentration | 3.2957 | 20.89 | \<0.001 |
-| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 1.6632 | 1.824 | 0.002   |
-| $R^2_{adj}$= 0.603                 | Particle Concentration | 0.7694 | 0.662 | 0.027   |
+| **All Rhizaria Epipelagic**        | Temperature            | 0.7950 | 0.492 | 0.0440  |
+| $R^2_{adj}$= 0.48                  | Salinity               | 2.4117 | 5.113 | \<0.001 |
+|                                    | Silica                 | 2.1511 | 8.601 | \<0.001 |
+|                                    | Bacteria Concentration | 0.8870 | 1.544 | \<0.001 |
+|                                    | Avg Mass Flux          | 1.4107 | 0.802 | 0.0211  |
+|                                    | Primary Productivity   | 1.9280 | 2.736 | \<0.001 |
+|                                    | Particle Concentration | 0.9151 | 2.136 | \<0.001 |
+| **All Rhizaria Upper Mesopelagic** | Silica                 | 0.6702 | 0.404 | 0.0814  |
+| $R^2_{adj}$= 0.40                  | Avg N Flux             | 2.9524 | 3.172 | \<0.001 |
+|                                    | Particle Concentration | 3.5568 | 11.82 | \<0.001 |
+| **All Rhizaria Lower Mesopelagic** | Avg Mass Flux          | 2.036  | 2.953 | \<0.001 |
+| $R^2_{adj}$= 0.61                  | Particle Concentration | 3.819  | 27.43 | \<0.001 |
 
-: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column.
+: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column. edf = estimated degrees of freedom, F = F-statistic, p = p-value.
+Models were selected using a backwards-stepwise approach with a retention threshold of p = 0.1 or a large enough reduction in model fit.
 
@@
 | Model                                | Term                   | edf   | F     | p       |
-|---------------|---------------|---------------|---------------|---------------|
-| **Acantharea Epipelagic**            | Salinity               | 2.264 | 4.113 | \<0.001 |
-| R2adj=0.53                           | O2                     | 2.579 | 6.712 | \<0.001 |
-|                                      | Avg Mass Flux          | 4.810 | 22.81 | \<0.001 |
-|                                      | Bacteria #/L           | 1.599 | 1.317 | 0.0117  |
-|                                      | Particle Concentration | 0.853 | 1.158 | 0.0087  |
-| **Acantharea Upper Mesopelagic**     | Avg C Flux             | 1.296 | 0.621 | 0.0439  |
-| R2adj=0.231                          | Avg N Flux             | 1.619 | 1.283 | 0.0026  |
-|                                      | Particle Concentration | 0.952 | 3.886 | \<0.001 |
-| **Acantharea Lower Mesopelagic**     | Temperature            | 2.076 | 4.155 | \<0.001 |
-| R2adj=0.509                          | Avg N Flux             | 1.494 | 1.216 | 0.0113  |
-|                                      | Primary Productivity   | 0.766 | 0.648 | 0.0253  |
-|                                      | Particle Concentration | 2.037 | 10.03 | \<0.001 |
-| **Aulacanthidae Epipelagic**         | Salinity               | 0.792 | 0.757 | 0.0241  |
-| R2adj=0.251                          | Avg N Flux             | 0.869 | 1.324 | 0.0018  |
-|                                      | Primary Productivity   | 2.312 | 7.704 | \<0.001 |
-|                                      | Particle Concentration | 2.008 | 2.200 | 0.0017  |
-| **Aulacanthidae Upper Mesopelagic**  | Particle Concentration | 2.832 | 9.802 | \<0.001 |
-| R2adj = 0.158                        |                        |       |       |         |
-| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.622 | 1.991 | 0.002   |
-| R2adj=0.298                          | Particle Concentration | 2.123 | 6.706 | \<0.001 |
-| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.653 | 13.06 | \<0.001 |
-| R2adj=0.2                            |                        |       |       |         |
-| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 0.972 | 6.248 | \<0.001 |
-| R2adj=.147                           |                        |       |       |         |
-| **Castanellidae Epipelagic**         | Avg C Flux             | 1.462 | 0.816 | 0.0421  |
-| R2adj = 0.124                        | Primary Productivity   | 0.771 | 0.665 | 0.0247  |
-|                                      | Particle Concentration | 3.956 | 4.623 | \<0.001 |
-| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.822 | 0.919 | 0.0183  |
-| R2adj=.113                           | Particle Concentration | 0.970 | 6.208 | \<0.001 |
-| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 1.773 | 4.873 | \<0.001 |
-| R2adj=.133                           |                        |       |       |         |
-| **Collodaria Epipelagic**            | Salinity               | 0.925 | 2.267 | \<0.001 |
-| R2adj=0.16                           | O2                     | 2.015 | 2.217 | 0.002   |
-|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.002   |
-| **Foraminifera Epipelagic**          | Temperature            | 4.414 | 8.789 | \<0.001 |
-| R2adj=0.445                          | O2                     | 1.603 | 1.287 | 0.0111  |
-|                                      | Avg N Flux             | 2.512 | 3.396 | \<0.001 |
-|                                      | Primary Productivity   | 0.824 | 0.920 | 0.0153  |
-|                                      | Particle Concentration | 0.919 | 2.257 | \<0.001 |
+|--------------------------------------|------------------------|-------|-------|---------|
+| **Acantharea Epipelagic**            | Temperature            | 0.868 | 0.380 | 0.0935  |
+| $R^2_{adj}$=0.539                    | Salinity               | 2.349 | 4.646 |\<0.001  |
+|                                      | O2                     | 2.501 | 5.493 |\<0.001  |
+|                                      | Avg Mass Flux          | 4.815 | 22.59 |\<0.001  |
+|                                      | Bacteria               | 1.608 | 1.360 | 0.00945 |
+|                                      | Particle Concentration | 0.877 | 1.414 | 0.00449 |
+| **Acantharea Upper Mesopelagic**     | Avg Mass Flux          | 1.866 | 1.220 | 0.0166  |
+| $R^2_{adj}$=0.197                    | Avg N Flux             | 2.208 | 3.100 |\<0.001  |
+|                                      | Primary Productivity   | 2.382 | 2.497 |\<0.001  |
+|                                      | Particle Concentration | 0.724 | 0.521 | 0.0437  |
+| **Acantharea Lower Mesopelagic**     | Temperature            | 1.958 | 2.986 |\<0.001  |
+| $R^2_{adj}$=0.496                    | Avg F Flux             | 1.513 | 1.565 | 0.0037  |
+|                                      | Primary Productivity   | 0.856 | 1.153 | 0.0045  |
+|                                      | Particle Concentration | 1.661 | 8.667 |\<0.001  |
+| **Aulacanthidae Epipelagic**         | Temperature            | 2.591 | 5.003 |\<0.001  |
+| $R^2_{adj}$=0.394                    | Salinity               | 1.523 | 1.789 | 0.002   |
+|                                      | RFU                    | 0.872 | 1.315 | 0.0042  |
+|                                      | Bacteria               | 0.808 | 0.833 | 0.0160  |
+|                                      | Primary Productivity   | 2.622 | 9.160 |\<0.001  |
+|                                      | Particle Concentration | 1.937 | 2.751 |\<0.001  |
+| **Aulacanthidae Upper Mesopelagic**  | Avg C Flux             | 1.675 | 1.408 | 0.0109  |
+| $R^2_{adj}$=0.132                    | Particle Concentration | 2.308 | 2.476 | 0.0013  |
+| **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.476 | 1.418 | 0.0076  |
+| $R^2_{adj}$=0.308                    | Particle Concentration | 2.457 | 7.756 | <0.001  |
+| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.778 | 12.76 | <0.001  |
+| $R^2_{adj}$=0.196                    |                        |       |       |         |
+| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 2.012 | 6.306 |\<0.001  |
+| $R^2_{adj}$=0.148                    |                         |      |       |         |
+| **Castanellidae Epipelagic**         | Salinity               | 0.915 | 0.400 | 0.0935  |
+| $R^2_{adj}$=0.228                    | RFU                    | 0.912 | 2.052 |\<0.001  |
+|                                      | Si                     | 2.018 | 3.656 |\<0.001  |
+|                                      | Bacteria               | 1.128 | 0.883 | 0.0170  |
+|                                      | Particle Concentration | 2.278 | 4.028 |\<0.001  |
+| **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.734 | 0.551 | 0.0527  |
+| $R^2_{adj}$=0.09                     | Particle Concentration | 0.965 | 5.102 |\<0.001  |
+| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 3.051 | 6.768 |\<0.001  |
+| $R^2_{adj}$=0.157                    |                        |       |       |         |
+| **Collodaria Epipelagic**.           | Salinity               | 0.926 | 2.267 |\<0.001  |
+| $R^2_{adj}$=0.16                     | O2                     | 2.015 | 2.217 |\<0.001  |
+|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.0024  |
+| **Foraminifera Epipelagic**          | Temperature            | 4.420 | 7.624 |\<0.001  |
+| $R^2_{adj}$=0.543                    | O2                     | 0.799 | 0.788 | 0.023   |
+|                                      | Bacteria.              | 0.000 | 0.000 | 0.523   |
+|                                      | Avg N Flux             | 1.847 | 3.022 |\<0.001  |
+|                                      | Primary Productivity   | 0.472 | 0.139 | 0.183   |
+|                                      | Particle Concentration | 1.685 | 2.955 |\<0.001  |
 
-: Taxa-specific generalized additive models for different regions of the water column.
+: Taxa-specific generalized additive models for different regions of the water column.edf = estimated degrees of freedom, F = F-statistic, p = p-value.
+Models were selected using a backwards-stepwise approach with a retention threshold of p = 0.1 or a large enough reduction in model fit.
 

### 2024-06-14 R/04_gam-intg.R

@@ -389,7 +389,28 @@ red_epi_Collodaria |>  summary()
 red_epi_Foraminifera |>  summary()
 
 
-
+## |- Save Models --------------------------
+saveRDS(
+  list(
+      tot_epi = red_epi_total,
+      tot_upmeso = red_upmeso_tot,
+      tot_lomeso = red_lomeso_tot,
+      acantharea_epi = red_epi_acantharea,
+      acantharea_upmeso = red_upmeso_acantharea,
+      acantharea_lomeso = red_lomeso_acantharea,
+      aulacanthidae_epi = red_epi_Aulacanthidae,
+      aulacanthidae_upmeso = red_upmeso_Aulacanthidae,
+      aulacanthidae_lomeso = red_lomeso_Aulacanthidae,
+      aulosphaeridae_upmeso = red_upmeso_Aulosphaeridae,
+      aulosphaeridae_lomeso = red_lomeso_Aulosphaeridae,
+      castanellidae_epi = red_epi_Castanellidae,
+      coelodendridae_upmeso = red_upmeso_Coelodendridae,
+      coelodendridae_lomeso = red_lomeso_Coelodendridae,
+      collodaria_epi = red_epi_Collodaria,
+      foraminifera_epi = red_epi_Foraminifera
+    ),
+  './data/04_gam-intg.rds'
+)
 
 
 

### 2024-06-14 R/plot_05_gam-partials.R

@@ -5,9 +5,18 @@
 rm(list = ls())
 
 
+###
+# Just to show particle relationships
 
 
-# |- Plot Function -----------------------
+
+
+###
+# Below is for first draft rendition #########
+###
+
+
+## |- Plot Function -----------------------
 
 make_gam_plot <- function(var, data) {
   plot <- ggplot() +
@@ -41,6 +50,7 @@ for(file in files) {
   name <- sub("^04_(.+?)-gam\\.rds$", '\\1', file)
   data[[name]] <- readRDS(paste0('./data/',file))
 }
+data = data[-which(names(data) == '04_gam-intg.rds')]
 
 save_plots <- function(name) {
   group_data <- data[[name]]
@@ -55,6 +65,70 @@ save_plots <- function(name) {
              units = 'in')
     }
   }
+}
+
+
+###
+# For Supplemental Plots ##########
+###
+
+taxa = names(data)
+taxa = taxa[-which(taxa == 'total')]
+
+zones = c(
+  `epi` = 'Epipelagic',
+  `upmeso` = 'Upper Mesopelagic',
+  `lomeso` = 'Lower Mesopelagic'
+)
+
+# need to make the plot into a function to avoid lazy load issue
+supp_plot <- function(taxon, zone, var) {
+  plot_data <- data[[taxon]][[zone]]
+  
+   outplot <- ggplot() +
+    geom_point(aes(x = plot_data[[var]]$obs[[var]], 
+                   y = plot_data[[var]]$obs$intg),
+               size = 3) +
+    geom_line(aes(x = plot_data[[var]]$pred[[var]],
+                  y = plot_data[[var]]$pred$fit)) +
+    geom_ribbon(aes(x = plot_data[[var]]$pred[[var]],
+                    ymin = plot_data[[var]]$pred$low,
+                    ymax = plot_data[[var]]$pred$high),
+                color = 'lightgrey',
+                alpha = 0.2) +
+    labs(x = var, y = 'Integrated Abundance',
+         subtitle = paste0(taxon, ' ', zones[zone])) +
+    theme_bw() +
+    theme(legend.position = 'none', 
+          axis.text = element_text(size = 8, face = 'bold'),
+          plot.subtitle = element_text(size = 8))
+   
+   return(outplot)
+}
+
+
+#### ERRORS HERE ####################
+
+for(zone in names(zones)) {
[…]

### 2024-06-14 R/plot_5b_model-fits.R

@@ -1,0 +1,63 @@
+######
+# Plotting Model Fit ########
+######
+
+rm(list = ls())
+graphics.off()
+library(ggplot2)
+library(ggpubr)
+
+# Load the data
+mods <- readRDS('./data/04_gam-intg.rds')
+
+
+plot_list = list()
+
+
+title_names <- c(
+  'tot_epi' = 'All Rhizaria Epipelagic',
+  'tot_upmeso' = 'All Rhizaria Upper Mesopelagic',
+  'tot_lomeso' = 'All Rhizaria Lower Mesopelagic',
+  'acantharea_epi' = 'Acantharea Epipelagic',
+  'acantharea_upmeso' = 'Acantharea Upper Mesopelagic',
+  'acantharea_lomeso' = 'Acantharea Lower Mesopelagic',
+  'aulacanthidae_epi' = 'Aulacanthidae Epipelagic',
+  'aulacanthidae_upmeso' = 'Aulacanthidae Upper Mesopelagic',
+  'aulacanthidae_lomeso' = 'Aulacanthidae Lower Mesopelagic',
+  'aulosphaeridae_upmeso' = 'Aulosphaeridae Upper Mesopelagic',
+  'aulosphaeridae_lomeso' = 'Aulosphaeridae Lower Mesopelagic',
+  'castanellidae_epi' = 'Castanellidae Epipelagic',
+  'coelodendridae_upmeso' = 'Coelodendridae Upper Mesopelagic',
+  'coelodendridae_lomeso' = 'Coelodendridae Lower Mesopelagic',
+  'collodaria_epi' = 'Collodaria Epipelagic',
+  'foraminifera_epi' = 'Foraminifera Epipelagic'
+)
+
+
+  
+plot_list <- lapply(
+  names(mods), function(name) {
+    ggplot() +
+      geom_point(aes(x = predict(mods[[name]]),
+                     y = mods[[name]]$y),
+                 size = 1, color = 'black') +
+      geom_abline(intercept = 0, slope = 1, color = 'grey', lty = 2) +
+      labs(x = 'Predicted', y = 'Observed',
+           subtitle = title_names[name]) +
+      scale_x_continuous(limits = c(
+        max(c(-50, min(predict(mods[[name]]))) - 0.1*min(predict(mods[[name]]))), 
+        max(predict(mods[[name]])) + 0.1*max(predict(mods[[name]]))
+      ))+
+      scale_y_continuous(limits = c(
+        0, 
[…]

### 2024-06-17 R/04_gam-intg.R
finalize output figures

@@ -19,9 +19,10 @@ tot_rhiz <- readRDS('./data/03_all-integrated.rds')
 tot_rhiz$epi <- tot_rhiz$epi[-which(tot_rhiz$epi$sal < 33),] # drop epi malfunction
 taxa_rhiz <- readRDS('./data/03_taxa-integrated.rds')
 
-# drop epi sal malfunction
+# drop epi sal malfunction and RFU -999
 for(taxa in names(taxa_rhiz$epi)) {
   taxa_rhiz$epi[[taxa]] <- taxa_rhiz$epi[[taxa]][-which(taxa_rhiz$epi[[taxa]]$sal < 33),]
+  taxa_rhiz$epi[[taxa]] <- taxa_rhiz$epi[[taxa]][-which(taxa_rhiz$epi[[taxa]]$RFU == -999),]
 }
 
 ###

### 2024-06-17 R/plot_05_gam-partials.R

@@ -3,7 +3,9 @@
 ###
 
 rm(list = ls())
-
+library(ggplot2)
+library(ggpubr)
+library(dplyr)
 
 ###
 # Just to show particle relationships
@@ -75,11 +77,38 @@ save_plots <- function(name) {
 taxa = names(data)
 taxa = taxa[-which(taxa == 'total')]
 
+
+## |- Name mapping for plots ------------------
 zones = c(
   `epi` = 'Epipelagic',
   `upmeso` = 'Upper Mesopelagic',
   `lomeso` = 'Lower Mesopelagic'
 )
+
+taxa_names = c(
+  `acanth` = 'Acantharea',
+  `aulacanthidae` = 'Aulacanthidae',
+  `aulosphaeridae` = 'Aulosphaeridae',
+  `castanellidae` = 'Castanellidae',
+  `coelodendridae` = 'Coelodendridae',
+  `collodaria` = 'Collodaria',
+  `foram` = 'Foraminifera',
+)
+
+var_names = c(
+  `temp` = 'Temperature [°C]',
+  `sal` = 'Salinity [ppt]',
+  `pp` = 'Primary Production [mg C m^-2 d^-1]',
+  `avg_mass_flux_200` = 'Average Mass Flux [mg C m^-2 d^-1]',
+  `Bact_enumb` = 'Bacterial Abundance [10^8 kg^-1]',
+  `RFU` = 'Relative Fluorescence Units',
+  `avg_fbn_200` = 'Average Nitrogen Flux [mg N m^-2 d^-1]',
+  `Si` = 'Silicate [µmol kg^-1]',
+  `o2` = 'Oxygen [µmol kg^-1]',
+  `par_conc` = 'Particle Concentration [L^-1]',
+  `avg_fbc_200` = 'Average Carbon Flux [mg C m^-2 d^-1]'
+)
+
 
 # need to make the plot into a function to avoid lazy load issue
 supp_plot <- function(taxon, zone, var) {
@@ -97,7 +126,7 @@ supp_plot <- function(taxon, zone, var) {
                 color = 'lightgrey',
                 alpha = 0.2) +
     labs(x = var, y = 'Integrated Abundance',
-         subtitle = paste0(taxon, ' ', zones[zone])) +
+         subtitle = paste0(taxa_names[taxon], ' ', zones[zone])) +
     theme_bw() +
     theme(legend.position = 'none', 
           axis.text = element_text(size = 8, face = 'bold'),

### 2024-06-17 R/plot_05_gam-partials.R
finalize core figure work

@@ -6,6 +6,7 @@ rm(list = ls())
 library(ggplot2)
 library(ggpubr)
 library(dplyr)
+graphics.off()
 
 ###
 # Just to show particle relationships
@@ -36,7 +37,11 @@ make_gam_plot <- function(var, data) {
     scale_color_gradient2(low = "#96e0fa", mid = "#B2182B", high = "#96e0fa", name = "Month",
                          midpoint = 6) +
     theme_bw() +
-    theme(legend.position = 'none', axis.text = element_text(size = 8, face = 'bold'))
+    theme(legend.position = 'none', 
+          axis.text = element_text(size = 18, face = 'bold'),
+          axis.title = element_blank(),
+          panel.border = element_rect(colour = 'black', fill = NA,
+                                      linewidth = 2))
   return(plot)
 }
 
@@ -92,7 +97,7 @@ taxa_names = c(
   `castanellidae` = 'Castanellidae',
   `coelodendridae` = 'Coelodendridae',
   `collodaria` = 'Collodaria',
-  `foram` = 'Foraminifera',
+  `foram` = 'Foraminifera'
 )
 
 var_names = c(
@@ -117,7 +122,7 @@ supp_plot <- function(taxon, zone, var) {
    outplot <- ggplot() +
     geom_point(aes(x = plot_data[[var]]$obs[[var]], 
                    y = plot_data[[var]]$obs$intg),
-               size = 3) +
+               size = 1) +
     geom_line(aes(x = plot_data[[var]]$pred[[var]],
                   y = plot_data[[var]]$pred$fit)) +
     geom_ribbon(aes(x = plot_data[[var]]$pred[[var]],
@@ -125,42 +130,43 @@ supp_plot <- function(taxon, zone, var) {
                     ymax = plot_data[[var]]$pred$high),
                 color = 'lightgrey',
                 alpha = 0.2) +
-    labs(x = var, y = 'Integrated Abundance',
+    labs(x = var_names[var], y = 'Integrated Abundance [indv. m^-2]',
          subtitle = paste0(taxa_names[taxon], ' ', zones[zone])) +
     theme_bw() +
     theme(legend.position = 'none', 
-          axis.text = element_text(size = 8, face = 'bold'),
-          plot.subtitle = element_text(size = 8))
+          axis.text = element_text(size = 6, face = 'bold'),
[…]

### 2024-06-17 R/01_size-vol-checks.R
move NBSS to own file

@@ -33,7 +33,7 @@ uvp_data <- uvp_data |>
 # Size threshold analysis ##########
 ####
 
-# |- Formatting for Rhiz only ----------------------------------
+## |- Formatting for Rhiz only ----------------------------------
 
 rhiz_only <- uvp_data |> 
   mod_zoo(func = 'names_keep', keep_names = 'Rhizaria',
@@ -56,34 +56,13 @@ rhiz_sizes <- rhiz_only$zoo_files |>
   list_to_tib('profileid')
 rhiz_sizes$esd <- rhiz_sizes$esd*unique(rhiz_only$meta$acq_pixel)
 
-# |- Density-based filter selection -------------------------
+## |- Density-based filter selection -------------------------
 rhiz_sizes$esd |> range()
 
 rhiz_sizes$esd |> 
   density() |> 
   plot()
 
-# |- NBSS option ------------------------------------
-
-calc_nbss <- function(df, metric = 'esd') {
-  
-  size_limits <- c(min_size = min(df[[metric]]), max_size = max(df[[metric]]))
-  
-  df$size_class <- df[[metric]] |>
-    assign_spectra_bins('log', min_size = size_limits[1], max_size = size_limits[2])
-  
-  db_num <- df |>
-    count_size_classes() |>
-    numeric_spectra(vol_L = 1, needs_format = T)
-  
-  return(db_num)
-}
-
-
-rhiz_nbss <- rhiz_sizes |>
-  calc_nbss()
-
-plot(log(n_s,2) ~ mp_size_class, data = rhiz_nbss)
 
 
 
@@ -91,7 +70,7 @@ plot(log(n_s,2) ~ mp_size_class, data = rhiz_nbss)
 # Sample Volume Selection ######################
 ####
 
-# |- get volume sampled in 25m bins ---------
+## |- get volume sampled in 25m bins ---------
 
 vol_by_cast <- uvp_data |> 
   get_ecopart_vol() |> 
@@ -113,7 +92,7 @@ for(r in 1:nrow(vol_by_cast)) {
 }
 
 
-# |- Detection Probabilities --------------------------
+## |- Detection Probabilities --------------------------
 # see barth review and benfield et al 1996 on VPR
 # euphotic zone
 
@@ -130,7 +109,7 @@ non_detect <- function(prob,vol) {
 non_detect(0.1, epi_avg)
 non_detect(0.1, meso_avg)
 
-# |- What about when integrating? ---------------------
+## |- What about when integrating? ---------------------
 
 # average volumes
[…]

### 2024-06-17 R/plot_s03_NBSS-calcs.R

@@ -1,0 +1,71 @@
+rm(list = ls())
+graphics.off()
+
+
+library(ggplot2)
+library(dplyr)
+library(EcotaxaTools)
+
+
+# Load the datar
+
+rhiz <- readRDS('./data/00_zoop-uvp.rds') |> 
+  mod_zoo(func = names_keep,
+          keep_names = 'Rhizaria', keep_children = TRUE)
+
+rhiz$zoo_files <- rhiz$zoo_files[-which(names(rhiz$zoo_files) == 'gf380c11')]
+rhiz$par_files <- rhiz$par_files[-which(names(rhiz$par_files) == 'gf380c11')]
+rhiz$meta <- rhiz$meta[which(rhiz$meta$profileid != 'gf380c11'),]
+
+rhiz = as_ecopart_obj(rhiz)
+
+vol_s = list()
+
+vol_s <- rhiz |> 
+  get_ecopart_vol() |> 
+  lapply(
+    function(df) {
+      sum(df$vol_sampled, na.rm = T)
+    }
+  ) |> 
+  list_to_tib('profileid')
+
+
+
+names(vol_s)[1] <- 'volume'
+
+
+## |- NBSS option ------------------------------------
+
+
+#' get the nbss for each cast.
+calc_nbss <- function(df, metric = 'esd') {
+  
+  size_limits <- c(min_size = min(df[[metric]]), max_size = max(df[[metric]]))
+  
+  df$size_class <- df[[metric]] |>
+    assign_spectra_bins('log', min_size = size_limits[1], max_size = size_limits[2])
+  
+  db_num <- df |>
+    count_size_classes() |>
+    numeric_spectra(vol_L = 1, needs_format = T)
+  
+  return(db_num)
+}
+
+
+# create bin sizes
+rhiz_sizes <- get_all(rhiz, 'esd', pixel_conv = T)
+min(rhiz_sizes)
+max(rhiz_sizes)
+
+
+## |-|- Create list of bins
+
+
+
+
+rhiz_nbss <- rhiz_sizes |>
+  calc_nbss()
+
+plot(log(n_s,2) ~ mp_size_class, data = rhiz_nbss)

### 2024-06-26 R/plot_01_epi-meso-size-comp.R
update size and nbss plots

@@ -71,12 +71,12 @@ for(r in 1:nrow(rhiz_sizes)) {
   }
 }
 
-# |- filter to more than 100 obs --------------
+##  filter to more than 100 obs --------------
 
 rhiz_sizes <- rhiz_sizes |> 
   filter(name %in% c('Acantharea', 'Aulacanthidae', 'Foraminifera', 'Aulosphaeridae'))
 
-# |- Plot sizes -------------------------
+## |- Plot sizes -------------------------
 
 ggplot(rhiz_sizes,aes(x = name, y = esd, fill = zone, color = zone)) +
   geom_point(position = position_jitterdodge(0.2), alpha = 0.25) +
@@ -96,7 +96,7 @@ ggplot(rhiz_sizes,aes(x = name, y = esd, fill = zone, color = zone)) +
 ggsave('./output/09_size-comp/size-comp.pdf',width = 85, units = 'mm')
 
 
-# |- Stats ----------------------
+## |- Stats ----------------------
 for(name in unique(rhiz_sizes$name)) {
   print(name)
   loop_data <- rhiz_sizes[which(rhiz_sizes$name == name),]

### 2024-06-26 R/plot_s03_NBSS-calcs.R

@@ -39,33 +39,111 @@ names(vol_s)[1] <- 'volume'
 
 
 #' get the nbss for each cast.
-calc_nbss <- function(df, metric = 'esd') {
+calc_nbss_width <- function(df,
+                      metric = 'esd',
+                      conv = 0.092,
+                      vol) {
   
-  size_limits <- c(min_size = min(df[[metric]]), max_size = max(df[[metric]]))
+  # assign size bins
+  # using width with prefixed sizes will make all bins the same regardless
+  # of the df input
+  size_classes <- (df[[metric]] * conv) |>
+    assign_spectra_bins('width',
+                        width = 1,
+                        min_size = size_limits[1], 
+                        max_size = size_limits[2])
   
-  df$size_class <- df[[metric]] |>
-    assign_spectra_bins('log', min_size = size_limits[1], max_size = size_limits[2])
+  ns = (table(size_classes) / vol)
   
-  db_num <- df |>
-    count_size_classes() |>
-    numeric_spectra(vol_L = 1, needs_format = T)
+  ns[which(ns == 0)] <- NA
   
-  return(db_num)
+  outdf <- data.frame(
+    size_bin = levels(size_classes),
+    count = as.vector(table(size_classes)),
+    ns = as.vector(ns)
+  ) |> 
+    bin_format(col = 'size_bin')
+  return(outdf)
 }
+
+
+calc_nbss_log <- function(df,
+                            metric = 'esd',
+                            conv = 0.092,
+                            vol) {
+  
+  # assign size bins
+  # using width with prefixed sizes will make all bins the same regardless
+  # of the df input
+  size_classes <- (df[[metric]] * conv) |>
+    assign_spectra_bins('log',
+                        min_size = size_limits[1], 
+                        max_size = size_limits[2])
+  
+  ns = (table(size_classes) / vol) |> 
+    log()
+  
+  ns[is.infinite(ns)] <- NA
+  
+  outdf <- data.frame(
+    size_bin = levels(size_classes),
+    count = as.vector(table(size_classes)),
+    ns = as.vector(ns)
+  ) |> 
+    bin_format(col = 'size_bin')
+  return(outdf)
+}
+
 
 
 # create bin sizes
[…]

### 2024-06-26 manuscript/01_intro.qmd
edit manuscript

@@
 
-Rhizaria are an extremely diverse super-group of single-celled organisms consisting of several phyla including Retaria (foraminifera and radiolaria) and Cercozoa.
-These organisms exist in a wide range of habitats and are widely represented in plankton communities throughout the global ocean.
+Rhizaria are an extremely diverse super-group of single-celled organisms which occupy a wide range of habitats.
+Planktonic Rhizaria include Foraminifera, Radiolaria (including Acantharea and Collodaria) and Phaeodaria, all of which are observed throughout the global ocean [@biard2016; @laget2024] .
 While the taxonomy of these organisms has recently undergone several reclassifications [@biard2022], their presence in ocean ecosystems has been long known to oceanographers.
@@
 Despite their noticeable morphology and global distribution, Rhizaria were largely understudied throughout the 20th century.
-The bulk of modern plankton research has focused on hard-bodied crustacea which are numerically dominant and easily sampled with nets and preservatives.
-Fragile organisms like Rhizaria were difficult to adequately study as they can be destroyed through standard zooplankton sampling techniques and preserve poorly.
-A number of studies in the late 1900s did employ alternative techniques to quantify Rhizaria including diaphragm pumps [@michaels1988] or blue-water SCUBA collections [@caron1984; @caron1995; @bijma1990].
+Fragile organisms like Rhizaria are difficult to adequately study because they can be destroyed through standard zooplankton sampling techniques and preserve poorly.
+A number of studies in the late 1900s did employ alternative techniques to quantify Rhizaria including diaphragm pumps [@michaels1988] or blue-water SCUBA collections [@caron1984; @bijma1990; @caron1995].
 However, the bulk of Rhizaria research was constrained to sediment traps or paleontological studies of sediment [@takahashi1983; @boltovskoy1993].
-Only recently has the advent of molecular techniques and in-situ imaging tools ignited a renewed focus on Rhizaria in pelagic ecosystems [@caron2016].
 
-The wave of new data on Rhizaria has facilitated an improved understanding of the significance in ocean ecosystem functions.
-Firstly, taxonomists have been able to greatly refine the understanding of evolutionary relationships amongst these diverse protists [@aurahs2009; @decelle2012; @decelle2013; @biard2015; @cavalier-smith2018; rev by @biard2022].
-DNA metabarcoding studies have revealed insights into the distributional patterns [@decelle2013; @biard2017; @marsbrisbin2020; @blanco-bercial2022; @llopismonferrer2022; @sogawa2022], ecological relationships [@decelle2012; @nakamura2023], and contribution to biogeochemical fluxes [@guidi2016; @gutierrez-rodriguez2019].
-Transcriptomic and proteomic approaches also have been used to quantify Rhizaria contribution to community metabolism [@cohen2023].
+Recently, the advent of molecular techniques and *in situ* imaging tools have ignited a renewed focus on Rhizaria in pelagic ecosystems [@caron2016].
+Molecular tools have shed light on Rhizaria taxonomy [@biard2022], distribution [@blanco-bercial2020; @sogawa2022], ecology [@decelle2012; @nakamura2023], biogeochemical roles [@gutierrez-rodriguez2019; @laget2024], and metabolism [@cohen2023].
 Yet, despite the excellent taxonomic resolution provided by molecular approaches, they do not provide a truly quantitative metric for estimating Rhizaria abundance or biomass.
-In-situ imaging tools however, offer the ability to observe organisms in the natural state and quantify their abundance [@ohman2019a; @barth-inpress].
-@biard2016 utilized in-situ imaging at a global scale to suggest Rhizaria were substantial contributors to the ocean carbon standing stock.
+*In situ* imaging tools however, offer the ability to observe organisms in the natural state and quantify their abundance [@ohman2019a; @barth2024].
+@biard2016 utilized *in situ* imaging at a global scale to suggest large (\>500$\mu$m diameter) Rhizaria were substantial contributors to the ocean carbon standing stock.
 While more recent calculations suggest lower carbon contribution [@laget2024], Rhizaria nonetheless have substantial influences on biogeochemical cycling.
-Due to their large sizes, ability to concentrate smaller particles and the unique structure of their mineral skeletons, Rhizaria have the potential to massively influence ocean biogeochemical cycling.
-A number of studies have made large advances in estimating the contribution of Rhizaria to ocean cycling of carbon [@lampitt2009; @stukel2018; @ikenoue2019; @gutierrez-rodriguez2019], silica [@biard2018; @llopismonferrer2021], and strontium [@decelle2013].
-Still, Rhizaria ecological roles are not well understood [@biard2022].
-This is a major challenge as it is critical to understand the ecological role of plankton to fully incorporate them into biological oceanographic models.
+This influence is due to their unique mineral skeletons and ability to repackage small particles into larger, fast sinking ones [@ikenoue2019].
+Thus, some Rhizaria may play substantial roles in ocean export through their interception of sinking particles [@stukel2018; @laget2024].
+However, Rhizaria ecology is poorly understood [@biard2022] and their ecological roles are likely more than diverse than simple particle interception.
 
-The ecological role of Rhizaria in plankton communities is complicated due to the fact different taxa can exhibit every different trophic modes.
-As zooplankton, rhizaria are predominately heterotrophic [@biard2022], yet their feeding modes can be quite varied.
-Phaeodarias (family Cercozoa) are largely thought to be flux-feeders, collecting and feeding on sinking particles [@nakamura2015; @stukel2019].
+The ecological role of Rhizaria in plankton communities is complicated due to the fact different taxa can exhibit very different trophic modes.
+Many Rhizaria are strictly heterotrophic [@biard2022], yet their feeding modes can be quite varied.
+Phaeodaria are largely thought to be flux-feeders, collecting and feeding on sinking particles [@gowing1989].
 Alternatively, Retaria can be either exclusively heterotrophic or mixotrophic, utilizing photosynthetic algal symbionts [@anderson2014; @decelle2015].
-Mixotrophic foraminifera host a variety of endosymbiont partners [@lee2006; @decelle2015], which are thought to support early and adult life stages and significantly contribute to total primary productivity [@kimoto2015].
-Still, foraminifera are omnivorous, possibly even predominately carnivorous, with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], mainly consuming live copepods [@caron1984].
-Radiolaria have several lineages, all of which have some taxa well known to host symbionts [@biard2022b].
-Amongst Radiolarias, arguably the most widespread are Collodaria who can be either large solitary cells or form massive colonies, up to several meters in length [@swanberg1981].
+Mixotrophic Foraminifera host a variety of endosymbiont partners [@lee2006; @decelle2015], which are thought to support early and adult life stages and significantly contribute to total primary productivity [@kimoto2015].
+Still, Foraminifera are omnivorous, possibly even predominately carnivorous, with several studies suggesting that they can be effective predators [@anderson1976; @gaskell2019], mainly consuming live copepods [@caron1984].
+Radiolaria have several lineages, many of which host symbionts [@biard2022b].
+Amongst Radiolaria, arguably the most widespread are Collodaria who can be either large solitary cells or form massive colonies, up to several meters in length [@swanberg1981].
 All known Collodaria species host dinoflagellate symbionts [@biard2022b] and can contribute substantially to primary productivity, particularly in oligotrophic ocean regions [@caron1995; @dennett2002].
@@
 A number of studies do suggest evidence for vertical zonation between Rhizaria groups according to various trophic strategies.
-Taxa-specific studies of Radiolarias suggest they may be restricted to the euphotic zone [@michaels1988; @boltovskoy2017].
-Although some studies report Acantharea in deeper waters [@decelle2013; @gutiérrez-rodríguez2022], Phaeodarias, alternatively, are generally found in the mesopelagic where photosynthesis cannot occur, but they can feed on sinking particles [@stukel2018].
+Taxon-specific studies of Radiolaria suggest they may be restricted to the euphotic zone [@michaels1988; @boltovskoy2017].
+Although some studies report Acantharea in deeper waters [@decelle2013; @gutiérrez-rodríguez2022], Phaeodaria, alternatively, are generally found beyond depths where light can sustain photosynthesis, but they can feed on sinking particles [@stukel2018; @laget2024].
 In an imaging-based study of the whole Rhizaria community, @biard2020 noted clear patterns in vertical zonation which largely corresponded to different trophic roles.
-In the oligotrophic ocean, @blanco-bercial2022 also noted that the protist community, including Rhizaria, partition along an autotroph and mixotroph to heterotroph gradient with increasing depth in the water column.
-Yet, few studies have made direct attempts to relate rhizaria abundances to environmental factors [@biard2020].
+Protists communities, including Rhizaria, have been described to partition along autotroph-mixotroph-heterotroph gradients with increasing depth in the water column [@blanco-bercial2022; @laget2024].
+Yet, few studies have made direct attempts to relate Rhizaria abundances to environmental factors [@biard2020].
 In part, this is due to the fact few studies have been able to sample Rhizaria in the same location over a consistent timeframe [@michaels1988; @boltovskoy1993; @michaels1995; @hull2011; @gutiérrez-rodríguez2022].
-Furthermore, no studies have utilized imaging, arguably the best method for quantifying rhizaria, consistently throughout the full mesopelagic.
+Furthermore, very few studies have utilized *in situ* imaging, arguably the best method for quantifying Rhizaria, to consistently throughout the full mesopelagic [@laget2024] .
 Given this lack of information, there are many unknowns with respect to Rhizaria ecology, seasonality and phenology across different groups.
 
-In this study, we present a comprehensive assessment of large Rhizaria measured for over a year from regularly occurring cruises at monthly intervals.
-We utilized an in-situ imaging approach to facilitate abundance calculations.
+In this study, we present a comprehensive assessment of large Rhizaria abundance measured for over a year from regularly occurring cruises at monthly intervals.
+We utilized an *in situ* imaging approach to facilitate abundance calculations.
 With this dataset, we address two critical aims.
-1) Quantification of large Rhizaria throughout the epipelagic (0-200m) and mesopelagic (200-1000m) over the course of an annual cycle.
+1) Quantifying of large Rhizaria throughout the epipelagic (0-200m) and mesopelagic (200-1000m) over the course of an annual cycle.
 These data were collected in the Sargasso Sea, and represents the first study of its kind in an oligotrophic system; and 2) We aim to test the hypothesis that Rhizaria exhibit niche partitioning according to trophic roles.
-This hypothesis makes several predictions, including vertical zonation, as seen in prior studies, but also that environmental variables related to trophic strategy will explain abundance patterns.
+This hypothesis has several implications, including vertical zonation, as seen in prior studies, but also that environmental variables related to trophic strategy will explain abundance patterns.
 Specifically, autotrophic/mixotrophic taxa will correspond to variables related to autotrophy (chl-a concentration, primary productivity, local DO maxima) and other rhizaria will correspond to factors which promote heterotrophy (particle concentration, flux, and local DO minima).

### 2024-06-26 manuscript/02_methods.qmd
edit manuscript

@@
 
-## Oceanographic Sampling
+## Oceanographic sampling
 
@@
 Cruises were conducted at approximately monthly intervals.
-Rhizaria individuals were sampled using the Underwater Vision Profiler 5 [UVP5; @picheral2010], a tool which is well established to accurately quantify large Rhizaria [@biard2016; @stukel2018; @stukel2019; @biard2020; @barth2022; @drago2022; @llopismonferrer2022; @panaïotis2023].
+Rhizaria individuals were sampled using the Underwater Vision Profiler 5 (UVP) [@picheral2010].
+The UVP5 is an in-situ camera used to capture plankton and particle images and has been well established to accurately quantify large Rhizaria abundance [@biard2016; @stukel2018; @stukel2019; @biard2020; @barth2022; @drago2022; @llopismonferrer2022; @panaïotis2023].
 The UVP5 was mounted to the sampling rosette and collected data autonomously on routine casts, from which only the downcast data are utilized.
-The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022, during which time the BATS region was sampled for 3-5 days at monthly intervals.
+The UVP5 was deployed from June-September 2019 then from October 2020 - January 2022, during which time the BATS region was sampled for 3-5 days at monthly intervals (see sampling details in Supplemental Table 1).
 Casts were filtered to only include data collected in the BATS region, far offshore of Bermuda in the Sargasso Sea (approximately 31.0$^oN$-32.5$^oN$, 64.25$^oW$-63$^oW$; Supplemental Figure 1).
@@
 A variety of biotic and abiotic data were collected during each BATS cruise.
-Briefly, we will explain the data utilized in this study.
 The UVP5 provided particle count data at a high-frequency from each cast.
-Particle concentration was calculated from this data for all particles 184$\mu m$ - 450$\mu m$.
-The lower size range was set by what could be reliably sampled by the UVP5's pixel resolution (\>2px; 0.092mm per pixel) and the upper size range is representative of a potential prey field for mesozooplankton [@whitmore2021].
-For each UVP cast supporting continuous profiles of the CTD parameters salinity, temperature, and auxiliary CTD channels; Dissolved Oxygen (DO), in-situ chlorophyll fluorescence were measured at 24Hz using the BATS CTD package.
+Particle concentration, as a proxy for prey field [@whitmore2021], was calculated from this data for all particles 184$\mu m$ - 900$\mu m$.
+The lower size range was set by what could be reliably sampled by the UVP5's pixel resolution (\>2px; 0.092mm per pixel) and the upper size range was set to below the size of observable Rhizaria yet large enough to be inclusive of particles on which some Rhizaria may feed.
+For each UVP cast supporting continuous profiles of the CTD parameters salinity, temperature, and auxiliary CTD channels; Dissolved Oxygen (DO), *in situ* chlorophyll fluorescence were measured at 24Hz using the BATS CTD package.
 On select casts, Niskin bottles were used to collect bacterial abundance estimates (via epifluorescence microscopy) as well as measure inorganic nutrients ($NO_3$, and $Si$ as silicate/sicilic acid) at discrete depths.
 On each cruise, flux estimates of total mass, organic carbon, and nitrogen were also collected using sediment traps; in the present study we utilized flux to the mesopelagic as the flux at 200m.
-Also primary productivity was estimated through measuring $^{14}C$ uptake rates from in-situ incubations.
+Also primary productivity was estimated through measuring $^{14}C$ uptake rates from *in situ* incubations.
 For full descriptions of the BATS sampling program and methods, see @knap1997 and @lomas2013 for a review.
@@
 
-Environmental data were processed in a variety of ways to match the format of the Rhizaria abundance estimates (see below).
+Environmental data were processed to match the format of the Rhizaria abundance estimates (see below).
 CTD data were collected at higher frequency than the UVP (24Hz vs 15Hz respectively), so these data were averaged within matching bins to the UVP5 data.
-Data from Niskin bottles were first linearly interpolated in depth at 1m resolution then time averaged over the cruise, then subsequently averaged into matching UVP5-sized bins.
-Primary productivity estimates were totaled within the euphotic zone to represent a "total euphotic productivity".
+Data from Niskin bottles were first linearly interpolated in depth at 1m resolution then time averaged over the cruise, then subsequently averaged into matching bins as the UVP5 data.
+Primary productivity estimates were depth integrated throughout the euphotic zone (0-140m).
 
-## Rhizaria imaging processing and quantification
+## Rhizaria imaging processing and abundance quantification
 
-Individual vignettes of Rhizaria images were identified using the classification platform Ecotaxa [@picheral].
+Individual vignettes of Rhizaria images were identified using the classification platform EcoTaxa [@picheral].
 Data were pre-sorted utilizing a random-forest classifier and pre-trained learning set.
-Taxonomic classification were done based on morphology exclusively.
-While there are sparse taxonomic guides for in-situ images of rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020].
-Using the aforementioned sources and publicly available ecotaxa projects, we constructed a guide accessible at: <https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf>.
-Broadly, Rhizaria were classified as Foraminifera, Radiolarias (Acantharea or Collodaria), or as a variety of Phaeodaria families (Figure 1).
+Taxonomic classification were done based on morphological parameters measured in Zooprocess [@gorsky2010], which includes parameters such as major axis, equivalent spherical diameter (ESD), and grey values.
+While there are sparse taxonomic guides for *in situ* images of Rhizaria, identification largely relied on descriptions in [@nakamura2015; @suzuki2015; and @biard2020].
+Using the aforementioned sources and publicly available EcoTaxa projects, we constructed a guide accessible at: <https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf>.
+Broadly, Rhizaria were classified as Foraminifera, Radiolaria (Acantharea or Collodaria), or as a variety of Phaeodaria families (Figure 1).
 When identification could not be confidently made between a few candidate taxa, a less specific label was used.
@@
 
-![Example images of different Rhizaria taxa. 4mm scale bar shown in lower right. All vignettes are same scale](images/01_taxa.pdf)
+![Example images of different Rhizaria taxa as collected by UVP5 imaging. 4mm scale bar shown in lower right. All vignettes are same scale](images/01_taxa.pdf)
 
-The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$.
-However, identified rhizaria ranged from a 934$\mu m$ Aulacanthidae cell to a Collodarian colony over 10mm in diameter.
-To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth-inpress].
+The UVP5 samples at \~15Hz rate as it descends the water column and records the exact position of each particle larger than 600$\mu m$ (ESD).
+However, identified rhizaria ranged from a 934$\mu m$ ESD Aulacanthidae cell to a Collodarian colony over 10mm in diameter.
+To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth2024].
 However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded.
@@
 $m^{-3}$ and 3.912 ind.
-$m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth-inpress].
-Because we typically observed many rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions.
+$m^{-3}$, in the epipelagic and mesopelagic respectively, to fall below a 10% non-detection risk [@benfield1996; @barth2024].
+Because we typically observed many Rhizaria taxa below these concentrations, we present the 25m binned data to visualize broad-scale average distributions.
 For quantifying and modelling Rhizaria abundances, we present integrated abundance estimates, with each cast.
@@
 
-## Modelling environmental controls of Rhizaria Abundance
+## Modelling environmental controls of Rhizaria abundance
 
@@
 Terms were fit using maximum likelihood with a double penalty approach on unnecessary smooths [@marra2011].
-The smoothness parameter was restricted (k = 6) to prevent overfitting the models.
-At each iteration of the backwards step-wise procedure, the model term with the lowest F score (least statistically significant) was removed.
+The smoothness parameter was restricted (k = 6) to prevent over fitting the models.
+Then, the full model was checked for concurvity (a metric indicating high parameter relatedness in GAMs).
+If two terms had a high concurvity value (\>0.8), then two alternative models were constructed each leaving out one of the related terms.
+The more parsimonious model of these two was selected using AIC.
+Once concurvity of terms was reduced, then the model was subjected to the backwards step-wise procedure.
+At each iteration, the model term with the lowest F score (least statistically significant) was removed.
 This was repeated until all model terms were statistically significant or the $R^2_{adj}$ was substantially reduced.
-Models were fit for each region; epipelagic, upper mesopelagic, and lower mesopelagic.
+Models were fit for each water column region; epipelagic, upper mesopelagic, and lower mesopelagic.
 In cases where observations were too sparse for a given taxonomic grouping, models were not run.
-All code and full models are available in code, as well as intermediate data products at <https://github.com/TheAlexBarth/RhizariaSeasonality>.
+All code, full models, and model selection tools are available in open-source scripts, as well as intermediate data products at <https://github.com/TheAlexBarth/RhizariaSeasonality>.

### 2024-06-26 manuscript/03_results.qmd
finalize output figures
edit manuscript

@@
 
-The BATS sampling region is southeast of Bermuda, situated in the oligotrophic North Atlantic Subtropical Gyre.
-Due to the sampling location, while the environmental conditions are generally low in variation and oligotrophic, there is considerable influence from deep winter-mixing and summer stratification as well as secondary influences from mesoscale eddies which result in spatiotemporal heterogeniety [@mcgillicuddy1998; @lomas2013].
+While the environmental conditions were generally low in variation and oligotrophic, there is considerable influence from deep winter-mixing and summer stratification as well as secondary influences from mesoscale eddies which result in spatiotemporal heterogeniety [@mcgillicuddy1998; @lomas2013].
 Variability in the water column structure was visible during the study period (Figure 2).
@@
 February 2021 saw a notable downwelling event, likely due to a passing anti-cyclonic eddy which impacted the local region during early 2021.
-During this phaes, warmer, oxic water was significantly depressed deeper into the mesopelagic.
+During this phase, warmer, oxic water was significantly depressed deeper into the mesopelagic.
 This process was reversed in the spring months (March, April) primary due to the interaction of convective mixing and the passing of a strong cyclonic eddy resulting in a deep cold mixed layer.
-Primary production was highest during the spring mixing period, evidenced both by in-situ fluorescence (Figure 2D) and productivity incubation experiments (Figure 3A).
+Primary production was highest during the spring mixing period, evidenced both by *in situ* fluorescence (Figure 2D) and productivity incubation experiments (Figure 3A).
 Originating near the surface, the productivity peak moved deeper throughout the spring and declined into the summer (Figure 2D).
 However, there was a notable, yet smaller productivity bump in the late summer and early fall (Figure 3A) which occurred deeper in the epipelagic (Figure 2D).
-The particle concentration (184$\mu m$ - 450$\mu m$) was closely coupled to chlorophyll-a patterns.
+The particle concentration (184$\mu m$ - 900$\mu m$) was closely coupled to chlorophyll-a patterns.
 
@@
 (C) Dissolved Oxygen.
-(D) In-situ chlorophyll fluorescence.
-(E) Particle concenctrion (184 - 450$\mu m$).
+(D) in situ chlorophyll fluorescence.
+(E) Particle concenctrion (184 - 900\$\\mu m\$).
 (F) Bacteria Abundance.
 (G) Silica.
-(H) Nitrate](images/02-environmental.pdf)
+(H) Nitrate](images/02_environmental.pdf)
 
@@
 Copepods were the second most abundant, comprising 35.5% and all other living mesozooplankton were 22%.
-The large contribution of Rhizaria to the mesozooplankton community is most prominent in the epipelagic, where they accounted for 47% of all mesozooplankton.
-In the mesopelagic rhizaria were a smaller (but still prominent fraction), at 38% in the upper layers (200-500m) and 37% in the deeper mesopelagic (500-1000m).
+The large contribution of Rhizaria to the mesozooplankton community is most prominent in the epipelagic, where they accounted for 47% of all mesozooplankton abundance.
+In the mesopelagic Rhizaria were a smaller fraction, at 38% in the upper layers (200-500m) and 37% in the deeper mesopelagic (500-1000m).
 
@@
 
-Total average Rhizaria abundance had a bimodal distribution with respect to depth.
+Average abundance of all Rhizaria had a bimodal distribution with respect to depth.
 Total abundance was highest just below the surface (0-100m), with secondary, wider peak occurring in the mid mesopelagic (Figure 4A).
@@
 The vertical distribution pattern and abundance varied considerably across taxonomic groups.
-Radiolarias were some of the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5B).
-This pattern was led by Collodaria, whose colonies were abundant in the upper epipelagic and declined into the top of the mesopelagic (Figure 4C).
+Radiolaria were some of the most abundant taxa observed, particularly in the epipelagic (Figure 4, Figure 5A).
 Acantharea displayed a bimodal distribution accounting for a large portion of the total Rhizaria pattern (Figure 4B, Figure 5).
-Foraminifera had a similar bimodal distribution, yet their overall average densities were much lower and spread wider throughout the mesopelagic (Figure 4E).
-Phaeodaria families exhibited a wide range of vertical distribution patterns.
-The most abundant, Aulacanthidae, also had a bimodal pattern but the density was highest in the lower mesopelagic (Figure 4D).
+The surface layers were largely comprised of Collodaria, whose colonies were abundant (Figure 4C).
+The most abundant Phaeodaria, Aulacanthidae, also had a bimodal pattern with the density was highest in the lower mesopelagic (Figure 4D).
+Foraminifera had a similar bimodal distribution, yet their overall average densities were much lower (Figure 4E).
 Aulosphaeridae had low average density and was nearly homogeneously distributed throughout the water column, although slightly lower in the epipelagic (Figure 4F).
@@
 
-![Seasonality of Rhizaria integrated abundance for the epipelagic (0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower mesopelagic (500-1000m) (E-F).
-Left panels (A,C,E) display total integrated abundance per monthly cruise colored by month.
-Right panels (B, D, F) display community composition of each total abundance.](images/05_seasonality.pdf)
+![Seasonality of Rhizaria integrated abundance for the epipelagic (0-200m) (A), upper mesopelagic (200-500m) (B), lower mesopelagic (500-1000m) (C).
+Data shown are the cruise mean integrated abundance in the water column region colored for each taxon.](images/05_seasonality.pdf)
 
@@
 The 2019 later summer - fall period also had much higher integrated abundance than similar months in 2021.
-While the majority of integrated abundance in the epipelagic was consistently attributable to Collodaria, Acanthrea abundance occurred sporadically and could account for a large portion of the total in some months (Figure 5B).
-The mesopelagic integrated abundance was much more consistent across monthly cruises, although average abundance was notably higher in 2019 (Figure 5C-F).
-The community composition in the mesopelagic was more diverse, mostly comprised of Phaeodarias.
-However, Acantharea and unidentified Rhizaria also were common members of the community (Figure 5D, 5F).
+While the majority of integrated abundance in the epipelagic was consistently attributable to Collodaria, Acanthrea abundance occurred sporadically and could account for a large portion of the total in some months (Figure 5A).
+The mesopelagic integrated abundance was much more consistent across monthly cruises, although average abundance was notably higher in 2019 (Figure 5B, 5C).
+The community composition in the mesopelagic was more diverse, mostly comprised of Phaeodaria families.
+However, Acantharea and unidentified Rhizaria also were common members of the community (Figure 5B, C).
 
@@
 Only Acanthrea, Foraminifera, Aulacanthidae, and Aulosphaeridae were consistently abundant in the epipelagic and mesopelagic.
-To investigate if the populations or morphologies shifted throughout the water column, we compared the sizes (Equivalent Spherical Diameters, ESD) between mesopelagic and epipelagic groups for each taxa.
+To investigate if the populations or morphologies shifted throughout the water column, we compared the sizes (ESD) between mesopelagic and epipelagic groups for each taxa.
 All groups were significantly different on average (Wilcox Rank Sum p-value \<0.001).
@@
 
-For total Rhizaria integrated abundance the GAMs produced moderate fits ($R^2_{adj}$ = 0.406-0.603) (Table 1).
-In the epipelagic, there were several significant predictor variables including inorganic nutrients ($NO_3$ and $Si$), water quality parameters (Salinity, DO), primary production, and particle concentration (Table 1).
-However, the upper and lower mesopelagic were exclusively explained by particle-related variables (concentration and mass flux) (Table 1).
+![Partial effect plots from Generalized Additive Models (GAMs) of all Rhizaria abundance.
+Models were constructed separately for depth integrated abundance through the Epipelagic (0-200m) (A-G), Upper Mesopelagic (200-500m) (H-J), and the Lower Mesopelagic (500-1000m) (K-L).
+Each panel represents the partial effect of a GAM model term, with smoothed relationship shown shaded with 95% confidence intervals.
+Individual points are depth integrated values for all Rhizaria from an individual cast, colored by month.](images/06_tot-partials.pdf)
+
+For integrated abundance of all Rhizaria, the GAMs produced moderate fits ($R^2_{adj}$ = 0.406-0.603) (Table 1).
+In the epipelagic, there were several significant predictor variables (Figure 7A-7G).
+This included terms which are likely indicators of changing water masses; Salinity (7E) and Temperature (7G).
+The model also had parameters related to both heterotrophy (Particle concentration, 7A) and autotrophy (Primary production, 7B).
+However the relationship to particle concentration is weakly increasing trend while the primary productivity effect appears to be influenced by a cruise with a high outlier for increased primary productivity.
+The models for the mesopelagic regions were much more reduced and the only significant effects were particle-related terms (particle concentration and mass flux).
+In the upper mesopelagic (Figure 7H-7J), final model included N flux, particle concentration and a non-significant effect for Silica.
+The lower mesopelagic (Figure 7K, 7L) only had model effects kept of total mass flux and particle concentration.
+For both the upper and lower mesopelagic, particle concentration had a clear, strong positive relationship (Figure 7I, 7L).
 
 | **Model**                          | Term                   | edf    | F     | p       |
-|--------------------|---------------|-------------|-------------|-------------|
+|---------------|---------------|---------------|---------------|---------------|
 | **All Rhizaria Epipelagic**        | Temperature            | 0.7950 | 0.492 | 0.0440  |
@@
 
-: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column. edf = estimated degrees of freedom, F = F-statistic, p = p-value.
+: Generalized Additive Model results for integrated total Rhizaria abundance in different regions of the water column.
+edf = estimated degrees of freedom, F = F-statistic, p = p-value.
 Models were selected using a backwards-stepwise approach with a retention threshold of p = 0.1 or a large enough reduction in model fit.
@@
 
-![Partial effects of smooth terms in taxa-specific GAM models from the epipelagic (0-200m).
-Effects are grouped by taxa; Acantharea (A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae (E).](images/06_epi-partials.pdf)
-
-Epipelagic Acantharea were explained by several predictor variables and had a good fit ($R^2_{adj}$ = 0.53, Table 2).
-Most notable smooths were mass flux and particle concentration, which had a weak positive association (Figure 7A), with July 2021 as a clear outlier where Acantharea abundances were high in the epipelagic despite lower fluxes and particle concentrations (Figure 5).
-Foraminifera had a good fitting GAM in the epipelagic ($R^2_{adj}$ = 0.445).
-There were several significant explanatory variables, although the clearest pattern was observed of high temperatures associated with more Foraminifera abundance (Table 2, Figure 7B).
+Epipelagic Acantharea were explained by several predictor variables and had a good fit ($R^2_{adj}$ = 0.539, Table 2, Supplemental FIgure 4).
+Foraminifera had a good fitting GAM in the epipelagic ($R^2_{adj}$ = 0.543).
+There were several significant explanatory variables, although the clearest pattern was observed of high temperatures associated with more Foraminifera abundance (Table 2, Supplemental Figure 4U).
 Epipelagic Aulacanthidae similarly had several predictor variables which were significant, including both water quality parameters and particle/flux predictors (Table 2).
-Interestingly, Aulacanthidae had primary production as a significant predictor, yet there was not a clear association (Figure 7C).
-There was a fit for Collodaria in the epipelagic ($R^2_{adj}$ = 0.16), although there was a logit-like relationship where higher abundances tended to occur during higher DO conditions in the surface waters (Figure 7D).
-Castanellidae also had similarly poor fits in the epipelagic ($R^2_{adj}$ = 0.124) (Table 2, Figure 7E).
+There was a weak fit for Collodaria in the epipelagic ($R^2_{adj}$ = 0.16), although there was a logit-like relationship where higher abundances tended to occur during higher DO conditions in the surface waters (Supplemental Figure 4).
+Castanellidae also had a weak model fit in the epipelagic ($R^2_{adj}$ = 0.228), although particle concentration was the strongest predictor term (Table 2, Supplemental Figure 4).
 
 | Model                                | Term                   | edf   | F     | p       |
-|--------------------------------------|------------------------|-------|-------|---------|
+|---------------|---------------|---------------|---------------|---------------|
 | **Acantharea Epipelagic**            | Temperature            | 0.868 | 0.380 | 0.0935  |
-| $R^2_{adj}$=0.539                    | Salinity               | 2.349 | 4.646 |\<0.001  |
-|                                      | O2                     | 2.501 | 5.493 |\<0.001  |
-|                                      | Avg Mass Flux          | 4.815 | 22.59 |\<0.001  |
+| $R^2_{adj}$=0.539                    | Salinity               | 2.349 | 4.646 | \<0.001 |
+|                                      | O2                     | 2.501 | 5.493 | \<0.001 |
+|                                      | Avg Mass Flux          | 4.815 | 22.59 | \<0.001 |
 |                                      | Bacteria               | 1.608 | 1.360 | 0.00945 |
@@
 | **Acantharea Upper Mesopelagic**     | Avg Mass Flux          | 1.866 | 1.220 | 0.0166  |
-| $R^2_{adj}$=0.197                    | Avg N Flux             | 2.208 | 3.100 |\<0.001  |
-|                                      | Primary Productivity   | 2.382 | 2.497 |\<0.001  |
+| $R^2_{adj}$=0.197                    | Avg N Flux             | 2.208 | 3.100 | \<0.001 |
+|                                      | Primary Productivity   | 2.382 | 2.497 | \<0.001 |
 |                                      | Particle Concentration | 0.724 | 0.521 | 0.0437  |
-| **Acantharea Lower Mesopelagic**     | Temperature            | 1.958 | 2.986 |\<0.001  |
-| $R^2_{adj}$=0.496                    | Avg F Flux             | 1.513 | 1.565 | 0.0037  |
+| **Acantharea Lower Mesopelagic**     | Temperature            | 1.958 | 2.986 | \<0.001 |
+| $R^2_{adj}$=0.496                    | Avg N Flux             | 1.513 | 1.565 | 0.0037  |
 |                                      | Primary Productivity   | 0.856 | 1.153 | 0.0045  |
-|                                      | Particle Concentration | 1.661 | 8.667 |\<0.001  |
-| **Aulacanthidae Epipelagic**         | Temperature            | 2.591 | 5.003 |\<0.001  |
+|                                      | Particle Concentration | 1.661 | 8.667 | \<0.001 |
+| **Aulacanthidae Epipelagic**         | Temperature            | 2.591 | 5.003 | \<0.001 |
 | $R^2_{adj}$=0.394                    | Salinity               | 1.523 | 1.789 | 0.002   |
@@
 |                                      | Bacteria               | 0.808 | 0.833 | 0.0160  |
-|                                      | Primary Productivity   | 2.622 | 9.160 |\<0.001  |
-|                                      | Particle Concentration | 1.937 | 2.751 |\<0.001  |
+|                                      | Primary Productivity   | 2.622 | 9.160 | \<0.001 |
+|                                      | Particle Concentration | 1.937 | 2.751 | \<0.001 |
 | **Aulacanthidae Upper Mesopelagic**  | Avg C Flux             | 1.675 | 1.408 | 0.0109  |
@@
 | **Aulacanthidae Lower Mesopelagic**  | Avg Mass Flux          | 1.476 | 1.418 | 0.0076  |
-| $R^2_{adj}$=0.308                    | Particle Concentration | 2.457 | 7.756 | <0.001  |
-| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.778 | 12.76 | <0.001  |
+| $R^2_{adj}$=0.308                    | Particle Concentration | 2.457 | 7.756 | \<0.001 |
+| **Aulosphaeridae Upper Mesopelagic** | Particle Concentration | 2.778 | 12.76 | \<0.001 |
 | $R^2_{adj}$=0.196                    |                        |       |       |         |
-| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 2.012 | 6.306 |\<0.001  |
+| **Aulosphaeridae Lower Mesopelagic** | Particle Concentration | 2.012 | 6.306 | \<0.001 |
 | $R^2_{adj}$=0.148                    |                        |       |       |         |
 | **Castanellidae Epipelagic**         | Salinity               | 0.915 | 0.400 | 0.0935  |
-| $R^2_{adj}$=0.228                    | RFU                    | 0.912 | 2.052 |\<0.001  |
-|                                      | Si                     | 2.018 | 3.656 |\<0.001  |
+| $R^2_{adj}$=0.228                    | RFU                    | 0.912 | 2.052 | \<0.001 |
+|                                      | Si                     | 2.018 | 3.656 | \<0.001 |
 |                                      | Bacteria               | 1.128 | 0.883 | 0.0170  |
-|                                      | Particle Concentration | 2.278 | 4.028 |\<0.001  |
+|                                      | Particle Concentration | 2.278 | 4.028 | \<0.001 |
 | **Coelodendridae Upper Mesopelagic** | Avg N Flux             | 0.734 | 0.551 | 0.0527  |
-| $R^2_{adj}$=0.09                     | Particle Concentration | 0.965 | 5.102 |\<0.001  |
-| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 3.051 | 6.768 |\<0.001  |
+| $R^2_{adj}$=0.09                     | Particle Concentration | 0.965 | 5.102 | \<0.001 |
+| **Coelodendridae Lower Mesopelagic** | Particle Concentration | 3.051 | 6.768 | \<0.001 |
 | $R^2_{adj}$=0.157                    |                        |       |       |         |
-| **Collodaria Epipelagic**.           | Salinity               | 0.926 | 2.267 |\<0.001  |
-| $R^2_{adj}$=0.16                     | O2                     | 2.015 | 2.217 |\<0.001  |
-|                                      | Bacteria #/L           | 1.843 | 2.100 | 0.0024  |
-| **Foraminifera Epipelagic**          | Temperature            | 4.420 | 7.624 |\<0.001  |
+| **Collodaria Epipelagic**            | Salinity               | 0.926 | 2.267 | \<0.001 |
+| $R^2_{adj}$=0.16                     | O2                     | 2.015 | 2.217 | \<0.001 |
+|                                      | Bacteria               | 1.843 | 2.100 | 0.0024  |
+| **Foraminifera Epipelagic**          | Temperature            | 4.420 | 7.624 | \<0.001 |
 | $R^2_{adj}$=0.543                    | O2                     | 0.799 | 0.788 | 0.023   |
-|                                      | Bacteria.              | 0.000 | 0.000 | 0.523   |
-|                                      | Avg N Flux             | 1.847 | 3.022 |\<0.001  |
+|                                      | Bacteria               | 0.000 | 0.000 | 0.523   |
+|                                      | Avg N Flux             | 1.847 | 3.022 | \<0.001 |
 |                                      | Primary Productivity   | 0.472 | 0.139 | 0.183   |
-|                                      | Particle Concentration | 1.685 | 2.955 |\<0.001  |
+|                                      | Particle Concentration | 1.685 | 2.955 | \<0.001 |
 
@@
 
-![Partial effects of smooth terms in taxa-specific GAM models from the upper mesopelagic (200-500m).
-Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/07_upmeso-partial.pdf)
-
 In the upper mesopelagic (200-500m), abundances were generally low (Figure 4) so GAMs were only constructed for Acantharea, Aulacanthidae, Aulosphaeridae, and Coelodendridae (Table 2).
-All these models had generally poor fits ($R^2_{adj}$ \< 0.25).
-Yet, for all upper mesopelagic models, particle concentration was a significant explanatory variable (Table 2, Figure 8).
-Carbon flux was significant for Acantharea and nitrogen flux was significant for both Acantharea and Coelodendridae (Table 2, Figure 8A,D).
-The lower mesopelagic also had generally poor GAM fits for taxa specific models ($R^2_{adj}$ \<0.3), with the exception of Acantharea ($R^2_{adj}$ = 0.509).
-Acantharea in the lower mesopelagic was most clearly positively associated with particle concentration and nitrogen flux, as well as temperature to a slight degree (Figure 9A).
-For all Phaeodarias with a significant model, particle concentration was a main predictor variable (Table 2, Figure 9B-D).
-Aulacanthidae had the best fitting model of the Phaeodarias ($R^2_{adj}$ = 0.298), which also included mass flux as a statistically significant smooth (Figure 9B).
-
-![Partial effects of smooth terms in taxa-specific GAM models from the lower mesopelagic (500-1000m).
-Effects are grouped by taxa; Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae (D).](images/08_lomeso-partials.pdf)
+All these models had generally poor fits ($R^2_{adj}$ \< 0.20).
+Yet, for all upper mesopelagic models, particle concentration was a significant explanatory variable (Table 2, Supplemental Figure 5).
+The lower mesopelagic also had generally poor GAM fits for taxon specific models ($R^2_{adj}$ \<0.31), with the exception of Acantharea ($R^2_{adj}$ = 0.496).
+Acantharea in the lower mesopelagic was most clearly positively associated with particle concentration (Supplemental Figure 6).
+For all Phaeodaria with a significant model, particle concentration was a main predictor variable (Table 2, Supplemental Figure 6).

### 2024-06-26 manuscript/04_discussion.qmd
edit manuscript

@@
 During a prior time period, @blanco-bercial2022 noted that there is considerable seasonality in the community composition of all protists.
-Despite the seasonality of total Rhizaria abundance, community composition was relatively consistent, with Collodaria representing the bulk of the community.
+Despite the seasonality of Rhizaria abundance, community composition was relatively consistent, with Collodaria representing the bulk of the community.
 It should be noted that the overall taxonomic resolution of the UVP5 is fairly low, so there may be a switching of species within the broad groups identified in this study which were not captured.
@@
 Thus, the percent contribution reported here (42.7%) of Rhizaria to the total mesozooplankton community may be inflated due to under sampling of organisms such as Euphausiids and Chaetognaths which have quick escape responses.
-Regardless, it is worth noting that in the same region, with data collected in 2012 and 2013 using similar calculation methods, @biard2016 estimated Rhizaria only contribute 15% of the total mesozooplankton community in the upper 500m.
+Regardless, it is worth noting that in the same region, with data collected in 2012 and 2013 using similar calculation methods, @biard2016 estimated large Rhizaria only contribute 15% of the total mesozooplankton community in the upper 500m.
 Likely, Rhizaria display considerable interannual variability.
 In the present study, we noticed considerably higher Rhizaria abundance throughout the water column in 2019 compared to 2021.
-While this may have been driven by increased mass flux, more information is needed to truly understand the magnitude by which Rhizaria can vary interannually.
+Similarly, in the California Current a multi-year study of Aulosphaeridae saw considerable variation between years [@biard2018].
+To truly understand the drivers of interannual variability however, sustained observations of Rhizaria over longer time periods are required.
 
@@
 However, variation with environmental parameters throughout the water column are likely, just not captured in the modelling aspect of this study.
+Future studies which are able to account for detection biases in finer vertical resolution of Rhizaria may improve ecological models of their abundance.
 One consistent parameter which had significant positive associations was particle concentration.
-This observation is not surprising as most rhizaria likely to some extent engage in flux feeding.
+This observation is not surprising as most Rhizaria likely to some extent engage in flux feeding.
 
@@
 Most prevalent in the epipelagic were Collodaria.
-These mixotrophic Radiolarians have long been reported to contribute to primary productivity in the euphotic zone [@michaels1995; @dennett2002].
+These mixotrophic Radiolaria have long been reported to contribute to primary productivity in the euphotic zone [@michaels1995; @dennett2002].
 Collodaria are thought to be particularly successful globally in oligotrophic regions due to their photosymbiotic relationships [@biard2016; @biard2017].
@@
 Furthermore, while Collodaria were primarily absent from below 250m, there were a few instances of deeper colonies being observed.
-Global investigations of polycystine flux, suggest that deep-Collodaria in Oligotrophic regions may be a consequence of isothermal submersion [@boltovskoy2017].
+Global investigations of polycystine flux, suggest that deep-Collodaria in oligotrophic regions may be a consequence of isothermal submersion [@boltovskoy2017].
 Alternatively, surface waters at BATS often mix into the mode water during the seasonal mixing, so Collodaria in the deeper waters may be a result of diapyncal mixing.
@@
 Given it's smaller size [@nakamura2015], Castanellidae does not need a large diameter to efficiently flux feed at the typically particle rich region of the lower epipelagic.
-Both Castanelldiae and Collodaria had poor fitting GAMs.
-This is somewhat of a surprise for Collodaria who had large abundance.
-However, given the consistency of their abundance, it may be that this study did not capture a wide enough range of conditions for describing Collodaria's preferred niche.
 
 The mesopelagic generally was home to known heterotrophic organisms, particularly for those which were constrained to exclusively occupy deeper waters.
-This is consistent with @blanco-bercial2022's observation of an auto-/mixotroph to heterotroph gradient in the protist community.
+This is consistent with @blanco-bercial2022's observation of an auto-/mixotroph to heterotroph gradient in local protist community as well as global patterns in Rhizaria ecology \[@laget2024\].
 The upper mesopelagic interestingly had relatively low total abundance.
@@
 So, while the base of the epipelagic may provide a rich feeding environment for Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F), the region from 200-500m might be otherwise food poor.
-Perhaps it is more advantageous for Rhizaria to situate deeper, in darker regions of the twilight zone.
+Perhaps it is more advantageous for Rhizaria to situate deeper, in darker regions of the twilight zone where visual predators and vertically migrating organisms may not feed on them.
 Also it should be noted that Phaeodaria utilize silica to build their opaline tests, and silica concentrations started to increase around 500m (Figure 2G).
-Although $Si$ was not a significant smooth for any taxa-specific model, this lack of association might be due to the overall lack of variation of $Si$ between integrated abundance of each cast.
+However, $Si$ was not a significant term for any taxon-specific model in the mesopelagic.
 Aulosphaeridae was only found to have significant relationships, although weak fits, to particle concentration in the mesopelagic.
 In our study, while consistently observed, overall abundances of Aulosphaeridae were very low.
-In the Pacific Ocean, on California's Coast, much higher abundances of Aulosphaeridae have been reported [@zasko2005; @biard2020] and they have massive potential to impact silica export [@biard2018].
+In the Pacific Ocean, on California's Coast, much higher abundances of Aulosphaeridae have been reported \[@zasko2005; @biard2020\] and they have massive potential to impact silica export [@biard2018].
 Coelodendridae were also seemingly restricted to the deeper section of the mesopelagic.
@@
 So it may be that only a few types of Coelodendridae were observed in this study, while the epipelagic variety was not.
-Alternatively, the lower epipelagic of the California Current may provide adequate habitat for Coelodendridae, which is not available in the oligotrophic Sargasso Sea.
+Alternatively, the lower epipelagic of the extremely productive California Current may provide adequate habitat for Coelodendridae, which is not available in the oligotrophic Sargasso Sea.
 
-A number of taxa were found to have a bimodal distribution, with considerable populations in both the epipelagic and mesopelagic.
+A number of taxa were found to have a bimodal distribution, with sizable populations in both the epipelagic and mesopelagic.
 Aulacanthidae had a bimodal distribution, although abundances were highest in the lower mesopelagic.
@@
 
-Acantharea also had a bimodal distribution, with much larger abundances than Aulacanthidae or Foraminifera.
+Acantharea also had a bimodal distribution, with much more sizable abundances than Aulacanthidae or Foraminifera.
 Most prior studies of Acantharea vertical distribution found them concentrated in near surface layers of the water column [@michaels1988 @zasko2005; @biard2020].
@@
 Recently @marsbrisbin2020 described apparent predator behavior amongst near-surface Acantharea.
-Thus it is likely that epipelagic Acantharea may commonly be heterotrophic.
+Thus it is likely that epipelagic Acantharea may commonly be heterotrophic or mixotrophic with an increased reliance on heterotrophy.
 Yet, it should be noted in the Sargasso Sea, both heterotrophic and symbiotic lineages of Acantharea have been reported [@blanco-bercial2022].
@@
 Although the authors also observed low abundances in the mesopelagic and noted that the smaller sizes may be due to lower food availability [@marsbrisbin2020].
-Since food is more scarce in the mesopelagi, nutritional quality lower, yet flux feeders would likely grow larger to increase their feeding range [@biard2020].
+Since food is more scarce in the mesopelagic, nutritional quality lower, yet flux feeders would likely grow larger to increase their feeding range [@stukel2018; @biard2020].
 In the data collecting in this study, Acanthareas in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth.
@@
 Considering the potential role of Rhizaria in the biological carbon pump, they may have a somewhat mixed role.
-In the shallower regions, where smaller Rhizaria are abundant, they may be an attenuating force on sinking particles [@stukel2019].
-It should be noted that in our study, we focused on the "small" particle concentration field (\<450$\mu m$), and these particles are generally slower sinking than large particles.
+In the shallower regions, they may be an attenuating force on sinking particles [@stukel2019].
 However, once consumed and repackaged by larger Rhizaria, they can sink quicker and contribute more to overall flux [@michaels1988].
 Thus, Rhizaria may act as an aggregation mechanism.
-However, this is largely speculation, to truly test this, more work is needed measuring Rhizaria flux.
+However, to truly test this, more work is needed measuring Rhizaria flux.
 
@@
 While this study represents a step-forward in our understanding of Rhizaria ecology, continued research on Rhizaria is much needed to better understand their ecology.
-Particularly extended descriptive work to capture interannual patterns.
+Particularly extended work describing interannual patterns in Rhizaria abundance.
 Also work defining biotic interactions, feeding rates, productivity, and life history are all rich fields of interest in Rhizaria.

### 2024-06-26 manuscript/main.tex
edit manuscript

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{setspace}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
 \fi
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \makeatother
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 % Add any tex header commands here
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
+
 \ifLuaTeX
@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \end{minipage}  \end{titlepage}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
 Rhizaria are an extremely diverse super-group of single-celled organisms
-consisting of several phyla including Retaria (foraminifera and
-radiolaria) and Cercozoa. These organisms exist in a wide range of
-habitats and are widely represented in plankton communities throughout
-the global ocean. While the taxonomy of these organisms has recently
-undergone several reclassifications (Biard, 2022a), their presence in
-ocean ecosystems has been long known to oceanographers. Some of the
-earliest records of their existence are from oceanographic expeditions
-in the 19th century (Haekel, 1887). Rhizaria are unique members of the
-plankton and protist community because they can reach large sizes (up to
-several mm in diameter) and they construct intricate mineral skeletons
-out of either silica, strontium, or calcium carbonate (Biard, 2022a;
-Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki and Not, 2015). Despite
-their noticeable morphology and global distribution, Rhizaria were
-largely understudied throughout the 20th century. The bulk of modern
-plankton research has focused on hard-bodied crustacea which are
-numerically dominant and easily sampled with nets and preservatives.
-Fragile organisms like Rhizaria were difficult to adequately study as
-they can be destroyed through standard zooplankton sampling techniques
-and preserve poorly. A number of studies in the late 1900s did employ
-alternative techniques to quantify Rhizaria including diaphragm pumps
-(Michaels, 1988) or blue-water SCUBA collections (Bijma et al., 1990;
-Caron et al., 1995; Caron and Be, 1984). However, the bulk of Rhizaria
-research was constrained to sediment traps or paleontological studies of
-sediment (Boltovskoy et al., 1993; Takahashi et al., 1983). Only
-recently has the advent of molecular techniques and in-situ imaging
-tools ignited a renewed focus on Rhizaria in pelagic ecosystems (Caron,
-2016).
-
-The wave of new data on Rhizaria has facilitated an improved
-understanding of the significance in ocean ecosystem functions. Firstly,
-taxonomists have been able to greatly refine the understanding of
-evolutionary relationships amongst these diverse protists (Aurahs et
-al., 2009; Biard et al., 2015; Cavalier-Smith et al., 2018; Decelle et
-al., 2013, 2012; rev by Biard, 2022a). DNA metabarcoding studies have
-revealed insights into the distributional patterns (Biard et al., 2017;
-Blanco-Bercial et al., 2022; Decelle et al., 2013; Llopis Monferrer et
-al., 2022; Mars Brisbin et al., 2020; Sogawa et al., 2022), ecological
-relationships (Decelle et al., 2012; Nakamura et al., 2023), and
-contribution to biogeochemical fluxes (Guidi et al., 2016;
-Gutierrez-Rodriguez et al., 2019). Transcriptomic and proteomic
-approaches also have been used to quantify Rhizaria contribution to
-community metabolism (Cohen et al., 2023). Yet, despite the excellent
+which occupy a wide range of habitats. Planktonic Rhizaria include
+Foraminifera, Radiolaria (including Acantharea and Collodaria) and
+Phaeodaria, all of which are observed throughout the global ocean (Biard
+et al., 2016; Laget et al., 2024) . While the taxonomy of these
+organisms has recently undergone several reclassifications (Biard,
+2022a), their presence in ocean ecosystems has been long known to
+oceanographers. Some of the earliest records of their existence are from
+oceanographic expeditions in the 19th century (Haekel, 1887). Rhizaria
+are unique members of the plankton and protist community because they
+can reach large sizes (up to several mm in diameter) and they construct
+intricate mineral skeletons out of either silica, strontium, or calcium
+carbonate (Biard, 2022a; Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki
+and Not, 2015). Despite their noticeable morphology and global
+distribution, Rhizaria were largely understudied throughout the 20th
+century. Fragile organisms like Rhizaria are difficult to adequately
+study because they can be destroyed through standard zooplankton
+sampling techniques and preserve poorly. A number of studies in the late
+1900s did employ alternative techniques to quantify Rhizaria including
+diaphragm pumps (Michaels, 1988) or blue-water SCUBA collections (Bijma
+et al., 1990; Caron et al., 1995; Caron and Be, 1984). However, the bulk
+of Rhizaria research was constrained to sediment traps or
+paleontological studies of sediment (Boltovskoy et al., 1993; Takahashi
+et al., 1983).
+
+Recently, the advent of molecular techniques and \emph{in situ} imaging
+tools have ignited a renewed focus on Rhizaria in pelagic ecosystems
+(Caron, 2016). Molecular tools have shed light on Rhizaria taxonomy
+(Biard, 2022a), distribution (Blanco-Bercial, 2020; Sogawa et al.,
+2022), ecology (Decelle et al., 2012; Nakamura et al., 2023),
+biogeochemical roles (Gutierrez-Rodriguez et al., 2019; Laget et al.,
+2024), and metabolism (Cohen et al., 2023). Yet, despite the excellent
 taxonomic resolution provided by molecular approaches, they do not
 provide a truly quantitative metric for estimating Rhizaria abundance or
-biomass. In-situ imaging tools however, offer the ability to observe
-organisms in the natural state and quantify their abundance (Barth and
-Stone, In press; Ohman, 2019). Biard et al. (2016) utilized in-situ
-imaging at a global scale to suggest Rhizaria were substantial
+biomass. \emph{In situ} imaging tools however, offer the ability to
+observe organisms in the natural state and quantify their abundance
+(Barth and Stone, 2024; Ohman, 2019). Biard et al. (2016) utilized
+\emph{in situ} imaging at a global scale to suggest large
+(\textgreater500\(\mu\)m diameter) Rhizaria were substantial
 contributors to the ocean carbon standing stock. While more recent
@@
 Rhizaria nonetheless have substantial influences on biogeochemical
-cycling. Due to their large sizes, ability to concentrate smaller
-particles and the unique structure of their mineral skeletons, Rhizaria
-have the potential to massively influence ocean biogeochemical cycling.
-A number of studies have made large advances in estimating the
-contribution of Rhizaria to ocean cycling of carbon (Gutierrez-Rodriguez
-et al., 2019; Ikenoue et al., 2019; Lampitt et al., 2009; Stukel et al.,
-2018), silica (Biard et al., 2018; Llopis Monferrer et al., 2021), and
-strontium (Decelle et al., 2013). Still, Rhizaria ecological roles are
-not well understood (Biard, 2022a). This is a major challenge as it is
-critical to understand the ecological role of plankton to fully
-incorporate them into biological oceanographic models.
+cycling. This influence is due to their unique mineral skeletons and
+ability to repackage small particles into larger, fast sinking ones
+(Ikenoue et al., 2019). Thus, some Rhizaria may play substantial roles
+in ocean export through their interception of sinking particles (Laget
+et al., 2024; Stukel et al., 2018). However, Rhizaria ecology is poorly
+understood (Biard, 2022a) and their ecological roles are likely more
+than diverse than simple particle interception.
 
 The ecological role of Rhizaria in plankton communities is complicated
-due to the fact different taxa can exhibit every different trophic
-modes. As zooplankton, rhizaria are predominately heterotrophic (Biard,
-2022a), yet their feeding modes can be quite varied. Phaeodarias (family
-Cercozoa) are largely thought to be flux-feeders, collecting and feeding
-on sinking particles (Nakamura and Suzuki, 2015; Stukel et al., 2019).
-Alternatively, Retaria can be either exclusively heterotrophic or
+due to the fact different taxa can exhibit very different trophic modes.
+Many Rhizaria are strictly heterotrophic (Biard, 2022a), yet their
+feeding modes can be quite varied. Phaeodaria are largely thought to be
+flux-feeders, collecting and feeding on sinking particles (Gowing,
+1989). Alternatively, Retaria can be either exclusively heterotrophic or
 mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
-Decelle et al., 2015). Mixotrophic foraminifera host a variety of
+Decelle et al., 2015). Mixotrophic Foraminifera host a variety of
 endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
@@
 contribute to total primary productivity (Kimoto, 2015). Still,
-foraminifera are omnivorous, possibly even predominately carnivorous,
+Foraminifera are omnivorous, possibly even predominately carnivorous,
 with several studies suggesting that they can be effective predators
 (Anderson and Bé, 1976; Gaskell et al., 2019), mainly consuming live
-copepods (Caron and Be, 1984). Radiolaria have several lineages, all of
-which have some taxa well known to host symbionts (Biard, 2022b).
-Amongst Radiolarias, arguably the most widespread are Collodaria who can
-be either large solitary cells or form massive colonies, up to several
-meters in length (Swanberg and Anderson, 1981). All known Collodaria
-species host dinoflagellate symbionts (Biard, 2022b) and can contribute
-substantially to primary productivity, particularly in oligotrophic
-ocean regions (Caron et al., 1995; Dennett et al., 2002). This
-Collodaria-symbiont association has been suggested as a reason for their
-high abundances throughout the photic zone of oligotrophic environments
-globally (Biard et al., 2017, 2016). A few Acantharea (Radiolaria order)
-clades host algal symbionts (Biard, 2022b; Decelle et al., 2012),
-notably with two clades forming an exclusive relationship with
-\emph{Phaeocystis}. However, globally, Acantharea are less abundant than
-Collodaria (Biard, 2022a) and contribute less to total primary
-productivity (Michaels et al., 1995). This may be due to the fact
-several clades of Acantharea are cyst-forming and strictly heterotrophs
-(Biard, 2022b; Decelle et al., 2013). Furthermore, Mars Brisbin et al.
-(2020) documented apparent predation behavior in Acantharea near the
-surface, suggesting that there may be a larger reliance on carnivory.
+copepods (Caron and Be, 1984). Radiolaria have several lineages, many of
+which host symbionts (Biard, 2022b). Amongst Radiolaria, arguably the
+most widespread are Collodaria who can be either large solitary cells or
+form massive colonies, up to several meters in length (Swanberg and
+Anderson, 1981). All known Collodaria species host dinoflagellate
+symbionts (Biard, 2022b) and can contribute substantially to primary
+productivity, particularly in oligotrophic ocean regions (Caron et al.,
+1995; Dennett et al., 2002). This Collodaria-symbiont association has
+been suggested as a reason for their high abundances throughout the
+photic zone of oligotrophic environments globally (Biard et al., 2017,
+2016). A few Acantharea (Radiolaria order) clades host algal symbionts
+(Biard, 2022b; Decelle et al., 2012), notably with two clades forming an
+exclusive relationship with \emph{Phaeocystis}. However, globally,
+Acantharea are less abundant than Collodaria (Biard, 2022a) and
+contribute less to total primary productivity (Michaels et al., 1995).
+This may be due to the fact several clades of Acantharea are
+cyst-forming and strictly heterotrophs (Biard, 2022b; Decelle et al.,
+2013). Furthermore, Mars Brisbin et al. (2020) documented apparent
+predation behavior in Acantharea near the surface, suggesting that there
+may be a larger reliance on carnivory.
 
@@
 zonation between Rhizaria groups according to various trophic
-strategies. Taxa-specific studies of Radiolarias suggest they may be
+strategies. Taxon-specific studies of Radiolaria suggest they may be
 restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
 Although some studies report Acantharea in deeper waters (Decelle et
-al., 2013; Gutiérrez-Rodríguez et al., 2022), Phaeodarias,
-alternatively, are generally found in the mesopelagic where
-photosynthesis cannot occur, but they can feed on sinking particles
-(Stukel et al., 2018). In an imaging-based study of the whole Rhizaria
-community, Biard and Ohman (2020) noted clear patterns in vertical
-zonation which largely corresponded to different trophic roles. In the
-oligotrophic ocean, Blanco-Bercial et al. (2022) also noted that the
-protist community, including Rhizaria, partition along an autotroph and
-mixotroph to heterotroph gradient with increasing depth in the water
-column. Yet, few studies have made direct attempts to relate rhizaria
-abundances to environmental factors (Biard and Ohman, 2020). In part,
-this is due to the fact few studies have been able to sample Rhizaria in
-the same location over a consistent timeframe (Boltovskoy et al., 1993;
-Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels et al.,
-1995; Michaels, 1988). Furthermore, no studies have utilized imaging,
-arguably the best method for quantifying rhizaria, consistently
-throughout the full mesopelagic. Given this lack of information, there
-are many unknowns with respect to Rhizaria ecology, seasonality and
-phenology across different groups.
+al., 2013; Gutiérrez-Rodríguez et al., 2022), Phaeodaria, alternatively,
+are generally found beyond depths where light can sustain
+photosynthesis, but they can feed on sinking particles (Laget et al.,
+2024; Stukel et al., 2018). In an imaging-based study of the whole
+Rhizaria community, Biard and Ohman (2020) noted clear patterns in
+vertical zonation which largely corresponded to different trophic roles.
+Protists communities, including Rhizaria, have been described to
+partition along autotroph-mixotroph-heterotroph gradients with
+increasing depth in the water column (Blanco-Bercial et al., 2022; Laget
+et al., 2024). Yet, few studies have made direct attempts to relate
+Rhizaria abundances to environmental factors (Biard and Ohman, 2020). In
+part, this is due to the fact few studies have been able to sample
+Rhizaria in the same location over a consistent timeframe (Boltovskoy et
+al., 1993; Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels
+et al., 1995; Michaels, 1988). Furthermore, very few studies have
+utilized \emph{in situ} imaging, arguably the best method for
+quantifying Rhizaria, to consistently throughout the full mesopelagic
+(Laget et al., 2024) . Given this lack of information, there are many
+unknowns with respect to Rhizaria ecology, seasonality and phenology
+across different groups.
 
 In this study, we present a comprehensive assessment of large Rhizaria
-measured for over a year from regularly occurring cruises at monthly
-intervals. We utilized an in-situ imaging approach to facilitate
-abundance calculations. With this dataset, we address two critical aims.
-1) Quantification of large Rhizaria throughout the epipelagic (0-200m)
-and mesopelagic (200-1000m) over the course of an annual cycle. These
-data were collected in the Sargasso Sea, and represents the first study
-of its kind in an oligotrophic system; and 2) We aim to test the
-hypothesis that Rhizaria exhibit niche partitioning according to trophic
-roles. This hypothesis makes several predictions, including vertical
-zonation, as seen in prior studies, but also that environmental
-variables related to trophic strategy will explain abundance patterns.
-Specifically, autotrophic/mixotrophic taxa will correspond to variables
-related to autotrophy (chl-a concentration, primary productivity, local
-DO maxima) and other rhizaria will correspond to factors which promote
-heterotrophy (particle concentration, flux, and local DO minima).
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{oceanographic-sampling}{%
-\subsection{Oceanographic Sampling}\label{oceanographic-sampling}}
+abundance measured for over a year from regularly occurring cruises at
+monthly intervals. We utilized an \emph{in situ} imaging approach to
+facilitate abundance calculations. With this dataset, we address two
+critical aims. 1) Quantifying of large Rhizaria throughout the
+epipelagic (0-200m) and mesopelagic (200-1000m) over the course of an
+annual cycle. These data were collected in the Sargasso Sea, and
+represents the first study of its kind in an oligotrophic system; and 2)
+We aim to test the hypothesis that Rhizaria exhibit niche partitioning
+according to trophic roles. This hypothesis has several implications,
+including vertical zonation, as seen in prior studies, but also that
+environmental variables related to trophic strategy will explain
+abundance patterns. Specifically, autotrophic/mixotrophic taxa will
+correspond to variables related to autotrophy (chl-a concentration,
+primary productivity, local DO maxima) and other rhizaria will
+correspond to factors which promote heterotrophy (particle
+concentration, flux, and local DO minima).
+
+\section{Methods}\label{methods}
+
+\subsection{Oceanographic sampling}\label{oceanographic-sampling}
 
@@
 monthly intervals. Rhizaria individuals were sampled using the
-Underwater Vision Profiler 5 {[}UVP5; Picheral et al. (2010){]}, a tool
-which is well established to accurately quantify large Rhizaria (Barth
-and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago et
-al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023; Stukel
-et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the sampling
-rosette and collected data autonomously on routine casts, from which
-only the downcast data are utilized. The UVP5 was deployed from
+Underwater Vision Profiler 5 (UVP) (Picheral et al., 2010). The UVP5 is
+an in-situ camera used to capture plankton and particle images and has
+been well established to accurately quantify large Rhizaria abundance
+(Barth and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago
+et al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023;
+Stukel et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the
+sampling rosette and collected data autonomously on routine casts, from
+which only the downcast data are utilized. The UVP5 was deployed from
 June-September 2019 then from October 2020 - January 2022, during which
-time the BATS region was sampled for 3-5 days at monthly intervals.
-Casts were filtered to only include data collected in the BATS region,
-far offshore of Bermuda in the Sargasso Sea (approximately
-31.0\(^oN\)-32.5\(^oN\), 64.25\(^oW\)-63\(^oW\); Supplemental Figure 1).
-In general, casts extended to either 200m, 500m, or 1200m deep, with a
-few extended into the bathypelagic (4500m). However, Rhizaria were only
-typically found in large abundances throughout the epipelagic and
-mesopelagic zones. As such, we limit this study to results from the
-upper 1000m of the water column.
+time the BATS region was sampled for 3-5 days at monthly intervals (see
+sampling details in Supplemental Table 1). Casts were filtered to only
+include data collected in the BATS region, far offshore of Bermuda in
+the Sargasso Sea (approximately 31.0\(^oN\)-32.5\(^oN\),
+64.25\(^oW\)-63\(^oW\); Supplemental Figure 1). In general, casts
+extended to either 200m, 500m, or 1200m deep, with a few extended into
+the bathypelagic (4500m). However, Rhizaria were only typically found in
+large abundances throughout the epipelagic and mesopelagic zones. As
+such, we limit this study to results from the upper 1000m of the water
+column.
 
 A variety of biotic and abiotic data were collected during each BATS
-cruise. Briefly, we will explain the data utilized in this study. The
-UVP5 provided particle count data at a high-frequency from each cast.
-Particle concentration was calculated from this data for all particles
-184\(\mu m\) - 450\(\mu m\). The lower size range was set by what could
+cruise. The UVP5 provided particle count data at a high-frequency from
+each cast. Particle concentration, as a proxy for prey field (Whitmore
+and Ohman, 2021), was calculated from this data for all particles
+184\(\mu m\) - 900\(\mu m\). The lower size range was set by what could
 be reliably sampled by the UVP5's pixel resolution (\textgreater2px;
-0.092mm per pixel) and the upper size range is representative of a
-potential prey field for mesozooplankton (Whitmore and Ohman, 2021). For
-each UVP cast supporting continuous profiles of the CTD parameters
-salinity, temperature, and auxiliary CTD channels; Dissolved Oxygen
-(DO), in-situ chlorophyll fluorescence were measured at 24Hz using the
-BATS CTD package. On select casts, Niskin bottles were used to collect
-bacterial abundance estimates (via epifluorescence microscopy) as well
-as measure inorganic nutrients (\(NO_3\), and \(Si\) as silicate/sicilic
-acid) at discrete depths. On each cruise, flux estimates of total mass,
-organic carbon, and nitrogen were also collected using sediment traps;
-in the present study we utilized flux to the mesopelagic as the flux at
-200m. Also primary productivity was estimated through measuring
-\(^{14}C\) uptake rates from in-situ incubations. For full descriptions
-of the BATS sampling program and methods, see Knap et al. (1997) and
-Lomas et al. (2013) for a review. Additionally, data can be viewed
-online
+0.092mm per pixel) and the upper size range was set to below the size of
+observable Rhizaria yet large enough to be inclusive of particles on
+which some Rhizaria may feed. For each UVP cast supporting continuous
+profiles of the CTD parameters salinity, temperature, and auxiliary CTD
+channels; Dissolved Oxygen (DO), \emph{in situ} chlorophyll fluorescence
+were measured at 24Hz using the BATS CTD package. On select casts,
+Niskin bottles were used to collect bacterial abundance estimates (via
+epifluorescence microscopy) as well as measure inorganic nutrients
+(\(NO_3\), and \(Si\) as silicate/sicilic acid) at discrete depths. On
+each cruise, flux estimates of total mass, organic carbon, and nitrogen
+were also collected using sediment traps; in the present study we
+utilized flux to the mesopelagic as the flux at 200m. Also primary
+productivity was estimated through measuring \(^{14}C\) uptake rates
+from \emph{in situ} incubations. For full descriptions of the BATS
+sampling program and methods, see Knap et al. (1997) and Lomas et al.
+(2013) for a review. Additionally, data can be viewed online
 \href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}.
 
-Environmental data were processed in a variety of ways to match the
-format of the Rhizaria abundance estimates (see below). CTD data were
-collected at higher frequency than the UVP (24Hz vs 15Hz respectively),
-so these data were averaged within matching bins to the UVP5 data. Data
-from Niskin bottles were first linearly interpolated in depth at 1m
-resolution then time averaged over the cruise, then subsequently
-averaged into matching UVP5-sized bins. Primary productivity estimates
-were totaled within the euphotic zone to represent a ``total euphotic
-productivity''.
-
-\hypertarget{rhizaria-imaging-processing-and-quantification}{%
-\subsection{Rhizaria imaging processing and
-quantification}\label{rhizaria-imaging-processing-and-quantification}}
+Environmental data were processed to match the format of the Rhizaria
+abundance estimates (see below). CTD data were collected at higher
+frequency than the UVP (24Hz vs 15Hz respectively), so these data were
+averaged within matching bins to the UVP5 data. Data from Niskin bottles
+were first linearly interpolated in depth at 1m resolution then time
+averaged over the cruise, then subsequently averaged into matching bins
+as the UVP5 data. Primary productivity estimates were depth integrated
+throughout the euphotic zone (0-140m).
+
+\subsection{Rhizaria imaging processing and abundance
+quantification}\label{rhizaria-imaging-processing-and-abundance-quantification}
 
 Individual vignettes of Rhizaria images were identified using the
-classification platform Ecotaxa (Picheral et al., n.d.). Data were
+classification platform EcoTaxa (Picheral et al., n.d.). Data were
 pre-sorted utilizing a random-forest classifier and pre-trained learning
-set. Taxonomic classification were done based on morphology exclusively.
-While there are sparse taxonomic guides for in-situ images of rhizaria,
-identification largely relied on descriptions in (Nakamura and Suzuki,
-2015; Suzuki and Not, 2015; and Biard and Ohman, 2020). Using the
-aforementioned sources and publicly available ecotaxa projects, we
-constructed a guide accessible at:
+set. Taxonomic classification were done based on morphological
+parameters measured in Zooprocess (Gorsky et al., 2010), which includes
+parameters such as major axis, equivalent spherical diameter (ESD), and
+grey values. While there are sparse taxonomic guides for \emph{in situ}
+images of Rhizaria, identification largely relied on descriptions in
+(Nakamura and Suzuki, 2015; Suzuki and Not, 2015; and Biard and Ohman,
+2020). Using the aforementioned sources and publicly available EcoTaxa
+projects, we constructed a guide accessible at:
 \url{https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf}.
-Broadly, Rhizaria were classified as Foraminifera, Radiolarias
+Broadly, Rhizaria were classified as Foraminifera, Radiolaria
 (Acantharea or Collodaria), or as a variety of Phaeodaria families
@@
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-\caption{Example images of different Rhizaria taxa. 4mm scale bar shown
-in lower right. All vignettes are same scale}
-
-\end{figure}
+\caption{Example images of different Rhizaria taxa as collected by UVP5
+imaging. 4mm scale bar shown in lower right. All vignettes are same
+scale}
+
+\end{figure}%
 
@@
 column and records the exact position of each particle larger than
-600\(\mu m\). However, identified rhizaria ranged from a 934\(\mu m\)
-Aulacanthidae cell to a Collodarian colony over 10mm in diameter. To
-confirm that the UVP5 was sampling adequately across all size ranges, an
-normalized biomass size spectrum (NBSS) slope was constructed to
-identify a drop-off which would indicate poor-sampling at the small size
-range (Barth and Stone, In press; Lombard et al., 2019). However, it was
-evident from this analysis that all size ranges were adequately sampled
-across the size range (Supplemental Figure 2) so no data were excluded.
-The UVP5 reports the exact depth at which a particle is recorded,
-however to estimate abundance, observations must be binned over fixed
-depth intervals. Our deployments had variable descent depths and speeds
-with more casts descending to 500m than 1000m and descents quicker
-through the epipelagic than the mesopelagic (see Barth and Stone (2022)
-for an extended discussion of UVP5 data processing). For the present
-study, Rhizaria abundances were estimated in 25m vertical bins, which
-offer a moderate sampling volume per bin (average 0.948\(m^3\) in the
-epipelagic and 0.589\(m^3\) in the mesopelagic; Supplemental Table 1)
-while still maintaining ecologically relevant widths. However,
+600\(\mu m\) (ESD). However, identified rhizaria ranged from a
+934\(\mu m\) ESD Aulacanthidae cell to a Collodarian colony over 10mm in
+diameter. To confirm that the UVP5 was sampling adequately across all
+size ranges, an normalized biomass size spectrum (NBSS) slope was
+constructed to identify a drop-off which would indicate poor-sampling at
+the small size range (Barth and Stone, 2024; Lombard et al., 2019).
+However, it was evident from this analysis that all size ranges were
+adequately sampled across the size range (Supplemental Figure 2) so no
+data were excluded. The UVP5 reports the exact depth at which a particle
+is recorded, however to estimate abundance, observations must be binned
+over fixed depth intervals. Our deployments had variable descent depths
+and speeds with more casts descending to 500m than 1000m and descents
+quicker through the epipelagic than the mesopelagic (see Barth and Stone
+(2022) for an extended discussion of UVP5 data processing). For the
+present study, Rhizaria abundances were estimated in 25m vertical bins,
+which offer a moderate sampling volume per bin (average 0.948\(m^3\) in
+the epipelagic and 0.589\(m^3\) in the mesopelagic; Supplemental Table
+1) while still maintaining ecologically relevant widths. However,
 concentrations in a 25m bin would need to be greater than 2.428 ind.
@@
 respectively, to fall below a 10\% non-detection risk (Barth and Stone,
-In press; Benfield et al., 1996). Because we typically observed many
-rhizaria taxa below these concentrations, we present the 25m binned data
+2024; Benfield et al., 1996). Because we typically observed many
+Rhizaria taxa below these concentrations, we present the 25m binned data
 to visualize broad-scale average distributions. For quantifying and
@@
 
-\hypertarget{modelling-environmental-controls-of-rhizaria-abundance}{%
 \subsection{Modelling environmental controls of Rhizaria
-Abundance}\label{modelling-environmental-controls-of-rhizaria-abundance}}
+abundance}\label{modelling-environmental-controls-of-rhizaria-abundance}
 
@@
 (Marra and Wood, 2011). The smoothness parameter was restricted (k = 6)
-to prevent overfitting the models. At each iteration of the backwards
-step-wise procedure, the model term with the lowest F score (least
-statistically significant) was removed. This was repeated until all
-model terms were statistically significant or the \(R^2_{adj}\) was
-substantially reduced. Models were fit for each region; epipelagic,
-upper mesopelagic, and lower mesopelagic. In cases where observations
-were too sparse for a given taxonomic grouping, models were not run. All
-code and full models are available in code, as well as intermediate data
-products at \url{https://github.com/TheAlexBarth/RhizariaSeasonality}.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{environmental-variability}{%
+to prevent over fitting the models. Then, the full model was checked for
+concurvity (a metric indicating high parameter relatedness in GAMs). If
+two terms had a high concurvity value (\textgreater0.8), then two
+alternative models were constructed each leaving out one of the related
+terms. The more parsimonious model of these two was selected using AIC.
+Once concurvity of terms was reduced, then the model was subjected to
+the backwards step-wise procedure. At each iteration, the model term
+with the lowest F score (least statistically significant) was removed.
+This was repeated until all model terms were statistically significant
+or the \(R^2_{adj}\) was substantially reduced. Models were fit for each
+water column region; epipelagic, upper mesopelagic, and lower
+mesopelagic. In cases where observations were too sparse for a given
+taxonomic grouping, models were not run. All code, full models, and
+model selection tools are available in open-source scripts, as well as
+intermediate data products at
+\url{https://github.com/TheAlexBarth/RhizariaSeasonality}.
+
+\section{Results}\label{results}
+
 \subsubsection{Environmental
-Variability}\label{environmental-variability}}
-
-The BATS sampling region is southeast of Bermuda, situated in the
-oligotrophic North Atlantic Subtropical Gyre. Due to the sampling
-location, while the environmental conditions are generally low in
-variation and oligotrophic, there is considerable influence from deep
-winter-mixing and summer stratification as well as secondary influences
-from mesoscale eddies which result in spatiotemporal heterogeniety
-(Lomas et al., 2013; McGillicuddy et al., 1998). Variability in the
-water column structure was visible during the study period (Figure 2).
-This is best evidenced through the temperature profiles; In the late
-summer and early fall there was a stratified water column with high
-temperatures in the surface (\textless75m) (Figure 2A) and slightly
-elevated salinity (Figure 2B). This warm, stratified period appeared
-more intense during the few months sampled in 2019. In 2021, we observed
-the stratified layer slowly dissipated into the winter months following
-mixed layer entrainment. There was a consistent oxygen minimum zone
-(OMZ) located at about 800m deep (Figure 2C). February 2021 saw a
-notable downwelling event, likely due to a passing anti-cyclonic eddy
-which impacted the local region during early 2021. During this phaes,
-warmer, oxic water was significantly depressed deeper into the
-mesopelagic. This process was reversed in the spring months (March,
-April) primary due to the interaction of convective mixing and the
-passing of a strong cyclonic eddy resulting in a deep cold mixed layer.
-Primary production was highest during the spring mixing period,
-evidenced both by in-situ fluorescence (Figure 2D) and productivity
-incubation experiments (Figure 3A). Originating near the surface, the
-productivity peak moved deeper throughout the spring and declined into
-the summer (Figure 2D). However, there was a notable, yet smaller
-productivity bump in the late summer and early fall (Figure 3A) which
-occurred deeper in the epipelagic (Figure 2D). The particle
-concentration (184\(\mu m\) - 450\(\mu m\)) was closely coupled to
-chlorophyll-a patterns.
-
-\begin{figure}
+Variability}\label{environmental-variability}
+
+While the environmental conditions were generally low in variation and
+oligotrophic, there is considerable influence from deep winter-mixing
+and summer stratification as well as secondary influences from mesoscale
+eddies which result in spatiotemporal heterogeniety (Lomas et al., 2013;
+McGillicuddy et al., 1998). Variability in the water column structure
+was visible during the study period (Figure 2). This is best evidenced
+through the temperature profiles; In the late summer and early fall
+there was a stratified water column with high temperatures in the
+surface (\textless75m) (Figure 2A) and slightly elevated salinity
+(Figure 2B). This warm, stratified period appeared more intense during
+the few months sampled in 2019. In 2021, we observed the stratified
+layer slowly dissipated into the winter months following mixed layer
+entrainment. There was a consistent oxygen minimum zone (OMZ) located at
+about 800m deep (Figure 2C). February 2021 saw a notable downwelling
+event, likely due to a passing anti-cyclonic eddy which impacted the
+local region during early 2021. During this phase, warmer, oxic water
+was significantly depressed deeper into the mesopelagic. This process
+was reversed in the spring months (March, April) primary due to the
+interaction of convective mixing and the passing of a strong cyclonic
+eddy resulting in a deep cold mixed layer. Primary production was
+highest during the spring mixing period, evidenced both by \emph{in
+situ} fluorescence (Figure 2D) and productivity incubation experiments
+(Figure 3A). Originating near the surface, the productivity peak moved
+deeper throughout the spring and declined into the summer (Figure 2D).
+However, there was a notable, yet smaller productivity bump in the late
+summer and early fall (Figure 3A) which occurred deeper in the
+epipelagic (Figure 2D). The particle concentration (184\(\mu m\) -
+900\(\mu m\)) was closely coupled to chlorophyll-a patterns.
+
+\begin{figure}[H]
 
@@
 axis shows depth in meters. (A) Temperature. (B) Salinity. (C) Dissolved
-Oxygen. (D) In-situ chlorophyll fluorescence. (E) Particle concenctrion
-(184 - 450\(\mu m\)). (F) Bacteria Abundance. (G) Silica. (H) Nitrate}
-
-\end{figure}
+Oxygen. (D) in situ chlorophyll fluorescence. (E) Particle concenctrion
+(184 - 900\$\textbackslash mu m\$). (F) Bacteria Abundance. (G) Silica.
+(H) Nitrate}
+
+\end{figure}%
 
@@
 
-\end{figure}
+\end{figure}%
 
@@
 
-\hypertarget{rhizaria-abundance-and-distribution}{%
 \subsubsection{Rhizaria abundance and
-distribution}\label{rhizaria-abundance-and-distribution}}
+distribution}\label{rhizaria-abundance-and-distribution}
 
@@
 Rhizaria to the mesozooplankton community is most prominent in the
-epipelagic, where they accounted for 47\% of all mesozooplankton. In the
-mesopelagic rhizaria were a smaller (but still prominent fraction), at
-38\% in the upper layers (200-500m) and 37\% in the deeper mesopelagic
+epipelagic, where they accounted for 47\% of all mesozooplankton
+abundance. In the mesopelagic Rhizaria were a smaller fraction, at 38\%
+in the upper layers (200-500m) and 37\% in the deeper mesopelagic
 (500-1000m).
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-\end{figure}
-
-Total average Rhizaria abundance had a bimodal distribution with respect
-to depth. Total abundance was highest just below the surface (0-100m),
-with secondary, wider peak occurring in the mid mesopelagic (Figure 4A).
-Variation in depth binned abundance was large, likely due to seasonal
-variability but also increased from the detection-risk described in the
-methods. The vertical distribution pattern and abundance varied
-considerably across taxonomic groups. Radiolarias were some of the most
-abundant taxa observed, particularly in the epipelagic (Figure 4, Figure
-5B). This pattern was led by Collodaria, whose colonies were abundant in
-the upper epipelagic and declined into the top of the mesopelagic
-(Figure 4C). Acantharea displayed a bimodal distribution accounting for
-a large portion of the total Rhizaria pattern (Figure 4B, Figure 5).
-Foraminifera had a similar bimodal distribution, yet their overall
-average densities were much lower and spread wider throughout the
-mesopelagic (Figure 4E). Phaeodaria families exhibited a wide range of
-vertical distribution patterns. The most abundant, Aulacanthidae, also
-had a bimodal pattern but the density was highest in the lower
-mesopelagic (Figure 4D). Aulosphaeridae had low average density and was
-nearly homogeneously distributed throughout the water column, although
-slightly lower in the epipelagic (Figure 4F). Castanellidae were the
-only Phaeodaria who appeared to be effectively restricted to the photic
-zone (Figure 4G). Alternatively, Coelodendridae primarily occurred in
-the lower mesopelagic (Figure 4H). A few individuals from the families
+\end{figure}%
+
+Average abundance of all Rhizaria had a bimodal distribution with
+respect to depth. Total abundance was highest just below the surface
+(0-100m), with secondary, wider peak occurring in the mid mesopelagic
+(Figure 4A). Variation in depth binned abundance was large, likely due
+to seasonal variability but also increased from the detection-risk
+described in the methods. The vertical distribution pattern and
+abundance varied considerably across taxonomic groups. Radiolaria were
+some of the most abundant taxa observed, particularly in the epipelagic
+(Figure 4, Figure 5A). Acantharea displayed a bimodal distribution
+accounting for a large portion of the total Rhizaria pattern (Figure 4B,
+Figure 5). The surface layers were largely comprised of Collodaria,
+whose colonies were abundant (Figure 4C). The most abundant Phaeodaria,
+Aulacanthidae, also had a bimodal pattern with the density was highest
+in the lower mesopelagic (Figure 4D). Foraminifera had a similar bimodal
+distribution, yet their overall average densities were much lower
+(Figure 4E). Aulosphaeridae had low average density and was nearly
+homogeneously distributed throughout the water column, although slightly
+lower in the epipelagic (Figure 4F). Castanellidae were the only
+Phaeodaria who appeared to be effectively restricted to the photic zone
+(Figure 4G). Alternatively, Coelodendridae primarily occurred in the
+lower mesopelagic (Figure 4H). A few individuals from the families
 Tuscaroridae and Medusettidae were also observed in the mesopelagic, yet
@@
 \caption{Seasonality of Rhizaria integrated abundance for the epipelagic
-(0-200m) (A-B), upper mesopelagic (200-500m) (C-D), lower mesopelagic
-(500-1000m) (E-F). Left panels (A,C,E) display total integrated
-abundance per monthly cruise colored by month. Right panels (B, D, F)
-display community composition of each total abundance.}
-
-\end{figure}
+(0-200m) (A), upper mesopelagic (200-500m) (B), lower mesopelagic
+(500-1000m) (C). Data shown are the cruise mean integrated abundance in
+the water column region colored for each taxon.}
+
+\end{figure}%
 
@@
 abundance occurred sporadically and could account for a large portion of
-the total in some months (Figure 5B). The mesopelagic integrated
+the total in some months (Figure 5A). The mesopelagic integrated
 abundance was much more consistent across monthly cruises, although
-average abundance was notably higher in 2019 (Figure 5C-F). The
+average abundance was notably higher in 2019 (Figure 5B, 5C). The
 community composition in the mesopelagic was more diverse, mostly
-comprised of Phaeodarias. However, Acantharea and unidentified Rhizaria
-also were common members of the community (Figure 5D, 5F).
-
-\hypertarget{body-size-throughout-the-water-column.}{%
+comprised of Phaeodaria families. However, Acantharea and unidentified
+Rhizaria also were common members of the community (Figure 5B, C).
+
 \subsubsection{Body size throughout the water
-column.}\label{body-size-throughout-the-water-column.}}
+column.}\label{body-size-throughout-the-water-column.}
 
@@
 if the populations or morphologies shifted throughout the water column,
-we compared the sizes (Equivalent Spherical Diameters, ESD) between
-mesopelagic and epipelagic groups for each taxa. All groups were
-significantly different on average (Wilcox Rank Sum p-value
-\textless0.001). Acantharea were smaller, on average in the mesopelagic
-while all other taxa tended to be larger (Figure 6).
-
-\begin{figure}
+we compared the sizes (ESD) between mesopelagic and epipelagic groups
+for each taxa. All groups were significantly different on average
+(Wilcox Rank Sum p-value \textless0.001). Acantharea were smaller, on
+average in the mesopelagic while all other taxa tended to be larger
+(Figure 6).
+
+\begin{figure}[H]
 
@@
 
-\end{figure}
-
-\hypertarget{environmental-drivers-of-rhizaria-abundance}{%
+\end{figure}%
+
 \subsubsection{Environmental Drivers of Rhizaria
-Abundance}\label{environmental-drivers-of-rhizaria-abundance}}
-
-For total Rhizaria integrated abundance the GAMs produced moderate fits
-(\(R^2_{adj}\) = 0.406-0.603) (Table 1). In the epipelagic, there were
-several significant predictor variables including inorganic nutrients
-(\(NO_3\) and \(Si\)), water quality parameters (Salinity, DO), primary
-production, and particle concentration (Table 1). However, the upper and
-lower mesopelagic were exclusively explained by particle-related
-variables (concentration and mass flux) (Table 1).
+Abundance}\label{environmental-drivers-of-rhizaria-abundance}
+
+For integrated abundance of all Rhizaria, the GAMs produced moderate
+fits (\(R^2_{adj}\) = 0.406-0.603) (Table 1). In the epipelagic, there
+were several significant predictor variables (Figure 7A-7G). This
+included terms which are likely indicators of changing water masses;
+Salinity (7E) and Temperature (7G). The model also had parameters
+related to both heterotrophy (Particle concentration, 7A) and autotrophy
+(Primary production, 7B). However the relationship to particle
+concentration is weakly increasing trend while the primary productivity
+effect appears to be influenced by a cruise with a high outlier for
+increased primary productivity. The models for the mesopelagic regions
+were much more reduced and the only significant effects were
+particle-related terms (particle concentration and mass flux). In the
+upper mesopelagic (Figure 7H-7J), final model included N flux, particle
+concentration and a non-significant effect for Silica. The lower
+mesopelagic (Figure 7K, 7L) only had model effects kept of total mass
+flux and particle concentration. For both the upper and lower
+mesopelagic, particle concentration had a clear, strong positive
+relationship (Figure 7I, 7L).
 
@@
 \caption{Generalized Additive Model results for integrated total
-Rhizaria abundance in different regions of the water
-column.}\tabularnewline
-\toprule()
+Rhizaria abundance in different regions of the water column. edf =
+estimated degrees of freedom, F = F-statistic, p = p-value. Models were
+selected using a backwards-stepwise approach with a retention threshold
+of p = 0.1 or a large enough reduction in model fit.}\tabularnewline
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endhead
-\textbf{All Rhizaria Epipelagic} & Salinity & 2.0818 & 4.654 &
+\bottomrule\noalign{}
+\endlastfoot
+\textbf{All Rhizaria Epipelagic} & Temperature & 0.7950 & 0.492 &
+0.0440 \\
+\(R^2_{adj}\)= 0.48 & Salinity & 2.4117 & 5.113 & \textless0.001 \\
+& Silica & 2.1511 & 8.601 & \textless0.001 \\
+& Bacteria Concentration & 0.8870 & 1.544 & \textless0.001 \\
+& Avg Mass Flux & 1.4107 & 0.802 & 0.0211 \\
+& Primary Productivity & 1.9280 & 2.736 & \textless0.001 \\
+& Particle Concentration & 0.9151 & 2.136 & \textless0.001 \\
+\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6702 & 0.404 &
+0.0814 \\
+\(R^2_{adj}\)= 0.40 & Avg N Flux & 2.9524 & 3.172 & \textless0.001 \\
+& Particle Concentration & 3.5568 & 11.82 & \textless0.001 \\
+\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 2.036 & 2.953
+& \textless0.001 \\
+\(R^2_{adj}\)= 0.61 & Particle Concentration & 3.819 & 27.43 &
 \textless0.001 \\
-\(R^2_{adj}\)= 0.42 & DO & 0.8272 & 0.945 & 0.0136 \\
-& Silica & 3.1206 & 15.75 & \textless0.001 \\
-& NO3 & 2.9970 & 6.268 & \textless0.001 \\
-& Primary Productivity & 1.8367 & 2.099 & \textless0.001 \\
-& Particle Concentration & 0.9282 & 2.566 & \textless0.001 \\
-\textbf{All Rhizaria Upper Mesopelagic} & Silica & 0.6847 & 0.431 &
-0.073 \\
-\(R^2_{adj}\)= 0.406 & Avg Mass Flux & 1.5426 & 0.923 & 0.045 \\
-& Particle Concentration & 3.2957 & 20.89 & \textless0.001 \\
-\textbf{All Rhizaria Lower Mesopelagic} & Avg Mass Flux & 1.6632 & 1.824
-& 0.002 \\
-\(R^2_{adj}\)= 0.603 & Particle Concentration & 0.7694 & 0.662 &
-0.027 \\
-\bottomrule()
 \end{longtable}
@@
 
-\begin{figure}
-
-{\centering \includegraphics{images/06_epi-partials.pdf}
-
-}
-
-\caption{Partial effects of smooth terms in taxa-specific GAM models
-from the epipelagic (0-200m). Effects are grouped by taxa; Acantharea
-(A), Foraminifera (B), Aulacanthidae (C), Collodaria (D), Castanellidae
-(E).}
-
-\end{figure}
-
 Epipelagic Acantharea were explained by several predictor variables and
-had a good fit (\(R^2_{adj}\) = 0.53, Table 2). Most notable smooths
-were mass flux and particle concentration, which had a weak positive
-association (Figure 7A), with July 2021 as a clear outlier where
-Acantharea abundances were high in the epipelagic despite lower fluxes
-and particle concentrations (Figure 5). Foraminifera had a good fitting
-GAM in the epipelagic (\(R^2_{adj}\) = 0.445). There were several
-significant explanatory variables, although the clearest pattern was
-observed of high temperatures associated with more Foraminifera
-abundance (Table 2, Figure 7B). Epipelagic Aulacanthidae similarly had
-several predictor variables which were significant, including both water
-quality parameters and particle/flux predictors (Table 2).
-Interestingly, Aulacanthidae had primary production as a significant
-predictor, yet there was not a clear association (Figure 7C). There was
-a fit for Collodaria in the epipelagic (\(R^2_{adj}\) = 0.16), although
-there was a logit-like relationship where higher abundances tended to
-occur during higher DO conditions in the surface waters (Figure 7D).
-Castanellidae also had similarly poor fits in the epipelagic
-(\(R^2_{adj}\) = 0.124) (Table 2, Figure 7E).
+had a good fit (\(R^2_{adj}\) = 0.539, Table 2, Supplemental FIgure 4).
+Foraminifera had a good fitting GAM in the epipelagic (\(R^2_{adj}\) =
+0.543). There were several significant explanatory variables, although
+the clearest pattern was observed of high temperatures associated with
+more Foraminifera abundance (Table 2, Supplemental Figure 4U).
+Epipelagic Aulacanthidae similarly had several predictor variables which
+were significant, including both water quality parameters and
+particle/flux predictors (Table 2). There was a weak fit for Collodaria
+in the epipelagic (\(R^2_{adj}\) = 0.16), although there was a
+logit-like relationship where higher abundances tended to occur during
+higher DO conditions in the surface waters (Supplemental Figure 4).
+Castanellidae also had a weak model fit in the epipelagic (\(R^2_{adj}\)
+= 0.228), although particle concentration was the strongest predictor
+term (Table 2, Supplemental Figure 4).
 
@@
 \caption{Taxa-specific generalized additive models for different regions
-of the water column.}\tabularnewline
-\toprule()
+of the water column.edf = estimated degrees of freedom, F = F-statistic,
+p = p-value. Models were selected using a backwards-stepwise approach
+with a retention threshold of p = 0.1 or a large enough reduction in
+model fit.}\tabularnewline
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endhead
-\textbf{Acantharea Epipelagic} & Salinity & 2.264 & 4.113 &
+\bottomrule\noalign{}
+\endlastfoot
+\textbf{Acantharea Epipelagic} & Temperature & 0.868 & 0.380 & 0.0935 \\
+\(R^2_{adj}\)=0.539 & Salinity & 2.349 & 4.646 & \textless0.001 \\
+& O2 & 2.501 & 5.493 & \textless0.001 \\
+& Avg Mass Flux & 4.815 & 22.59 & \textless0.001 \\
+& Bacteria & 1.608 & 1.360 & 0.00945 \\
+& Particle Concentration & 0.877 & 1.414 & 0.00449 \\
+\textbf{Acantharea Upper Mesopelagic} & Avg Mass Flux & 1.866 & 1.220 &
+0.0166 \\
+\(R^2_{adj}\)=0.197 & Avg N Flux & 2.208 & 3.100 & \textless0.001 \\
+& Primary Productivity & 2.382 & 2.497 & \textless0.001 \\
+& Particle Concentration & 0.724 & 0.521 & 0.0437 \\
+\textbf{Acantharea Lower Mesopelagic} & Temperature & 1.958 & 2.986 &
 \textless0.001 \\
-R2adj=0.53 & O2 & 2.579 & 6.712 & \textless0.001 \\
-& Avg Mass Flux & 4.810 & 22.81 & \textless0.001 \\
-& Bacteria \#/L & 1.599 & 1.317 & 0.0117 \\
-& Particle Concentration & 0.853 & 1.158 & 0.0087 \\
-\textbf{Acantharea Upper Mesopelagic} & Avg C Flux & 1.296 & 0.621 &
-0.0439 \\
-R2adj=0.231 & Avg N Flux & 1.619 & 1.283 & 0.0026 \\
-& Particle Concentration & 0.952 & 3.886 & \textless0.001 \\
-\textbf{Acantharea Lower Mesopelagic} & Temperature & 2.076 & 4.155 &
+\(R^2_{adj}\)=0.496 & Avg N Flux & 1.513 & 1.565 & 0.0037 \\
+& Primary Productivity & 0.856 & 1.153 & 0.0045 \\
+& Particle Concentration & 1.661 & 8.667 & \textless0.001 \\
+\textbf{Aulacanthidae Epipelagic} & Temperature & 2.591 & 5.003 &
 \textless0.001 \\
-R2adj=0.509 & Avg N Flux & 1.494 & 1.216 & 0.0113 \\
-& Primary Productivity & 0.766 & 0.648 & 0.0253 \\
-& Particle Concentration & 2.037 & 10.03 & \textless0.001 \\
-\textbf{Aulacanthidae Epipelagic} & Salinity & 0.792 & 0.757 & 0.0241 \\
-R2adj=0.251 & Avg N Flux & 0.869 & 1.324 & 0.0018 \\
-& Primary Productivity & 2.312 & 7.704 & \textless0.001 \\
-& Particle Concentration & 2.008 & 2.200 & 0.0017 \\
-\textbf{Aulacanthidae Upper Mesopelagic} & Particle Concentration &
-2.832 & 9.802 & \textless0.001 \\
-R2adj = 0.158 & & & & \\
-\textbf{Aulacanthidae Lower Mesopelagic} & Avg Mass Flux & 1.622 & 1.991
-& 0.002 \\
-R2adj=0.298 & Particle Concentration & 2.123 & 6.706 & \textless0.001 \\
+\(R^2_{adj}\)=0.394 & Salinity & 1.523 & 1.789 & 0.002 \\
+& RFU & 0.872 & 1.315 & 0.0042 \\
+& Bacteria & 0.808 & 0.833 & 0.0160 \\
+& Primary Productivity & 2.622 & 9.160 & \textless0.001 \\
+& Particle Concentration & 1.937 & 2.751 & \textless0.001 \\
+\textbf{Aulacanthidae Upper Mesopelagic} & Avg C Flux & 1.675 & 1.408 &
+0.0109 \\
+\(R^2_{adj}\)=0.132 & Particle Concentration & 2.308 & 2.476 & 0.0013 \\
+\textbf{Aulacanthidae Lower Mesopelagic} & Avg Mass Flux & 1.476 & 1.418
+& 0.0076 \\
+\(R^2_{adj}\)=0.308 & Particle Concentration & 2.457 & 7.756 &
+\textless0.001 \\
 \textbf{Aulosphaeridae Upper Mesopelagic} & Particle Concentration &
-2.653 & 13.06 & \textless0.001 \\
-R2adj=0.2 & & & & \\
+2.778 & 12.76 & \textless0.001 \\
+\(R^2_{adj}\)=0.196 & & & & \\
 \textbf{Aulosphaeridae Lower Mesopelagic} & Particle Concentration &
-0.972 & 6.248 & \textless0.001 \\
-R2adj=.147 & & & & \\
-\textbf{Castanellidae Epipelagic} & Avg C Flux & 1.462 & 0.816 &
-0.0421 \\
-R2adj = 0.124 & Primary Productivity & 0.771 & 0.665 & 0.0247 \\
-& Particle Concentration & 3.956 & 4.623 & \textless0.001 \\
-\textbf{Coelodendridae Upper Mesopelagic} & Avg N Flux & 0.822 & 0.919 &
-0.0183 \\
-R2adj=.113 & Particle Concentration & 0.970 & 6.208 & \textless0.001 \\
+2.012 & 6.306 & \textless0.001 \\
+\(R^2_{adj}\)=0.148 & & & & \\
+\textbf{Castanellidae Epipelagic} & Salinity & 0.915 & 0.400 & 0.0935 \\
+\(R^2_{adj}\)=0.228 & RFU & 0.912 & 2.052 & \textless0.001 \\
+& Si & 2.018 & 3.656 & \textless0.001 \\
+& Bacteria & 1.128 & 0.883 & 0.0170 \\
+& Particle Concentration & 2.278 & 4.028 & \textless0.001 \\
+\textbf{Coelodendridae Upper Mesopelagic} & Avg N Flux & 0.734 & 0.551 &
+0.0527 \\
+\(R^2_{adj}\)=0.09 & Particle Concentration & 0.965 & 5.102 &
+\textless0.001 \\
 \textbf{Coelodendridae Lower Mesopelagic} & Particle Concentration &
-1.773 & 4.873 & \textless0.001 \\
-R2adj=.133 & & & & \\
-\textbf{Collodaria Epipelagic} & Salinity & 0.925 & 2.267 &
+3.051 & 6.768 & \textless0.001 \\
+\(R^2_{adj}\)=0.157 & & & & \\
+\textbf{Collodaria Epipelagic} & Salinity & 0.926 & 2.267 &
 \textless0.001 \\
-R2adj=0.16 & O2 & 2.015 & 2.217 & 0.002 \\
-& Bacteria \#/L & 1.843 & 2.100 & 0.002 \\
-\textbf{Foraminifera Epipelagic} & Temperature & 4.414 & 8.789 &
+\(R^2_{adj}\)=0.16 & O2 & 2.015 & 2.217 & \textless0.001 \\
+& Bacteria & 1.843 & 2.100 & 0.0024 \\
+\textbf{Foraminifera Epipelagic} & Temperature & 4.420 & 7.624 &
 \textless0.001 \\
-R2adj=0.445 & O2 & 1.603 & 1.287 & 0.0111 \\
-& Avg N Flux & 2.512 & 3.396 & \textless0.001 \\
-& Primary Productivity & 0.824 & 0.920 & 0.0153 \\
-& Particle Concentration & 0.919 & 2.257 & \textless0.001 \\
-\bottomrule()
+\(R^2_{adj}\)=0.543 & O2 & 0.799 & 0.788 & 0.023 \\
+& Bacteria & 0.000 & 0.000 & 0.523 \\
+& Avg N Flux & 1.847 & 3.022 & \textless0.001 \\
+& Primary Productivity & 0.472 & 0.139 & 0.183 \\
+& Particle Concentration & 1.685 & 2.955 & \textless0.001 \\
 \end{longtable}
-
-\begin{figure}
-
-{\centering \includegraphics{images/07_upmeso-partial.pdf}
-
-}
-
-\caption{Partial effects of smooth terms in taxa-specific GAM models
-from the upper mesopelagic (200-500m). Effects are grouped by taxa;
-Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
-(D).}
-
-\end{figure}
 
@@
 Aulosphaeridae, and Coelodendridae (Table 2). All these models had
-generally poor fits (\(R^2_{adj}\) \textless{} 0.25). Yet, for all upper
+generally poor fits (\(R^2_{adj}\) \textless{} 0.20). Yet, for all upper
 mesopelagic models, particle concentration was a significant explanatory
-variable (Table 2, Figure 8). Carbon flux was significant for Acantharea
-and nitrogen flux was significant for both Acantharea and Coelodendridae
-(Table 2, Figure 8A,D). The lower mesopelagic also had generally poor
-GAM fits for taxa specific models (\(R^2_{adj}\) \textless0.3), with the
-exception of Acantharea (\(R^2_{adj}\) = 0.509). Acantharea in the lower
-mesopelagic was most clearly positively associated with particle
-concentration and nitrogen flux, as well as temperature to a slight
-degree (Figure 9A). For all Phaeodarias with a significant model,
-particle concentration was a main predictor variable (Table 2, Figure
-9B-D). Aulacanthidae had the best fitting model of the Phaeodarias
-(\(R^2_{adj}\) = 0.298), which also included mass flux as a
-statistically significant smooth (Figure 9B).
-
-\begin{figure}
-
-{\centering \includegraphics{images/08_lomeso-partials.pdf}
-
-}
-
-\caption{Partial effects of smooth terms in taxa-specific GAM models
-from the lower mesopelagic (500-1000m). Effects are grouped by taxa;
-Acantharea (A), Aulacanthidae (B), Aulosphaeridae (C), Coelodendridae
-(D).}
-
-\end{figure}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\hypertarget{overall-rhizaria-abundance-and-patterns}{%
+variable (Table 2, Supplemental Figure 5). The lower mesopelagic also
+had generally poor GAM fits for taxon specific models (\(R^2_{adj}\)
+\textless0.31), with the exception of Acantharea (\(R^2_{adj}\) =
+0.496). Acantharea in the lower mesopelagic was most clearly positively
+associated with particle concentration (Supplemental Figure 6). For all
+Phaeodaria with a significant model, particle concentration was a main
+predictor variable (Table 2, Supplemental Figure 6).
+
+\section{Discussion}\label{discussion}
+
 \subsubsection{Overall Rhizaria abundance and
-patterns}\label{overall-rhizaria-abundance-and-patterns}}
+patterns}\label{overall-rhizaria-abundance-and-patterns}
 
@@
 noted that there is considerable seasonality in the community
-composition of all protists. Despite the seasonality of total Rhizaria
+composition of all protists. Despite the seasonality of Rhizaria
 abundance, community composition was relatively consistent, with
@@
 noting that in the same region, with data collected in 2012 and 2013
-using similar calculation methods, Biard et al. (2016) estimated
+using similar calculation methods, Biard et al. (2016) estimated large
 Rhizaria only contribute 15\% of the total mesozooplankton community in
@@
 Rhizaria abundance throughout the water column in 2019 compared to 2021.
-While this may have been driven by increased mass flux, more information
-is needed to truly understand the magnitude by which Rhizaria can vary
-interannually.
-
-\hypertarget{relationship-to-environmental-parameters}{%
+Similarly, in the California Current a multi-year study of
+Aulosphaeridae saw considerable variation between years (Biard et al.,
+2018). To truly understand the drivers of interannual variability
+however, sustained observations of Rhizaria over longer time periods are
+required.
+
 \subsubsection{Relationship to environmental
-parameters}\label{relationship-to-environmental-parameters}}
+parameters}\label{relationship-to-environmental-parameters}
 
@@
 parameters throughout the water column are likely, just not captured in
-the modelling aspect of this study. One consistent parameter which had
-significant positive associations was particle concentration. This
-observation is not surprising as most rhizaria likely to some extent
-engage in flux feeding.
-
-\hypertarget{vertical-structure-and-trophic-roles}{%
+the modelling aspect of this study. Future studies which are able to
+account for detection biases in finer vertical resolution of Rhizaria
+may improve ecological models of their abundance. One consistent
+parameter which had significant positive associations was particle
+concentration. This observation is not surprising as most Rhizaria
+likely to some extent engage in flux feeding.
+
 \subsubsection{Vertical Structure and Trophic
-Roles}\label{vertical-structure-and-trophic-roles}}
+Roles}\label{vertical-structure-and-trophic-roles}
 
@@
 Most prevalent in the epipelagic were Collodaria. These mixotrophic
-Radiolarians have long been reported to contribute to primary
-productivity in the euphotic zone (Dennett et al., 2002; Michaels et
-al., 1995). Collodaria are thought to be particularly successful
-globally in oligotrophic regions due to their photosymbiotic
-relationships (Biard et al., 2017, 2016). We observed the highest
-abundance of Collodaria during June 2021, supporting the notion they can
-thrive during the typically low-nutrient conditions of summer
-stratification. However, Collodaria also increased during the spring
-mixing period, suggesting that they can thrive during conditions which
-may typically be thought to favor autotrophs. Furthermore, while
-Collodaria were primarily absent from below 250m, there were a few
-instances of deeper colonies being observed. Global investigations of
-polycystine flux, suggest that deep-Collodaria in Oligotrophic regions
-may be a consequence of isothermal submersion (Boltovskoy, 2017).
-Alternatively, surface waters at BATS often mix into the mode water
-during the seasonal mixing, so Collodaria in the deeper waters may be a
-result of diapyncal mixing. Another effectively exclusively epipelagic
-Rhizaria was the Phaeodaria family of Castanellidae. All Phaeodaria are
-thought to be fully heterotrophic (Nakamura and Suzuki, 2015),
-nonetheless a number of studies, including this one, report
-Castanellidae to be typically found in the lower epipelagic (Biard et
-al., 2018; Biard and Ohman, 2020; Zasko and Rusanov, 2005). It should be
-considered that perhaps Castanellidae specializes in feeding on sinking
-particles directly at the base of the epipelagic. Given it's smaller
-size (Nakamura and Suzuki, 2015), Castanellidae does not need a large
-diameter to efficiently flux feed at the typically particle rich region
-of the lower epipelagic. Both Castanelldiae and Collodaria had poor
-fitting GAMs. This is somewhat of a surprise for Collodaria who had
-large abundance. However, given the consistency of their abundance, it
-may be that this study did not capture a wide enough range of conditions
-for describing Collodaria's preferred niche.
+Radiolaria have long been reported to contribute to primary productivity
+in the euphotic zone (Dennett et al., 2002; Michaels et al., 1995).
+Collodaria are thought to be particularly successful globally in
+oligotrophic regions due to their photosymbiotic relationships (Biard et
+al., 2017, 2016). We observed the highest abundance of Collodaria during
+June 2021, supporting the notion they can thrive during the typically
+low-nutrient conditions of summer stratification. However, Collodaria
+also increased during the spring mixing period, suggesting that they can
+thrive during conditions which may typically be thought to favor
+autotrophs. Furthermore, while Collodaria were primarily absent from
+below 250m, there were a few instances of deeper colonies being
+observed. Global investigations of polycystine flux, suggest that
+deep-Collodaria in oligotrophic regions may be a consequence of
+isothermal submersion (Boltovskoy, 2017). Alternatively, surface waters
+at BATS often mix into the mode water during the seasonal mixing, so
+Collodaria in the deeper waters may be a result of diapyncal mixing.
+Another effectively exclusively epipelagic Rhizaria was the Phaeodaria
+family of Castanellidae. All Phaeodaria are thought to be fully
+heterotrophic (Nakamura and Suzuki, 2015), nonetheless a number of
+studies, including this one, report Castanellidae to be typically found
+in the lower epipelagic (Biard et al., 2018; Biard and Ohman, 2020;
+Zasko and Rusanov, 2005). It should be considered that perhaps
+Castanellidae specializes in feeding on sinking particles directly at
+the base of the epipelagic. Given it's smaller size (Nakamura and
+Suzuki, 2015), Castanellidae does not need a large diameter to
+efficiently flux feed at the typically particle rich region of the lower
+epipelagic.
 
@@
 deeper waters. This is consistent with Blanco-Bercial et al. (2022)'s
-observation of an auto-/mixotroph to heterotroph gradient in the protist
-community. The upper mesopelagic interestingly had relatively low total
-abundance. This low-abundance region likely reflects the dynamics of
-productivity and export throughout the water column. While productivity
-and thus sinking particles for flux feeders are high in the euphotic
-zone, much of this is attenuated throughout the epipelagic. So, while
-the base of the epipelagic may provide a rich feeding environment for
-Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F),
-the region from 200-500m might be otherwise food poor. Perhaps it is
-more advantageous for Rhizaria to situate deeper, in darker regions of
-the twilight zone. Also it should be noted that Phaeodaria utilize
-silica to build their opaline tests, and silica concentrations started
-to increase around 500m (Figure 2G). Although \(Si\) was not a
-significant smooth for any taxa-specific model, this lack of association
-might be due to the overall lack of variation of \(Si\) between
-integrated abundance of each cast. Aulosphaeridae was only found to have
+observation of an auto-/mixotroph to heterotroph gradient in local
+protist community as well as global patterns in Rhizaria ecology
+{[}Laget et al. (2024){]}. The upper mesopelagic interestingly had
+relatively low total abundance. This low-abundance region likely
+reflects the dynamics of productivity and export throughout the water
+column. While productivity and thus sinking particles for flux feeders
+are high in the euphotic zone, much of this is attenuated throughout the
+epipelagic. So, while the base of the epipelagic may provide a rich
+feeding environment for Castanellidae, smaller protists, or
+heterotrophic bacteria (Figure 2F), the region from 200-500m might be
+otherwise food poor. Perhaps it is more advantageous for Rhizaria to
+situate deeper, in darker regions of the twilight zone where visual
+predators and vertically migrating organisms may not feed on them. Also
+it should be noted that Phaeodaria utilize silica to build their opaline
+tests, and silica concentrations started to increase around 500m (Figure
+2G). However, \(Si\) was not a significant term for any taxon-specific
+model in the mesopelagic. Aulosphaeridae was only found to have
 significant relationships, although weak fits, to particle concentration
@@
 California's Coast, much higher abundances of Aulosphaeridae have been
-reported (Biard and Ohman, 2020; Zasko and Rusanov, 2005) and they have
-massive potential to impact silica export (Biard et al., 2018).
+reported {[}Zasko and Rusanov (2005); Biard and Ohman (2020){]} and they
+have massive potential to impact silica export (Biard et al., 2018).
 Coelodendridae were also seemingly restricted to the deeper section of
@@
 observed in this study, while the epipelagic variety was not.
-Alternatively, the lower epipelagic of the California Current may
-provide adequate habitat for Coelodendridae, which is not available in
-the oligotrophic Sargasso Sea.
-
-A number of taxa were found to have a bimodal distribution, with
-considerable populations in both the epipelagic and mesopelagic.
-Aulacanthidae had a bimodal distribution, although abundances were
-highest in the lower mesopelagic. Foraminifera also had a bimodal
-distribution. Some lineages of Foraminifera are known to host
-photosymbionts (Biard, 2022a; Kimoto, 2015), however they are also
-efficient predators commonly seen throughout the mesopelagic (Caron and
-Be, 1984; Gaskell et al., 2019). Thus it is not surprising to find their
-presence in both locations of the water column. Foraminifera are also
-known to vary their vertical distribution across their life cycle in
-phase with lunar cycles (Biard, 2022a; Bijma et al., 1990; Gaskell et
-al., 2019; Kimoto, 2015). However, the sampling scheme of the BATS
-program does not capture this frequency and was not investigated in the
-present study.
-
-Acantharea also had a bimodal distribution, with much larger abundances
-than Aulacanthidae or Foraminifera. Most prior studies of Acantharea
-vertical distribution found them concentrated in near surface layers of
-the water column Zasko and Rusanov (2005). This would support the
-paradigm that large Acantharea abundances may be supported by their
+Alternatively, the lower epipelagic of the extremely productive
+California Current may provide adequate habitat for Coelodendridae,
+which is not available in the oligotrophic Sargasso Sea.
+
+A number of taxa were found to have a bimodal distribution, with sizable
+populations in both the epipelagic and mesopelagic. Aulacanthidae had a
+bimodal distribution, although abundances were highest in the lower
+mesopelagic. Foraminifera also had a bimodal distribution. Some lineages
+of Foraminifera are known to host photosymbionts (Biard, 2022a; Kimoto,
+2015), however they are also efficient predators commonly seen
+throughout the mesopelagic (Caron and Be, 1984; Gaskell et al., 2019).
+Thus it is not surprising to find their presence in both locations of
+the water column. Foraminifera are also known to vary their vertical
+distribution across their life cycle in phase with lunar cycles (Biard,
+2022a; Bijma et al., 1990; Gaskell et al., 2019; Kimoto, 2015). However,
+the sampling scheme of the BATS program does not capture this frequency
+and was not investigated in the present study.
+
+Acantharea also had a bimodal distribution, with much more sizable
+abundances than Aulacanthidae or Foraminifera. Most prior studies of
+Acantharea vertical distribution found them concentrated in near surface
+layers of the water column Zasko and Rusanov (2005). This would support
+the paradigm that large Acantharea abundances may be supported by their
 mixotrophic abilities (Michaels et al., 1995; Suzuki and Not, 2015).
@@
 Acantharea. Thus it is likely that epipelagic Acantharea may commonly be
-heterotrophic. Yet, it should be noted in the Sargasso Sea, both
-heterotrophic and symbiotic lineages of Acantharea have been reported
-(Blanco-Bercial et al., 2022). Additionally, Michaels (1988) noted that
-the majority of Acantharea (by abundance) were smaller than
-160\(\mu m\). While that estimate may be inflated due to inability to
-capture larger cells, small Acantharea were not captured in the present
-study. Thus, trophic strategy may shift based on sizes of Acantharea.
+heterotrophic or mixotrophic with an increased reliance on heterotrophy.
+Yet, it should be noted in the Sargasso Sea, both heterotrophic and
+symbiotic lineages of Acantharea have been reported (Blanco-Bercial et
+al., 2022). Additionally, Michaels (1988) noted that the majority of
+Acantharea (by abundance) were smaller than 160\(\mu m\). While that
+estimate may be inflated due to inability to capture larger cells, small
+Acantharea were not captured in the present study. Thus, trophic
+strategy may shift based on sizes of Acantharea.
 
@@
 availability (Mars Brisbin et al., 2020). Since food is more scarce in
-the mesopelagi, nutritional quality lower, yet flux feeders would likely
-grow larger to increase their feeding range (Biard and Ohman, 2020). In
-the data collecting in this study, Acanthareas in the mesopelagic were
-significantly smaller than the epipelagic, despite the other bimodal
-taxa (Foraminifera and Aulacanthidae) being significantly larger with
-depth. This provides added support for the hypothesis that cyst-forming
-Acantharea may utilize different sections of the water column throughout
-their life cycle. However to further investigate this, more work is
-needed with higher temporal and taxonomic resolution.
-
-\hypertarget{conclusions-and-considerations}{%
+the mesopelagic, nutritional quality lower, yet flux feeders would
+likely grow larger to increase their feeding range (Biard and Ohman,
+2020; Stukel et al., 2018). In the data collecting in this study,
+Acanthareas in the mesopelagic were significantly smaller than the
+epipelagic, despite the other bimodal taxa (Foraminifera and
+Aulacanthidae) being significantly larger with depth. This provides
+added support for the hypothesis that cyst-forming Acantharea may
+utilize different sections of the water column throughout their life
+cycle. However to further investigate this, more work is needed with
+higher temporal and taxonomic resolution.
+
 \subsubsection{Conclusions and
-Considerations}\label{conclusions-and-considerations}}
+Considerations}\label{conclusions-and-considerations}
 
@@
 biological carbon pump, they may have a somewhat mixed role. In the
-shallower regions, where smaller Rhizaria are abundant, they may be an
-attenuating force on sinking particles (Stukel et al., 2019). It should
-be noted that in our study, we focused on the ``small'' particle
-concentration field (\textless450\(\mu m\)), and these particles are
-generally slower sinking than large particles. However, once consumed
-and repackaged by larger Rhizaria, they can sink quicker and contribute
-more to overall flux (Michaels, 1988). Thus, Rhizaria may act as an
-aggregation mechanism. However, this is largely speculation, to truly
-test this, more work is needed measuring Rhizaria flux.
+shallower regions, they may be an attenuating force on sinking particles
+(Stukel et al., 2019). However, once consumed and repackaged by larger
+Rhizaria, they can sink quicker and contribute more to overall flux
+(Michaels, 1988). Thus, Rhizaria may act as an aggregation mechanism.
+However, to truly test this, more work is needed measuring Rhizaria
+flux.
 
@@
 much needed to better understand their ecology. Particularly extended
-descriptive work to capture interannual patterns. Also work defining
-biotic interactions, feeding rates, productivity, and life history are
-all rich fields of interest in Rhizaria.
-
-\hypertarget{references}{%
-\section{References:}\label{references}}
-
-\hypertarget{refs}{}
+work describing interannual patterns in Rhizaria abundance. Also work
+defining biotic interactions, feeding rates, productivity, and life
+history are all rich fields of interest in Rhizaria.
+
+\section{References:}\label{references}
+
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-anderson2014}{}}%
+\bibitem[\citeproctext]{ref-anderson2014}
 Anderson, O.R., 2014. Living Together in the Plankton: A Survey of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-anderson1976}{}}%
+\bibitem[\citeproctext]{ref-anderson1976}
 Anderson, O.R., Bé, A.W.H., 1976. A CYTOCHEMICAL FINE STRUCTURE STUDY OF
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-aurahs2009}{}}%
-Aurahs, R., Grimm, G.W., Hemleben, V., Hemleben, C., Kucera, M., 2009.
-Geographical distribution of cryptic genetic types in the planktonic
-foraminifer Globigerinoides ruber. Molecular Ecology 18, 1692--1706.
-\url{https://doi.org/10.1111/j.1365-294X.2009.04136.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-barth-inpress}{}}%
-Barth, A., Stone, J., In press. Understanding the picture: The promis
-and challenges of in-situ imagery data in the study of plankton ecology.
-Journal of Plankton Research: Horizons.
-
-\leavevmode\vadjust pre{\hypertarget{ref-barth2022}{}}%
+\bibitem[\citeproctext]{ref-barth2024}
+Barth, A., Stone, J., 2024. Understanding the picture: The promis and
+challenges of in-situ imagery data in the study of plankton ecology.
+Journal of Plankton Research: Horizons 1--15.
+\url{https://doi.org/10.1093/plankt/fbae023}
+
+\bibitem[\citeproctext]{ref-barth2022}
 Barth, A., Stone, J., 2022.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-benfield1996}{}}%
+\bibitem[\citeproctext]{ref-benfield1996}
 Benfield, M.C., Davis, C.S., Wiebe, P.H., Gallager, S.M., Gregory Lough,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
+\bibitem[\citeproctext]{ref-biard2022}
 Biard, T., 2022a. Rhizaria. Encyclopeida of Life Sciences, pp. 1--11.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biard2022b}{}}%
+\bibitem[\citeproctext]{ref-biard2022b}
 Biard, T., 2022b. Diversity and ecology of Radiolaria in modern oceans.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biard2017}{}}%
+\bibitem[\citeproctext]{ref-biard2017}
 Biard, T., Bigeard, E., Audic, S., Poulain, J., Gutierrez-Rodriguez, A.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biard2018}{}}%
+\bibitem[\citeproctext]{ref-biard2018}
 Biard, T., Krause, J.W., Stukel, M.R., Ohman, M.D., 2018. The
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biard2020}{}}%
+\bibitem[\citeproctext]{ref-biard2020}
 Biard, T., Ohman, M.D., 2020. Vertical niche definition of test-bearing
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-biard2015}{}}%
-Biard, T., Pillet, L., Decelle, J., Poirier, C., Suzuki, N., Not, F.,
-2015. Towards an integrative morpho-molecular classification of the
-collodaria (polycystinea, radiolaria). Protist 166, 374--388.
-\url{https://doi.org/10.1016/j.protis.2015.05.002}
-
-\leavevmode\vadjust pre{\hypertarget{ref-biard2016}{}}%
+\bibitem[\citeproctext]{ref-biard2016}
 Biard, T., Stemmann, L., Picheral, M., Mayot, N., Vandromme, P., Hauss,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bijma1990}{}}%
+\bibitem[\citeproctext]{ref-bijma1990}
 Bijma, J., Erez, J., Hemleben, C., 1990.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-blanco-bercial2022}{}}%
+\bibitem[\citeproctext]{ref-blanco-bercial2020}
+Blanco-Bercial, L., 2020.
+\href{https://www.frontiersin.org/articles/10.3389/fmars.2020.00173}{Metabarcoding
+analyses and seasonality of the zooplankton community at BATS}.
+Frontiers in Marine Science 7.
+
+\bibitem[\citeproctext]{ref-blanco-bercial2022}
 Blanco-Bercial, L., Parsons, R., Bolaños, L.M., Johnson, R., Giovannoni,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy2017}{}}%
+\bibitem[\citeproctext]{ref-boltovskoy2017}
 Boltovskoy, D., 2017. Vertical distribution patterns of radiolaria
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy1993}{}}%
+\bibitem[\citeproctext]{ref-boltovskoy1993}
 Boltovskoy, D., Alder, V.A., Abelmann, A., 1993. Annual flux of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-caron2016}{}}%
+\bibitem[\citeproctext]{ref-caron2016}
 Caron, D.A., 2016.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-caron1984}{}}%
+\bibitem[\citeproctext]{ref-caron1984}
 Caron, D.A., Be, A.W.H., 1984. PREDICTED AND OBSERVED FEEDING RATES OF
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-caron1995}{}}%
+\bibitem[\citeproctext]{ref-caron1995}
 Caron, D.A., Michaels, A.F., Swanberg, N.R., Howse, F.A., 1995. Primary
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cavalier-smith2018}{}}%
-Cavalier-Smith, T., Chao, E.E., Lewis, R., 2018. Multigene phylogeny and
-cell evolution of chromist infrakingdom Rhizaria: contrasting cell
-organisation~of sister~phyla Cercozoa and Retaria. Protoplasma 255,
-1517--1574. \url{https://doi.org/10.1007/s00709-018-1241-1}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cohen2023}{}}%
+\bibitem[\citeproctext]{ref-cohen2023}
 Cohen, N.R., Krinos, A.I., Kell, R.M., Chmiel, R.J., Moran, D.M.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-decelle2015}{}}%
+\bibitem[\citeproctext]{ref-decelle2015}
 Decelle, J., Colin, S., Foster, R.A., 2015. Photosymbiosis in Marine
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-decelle2013}{}}%
+\bibitem[\citeproctext]{ref-decelle2013}
 Decelle, J., Martin, P., Paborstava, K., Pond, D.W., Tarling, G., Mahé,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-decelle2012}{}}%
+\bibitem[\citeproctext]{ref-decelle2012}
 Decelle, J., Siano, R., Probert, I., Poirier, C., Not, F., 2012.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dennett2002}{}}%
+\bibitem[\citeproctext]{ref-dennett2002}
 Dennett, M.R., Caron, D.A., Michaels, A.F., Gallager, S.M., Davis, C.S.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-drago2022}{}}%
+\bibitem[\citeproctext]{ref-drago2022}
 Drago, L., Panaïotis, T., Irisson, J.-O., Babin, M., Biard, T.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gaskell2019}{}}%
+\bibitem[\citeproctext]{ref-gaskell2019}
 Gaskell, D.E., Ohman, M.D., Hull, P.M., 2019. Zooglider-based
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-guidi2016}{}}%
-Guidi, L., Chaffron, S., Bittner, L., Eveillard, D., Larhlimi, A., Roux,
-S., Darzi, Y., Audic, S., Berline, L., Brum, J.R., Coelho, L.P.,
-Espinoza, J.C.I., Malviya, S., Sunagawa, S., Dimier, C., Kandels-Lewis,
-S., Picheral, M., Poulain, J., Searson, S., Stemmann, L., Not, F.,
-Hingamp, P., Speich, S., Follows, M., Karp-Boss, L., Boss, E., Ogata,
-H., Pesant, S., Weissenbach, J., Wincker, P., Acinas, S.G., Bork, P.,
-Vargas, C. de, Iudicone, D., Sullivan, M.B., Raes, J., Karsenti, E.,
-Bowler, C., Gorsky, G., 2016. Plankton networks driving carbon export in
-the oligotrophic ocean. Nature 532, 465--470.
-\url{https://doi.org/10.1038/nature16942}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gutierrez-rodriguez2019}{}}%
+\bibitem[\citeproctext]{ref-gorsky2010}
+Gorsky, G., Ohman, M.D., Picheral, M., Gasparini, S., Stemmann, L.,
+Romagnan, J.-B., Cawood, A., Pesant, S., García-Comas, C., Prejger, F.,
+2010. Digital zooplankton image analysis using the ZooScan integrated
+system. Journal of Plankton Research 32, 285--303.
+\url{https://doi.org/10.1093/plankt/fbp124}
+
+\bibitem[\citeproctext]{ref-gowing1989}
+Gowing, M.M., 1989. Abundance and feeding ecology of Antarctic
+phaeodarian radiolarians. Marine Biology 103, 107--118.
+\url{https://doi.org/10.1007/BF00391069}
+
+\bibitem[\citeproctext]{ref-gutierrez-rodriguez2019}
 Gutierrez-Rodriguez, A., Stukel, M.R., Lopes dos Santos, A., Biard, T.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gutiuxe9rrez-rodruxedguez2022}{}}%
+\bibitem[\citeproctext]{ref-gutiuxe9rrez-rodruxedguez2022}
 Gutiérrez-Rodríguez, A., Lopes dos Santos, A., Safi, K., Probert, I.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-haekel1887}{}}%
+\bibitem[\citeproctext]{ref-haekel1887}
 Haekel, E., 1887.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hull2011}{}}%
+\bibitem[\citeproctext]{ref-hull2011}
 Hull, P.M., Osborn, K.J., Norris, R.D., Robison, B.H., 2011. Seasonality
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ikenoue2019}{}}%
+\bibitem[\citeproctext]{ref-ikenoue2019}
 Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M.C., Takahashi,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kimoto2015}{}}%
+\bibitem[\citeproctext]{ref-kimoto2015}
 Kimoto, K., 2015. Planktic foraminifera. pp. 129--178.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-knap1997}{}}%
+\bibitem[\citeproctext]{ref-knap1997}
 Knap, A.H., Michaels, A.F., Steinberg, D.K., Bahr, F., Bates, N.R.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-laget2024}{}}%
+\bibitem[\citeproctext]{ref-laget2024}
 Laget, M., Drago, L., Panaïotis, T., Kiko, R., Stemmann, L., Rogge, A.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lampitt2009}{}}%
-Lampitt, R.S., Salter, I., Johns, D., 2009. Radiolaria: Major exporters
-of organic carbon to the deep ocean. Global Biogeochemical Cycles 23.
-\url{https://doi.org/10.1029/2008GB003221}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lee2006}{}}%
+\bibitem[\citeproctext]{ref-lee2006}
 Lee, J.J., 2006.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2022}{}}%
+\bibitem[\citeproctext]{ref-llopismonferrer2022}
 Llopis Monferrer, N., Biard, T., Sandin, M.M., Lombard, F., Picheral,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2021}{}}%
-Llopis Monferrer, N., Leynaert, A., Tréguer, P., Gutiérrez-Rodríguez,
-A., Moriceau, B., Gallinari, M., Latasa, M., L'Helguen, S., Maguer,
-J.-F., Safi, K., Pinkerton, M.H., Not, F., 2021. Role of small Rhizaria
-and diatoms in the pelagic silica production of the Southern Ocean.
-Limnology and Oceanography 66, 2187--2202.
-\url{https://doi.org/10.1002/lno.11743}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lomas2013}{}}%
+\bibitem[\citeproctext]{ref-lomas2013}
 Lomas, M.W., Bates, N.R., Johnson, R.J., Knap, A.H., Steinberg, D.K.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lombard2019}{}}%
+\bibitem[\citeproctext]{ref-lombard2019}
 Lombard, F., Boss, E., Waite, A.M., Vogt, M., Uitz, J., Stemmann, L.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-marra2011}{}}%
+\bibitem[\citeproctext]{ref-marra2011}
 Marra, G., Wood, S.N., 2011. Practical variable selection for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-marsbrisbin2020}{}}%
+\bibitem[\citeproctext]{ref-marsbrisbin2020}
 Mars Brisbin, M., Brunner, O.D., Grossmann, M.M., Mitarai, S., 2020.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mcgillicuddy1998}{}}%
+\bibitem[\citeproctext]{ref-mcgillicuddy1998}
 McGillicuddy, D.J., Robinson, A.R., Siegel, D.A., Jannasch, H.W.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-michaels1988}{}}%
+\bibitem[\citeproctext]{ref-michaels1988}
 Michaels, A.F., 1988. Vertical distribution and abundance of Acantharia
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-michaels1995}{}}%
+\bibitem[\citeproctext]{ref-michaels1995}
 Michaels, A.F., Caron, D.A., Swanberg, N.R., Howse, F.A., Michaels,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-michaels1996}{}}%
+\bibitem[\citeproctext]{ref-michaels1996}
 Michaels, A.F., Knap, A.H., 1996. Overview of the U.S. JGOFS Bermuda
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nakamura2023}{}}%
+\bibitem[\citeproctext]{ref-nakamura2023}
 Nakamura, Y., Itagaki, H., Tuji, A., Shimode, S., Yamaguchi, A., Hidaka,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-nakamura2015}{}}%
+\bibitem[\citeproctext]{ref-nakamura2015}
 Nakamura, Y., Suzuki, N., 2015.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ohman2019a}{}}%
+\bibitem[\citeproctext]{ref-ohman2019a}
 Ohman, M.D., 2019. A sea of tentacles: optically discernible traits
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-panauxefotis2023}{}}%
+\bibitem[\citeproctext]{ref-panauxefotis2023}
 Panaïotis, T., Babin, M., Biard, T., Carlotti, F., Coppola, L., Guidi,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-picheral}{}}%
+\bibitem[\citeproctext]{ref-picheral}
 Picheral, M., Colin, S.P., Irisson, J.-O., n.d.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-picheral2010}{}}%
+\bibitem[\citeproctext]{ref-picheral2010}
 Picheral, M., Guidi, L., Stemmann, L., Karl, D.M., Iddaoud, G., Gorsky,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sogawa2022}{}}%
+\bibitem[\citeproctext]{ref-sogawa2022}
 Sogawa, S., Nakamura, Y., Nagai, S., Nishi, N., Hidaka, K., Shimizu, Y.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-stukel2018}{}}%
+\bibitem[\citeproctext]{ref-stukel2018}
 Stukel, M.R., Biard, T., Krause, J., Ohman, M.D., 2018. Large Phaeodaria
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-stukel2019}{}}%
+\bibitem[\citeproctext]{ref-stukel2019}
 Stukel, M.R., Ohman, M.D., Kelly, T.B., Biard, T., 2019.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-suzuki2015}{}}%
+\bibitem[\citeproctext]{ref-suzuki2015}
 Suzuki, N., Not, F., 2015.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-swanberg1981}{}}%
+\bibitem[\citeproctext]{ref-swanberg1981}
 Swanberg, N.R., Anderson, O.R., 1981. Collozoum caudatum sp. nov.: A
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-takahashi1983}{}}%
+\bibitem[\citeproctext]{ref-takahashi1983}
 Takahashi, K., Hurd, D.C., Honjo, S., 1983. Phaeodarian skeletons: Their
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-whitmore2021}{}}%
+\bibitem[\citeproctext]{ref-whitmore2021}
 Whitmore, B.M., Ohman, M.D., 2021. Zooglider-measured association of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wood2017}{}}%
+\bibitem[\citeproctext]{ref-wood2017}
 Wood, S.N., 2017. Generalized additive models: an introduction with R,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wood2001}{}}%
+\bibitem[\citeproctext]{ref-wood2001}
 Wood, S.N., 2001. mgcv: GAMs and Generalized Ridge Regression for R 1.
 
-\leavevmode\vadjust pre{\hypertarget{ref-zasko2005}{}}%
+\bibitem[\citeproctext]{ref-zasko2005}
 Zasko, D.N., Rusanov, I.I., 2005. Vertical Distribution of Radiolarians

### 2024-06-26 manuscript/main.tex
render manuscript with revisions

@@
 Phaeodaria, all of which are observed throughout the global ocean (Biard
-et al., 2016; Laget et al., 2024) . While the taxonomy of these
-organisms has recently undergone several reclassifications (Biard,
-2022a), their presence in ocean ecosystems has been long known to
+\emph{et al.}, 2016; Laget \emph{et al.}, 2024) . While the taxonomy of
+these organisms has recently undergone several reclassifications (Biard,
+2022b), their presence in ocean ecosystems has been long known to
 oceanographers. Some of the earliest records of their existence are from
@@
 intricate mineral skeletons out of either silica, strontium, or calcium
-carbonate (Biard, 2022a; Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki
-and Not, 2015). Despite their noticeable morphology and global
+carbonate (Kimoto, 2015; Nakamura and Suzuki, 2015; Suzuki and Not,
+2015; Biard, 2022b). Despite their noticeable morphology and global
 distribution, Rhizaria were largely understudied throughout the 20th
@@
 1900s did employ alternative techniques to quantify Rhizaria including
-diaphragm pumps (Michaels, 1988) or blue-water SCUBA collections (Bijma
-et al., 1990; Caron et al., 1995; Caron and Be, 1984). However, the bulk
-of Rhizaria research was constrained to sediment traps or
-paleontological studies of sediment (Boltovskoy et al., 1993; Takahashi
-et al., 1983).
+diaphragm pumps (Michaels, 1988) or blue-water SCUBA collections (Caron
+and Be, 1984; Bijma \emph{et al.}, 1990; Caron \emph{et al.}, 1995).
+However, the bulk of Rhizaria research was constrained to sediment traps
+or paleontological studies of sediment (Takahashi \emph{et al.}, 1983;
+Boltovskoy \emph{et al.}, 1993).
 
@@
 (Caron, 2016). Molecular tools have shed light on Rhizaria taxonomy
-(Biard, 2022a), distribution (Blanco-Bercial, 2020; Sogawa et al.,
-2022), ecology (Decelle et al., 2012; Nakamura et al., 2023),
-biogeochemical roles (Gutierrez-Rodriguez et al., 2019; Laget et al.,
-2024), and metabolism (Cohen et al., 2023). Yet, despite the excellent
-taxonomic resolution provided by molecular approaches, they do not
-provide a truly quantitative metric for estimating Rhizaria abundance or
-biomass. \emph{In situ} imaging tools however, offer the ability to
-observe organisms in the natural state and quantify their abundance
-(Barth and Stone, 2024; Ohman, 2019). Biard et al. (2016) utilized
-\emph{in situ} imaging at a global scale to suggest large
-(\textgreater500\(\mu\)m diameter) Rhizaria were substantial
-contributors to the ocean carbon standing stock. While more recent
-calculations suggest lower carbon contribution (Laget et al., 2024),
-Rhizaria nonetheless have substantial influences on biogeochemical
-cycling. This influence is due to their unique mineral skeletons and
-ability to repackage small particles into larger, fast sinking ones
-(Ikenoue et al., 2019). Thus, some Rhizaria may play substantial roles
-in ocean export through their interception of sinking particles (Laget
-et al., 2024; Stukel et al., 2018). However, Rhizaria ecology is poorly
-understood (Biard, 2022a) and their ecological roles are likely more
-than diverse than simple particle interception.
+(Biard, 2022b), distribution (Blanco-Bercial, 2020; Sogawa \emph{et
+al.}, 2022), ecology (Decelle \emph{et al.}, 2012; Nakamura \emph{et
+al.}, 2023), biogeochemical roles (Gutierrez-Rodriguez \emph{et al.},
+2019; Laget \emph{et al.}, 2024), and metabolism (Cohen \emph{et al.},
+2023). Yet, despite the excellent taxonomic resolution provided by
+molecular approaches, they do not provide a truly quantitative metric
+for estimating Rhizaria abundance or biomass. \emph{In situ} imaging
+tools however, offer the ability to observe organisms in the natural
+state and quantify their abundance (Ohman, 2019; Barth and Stone, 2024).
+Biard \emph{et al.} (2016) utilized \emph{in situ} imaging at a global
+scale to suggest large (\textgreater500\(\mu\)m diameter) Rhizaria were
+substantial contributors to the ocean carbon standing stock. While more
+recent calculations suggest lower carbon contribution (Laget \emph{et
+al.}, 2024), Rhizaria nonetheless have substantial influences on
+biogeochemical cycling. This influence is due to their unique mineral
+skeletons and ability to repackage small particles into larger, fast
+sinking ones (Ikenoue \emph{et al.}, 2019). Thus, some Rhizaria may play
+substantial roles in ocean export through their interception of sinking
+particles (Stukel \emph{et al.}, 2018; Laget \emph{et al.}, 2024).
+However, Rhizaria ecology is poorly understood (Biard, 2022b) and their
+ecological roles are likely more than diverse than simple particle
+interception.
 
@@
 due to the fact different taxa can exhibit very different trophic modes.
-Many Rhizaria are strictly heterotrophic (Biard, 2022a), yet their
+Many Rhizaria are strictly heterotrophic (Biard, 2022b), yet their
 feeding modes can be quite varied. Phaeodaria are largely thought to be
@@
 mixotrophic, utilizing photosynthetic algal symbionts (Anderson, 2014;
-Decelle et al., 2015). Mixotrophic Foraminifera host a variety of
-endosymbiont partners (Decelle et al., 2015; Lee, 2006), which are
-thought to support early and adult life stages and significantly
+Decelle \emph{et al.}, 2015). Mixotrophic Foraminifera host a variety of
+endosymbiont partners (Lee, 2006; Decelle \emph{et al.}, 2015), which
+are thought to support early and adult life stages and significantly
 contribute to total primary productivity (Kimoto, 2015). Still,
@@
 with several studies suggesting that they can be effective predators
-(Anderson and Bé, 1976; Gaskell et al., 2019), mainly consuming live
-copepods (Caron and Be, 1984). Radiolaria have several lineages, many of
-which host symbionts (Biard, 2022b). Amongst Radiolaria, arguably the
-most widespread are Collodaria who can be either large solitary cells or
-form massive colonies, up to several meters in length (Swanberg and
-Anderson, 1981). All known Collodaria species host dinoflagellate
-symbionts (Biard, 2022b) and can contribute substantially to primary
-productivity, particularly in oligotrophic ocean regions (Caron et al.,
-1995; Dennett et al., 2002). This Collodaria-symbiont association has
-been suggested as a reason for their high abundances throughout the
-photic zone of oligotrophic environments globally (Biard et al., 2017,
-2016). A few Acantharea (Radiolaria order) clades host algal symbionts
-(Biard, 2022b; Decelle et al., 2012), notably with two clades forming an
-exclusive relationship with \emph{Phaeocystis}. However, globally,
-Acantharea are less abundant than Collodaria (Biard, 2022a) and
-contribute less to total primary productivity (Michaels et al., 1995).
-This may be due to the fact several clades of Acantharea are
-cyst-forming and strictly heterotrophs (Biard, 2022b; Decelle et al.,
-2013). Furthermore, Mars Brisbin et al. (2020) documented apparent
-predation behavior in Acantharea near the surface, suggesting that there
-may be a larger reliance on carnivory.
+(Anderson and Bé, 1976; Gaskell \emph{et al.}, 2019), mainly consuming
+live copepods (Caron and Be, 1984). Radiolaria have several lineages,
+many of which host symbionts (Biard, 2022a). Amongst Radiolaria,
+arguably the most widespread are Collodaria who can be either large
+solitary cells or form massive colonies, up to several meters in length
+(Swanberg and Anderson, 1981). All known Collodaria species host
+dinoflagellate symbionts (Biard, 2022a) and can contribute substantially
+to primary productivity, particularly in oligotrophic ocean regions
+(Caron \emph{et al.}, 1995; Dennett \emph{et al.}, 2002). This
+Collodaria-symbiont association has been suggested as a reason for their
+high abundances throughout the photic zone of oligotrophic environments
+globally (Biard \emph{et al.}, 2016, 2017). A few Acantharea (Radiolaria
+order) clades host algal symbionts (Decelle \emph{et al.}, 2012; Biard,
+2022a), notably with two clades forming an exclusive relationship with
+\emph{Phaeocystis}. However, globally, Acantharea are less abundant than
+Collodaria (Biard, 2022b) and contribute less to total primary
+productivity (Michaels \emph{et al.}, 1995). This may be due to the fact
+several clades of Acantharea are cyst-forming and strictly heterotrophs
+(Decelle \emph{et al.}, 2013; Biard, 2022a). Furthermore, Mars Brisbin
+\emph{et al.} (2020) documented apparent predation behavior in
+Acantharea near the surface, suggesting that there may be a larger
+reliance on carnivory.
 
@@
 strategies. Taxon-specific studies of Radiolaria suggest they may be
-restricted to the euphotic zone (Boltovskoy, 2017; Michaels, 1988).
-Although some studies report Acantharea in deeper waters (Decelle et
-al., 2013; Gutiérrez-Rodríguez et al., 2022), Phaeodaria, alternatively,
-are generally found beyond depths where light can sustain
-photosynthesis, but they can feed on sinking particles (Laget et al.,
-2024; Stukel et al., 2018). In an imaging-based study of the whole
-Rhizaria community, Biard and Ohman (2020) noted clear patterns in
-vertical zonation which largely corresponded to different trophic roles.
-Protists communities, including Rhizaria, have been described to
-partition along autotroph-mixotroph-heterotroph gradients with
-increasing depth in the water column (Blanco-Bercial et al., 2022; Laget
-et al., 2024). Yet, few studies have made direct attempts to relate
-Rhizaria abundances to environmental factors (Biard and Ohman, 2020). In
-part, this is due to the fact few studies have been able to sample
-Rhizaria in the same location over a consistent timeframe (Boltovskoy et
-al., 1993; Gutiérrez-Rodríguez et al., 2022; Hull et al., 2011; Michaels
-et al., 1995; Michaels, 1988). Furthermore, very few studies have
-utilized \emph{in situ} imaging, arguably the best method for
-quantifying Rhizaria, to consistently throughout the full mesopelagic
-(Laget et al., 2024) . Given this lack of information, there are many
-unknowns with respect to Rhizaria ecology, seasonality and phenology
-across different groups.
+restricted to the euphotic zone (Michaels, 1988; Boltovskoy, 2017).
+Although some studies report Acantharea in deeper waters (Decelle
+\emph{et al.}, 2013; Gutiérrez-Rodríguez \emph{et al.}, 2022),
+Phaeodaria, alternatively, are generally found beyond depths where light
+can sustain photosynthesis, but they can feed on sinking particles
+(Stukel \emph{et al.}, 2018; Laget \emph{et al.}, 2024). In an
+imaging-based study of the whole Rhizaria community, Biard and Ohman
+(2020) noted clear patterns in vertical zonation which largely
+corresponded to different trophic roles. Protists communities, including
+Rhizaria, have been described to partition along
+autotroph-mixotroph-heterotroph gradients with increasing depth in the
+water column (Blanco-Bercial \emph{et al.}, 2022; Laget \emph{et al.},
+2024). Yet, few studies have made direct attempts to relate Rhizaria
+abundances to environmental factors (Biard and Ohman, 2020). In part,
+this is due to the fact few studies have been able to sample Rhizaria in
+the same location over a consistent timeframe (Michaels, 1988;
+Boltovskoy \emph{et al.}, 1993; Michaels \emph{et al.}, 1995; Hull
+\emph{et al.}, 2011; Gutiérrez-Rodríguez \emph{et al.}, 2022).
+Furthermore, very few studies have utilized \emph{in situ} imaging,
+arguably the best method for quantifying Rhizaria, to consistently
+throughout the full mesopelagic (Laget \emph{et al.}, 2024) . Given this
+lack of information, there are many unknowns with respect to Rhizaria
+ecology, seasonality and phenology across different groups.
 
@@
 Data were collected in collaboration with the Bermuda Atlantic
-Time-series Study (Lomas et al., 2013; Michaels and Knap, 1996) on board
-the R/V Atlantic Explorer. Cruises were conducted at approximately
-monthly intervals. Rhizaria individuals were sampled using the
-Underwater Vision Profiler 5 (UVP) (Picheral et al., 2010). The UVP5 is
-an in-situ camera used to capture plankton and particle images and has
-been well established to accurately quantify large Rhizaria abundance
-(Barth and Stone, 2022; Biard et al., 2016; Biard and Ohman, 2020; Drago
-et al., 2022; Llopis Monferrer et al., 2022; Panaïotis et al., 2023;
-Stukel et al., 2019; Stukel et al., 2018). The UVP5 was mounted to the
+Time-series Study (Michaels and Knap, 1996; Lomas \emph{et al.}, 2013)
+on board the R/V Atlantic Explorer. Cruises were conducted at
+approximately monthly intervals. Rhizaria individuals were sampled using
+the Underwater Vision Profiler 5 (UVP) (Picheral \emph{et al.}, 2010).
+The UVP5 is an in-situ camera used to capture plankton and particle
+images and has been well established to accurately quantify large
+Rhizaria abundance (Biard \emph{et al.}, 2016; Stukel \emph{et al.},
+2018; Stukel \emph{et al.}, 2019; Biard and Ohman, 2020; Barth and
+Stone, 2022; Drago \emph{et al.}, 2022; Llopis Monferrer \emph{et al.},
+2022; Panaïotis \emph{et al.}, 2023). The UVP5 was mounted to the
 sampling rosette and collected data autonomously on routine casts, from
@@
 from \emph{in situ} incubations. For full descriptions of the BATS
-sampling program and methods, see Knap et al. (1997) and Lomas et al.
-(2013) for a review. Additionally, data can be viewed online
+sampling program and methods, see Knap \emph{et al.} (1997) and Lomas
+\emph{et al.} (2013) for a review. Additionally, data can be viewed
+online
 \href{https://bats.bios.asu.edu/bats-data/}{(https://bats.bios.asu.edu/bats-data/)}.
@@
 Individual vignettes of Rhizaria images were identified using the
-classification platform EcoTaxa (Picheral et al., n.d.). Data were
+classification platform EcoTaxa (Picheral \emph{et al.}). Data were
 pre-sorted utilizing a random-forest classifier and pre-trained learning
 set. Taxonomic classification were done based on morphological
-parameters measured in Zooprocess (Gorsky et al., 2010), which includes
-parameters such as major axis, equivalent spherical diameter (ESD), and
-grey values. While there are sparse taxonomic guides for \emph{in situ}
-images of Rhizaria, identification largely relied on descriptions in
-(Nakamura and Suzuki, 2015; Suzuki and Not, 2015; and Biard and Ohman,
-2020). Using the aforementioned sources and publicly available EcoTaxa
-projects, we constructed a guide accessible at:
+parameters measured in Zooprocess (Gorsky \emph{et al.}, 2010), which
+includes parameters such as major axis, equivalent spherical diameter
+(ESD), and grey values. While there are sparse taxonomic guides for
+\emph{in situ} images of Rhizaria, identification largely relied on
+descriptions in (Nakamura and Suzuki, 2015; Suzuki and Not, 2015; and
+Biard and Ohman, 2020). Using the aforementioned sources and publicly
+available EcoTaxa projects, we constructed a guide accessible at:
 \url{https://thealexbarth.github.io/media//Project_Items/Oligotrophic_Community/ecotaxa_UVP-guide-stone-lab.pdf}.
@@
 constructed to identify a drop-off which would indicate poor-sampling at
-the small size range (Barth and Stone, 2024; Lombard et al., 2019).
-However, it was evident from this analysis that all size ranges were
-adequately sampled across the size range (Supplemental Figure 2) so no
-data were excluded. The UVP5 reports the exact depth at which a particle
-is recorded, however to estimate abundance, observations must be binned
-over fixed depth intervals. Our deployments had variable descent depths
-and speeds with more casts descending to 500m than 1000m and descents
-quicker through the epipelagic than the mesopelagic (see Barth and Stone
-(2022) for an extended discussion of UVP5 data processing). For the
-present study, Rhizaria abundances were estimated in 25m vertical bins,
-which offer a moderate sampling volume per bin (average 0.948\(m^3\) in
-the epipelagic and 0.589\(m^3\) in the mesopelagic; Supplemental Table
-1) while still maintaining ecologically relevant widths. However,
-concentrations in a 25m bin would need to be greater than 2.428 ind.
-\(m^{-3}\) and 3.912 ind. \(m^{-3}\), in the epipelagic and mesopelagic
-respectively, to fall below a 10\% non-detection risk (Barth and Stone,
-2024; Benfield et al., 1996). Because we typically observed many
-Rhizaria taxa below these concentrations, we present the 25m binned data
-to visualize broad-scale average distributions. For quantifying and
-modelling Rhizaria abundances, we present integrated abundance
-estimates, with each cast. Due to the variable descent depths of the
-UVP, data are categorized as epipelagic (0-200m), upper mesopelagic
-(200-500m), and lower mesopelagic (500-1000m). The average sampling
-volume integrated through these regions were 7.59\(m^2\), 7.06\(m^2\),
-and 11.77\(m^2\), with non-detection thresholds at 0.30 ind. \(m^{-2}\),
-0.33 ind. \(m^{-2}\), and 0.20 ind. \(m^{-2}\) respectively. All UVP
-data processing was done using the \texttt{EcotaxaTools} package in R
-(Barth 2023).
+the small size range (Lombard \emph{et al.}, 2019; Barth and Stone,
+2024). However, it was evident from this analysis that all size ranges
+were adequately sampled across the size range (Supplemental Figure 2) so
+no data were excluded. The UVP5 reports the exact depth at which a
+particle is recorded, however to estimate abundance, observations must
+be binned over fixed depth intervals. Our deployments had variable
+descent depths and speeds with more casts descending to 500m than 1000m
+and descents quicker through the epipelagic than the mesopelagic (see
+Barth and Stone (2022) for an extended discussion of UVP5 data
+processing). For the present study, Rhizaria abundances were estimated
+in 25m vertical bins, which offer a moderate sampling volume per bin
+(average 0.948\(m^3\) in the epipelagic and 0.589\(m^3\) in the
+mesopelagic; Supplemental Table 1) while still maintaining ecologically
+relevant widths. However, concentrations in a 25m bin would need to be
+greater than 2.428 ind. \(m^{-3}\) and 3.912 ind. \(m^{-3}\), in the
+epipelagic and mesopelagic respectively, to fall below a 10\%
+non-detection risk (Benfield \emph{et al.}, 1996; Barth and Stone,
+2024). Because we typically observed many Rhizaria taxa below these
+concentrations, we present the 25m binned data to visualize broad-scale
+average distributions. For quantifying and modelling Rhizaria
+abundances, we present integrated abundance estimates, with each cast.
+Due to the variable descent depths of the UVP, data are categorized as
+epipelagic (0-200m), upper mesopelagic (200-500m), and lower mesopelagic
+(500-1000m). The average sampling volume integrated through these
+regions were 7.59\(m^2\), 7.06\(m^2\), and 11.77\(m^2\), with
+non-detection thresholds at 0.30 ind. \(m^{-2}\), 0.33 ind. \(m^{-2}\),
+and 0.20 ind. \(m^{-2}\) respectively. All UVP data processing was done
+using the \texttt{EcotaxaTools} package in R (Barth 2023).
 
@@
 and summer stratification as well as secondary influences from mesoscale
-eddies which result in spatiotemporal heterogeniety (Lomas et al., 2013;
-McGillicuddy et al., 1998). Variability in the water column structure
-was visible during the study period (Figure 2). This is best evidenced
-through the temperature profiles; In the late summer and early fall
-there was a stratified water column with high temperatures in the
-surface (\textless75m) (Figure 2A) and slightly elevated salinity
-(Figure 2B). This warm, stratified period appeared more intense during
-the few months sampled in 2019. In 2021, we observed the stratified
-layer slowly dissipated into the winter months following mixed layer
-entrainment. There was a consistent oxygen minimum zone (OMZ) located at
-about 800m deep (Figure 2C). February 2021 saw a notable downwelling
-event, likely due to a passing anti-cyclonic eddy which impacted the
-local region during early 2021. During this phase, warmer, oxic water
-was significantly depressed deeper into the mesopelagic. This process
-was reversed in the spring months (March, April) primary due to the
-interaction of convective mixing and the passing of a strong cyclonic
-eddy resulting in a deep cold mixed layer. Primary production was
-highest during the spring mixing period, evidenced both by \emph{in
-situ} fluorescence (Figure 2D) and productivity incubation experiments
-(Figure 3A). Originating near the surface, the productivity peak moved
-deeper throughout the spring and declined into the summer (Figure 2D).
-However, there was a notable, yet smaller productivity bump in the late
-summer and early fall (Figure 3A) which occurred deeper in the
-epipelagic (Figure 2D). The particle concentration (184\(\mu m\) -
-900\(\mu m\)) was closely coupled to chlorophyll-a patterns.
+eddies which result in spatiotemporal heterogeniety (McGillicuddy
+\emph{et al.}, 1998; Lomas \emph{et al.}, 2013). Variability in the
+water column structure was visible during the study period (Figure 2).
+This is best evidenced through the temperature profiles; In the late
+summer and early fall there was a stratified water column with high
+temperatures in the surface (\textless75m) (Figure 2A) and slightly
+elevated salinity (Figure 2B). This warm, stratified period appeared
+more intense during the few months sampled in 2019. In 2021, we observed
+the stratified layer slowly dissipated into the winter months following
+mixed layer entrainment. There was a consistent oxygen minimum zone
+(OMZ) located at about 800m deep (Figure 2C). February 2021 saw a
+notable downwelling event, likely due to a passing anti-cyclonic eddy
+which impacted the local region during early 2021. During this phase,
+warmer, oxic water was significantly depressed deeper into the
+mesopelagic. This process was reversed in the spring months (March,
+April) primary due to the interaction of convective mixing and the
+passing of a strong cyclonic eddy resulting in a deep cold mixed layer.
+Primary production was highest during the spring mixing period,
+evidenced both by \emph{in situ} fluorescence (Figure 2D) and
+productivity incubation experiments (Figure 3A). Originating near the
+surface, the productivity peak moved deeper throughout the spring and
+declined into the summer (Figure 2D). However, there was a notable, yet
+smaller productivity bump in the late summer and early fall (Figure 3A)
+which occurred deeper in the epipelagic (Figure 2D). The particle
+concentration (184\(\mu m\) - 900\(\mu m\)) was closely coupled to
+chlorophyll-a patterns.
 
@@
 
-{\centering \includegraphics{images/02-environmental.pdf}
+{\centering \includegraphics{images/02_environmental.pdf}
 
@@
 Abundance}\label{environmental-drivers-of-rhizaria-abundance}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{images/06_tot-partials.pdf}
+
+}
+
+\caption{Partial effect plots from Generalized Additive Models (GAMs) of
+all Rhizaria abundance. Models were constructed separately for depth
+integrated abundance through the Epipelagic (0-200m) (A-G), Upper
+Mesopelagic (200-500m) (H-J), and the Lower Mesopelagic (500-1000m)
+(K-L). Each panel represents the partial effect of a GAM model term,
+with smoothed relationship shown shaded with 95\% confidence intervals.
+Individual points are depth integrated values for all Rhizaria from an
+individual cast, colored by month.}
+
+\end{figure}%
 
@@
 Rhizaria abundances were higher in the summer months and lower during
-the winter. During a prior time period, Blanco-Bercial et al. (2022)
-noted that there is considerable seasonality in the community
+the winter. During a prior time period, Blanco-Bercial \emph{et al.}
+(2022) noted that there is considerable seasonality in the community
 composition of all protists. Despite the seasonality of Rhizaria
@@
 consistent with observations from metabarcoding of the whole protist
-community in the same study region (Blanco-Bercial et al., 2022). This
-finding is not surprising as the overall seasonal variation in
+community in the same study region (Blanco-Bercial \emph{et al.}, 2022).
+This finding is not surprising as the overall seasonal variation in
 environmental conditions in this region were low.
@@
 noting that in the same region, with data collected in 2012 and 2013
-using similar calculation methods, Biard et al. (2016) estimated large
-Rhizaria only contribute 15\% of the total mesozooplankton community in
-the upper 500m. Likely, Rhizaria display considerable interannual
-variability. In the present study, we noticed considerably higher
-Rhizaria abundance throughout the water column in 2019 compared to 2021.
-Similarly, in the California Current a multi-year study of
-Aulosphaeridae saw considerable variation between years (Biard et al.,
-2018). To truly understand the drivers of interannual variability
+using similar calculation methods, Biard \emph{et al.} (2016) estimated
+large Rhizaria only contribute 15\% of the total mesozooplankton
+community in the upper 500m. Likely, Rhizaria display considerable
+interannual variability. In the present study, we noticed considerably
+higher Rhizaria abundance throughout the water column in 2019 compared
+to 2021. Similarly, in the California Current a multi-year study of
+Aulosphaeridae saw considerable variation between years (Biard \emph{et
+al.}, 2018). To truly understand the drivers of interannual variability
 however, sustained observations of Rhizaria over longer time periods are
@@
 strongly vary with changes in parameters such as temperature,
-chlorophyll-a, mixing, and water clarity (Biard et al., 2017; Biard and
-Ohman, 2020). Alternatively, Acantharea had relatively good fitting
-GAMs. These taxa also had some of the largest variation from month to
-month on cruises. Thus, it may be that in the oligotrophic, the
+chlorophyll-a, mixing, and water clarity (Biard \emph{et al.}, 2017;
+Biard and Ohman, 2020). Alternatively, Acantharea had relatively good
+fitting GAMs. These taxa also had some of the largest variation from
+month to month on cruises. Thus, it may be that in the oligotrophic, the
 relatively stable conditions can support certain taxa while others are
@@
 Radiolaria have long been reported to contribute to primary productivity
-in the euphotic zone (Dennett et al., 2002; Michaels et al., 1995).
-Collodaria are thought to be particularly successful globally in
-oligotrophic regions due to their photosymbiotic relationships (Biard et
-al., 2017, 2016). We observed the highest abundance of Collodaria during
-June 2021, supporting the notion they can thrive during the typically
-low-nutrient conditions of summer stratification. However, Collodaria
-also increased during the spring mixing period, suggesting that they can
-thrive during conditions which may typically be thought to favor
-autotrophs. Furthermore, while Collodaria were primarily absent from
-below 250m, there were a few instances of deeper colonies being
-observed. Global investigations of polycystine flux, suggest that
-deep-Collodaria in oligotrophic regions may be a consequence of
-isothermal submersion (Boltovskoy, 2017). Alternatively, surface waters
-at BATS often mix into the mode water during the seasonal mixing, so
-Collodaria in the deeper waters may be a result of diapyncal mixing.
-Another effectively exclusively epipelagic Rhizaria was the Phaeodaria
-family of Castanellidae. All Phaeodaria are thought to be fully
-heterotrophic (Nakamura and Suzuki, 2015), nonetheless a number of
-studies, including this one, report Castanellidae to be typically found
-in the lower epipelagic (Biard et al., 2018; Biard and Ohman, 2020;
-Zasko and Rusanov, 2005). It should be considered that perhaps
-Castanellidae specializes in feeding on sinking particles directly at
-the base of the epipelagic. Given it's smaller size (Nakamura and
-Suzuki, 2015), Castanellidae does not need a large diameter to
-efficiently flux feed at the typically particle rich region of the lower
-epipelagic.
+in the euphotic zone (Michaels \emph{et al.}, 1995; Dennett \emph{et
+al.}, 2002). Collodaria are thought to be particularly successful
+globally in oligotrophic regions due to their photosymbiotic
+relationships (Biard \emph{et al.}, 2016, 2017). We observed the highest
+abundance of Collodaria during June 2021, supporting the notion they can
+thrive during the typically low-nutrient conditions of summer
+stratification. However, Collodaria also increased during the spring
+mixing period, suggesting that they can thrive during conditions which
+may typically be thought to favor autotrophs. Furthermore, while
+Collodaria were primarily absent from below 250m, there were a few
+instances of deeper colonies being observed. Global investigations of
+polycystine flux, suggest that deep-Collodaria in oligotrophic regions
+may be a consequence of isothermal submersion (Boltovskoy, 2017).
+Alternatively, surface waters at BATS often mix into the mode water
+during the seasonal mixing, so Collodaria in the deeper waters may be a
+result of diapyncal mixing. Another effectively exclusively epipelagic
+Rhizaria was the Phaeodaria family of Castanellidae. All Phaeodaria are
+thought to be fully heterotrophic (Nakamura and Suzuki, 2015),
+nonetheless a number of studies, including this one, report
+Castanellidae to be typically found in the lower epipelagic (Zasko and
+Rusanov, 2005; Biard \emph{et al.}, 2018; Biard and Ohman, 2020). It
+should be considered that perhaps Castanellidae specializes in feeding
+on sinking particles directly at the base of the epipelagic. Given it's
+smaller size (Nakamura and Suzuki, 2015), Castanellidae does not need a
+large diameter to efficiently flux feed at the typically particle rich
+region of the lower epipelagic.
 
@@
 particularly for those which were constrained to exclusively occupy
-deeper waters. This is consistent with Blanco-Bercial et al. (2022)'s
-observation of an auto-/mixotroph to heterotroph gradient in local
-protist community as well as global patterns in Rhizaria ecology
-{[}Laget et al. (2024){]}. The upper mesopelagic interestingly had
-relatively low total abundance. This low-abundance region likely
+deeper waters. This is consistent with Blanco-Bercial \emph{et al.}
+(2022)'s observation of an auto-/mixotroph to heterotroph gradient in
+local protist community as well as global patterns in Rhizaria ecology
+{[}Laget \emph{et al.} (2024){]}. The upper mesopelagic interestingly
+had relatively low total abundance. This low-abundance region likely
 reflects the dynamics of productivity and export throughout the water
@@
 reported {[}Zasko and Rusanov (2005); Biard and Ohman (2020){]} and they
-have massive potential to impact silica export (Biard et al., 2018).
-Coelodendridae were also seemingly restricted to the deeper section of
-the mesopelagic. This is interesting given that in the California
-Current, (Biard and Ohman, 2020) found a bimodal distribution in
-Coelodendridae. There are several morphotypes corresponding to different
-taxa of Coelodendridae (Biard and Ohman, 2020; Nakamura and Suzuki,
-2015). So it may be that only a few types of Coelodendridae were
+have massive potential to impact silica export (Biard \emph{et al.},
+2018). Coelodendridae were also seemingly restricted to the deeper
+section of the mesopelagic. This is interesting given that in the
+California Current, (Biard and Ohman, 2020) found a bimodal distribution
+in Coelodendridae. There are several morphotypes corresponding to
+different taxa of Coelodendridae (Nakamura and Suzuki, 2015; Biard and
+Ohman, 2020). So it may be that only a few types of Coelodendridae were
 observed in this study, while the epipelagic variety was not.
@@
 mesopelagic. Foraminifera also had a bimodal distribution. Some lineages
-of Foraminifera are known to host photosymbionts (Biard, 2022a; Kimoto,
-2015), however they are also efficient predators commonly seen
-throughout the mesopelagic (Caron and Be, 1984; Gaskell et al., 2019).
-Thus it is not surprising to find their presence in both locations of
-the water column. Foraminifera are also known to vary their vertical
-distribution across their life cycle in phase with lunar cycles (Biard,
-2022a; Bijma et al., 1990; Gaskell et al., 2019; Kimoto, 2015). However,
-the sampling scheme of the BATS program does not capture this frequency
-and was not investigated in the present study.
+of Foraminifera are known to host photosymbionts (Kimoto, 2015; Biard,
+2022b), however they are also efficient predators commonly seen
+throughout the mesopelagic (Caron and Be, 1984; Gaskell \emph{et al.},
+2019). Thus it is not surprising to find their presence in both
+locations of the water column. Foraminifera are also known to vary their
+vertical distribution across their life cycle in phase with lunar cycles
+(Bijma \emph{et al.}, 1990; Kimoto, 2015; Gaskell \emph{et al.}, 2019;
+Biard, 2022b). However, the sampling scheme of the BATS program does not
+capture this frequency and was not investigated in the present study.
 
@@
 the paradigm that large Acantharea abundances may be supported by their
-mixotrophic abilities (Michaels et al., 1995; Suzuki and Not, 2015).
-While the UVP5 images cannot distinguish between mixotrophic and
+mixotrophic abilities (Michaels \emph{et al.}, 1995; Suzuki and Not,
+2015). While the UVP5 images cannot distinguish between mixotrophic and
 heterotrophic Acantharea, the GAMs constructed for Acantharea abundance
 found positive associations with particle concentration and mass flux,
-suggesting a higher reliance on heterotrophy. Recently Mars Brisbin et
-al. (2020) described apparent predator behavior amongst near-surface
-Acantharea. Thus it is likely that epipelagic Acantharea may commonly be
-heterotrophic or mixotrophic with an increased reliance on heterotrophy.
-Yet, it should be noted in the Sargasso Sea, both heterotrophic and
-symbiotic lineages of Acantharea have been reported (Blanco-Bercial et
-al., 2022). Additionally, Michaels (1988) noted that the majority of
-Acantharea (by abundance) were smaller than 160\(\mu m\). While that
-estimate may be inflated due to inability to capture larger cells, small
-Acantharea were not captured in the present study. Thus, trophic
-strategy may shift based on sizes of Acantharea.
-
-Decelle et al. (2013) proposed a hypothetical life cycle for
+suggesting a higher reliance on heterotrophy. Recently Mars Brisbin
+\emph{et al.} (2020) described apparent predator behavior amongst
+near-surface Acantharea. Thus it is likely that epipelagic Acantharea
+may commonly be heterotrophic or mixotrophic with an increased reliance
+on heterotrophy. Yet, it should be noted in the Sargasso Sea, both
+heterotrophic and symbiotic lineages of Acantharea have been reported
+(Blanco-Bercial \emph{et al.}, 2022). Additionally, Michaels (1988)
+noted that the majority of Acantharea (by abundance) were smaller than
+160\(\mu m\). While that estimate may be inflated due to inability to
+capture larger cells, small Acantharea were not captured in the present
+study. Thus, trophic strategy may shift based on sizes of Acantharea.
+
+Decelle \emph{et al.} (2013) proposed a hypothetical life cycle for
 cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life
@@
 the newly spawned mesopelagic Acantharea are not necessarily required to
-rapidly return to the photic zone (Decelle et al., 2013, 2012). This
-hypothesis predicts that Acanthareas in the mesopelagic would be smaller
-(Decelle et al., 2013). Mars Brisbin et al. (2020) provided some support
-for this hypothesis, with a significant decrease in Acantharea sizes
-with depth. Although the authors also observed low abundances in the
-mesopelagic and noted that the smaller sizes may be due to lower food
-availability (Mars Brisbin et al., 2020). Since food is more scarce in
-the mesopelagic, nutritional quality lower, yet flux feeders would
-likely grow larger to increase their feeding range (Biard and Ohman,
-2020; Stukel et al., 2018). In the data collecting in this study,
-Acanthareas in the mesopelagic were significantly smaller than the
-epipelagic, despite the other bimodal taxa (Foraminifera and
-Aulacanthidae) being significantly larger with depth. This provides
-added support for the hypothesis that cyst-forming Acantharea may
-utilize different sections of the water column throughout their life
-cycle. However to further investigate this, more work is needed with
-higher temporal and taxonomic resolution.
+rapidly return to the photic zone (Decelle \emph{et al.}, 2012, 2013).
+This hypothesis predicts that Acanthareas in the mesopelagic would be
+smaller (Decelle \emph{et al.}, 2013). Mars Brisbin \emph{et al.} (2020)
+provided some support for this hypothesis, with a significant decrease
+in Acantharea sizes with depth. Although the authors also observed low
+abundances in the mesopelagic and noted that the smaller sizes may be
+due to lower food availability (Mars Brisbin \emph{et al.}, 2020). Since
+food is more scarce in the mesopelagic, nutritional quality lower, yet
+flux feeders would likely grow larger to increase their feeding range
+(Stukel \emph{et al.}, 2018; Biard and Ohman, 2020). In the data
+collecting in this study, Acanthareas in the mesopelagic were
+significantly smaller than the epipelagic, despite the other bimodal
+taxa (Foraminifera and Aulacanthidae) being significantly larger with
+depth. This provides added support for the hypothesis that cyst-forming
+Acantharea may utilize different sections of the water column throughout
+their life cycle. However to further investigate this, more work is
+needed with higher temporal and taxonomic resolution.
 
@@
 shallower regions, they may be an attenuating force on sinking particles
-(Stukel et al., 2019). However, once consumed and repackaged by larger
-Rhizaria, they can sink quicker and contribute more to overall flux
-(Michaels, 1988). Thus, Rhizaria may act as an aggregation mechanism.
-However, to truly test this, more work is needed measuring Rhizaria
-flux.
+(Stukel \emph{et al.}, 2019). However, once consumed and repackaged by
+larger Rhizaria, they can sink quicker and contribute more to overall
+flux (Michaels, 1988). Thus, Rhizaria may act as an aggregation
+mechanism. However, to truly test this, more work is needed measuring
+Rhizaria flux.
 
@@
 \bibitem[\citeproctext]{ref-anderson2014}
-Anderson, O.R., 2014. Living Together in the Plankton: A Survey of
-Marine Protist Symbioses. Acta Protozoologica 53, 29--38.
-\url{https://doi.org/10.4467/16890027AP.13.0019.1116}
+Anderson, O. R. (2014)
+\href{https://doi.org/10.4467/16890027AP.13.0019.1116}{Living Together
+in the Plankton: A Survey of Marine Protist Symbioses}. \emph{Acta
+Protozoologica}, \textbf{53}, 29--38.
 
 \bibitem[\citeproctext]{ref-anderson1976}
-Anderson, O.R., Bé, A.W.H., 1976. A CYTOCHEMICAL FINE STRUCTURE STUDY OF
-PHAGOTROPHY IN A PLANKTONIC FORAMINIFER, {\emph{HASTIGERINA PELAGICA}}
-(d'ORBIGNY). The Biological Bulletin 151, 437--449.
-\url{https://doi.org/10.2307/1540498}
-
-\bibitem[\citeproctext]{ref-barth2024}
-Barth, A., Stone, J., 2024. Understanding the picture: The promis and
-challenges of in-situ imagery data in the study of plankton ecology.
-Journal of Plankton Research: Horizons 1--15.
-\url{https://doi.org/10.1093/plankt/fbae023}
+Anderson, O. R. and Bé, A. W. H. (1976)
+\href{https://doi.org/10.2307/1540498}{A CYTOCHEMICAL FINE STRUCTURE
+STUDY OF PHAGOTROPHY IN A PLANKTONIC FORAMINIFER, {\emph{HASTIGERINA
+PELAGICA}} (d'ORBIGNY)}. \emph{The Biological Bulletin}, \textbf{151},
+437--449.
 
 \bibitem[\citeproctext]{ref-barth2022}
-Barth, A., Stone, J., 2022.
+Barth, A. and Stone, J. (2022)
 \href{https://www.frontiersin.org/articles/10.3389/fmars.2022.898057}{Comparison
 of an in situ imaging device and net-based method to study
-mesozooplankton communities in an oligotrophic system}. Frontiers in
-Marine Science 9.
+mesozooplankton communities in an oligotrophic system}. \emph{Frontiers
+in Marine Science}, \textbf{9}.
+
+\bibitem[\citeproctext]{ref-barth2024}
+Barth, A. and Stone, J. (2024)
+\href{https://doi.org/10.1093/plankt/fbae023}{Understanding the picture:
+The promis and challenges of in-situ imagery data in the study of
+plankton ecology}. \emph{Journal of Plankton Research: Horizons}, 1--15.
 
 \bibitem[\citeproctext]{ref-benfield1996}
-Benfield, M.C., Davis, C.S., Wiebe, P.H., Gallager, S.M., Gregory Lough,
-R., Copley, N.J., 1996. Video Plankton Recorder estimates of copepod,
-pteropod and larvacean distributions from a stratified region of Georges
-Bank with comparative measurements from a MOCNESS sampler. Deep Sea
-Research Part II: Topical Studies in Oceanography 43, 1925--1945.
-\url{https://doi.org/10.1016/S0967-0645(96)00044-6}
+Benfield, M. C., Davis, C. S., Wiebe, P. H., Gallager, S. M., Gregory
+Lough, R., and Copley, N. J. (1996)
+\href{https://doi.org/10.1016/S0967-0645(96)00044-6}{Video Plankton
+Recorder estimates of copepod, pteropod and larvacean distributions from
+a stratified region of Georges Bank with comparative measurements from a
+MOCNESS sampler}. \emph{Deep Sea Research Part II: Topical Studies in
+Oceanography}, \textbf{43}, 1925--1945.
+
+\bibitem[\citeproctext]{ref-biard2022b}
+Biard, T. (2022a)
+\href{https://doi.org/10.1111/1462-2920.16004}{Diversity and ecology of
+Radiolaria in modern oceans}. \emph{Environmental Microbiology},
+\textbf{24}, 2179--2200.
 
 \bibitem[\citeproctext]{ref-biard2022}
-Biard, T., 2022a. Rhizaria. Encyclopeida of Life Sciences, pp. 1--11.
-\url{https://doi.org/10.1002/9780470015902.a0029469}
-
-\bibitem[\citeproctext]{ref-biard2022b}
-Biard, T., 2022b. Diversity and ecology of Radiolaria in modern oceans.
-Environmental Microbiology 24, 2179--2200.
-\url{https://doi.org/10.1111/1462-2920.16004}
+Biard, T. (2022b)
+\href{https://doi.org/10.1002/9780470015902.a0029469}{Rhizaria}.
+Encyclopeida of Life Sciences, pp. 1--11.
 
@@
 Biard, T., Bigeard, E., Audic, S., Poulain, J., Gutierrez-Rodriguez, A.,
-Pesant, S., Stemmann, L., Not, F., 2017. Biogeography and diversity of
-Collodaria (Radiolaria) in the global ocean. The ISME Journal 11,
-1331--1344. \url{https://doi.org/10.1038/ismej.2017.12}
+Pesant, S., Stemmann, L., and Not, F. (2017)
+\href{https://doi.org/10.1038/ismej.2017.12}{Biogeography and diversity
+of Collodaria (Radiolaria) in the global ocean}. \emph{The ISME
+Journal}, \textbf{11}, 1331--1344.
 
 \bibitem[\citeproctext]{ref-biard2018}
-Biard, T., Krause, J.W., Stukel, M.R., Ohman, M.D., 2018. The
-Significance of Giant Phaeodarians (Rhizaria) to Biogenic Silica Export
-in the California Current Ecosystem. Global Biogeochemical Cycles 32,
-987--1004. \url{https://doi.org/10.1029/2018GB005877}
+Biard, T., Krause, J. W., Stukel, M. R., and Ohman, M. D. (2018)
+\href{https://doi.org/10.1029/2018GB005877}{The Significance of Giant
+Phaeodarians (Rhizaria) to Biogenic Silica Export in the California
+Current Ecosystem}. \emph{Global Biogeochemical Cycles}, \textbf{32},
+987--1004.
 
 \bibitem[\citeproctext]{ref-biard2020}
-Biard, T., Ohman, M.D., 2020. Vertical niche definition of test-bearing
-protists (Rhizaria) into the twilight zone revealed by in situ imaging.
-Limnology and Oceanography 65, 2583--2602.
-\url{https://doi.org/10.1002/lno.11472}
+Biard, T. and Ohman, M. D. (2020)
+\href{https://doi.org/10.1002/lno.11472}{Vertical niche definition of
+test-bearing protists (Rhizaria) into the twilight zone revealed by in
+situ imaging}. \emph{Limnology and Oceanography}, \textbf{65},
+2583--2602.
 
@@
 Biard, T., Stemmann, L., Picheral, M., Mayot, N., Vandromme, P., Hauss,
-H., Gorsky, G., Guidi, L., Kiko, R., Not, F., 2016. In situ imaging
-reveals the biomass of giant protists in the global ocean. Nature 532,
-504--507. \url{https://doi.org/10.1038/nature17652}
+H., Gorsky, G., Guidi, L., \emph{et al.} (2016)
+\href{https://doi.org/10.1038/nature17652}{In situ imaging reveals the
+biomass of giant protists in the global ocean}. \emph{Nature},
+\textbf{532}, 504--507.
 
 \bibitem[\citeproctext]{ref-bijma1990}
-Bijma, J., Erez, J., Hemleben, C., 1990.
+Bijma, J., Erez, J., and Hemleben, C. (1990)
 \href{https://epic.awi.de/id/eprint/6096/1/Bij1990a.pdf}{Lunar and
 semi-luan reproductive cycles in some spinose planktonic foraminifers}.
-Journal of Foraminiferal Research 20, 117--127.
+\emph{Journal of Foraminiferal Research}, \textbf{20}, 117--127.
 
 \bibitem[\citeproctext]{ref-blanco-bercial2020}
-Blanco-Bercial, L., 2020.
+Blanco-Bercial, L. (2020)
 \href{https://www.frontiersin.org/articles/10.3389/fmars.2020.00173}{Metabarcoding
 analyses and seasonality of the zooplankton community at BATS}.
-Frontiers in Marine Science 7.
+\emph{Frontiers in Marine Science}, \textbf{7}.
 
 \bibitem[\citeproctext]{ref-blanco-bercial2022}
-Blanco-Bercial, L., Parsons, R., Bolaños, L.M., Johnson, R., Giovannoni,
-S.J., Curry, R., 2022.
+Blanco-Bercial, L., Parsons, R., Bolaños, L. M., Johnson, R.,
+Giovannoni, S. J., and Curry, R. (2022)
 \href{https://www.frontiersin.org/articles/10.3389/fmars.2022.897140}{The
 protist community traces seasonality and mesoscale hydrographic features
-in the oligotrophic sargasso sea}. Frontiers in Marine Science 9.
+in the oligotrophic sargasso sea}. \emph{Frontiers in Marine Science},
+\textbf{9}.
 
 \bibitem[\citeproctext]{ref-boltovskoy2017}
-Boltovskoy, D., 2017. Vertical distribution patterns of radiolaria
-polycystina (protista) in the world ocean: Living ranges, isothermal
-submersion and settling shells. Journal of Plankton Research 39,
-330--349. \url{https://doi.org/10.1093/plankt/fbx003}
+Boltovskoy, D. (2017)
+\href{https://doi.org/10.1093/plankt/fbx003}{Vertical distribution
+patterns of radiolaria polycystina (protista) in the world ocean: Living
+ranges, isothermal submersion and settling shells}. \emph{Journal of
+Plankton Research}, \textbf{39}, 330--349.
 
 \bibitem[\citeproctext]{ref-boltovskoy1993}
-Boltovskoy, D., Alder, V.A., Abelmann, A., 1993. Annual flux of
+Boltovskoy, D., Alder, V. A., and Abelmann, A. (1993)
+\href{https://doi.org/10.1016/0967-0637(93)90036-3}{Annual flux of
 radiolaria and other shelled plankters in the eastern equatorial
 atlantic at 853 m: seasonal variations and polycystine species-specific
-responses. Deep Sea Research Part I: Oceanographic Research Papers 40,
-1863--1895. \url{https://doi.org/10.1016/0967-0637(93)90036-3}
+responses}. \emph{Deep Sea Research Part I: Oceanographic Research
+Papers}, \textbf{40}, 1863--1895.
 
 \bibitem[\citeproctext]{ref-caron2016}
-Caron, D.A., 2016.
+Caron, D. A. (2016)
 \href{https://www.nature.com/articles/nature17892}{The rise of rhizaria
-\textbar{} nature}. Nature 532, 444--445.
+\textbar{} nature}. \emph{Nature}, \textbf{532}, 444--445.
 
 \bibitem[\citeproctext]{ref-caron1984}
-Caron, D.A., Be, A.W.H., 1984. PREDICTED AND OBSERVED FEEDING RATES OF
-THE SPINOSE PLANKTONIC FORAMINIFER. BULLETIN OF MARINE SCIENCE 35.
+Caron, D. A. and Be, A. W. H. (1984) PREDICTED AND OBSERVED FEEDING
+RATES OF THE SPINOSE PLANKTONIC FORAMINIFER. \emph{BULLETIN OF MARINE
+SCIENCE}, \textbf{35}.
 
 \bibitem[\citeproctext]{ref-caron1995}
-Caron, D.A., Michaels, A.F., Swanberg, N.R., Howse, F.A., 1995. Primary
-productivity by symbiont-bearing planktonic sarcodines (acantharia,
-radiolaria, foraminifera) in surface waters near bermuda. Journal of
-Plankton Research 17, 103--129.
-\url{https://doi.org/10.1093/plankt/17.1.103}
+Caron, D. A., Michaels, A. F., Swanberg, N. R., and Howse, F. A. (1995)
+\href{https://doi.org/10.1093/plankt/17.1.103}{Primary productivity by
+symbiont-bearing planktonic sarcodines (acantharia, radiolaria,
+foraminifera) in surface waters near bermuda}. \emph{Journal of Plankton
+Research}, \textbf{17}, 103--129.
 
 \bibitem[\citeproctext]{ref-cohen2023}
-Cohen, N.R., Krinos, A.I., Kell, R.M., Chmiel, R.J., Moran, D.M.,
-McIlvin, M.R., Lopez, P.Z., Barth, A., Stone, J., Alanis, B.A., Chan,
-E.W., Breier, J.A., Jakuba, M.V., Johnson, R., Alexander, H., Saito,
-M.A., 2023. Microeukaryote metabolism across the western North Atlantic
-Ocean revealed through autonomous underwater profiling.
-\url{https://doi.org/10.1101/2023.11.20.567900}
+Cohen, N. R., Krinos, A. I., Kell, R. M., Chmiel, R. J., Moran, D. M.,
+McIlvin, M. R., Lopez, P. Z., Barth, A., \emph{et al.} (2023)
+\href{https://doi.org/10.1101/2023.11.20.567900}{Microeukaryote
+metabolism across the western North Atlantic Ocean revealed through
+autonomous underwater profiling}.
 
 \bibitem[\citeproctext]{ref-decelle2015}
-Decelle, J., Colin, S., Foster, R.A., 2015. Photosymbiosis in Marine
-Planktonic Protists, in: Ohtsuka, S., Suzaki, T., Horiguchi, T., Suzuki,
-N., Not, F. (Eds.),. Springer Japan, Tokyo, pp. 465--500.
-\url{https://doi.org/10.1007/978-4-431-55130-0_19}
+Decelle, J., Colin, S., and Foster, R. A. (2015)
+\href{https://doi.org/10.1007/978-4-431-55130-0_19}{Photosymbiosis in
+Marine Planktonic Protists}. In Ohtsuka, S., Suzaki, T., Horiguchi, T.,
+Suzuki, N., and Not, F. (eds). Springer Japan, Tokyo, pp. 465--500.
 
 \bibitem[\citeproctext]{ref-decelle2013}
-Decelle, J., Martin, P., Paborstava, K., Pond, D.W., Tarling, G., Mahé,
-F., De Vargas, C., Lampitt, R., Not, F., 2013. Diversity, Ecology and
-Biogeochemistry of Cyst-Forming Acantharia (Radiolaria) in the Oceans.
-PLoS ONE 8, e53598. \url{https://doi.org/10.1371/journal.pone.0053598}
+Decelle, J., Martin, P., Paborstava, K., Pond, D. W., Tarling, G., Mahé,
+F., De Vargas, C., Lampitt, R., \emph{et al.} (2013)
+\href{https://doi.org/10.1371/journal.pone.0053598}{Diversity, Ecology
+and Biogeochemistry of Cyst-Forming Acantharia (Radiolaria) in the
+Oceans}. \emph{PLoS ONE}, \textbf{8}, e53598.
 
 \bibitem[\citeproctext]{ref-decelle2012}
-Decelle, J., Siano, R., Probert, I., Poirier, C., Not, F., 2012.
-Multiple microalgal partners in symbiosis with the acantharian
-Acanthochiasma sp. (Radiolaria). Symbiosis 58, 233--244.
-\url{https://doi.org/10.1007/s13199-012-0195-x}
+Decelle, J., Siano, R., Probert, I., Poirier, C., and Not, F. (2012)
+\href{https://doi.org/10.1007/s13199-012-0195-x}{Multiple microalgal
+partners in symbiosis with the acantharian Acanthochiasma sp.
+(Radiolaria)}. \emph{Symbiosis}, \textbf{58}, 233--244.
 
 \bibitem[\citeproctext]{ref-dennett2002}
-Dennett, M.R., Caron, D.A., Michaels, A.F., Gallager, S.M., Davis, C.S.,
-2002. Video plankton recorder reveals high abundances of colonial
-radiolaria in surface waters of the central north pacific. Journal of
-Plankton Research 24, 797--805.
-\url{https://doi.org/10.1093/plankt/24.8.797}
+Dennett, M. R., Caron, D. A., Michaels, A. F., Gallager, S. M., and
+Davis, C. S. (2002) \href{https://doi.org/10.1093/plankt/24.8.797}{Video
+plankton recorder reveals high abundances of colonial radiolaria in
+surface waters of the central north pacific}. \emph{Journal of Plankton
+Research}, \textbf{24}, 797--805.
 
@@
 Drago, L., Panaïotis, T., Irisson, J.-O., Babin, M., Biard, T.,
-Carlotti, F., Coppola, L., Guidi, L., Hauss, H., Karp-Boss, L., Lombard,
-F., Mcdonnell, A.M.P., Picheral, M., Rogge, A., Waite, A.M., Stemmann,
-L., Kiko, R., 2022. Global Distribution of Zooplankton Biomass Estimated
-by In Situ Imaging and Machine Learning. Frontiers in Marine Science 9.
-\url{https://doi.org/10.3389/fmars.2022.894372}
+Carlotti, F., Coppola, L., Guidi, L., \emph{et al.} (2022)
+\href{https://doi.org/10.3389/fmars.2022.894372}{Global Distribution of
+Zooplankton Biomass Estimated by In Situ Imaging and Machine Learning}.
+\emph{Frontiers in Marine Science}, \textbf{9}.
 
 \bibitem[\citeproctext]{ref-gaskell2019}
-Gaskell, D.E., Ohman, M.D., Hull, P.M., 2019. Zooglider-based
+Gaskell, D. E., Ohman, M. D., and Hull, P. M. (2019)
+\href{https://doi.org/10.2113/gsjfr.49.4.390}{Zooglider-based
 measurements of planktonic foraminifera in the california current
-system. Journal of Foraminiferal Research 49, 390--404.
-\url{https://doi.org/10.2113/gsjfr.49.4.390}
+system}. \emph{Journal of Foraminiferal Research}, \textbf{49},
+390--404.
 
 \bibitem[\citeproctext]{ref-gorsky2010}
-Gorsky, G., Ohman, M.D., Picheral, M., Gasparini, S., Stemmann, L.,
-Romagnan, J.-B., Cawood, A., Pesant, S., García-Comas, C., Prejger, F.,
-2010. Digital zooplankton image analysis using the ZooScan integrated
-system. Journal of Plankton Research 32, 285--303.
-\url{https://doi.org/10.1093/plankt/fbp124}
+Gorsky, G., Ohman, M. D., Picheral, M., Gasparini, S., Stemmann, L.,
+Romagnan, J.-B., Cawood, A., Pesant, S., \emph{et al.} (2010)
+\href{https://doi.org/10.1093/plankt/fbp124}{Digital zooplankton image
+analysis using the ZooScan integrated system}. \emph{Journal of Plankton
+Research}, \textbf{32}, 285--303.
 
 \bibitem[\citeproctext]{ref-gowing1989}
-Gowing, M.M., 1989. Abundance and feeding ecology of Antarctic
-phaeodarian radiolarians. Marine Biology 103, 107--118.
-\url{https://doi.org/10.1007/BF00391069}
+Gowing, M. M. (1989) \href{https://doi.org/10.1007/BF00391069}{Abundance
+and feeding ecology of Antarctic phaeodarian radiolarians}. \emph{Marine
+Biology}, \textbf{103}, 107--118.
 
 \bibitem[\citeproctext]{ref-gutierrez-rodriguez2019}
-Gutierrez-Rodriguez, A., Stukel, M.R., Lopes dos Santos, A., Biard, T.,
-Scharek, R., Vaulot, D., Landry, M.R., Not, F., 2019. High contribution
-of Rhizaria (Radiolaria) to vertical export in the California Current
-Ecosystem revealed by DNA metabarcoding. The ISME Journal 13, 964--976.
-\url{https://doi.org/10.1038/s41396-018-0322-7}
+Gutierrez-Rodriguez, A., Stukel, M. R., Lopes dos Santos, A., Biard, T.,
+Scharek, R., Vaulot, D., Landry, M. R., and Not, F. (2019)
+\href{https://doi.org/10.1038/s41396-018-0322-7}{High contribution of
+Rhizaria (Radiolaria) to vertical export in the California Current
+Ecosystem revealed by DNA metabarcoding}. \emph{The ISME Journal},
+\textbf{13}, 964--976.
 
@@
 Gutiérrez-Rodríguez, A., Lopes dos Santos, A., Safi, K., Probert, I.,
-Not, F., Fernández, D., Gourvil, P., Bilewitch, J., Hulston, D.,
-Pinkerton, M., Nodder, S.D., 2022. Planktonic protist diversity across
-contrasting subtropical and subantarctic waters of the southwest
-pacific. Progress in Oceanography 206, 102809.
-\url{https://doi.org/10.1016/j.pocean.2022.102809}
+Not, F., Fernández, D., Gourvil, P., Bilewitch, J., \emph{et al.} (2022)
+\href{https://doi.org/10.1016/j.pocean.2022.102809}{Planktonic protist
+diversity across contrasting subtropical and subantarctic waters of the
+southwest pacific}. \emph{Progress in Oceanography}, \textbf{206},
+102809.
 
 \bibitem[\citeproctext]{ref-haekel1887}
-Haekel, E., 1887.
-\href{https://www.gutenberg.org/cache/epub/44527/pg44527-images.html}{Report
+Haekel, E. (1887)
+\emph{\href{https://www.gutenberg.org/cache/epub/44527/pg44527-images.html}{Report
 on the radiolaria collected by h.m.s. Challenger during the years
-1873-1876, plates, by ernst haeckel}.
+1873-1876, plates, by ernst haeckel}}.
 
 \bibitem[\citeproctext]{ref-hull2011}
-Hull, P.M., Osborn, K.J., Norris, R.D., Robison, B.H., 2011. Seasonality
-and depth distribution of a mesopelagic foraminifer, Hastigerinella
-digitata, in Monterey Bay, California. Limnology and Oceanography 56,
-562--576. \url{https://doi.org/10.4319/lo.2011.56.2.0562}
+Hull, P. M., Osborn, K. J., Norris, R. D., and Robison, B. H. (2011)
+\href{https://doi.org/10.4319/lo.2011.56.2.0562}{Seasonality and depth
+distribution of a mesopelagic foraminifer, Hastigerinella digitata, in
+Monterey Bay, California}. \emph{Limnology and Oceanography},
+\textbf{56}, 562--576.
 
 \bibitem[\citeproctext]{ref-ikenoue2019}
-Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M.C., Takahashi,
-K., Harada, N., Fujiki, T., 2019. Phaeodaria: An Important Carrier of
-Particulate Organic Carbon in the Mesopelagic Twilight Zone of the North
-Pacific Ocean. Global Biogeochemical Cycles 33, 1146--1160.
-\url{https://doi.org/10.1029/2019GB006258}
+Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M. C., Takahashi,
+K., Harada, N., and Fujiki, T. (2019)
+\href{https://doi.org/10.1029/2019GB006258}{Phaeodaria: An Important
+Carrier of Particulate Organic Carbon in the Mesopelagic Twilight Zone
+of the North Pacific Ocean}. \emph{Global Biogeochemical Cycles},
+\textbf{33}, 1146--1160.
 
 \bibitem[\citeproctext]{ref-kimoto2015}
-Kimoto, K., 2015. Planktic foraminifera. pp. 129--178.
-\url{https://doi.org/10.1007/978-4-431-55130-0_7}
+Kimoto, K. (2015)
+\href{https://doi.org/10.1007/978-4-431-55130-0_7}{Planktic
+foraminifera}. pp. 129--178.
 
 \bibitem[\citeproctext]{ref-knap1997}
-Knap, A.H., Michaels, A.F., Steinberg, D.K., Bahr, F., Bates, N.R.,
-Bell, S., Countway, P., Close, A.R., Doyle, A.P., Dow, R.L., Howse,
-F.A., Gundersen, K., Johnson, R.J., Kelly, R., Little, R., Orcutt, K.,
-Parsons, R., Rathburn, C., Sanderson, M., Stone, S., 1997.
+Knap, A. H., Michaels, A. F., Steinberg, D. K., Bahr, F., Bates, N. R.,
+Bell, S., Countway, P., Close, A. R., \emph{et al.} (1997)
 \href{https://eprints.soton.ac.uk/361194/}{BATS methods manual, version
@@
 Laget, M., Drago, L., Panaïotis, T., Kiko, R., Stemmann, L., Rogge, A.,
-Llopis-Monferrer, N., Leynaert, A., Irisson, J.-O., Biard, T., 2024.
-Global census of the significance of giant mesopelagic protists to the
-marine carbon and silicon cycles. Nature Communications 15, 3341.
-\url{https://doi.org/10.1038/s41467-024-47651-4}
+Llopis-Monferrer, N., Leynaert, A., \emph{et al.} (2024)
+\href{https://doi.org/10.1038/s41467-024-47651-4}{Global census of the
+significance of giant mesopelagic protists to the marine carbon and
+silicon cycles}. \emph{Nature Communications}, \textbf{15}, 3341.
 
 \bibitem[\citeproctext]{ref-lee2006}
-Lee, J.J., 2006.
+Lee, J. J. (2006)
 \href{https://dalspace.library.dal.ca/bitstream/handle/10222/78255/VOLUME\%2042-NUMBER\%202-2006-PAGE\%2063.pdf?sequence=1}{Algal
-symbiosis in larger foraminifera}. Symbiosis 42, 63--75.
+symbiosis in larger foraminifera}. \emph{Symbiosis}, \textbf{42},
+63--75.
 
 \bibitem[\citeproctext]{ref-llopismonferrer2022}
-Llopis Monferrer, N., Biard, T., Sandin, M.M., Lombard, F., Picheral,
-M., Elineau, A., Guidi, L., Leynaert, A., Tréguer, P.J., Not, F., 2022.
+Llopis Monferrer, N., Biard, T., Sandin, M. M., Lombard, F., Picheral,
+M., Elineau, A., Guidi, L., Leynaert, A., \emph{et al.} (2022)
 \href{https://www.frontiersin.org/articles/10.3389/fmars.2022.895995}{Siliceous
 rhizaria abundances and diversity in the mediterranean sea assessed by
-combined imaging and metabarcoding approaches}. Frontiers in Marine
-Science 9.
+combined imaging and metabarcoding approaches}. \emph{Frontiers in
+Marine Science}, \textbf{9}.
 
 \bibitem[\citeproctext]{ref-lomas2013}
-Lomas, M.W., Bates, N.R., Johnson, R.J., Knap, A.H., Steinberg, D.K.,
-Carlson, C.A., 2013. Two decades and counting: 24-years of sustained
-open ocean biogeochemical measurements in the Sargasso Sea. Deep Sea
-Research Part II: Topical Studies in Oceanography 93, 16--32.
-\url{https://doi.org/10.1016/j.dsr2.2013.01.008}
+Lomas, M. W., Bates, N. R., Johnson, R. J., Knap, A. H., Steinberg, D.
+K., and Carlson, C. A. (2013)
+\href{https://doi.org/10.1016/j.dsr2.2013.01.008}{Two decades and
+counting: 24-years of sustained open ocean biogeochemical measurements
+in the Sargasso Sea}. \emph{Deep Sea Research Part II: Topical Studies
+in Oceanography}, \textbf{93}, 16--32.
 
 \bibitem[\citeproctext]{ref-lombard2019}
-Lombard, F., Boss, E., Waite, A.M., Vogt, M., Uitz, J., Stemmann, L.,
-Sosik, H.M., Schulz, J., Romagnan, J.-B., Picheral, M., Pearlman, J.,
-Ohman, M.D., Niehoff, B., Möller, K.O., Miloslavich, P., Lara-Lpez, A.,
-Kudela, R., Lopes, R.M., Kiko, R., Karp-Boss, L., Jaffe, J.S., Iversen,
-M.H., Irisson, J.-O., Fennel, K., Hauss, H., Guidi, L., Gorsky, G.,
-Giering, S.L.C., Gaube, P., Gallager, S., Dubelaar, G., Cowen, R.K.,
-Carlotti, F., Briseño-Avena, C., Berline, L., Benoit-Bird, K., Bax, N.,
-Batten, S., Ayata, S.D., Artigas, L.F., Appeltans, W., 2019.
+Lombard, F., Boss, E., Waite, A. M., Vogt, M., Uitz, J., Stemmann, L.,
+Sosik, H. M., Schulz, J., \emph{et al.} (2019)
 \href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00196}{Globally
 consistent quantitative observations of planktonic ecosystems}.
-Frontiers in Marine Science 6.
+\emph{Frontiers in Marine Science}, \textbf{6}.
 
 \bibitem[\citeproctext]{ref-marra2011}
-Marra, G., Wood, S.N., 2011. Practical variable selection for
-generalized additive models. Computational Statistics \& Data Analysis
-55, 2372--2387. \url{https://doi.org/10.1016/j.csda.2011.02.004}
+Marra, G. and Wood, S. N. (2011)
+\href{https://doi.org/10.1016/j.csda.2011.02.004}{Practical variable
+selection for generalized additive models}. \emph{Computational
+Statistics \& Data Analysis}, \textbf{55}, 2372--2387.
 
 \bibitem[\citeproctext]{ref-marsbrisbin2020}
-Mars Brisbin, M., Brunner, O.D., Grossmann, M.M., Mitarai, S., 2020.
-Paired high-throughput, in situ imaging and high-throughput sequencing
-illuminate acantharian abundance and vertical distribution. Limnology
-and Oceanography 65, 2953--2965. \url{https://doi.org/10.1002/lno.11567}
+Mars Brisbin, M., Brunner, O. D., Grossmann, M. M., and Mitarai, S.
+(2020) \href{https://doi.org/10.1002/lno.11567}{Paired high-throughput,
+in situ imaging and high-throughput sequencing illuminate acantharian
+abundance and vertical distribution}. \emph{Limnology and Oceanography},
+\textbf{65}, 2953--2965.
 
 \bibitem[\citeproctext]{ref-mcgillicuddy1998}
-McGillicuddy, D.J., Robinson, A.R., Siegel, D.A., Jannasch, H.W.,
-Johnson, R., Dickey, T.D., McNeil, J., Michaels, A.F., Knap, A.H., 1998.
-Influence of mesoscale eddies on new production in the Sargasso Sea.
-Nature 394, 263--266. \url{https://doi.org/10.1038/28367}
+McGillicuddy, D. J., Robinson, A. R., Siegel, D. A., Jannasch, H. W.,
+Johnson, R., Dickey, T. D., McNeil, J., Michaels, A. F., \emph{et al.}
+(1998) \href{https://doi.org/10.1038/28367}{Influence of mesoscale
+eddies on new production in the Sargasso Sea}. \emph{Nature},
+\textbf{394}, 263--266.
 
 \bibitem[\citeproctext]{ref-michaels1988}
-Michaels, A.F., 1988. Vertical distribution and abundance of Acantharia
-and their symbionts. Marine Biology 97, 559--569.
-\url{https://doi.org/10.1007/BF00391052}
+Michaels, A. F. (1988)
+\href{https://doi.org/10.1007/BF00391052}{Vertical distribution and
+abundance of Acantharia and their symbionts}. \emph{Marine Biology},
+\textbf{97}, 559--569.
 
 \bibitem[\citeproctext]{ref-michaels1995}
-Michaels, A.F., Caron, D.A., Swanberg, N.R., Howse, F.A., Michaels,
-C.M., 1995. Planktonic sarcodines (Acantharia, Radiolaria, Foraminifera)
-in surface waters near Bermuda: abundance, biomass and vertical flux.
-Journal of Plankton Research 17, 131--163.
-\url{https://doi.org/10.1093/plankt/17.1.131}
+Michaels, A. F., Caron, D. A., Swanberg, N. R., Howse, F. A., and
+Michaels, C. M. (1995)
+\href{https://doi.org/10.1093/plankt/17.1.131}{Planktonic sarcodines
+(Acantharia, Radiolaria, Foraminifera) in surface waters near Bermuda:
+abundance, biomass and vertical flux}. \emph{Journal of Plankton
+Research}, \textbf{17}, 131--163.
 
 \bibitem[\citeproctext]{ref-michaels1996}
-Michaels, A.F., Knap, A.H., 1996. Overview of the U.S. JGOFS Bermuda
-Atlantic Time-series Study and the Hydrostation S program. Deep Sea
-Research Part II: Topical Studies in Oceanography 43, 157--198.
-\url{https://doi.org/10.1016/0967-0645(96)00004-5}
+Michaels, A. F. and Knap, A. H. (1996)
+\href{https://doi.org/10.1016/0967-0645(96)00004-5}{Overview of the U.S.
+JGOFS Bermuda Atlantic Time-series Study and the Hydrostation S
+program}. \emph{Deep Sea Research Part II: Topical Studies in
+Oceanography}, \textbf{43}, 157--198.
 
@@
 Nakamura, Y., Itagaki, H., Tuji, A., Shimode, S., Yamaguchi, A., Hidaka,
-K., Ogiso-Tanaka, E., 2023. DNA metabarcoding focused on
-difficult-to-culture protists: An effective approach to clarify
-biological interactions. Environmental Microbiology 25, 3630--3638.
-\url{https://doi.org/10.1111/1462-2920.16524}
+K., and Ogiso-Tanaka, E. (2023)
+\href{https://doi.org/10.1111/1462-2920.16524}{DNA metabarcoding focused
+on difficult-to-culture protists: An effective approach to clarify
+biological interactions}. \emph{Environmental Microbiology},
+\textbf{25}, 3630--3638.
 
 \bibitem[\citeproctext]{ref-nakamura2015}
-Nakamura, Y., Suzuki, N., 2015.
+Nakamura, Y. and Suzuki, N. (2015)
 \href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_9}{Phaeodaria:
@@
 \bibitem[\citeproctext]{ref-ohman2019a}
-Ohman, M.D., 2019. A sea of tentacles: optically discernible traits
-resolved from planktonic organisms in situ. ICES Journal of Marine
-Science 76, 1959--1972. \url{https://doi.org/10.1093/icesjms/fsz184}
+Ohman, M. D. (2019) \href{https://doi.org/10.1093/icesjms/fsz184}{A sea
+of tentacles: optically discernible traits resolved from planktonic
+organisms in situ}. \emph{ICES Journal of Marine Science}, \textbf{76},
+1959--1972.
 
@@
 Panaïotis, T., Babin, M., Biard, T., Carlotti, F., Coppola, L., Guidi,
-L., Hauss, H., Karp-Boss, L., Kiko, R., Lombard, F., McDonnell, A.M.P.,
-Picheral, M., Rogge, A., Waite, A.M., Stemmann, L., Irisson, J.-O.,
-2023. Three major mesoplanktonic communities resolved by in situ imaging
-in the upper 500 m of the global ocean. Global Ecology and Biogeography
-32, 1991--2005. \url{https://doi.org/10.1111/geb.13741}
+L., Hauss, H., Karp-Boss, L., \emph{et al.} (2023)
+\href{https://doi.org/10.1111/geb.13741}{Three major mesoplanktonic
+communities resolved by in situ imaging in the upper 500 m of the global
+ocean}. \emph{Global Ecology and Biogeography}, \textbf{32}, 1991--2005.
 
 \bibitem[\citeproctext]{ref-picheral}
-Picheral, M., Colin, S.P., Irisson, J.-O., n.d.
+Picheral, M., Colin, S. P., and Irisson, J.-O.
 \href{https://ecotaxa.obs-vlfr.fr/}{Ecotaxa}.
@@
 \bibitem[\citeproctext]{ref-picheral2010}
-Picheral, M., Guidi, L., Stemmann, L., Karl, D.M., Iddaoud, G., Gorsky,
-G., 2010. The Underwater Vision Profiler 5: An advanced instrument for
-high spatial resolution studies of particle size spectra and
-zooplankton. Limnology and Oceanography: Methods 8, 462--473.
-\url{https://doi.org/10.4319/lom.2010.8.462}
+Picheral, M., Guidi, L., Stemmann, L., Karl, D. M., Iddaoud, G., and
+Gorsky, G. (2010) \href{https://doi.org/10.4319/lom.2010.8.462}{The
+Underwater Vision Profiler 5: An advanced instrument for high spatial
+resolution studies of particle size spectra and zooplankton}.
+\emph{Limnology and Oceanography: Methods}, \textbf{8}, 462--473.
 
@@
 Sogawa, S., Nakamura, Y., Nagai, S., Nishi, N., Hidaka, K., Shimizu, Y.,
-Setou, T., 2022. DNA metabarcoding reveals vertical variation and hidden
-diversity of alveolata and rhizaria communities in the western north
-pacific. Deep Sea Research Part I: Oceanographic Research Papers 183,
-103765. \url{https://doi.org/10.1016/j.dsr.2022.103765}
+and Setou, T. (2022)
+\href{https://doi.org/10.1016/j.dsr.2022.103765}{DNA metabarcoding
+reveals vertical variation and hidden diversity of alveolata and
+rhizaria communities in the western north pacific}. \emph{Deep Sea
+Research Part I: Oceanographic Research Papers}, \textbf{183}, 103765.
 
 \bibitem[\citeproctext]{ref-stukel2018}
-Stukel, M.R., Biard, T., Krause, J., Ohman, M.D., 2018. Large Phaeodaria
-in the twilight zone: Their role in the carbon cycle. Limnology and
-Oceanography 63, 2579--2594. \url{https://doi.org/10.1002/lno.10961}
+Stukel, M. R., Biard, T., Krause, J., and Ohman, M. D. (2018)
+\href{https://doi.org/10.1002/lno.10961}{Large Phaeodaria in the
+twilight zone: Their role in the carbon cycle}. \emph{Limnology and
+Oceanography}, \textbf{63}, 2579--2594.
 
 \bibitem[\citeproctext]{ref-stukel2019}
-Stukel, M.R., Ohman, M.D., Kelly, T.B., Biard, T., 2019.
+Stukel, M. R., Ohman, M. D., Kelly, T. B., and Biard, T. (2019)
 \href{https://www.frontiersin.org/articles/10.3389/fmars.2019.00397}{The
@@
 of particle flux into the mesopelagic ocean in the northeast pacific}.
-Frontiers in Marine Science 6.
+\emph{Frontiers in Marine Science}, \textbf{6}.
 
 \bibitem[\citeproctext]{ref-suzuki2015}
-Suzuki, N., Not, F., 2015.
+Suzuki, N. and Not, F. (2015)
 \href{https://link.springer.com/chapter/10.1007/978-4-431-55130-0_8}{Biology
@@
 \bibitem[\citeproctext]{ref-swanberg1981}
-Swanberg, N.R., Anderson, O.R., 1981. Collozoum caudatum sp. nov.: A
-giant colonial radiolarian from equatorial and Gulf Stream waters. Deep
-Sea Research Part A. Oceanographic Research Papers 28, 1033--1047.
-\url{https://doi.org/10.1016/0198-0149(81)90016-9}
+Swanberg, N. R. and Anderson, O. R. (1981)
+\href{https://doi.org/10.1016/0198-0149(81)90016-9}{Collozoum caudatum
+sp. nov.: A giant colonial radiolarian from equatorial and Gulf Stream
+waters}. \emph{Deep Sea Research Part A. Oceanographic Research Papers},
+\textbf{28}, 1033--1047.
 
 \bibitem[\citeproctext]{ref-takahashi1983}
-Takahashi, K., Hurd, D.C., Honjo, S., 1983. Phaeodarian skeletons: Their
-role in silica transport to the deep sea. Science 222, 616--618.
-\url{https://doi.org/10.1126/science.222.4624.616}
+Takahashi, K., Hurd, D. C., and Honjo, S. (1983)
+\href{https://doi.org/10.1126/science.222.4624.616}{Phaeodarian
+skeletons: Their role in silica transport to the deep sea}.
+\emph{Science}, \textbf{222}, 616--618.
 
 \bibitem[\citeproctext]{ref-whitmore2021}
-Whitmore, B.M., Ohman, M.D., 2021. Zooglider-measured association of
-zooplankton with the fine-scale vertical prey field. Limnology and
-Oceanography 66, 3811--3827. \url{https://doi.org/10.1002/lno.11920}
+Whitmore, B. M. and Ohman, M. D. (2021)
+\href{https://doi.org/10.1002/lno.11920}{Zooglider-measured association
+of zooplankton with the fine-scale vertical prey field}. \emph{Limnology
+and Oceanography}, \textbf{66}, 3811--3827.
 
 \bibitem[\citeproctext]{ref-wood2017}
-Wood, S.N., 2017. Generalized additive models: an introduction with R,
-Second edition. ed, Texts in statistical science. CRC Press/Taylor \&
-Francis Group, Boca Raton London New York.
+Wood, S. N. (2017) \emph{Generalized additive models: an introduction
+with R}. Second edition. CRC Press/Taylor \& Francis Group, Boca Raton
+London New York.
 
 \bibitem[\citeproctext]{ref-wood2001}
-Wood, S.N., 2001. mgcv: GAMs and Generalized Ridge Regression for R 1.
+Wood, S. N. (2001) mgcv: GAMs and Generalized Ridge Regression for R.
+\textbf{1}.
 
 \bibitem[\citeproctext]{ref-zasko2005}
-Zasko, D.N., Rusanov, I.I., 2005. Vertical Distribution of Radiolarians
-and Their Role in Epipelagic Communities of the East Pacific Rise and
-the Gulf of California. Biology Bulletin 32, 279--287.
-\url{https://doi.org/10.1007/s10525-005-0103-5}
+Zasko, D. N. and Rusanov, I. I. (2005)
+\href{https://doi.org/10.1007/s10525-005-0103-5}{Vertical Distribution
+of Radiolarians and Their Role in Epipelagic Communities of the East
+Pacific Rise and the Gulf of California}. \emph{Biology Bulletin},
+\textbf{32}, 279--287.
 

### 2024-07-06 manuscript/04_discussion.qmd
change last sentence and add kim ref

@@
 The mesopelagic generally was home to known heterotrophic organisms, particularly for those which were constrained to exclusively occupy deeper waters.
-This is consistent with @blanco-bercial2022's observation of an auto-/mixotroph to heterotroph gradient in local protist community as well as global patterns in Rhizaria ecology \[@laget2024\].
+This is consistent with @blanco-bercial2022's observation of an auto-/mixotroph to heterotroph gradient in local protist community as well as global patterns in Rhizaria ecology [@laget2024].
 The upper mesopelagic interestingly had relatively low total abundance.
@@
 In our study, while consistently observed, overall abundances of Aulosphaeridae were very low.
-In the Pacific Ocean, on California's Coast, much higher abundances of Aulosphaeridae have been reported \[@zasko2005; @biard2020\] and they have massive potential to impact silica export [@biard2018].
+In the Pacific Ocean, on California's Coast, much higher abundances of Aulosphaeridae have been reported [@zasko2005; @biard2020] and they have massive potential to impact silica export [@biard2018].
 Coelodendridae were also seemingly restricted to the deeper section of the mesopelagic.
@@
 Furthermore, given that horizontal transfer of symbionts between generations of Acantharea is unlikely due to their spawning behavior, the newly spawned mesopelagic Acantharea are not necessarily required to rapidly return to the photic zone [@decelle2012; @decelle2013].
-This hypothesis predicts that Acanthareas in the mesopelagic would be smaller [@decelle2013].
+This hypothesis predicts that Acantharea in the mesopelagic would be smaller [@decelle2013].
 @marsbrisbin2020 provided some support for this hypothesis, with a significant decrease in Acantharea sizes with depth.
 Although the authors also observed low abundances in the mesopelagic and noted that the smaller sizes may be due to lower food availability [@marsbrisbin2020].
-Since food is more scarce in the mesopelagic, nutritional quality lower, yet flux feeders would likely grow larger to increase their feeding range [@stukel2018; @biard2020].
-In the data collecting in this study, Acanthareas in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth.
+Since food is more scarce in the mesopelagic, nutritional quality lower [@kim2018], yet flux feeders would likely grow larger to increase their feeding range [@stukel2018; @biard2020].
+In the data collecting in this study, Acantharea in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth.
 This provides added support for the hypothesis that cyst-forming Acantharea may utilize different sections of the water column throughout their life cycle.
@@
 However, their role in food webs is not well understood.
-While this study represents a step-forward in our understanding of Rhizaria ecology, continued research on Rhizaria is much needed to better understand their ecology.
-Particularly extended work describing interannual patterns in Rhizaria abundance.
-Also work defining biotic interactions, feeding rates, productivity, and life history are all rich fields of interest in Rhizaria.
+While this study represents a step forward in our understanding of Rhizaria, further research is needed to better characterize their ecology and role in biogeochemical cycling.
+To fully understand the functioning of the oceans, not only carbon or nitrogen, but other elements such as silica or strontium, extended work addressing the interannual patterns in Rhizaria, their biotic interactions, and the key biological rates that impact ocean biogeochemistry, will be required.

### 2024-07-06 manuscript/02_methods.qmd
minor edit and render

@@
 However, identified rhizaria ranged from a 934$\mu m$ ESD Aulacanthidae cell to a Collodarian colony over 10mm in diameter.
-To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the small size range [@lombard2019; @barth2024].
+To confirm that the UVP5 was sampling adequately across all size ranges, an normalized biomass size spectrum (NBSS) slope was constructed to identify a drop-off which would indicate poor-sampling at the smaller end of the size range [@lombard2019; @barth2024].
 However, it was evident from this analysis that all size ranges were adequately sampled across the size range (Supplemental Figure 2) so no data were excluded.
@@
 $m^{-2}$ respectively.
-All UVP data processing was done using the `EcotaxaTools` package in R (Barth 2023).
+All UVP data processing was done using the `EcotaxaTools` package in R (<https://github.com/TheAlexBarth/EcotaxaTools?tab=readme-ov-file>).
 

### 2024-07-06 manuscript/main.tex
minor edit and render

@@
 \usepackage{amsmath,amssymb}
+\usepackage{lmodern}
 \usepackage{setspace}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
 \fi
@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
@@
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
   % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 % Add any tex header commands here
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
 \makeatletter
@@
 \makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
 \ifLuaTeX
@@
 \fi
-\usepackage{bookmark}
-
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \end{minipage}  \end{titlepage}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\section{Methods}\label{methods}
-
-\subsection{Oceanographic sampling}\label{oceanographic-sampling}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+\hypertarget{oceanographic-sampling}{%
+\subsection{Oceanographic sampling}\label{oceanographic-sampling}}
 
@@
 
+\hypertarget{rhizaria-imaging-processing-and-abundance-quantification}{%
 \subsection{Rhizaria imaging processing and abundance
-quantification}\label{rhizaria-imaging-processing-and-abundance-quantification}
+quantification}\label{rhizaria-imaging-processing-and-abundance-quantification}}
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 
@@
 
-\end{figure}%
+\end{figure}
 
@@
 constructed to identify a drop-off which would indicate poor-sampling at
-the small size range (Lombard \emph{et al.}, 2019; Barth and Stone,
-2024). However, it was evident from this analysis that all size ranges
-were adequately sampled across the size range (Supplemental Figure 2) so
-no data were excluded. The UVP5 reports the exact depth at which a
-particle is recorded, however to estimate abundance, observations must
-be binned over fixed depth intervals. Our deployments had variable
-descent depths and speeds with more casts descending to 500m than 1000m
-and descents quicker through the epipelagic than the mesopelagic (see
-Barth and Stone (2022) for an extended discussion of UVP5 data
-processing). For the present study, Rhizaria abundances were estimated
-in 25m vertical bins, which offer a moderate sampling volume per bin
-(average 0.948\(m^3\) in the epipelagic and 0.589\(m^3\) in the
+the smaller end of the size range (Lombard \emph{et al.}, 2019; Barth
+and Stone, 2024). However, it was evident from this analysis that all
+size ranges were adequately sampled across the size range (Supplemental
+Figure 2) so no data were excluded. The UVP5 reports the exact depth at
+which a particle is recorded, however to estimate abundance,
+observations must be binned over fixed depth intervals. Our deployments
+had variable descent depths and speeds with more casts descending to
+500m than 1000m and descents quicker through the epipelagic than the
+mesopelagic (see Barth and Stone (2022) for an extended discussion of
+UVP5 data processing). For the present study, Rhizaria abundances were
+estimated in 25m vertical bins, which offer a moderate sampling volume
+per bin (average 0.948\(m^3\) in the epipelagic and 0.589\(m^3\) in the
 mesopelagic; Supplemental Table 1) while still maintaining ecologically
@@
 and 0.20 ind. \(m^{-2}\) respectively. All UVP data processing was done
-using the \texttt{EcotaxaTools} package in R (Barth 2023).
-
+using the \texttt{EcotaxaTools} package in R
+(\url{https://github.com/TheAlexBarth/EcotaxaTools?tab=readme-ov-file}).
+
+\hypertarget{modelling-environmental-controls-of-rhizaria-abundance}{%
 \subsection{Modelling environmental controls of Rhizaria
-abundance}\label{modelling-environmental-controls-of-rhizaria-abundance}
+abundance}\label{modelling-environmental-controls-of-rhizaria-abundance}}
 
@@
 
-\section{Results}\label{results}
-
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{environmental-variability}{%
 \subsubsection{Environmental
-Variability}\label{environmental-variability}
+Variability}\label{environmental-variability}}
 
@@
 
+\hypertarget{rhizaria-abundance-and-distribution}{%
 \subsubsection{Rhizaria abundance and
-distribution}\label{rhizaria-abundance-and-distribution}
+distribution}\label{rhizaria-abundance-and-distribution}}
 
@@
 
+\hypertarget{body-size-throughout-the-water-column.}{%
 \subsubsection{Body size throughout the water
-column.}\label{body-size-throughout-the-water-column.}
+column.}\label{body-size-throughout-the-water-column.}}
 
@@
 
-\end{figure}%
-
+\end{figure}
+
+\hypertarget{environmental-drivers-of-rhizaria-abundance}{%
 \subsubsection{Environmental Drivers of Rhizaria
-Abundance}\label{environmental-drivers-of-rhizaria-abundance}
-
-\begin{figure}[H]
+Abundance}\label{environmental-drivers-of-rhizaria-abundance}}
+
+\begin{figure}
 
@@
 of p = 0.1 or a large enough reduction in model fit.}\tabularnewline
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endfirsthead
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 \textbf{All Rhizaria Epipelagic} & Temperature & 0.7950 & 0.492 &
@@
 \textless0.001 \\
+\bottomrule()
 \end{longtable}
@@
 model fit.}\tabularnewline
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 \textbf{Acantharea Epipelagic} & Temperature & 0.868 & 0.380 & 0.0935 \\
@@
 & Particle Concentration & 1.685 & 2.955 & \textless0.001 \\
+\bottomrule()
 \end{longtable}
@@
 
-\section{Discussion}\label{discussion}
-
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{overall-rhizaria-abundance-and-patterns}{%
 \subsubsection{Overall Rhizaria abundance and
-patterns}\label{overall-rhizaria-abundance-and-patterns}
+patterns}\label{overall-rhizaria-abundance-and-patterns}}
 
@@
 
+\hypertarget{relationship-to-environmental-parameters}{%
 \subsubsection{Relationship to environmental
-parameters}\label{relationship-to-environmental-parameters}
+parameters}\label{relationship-to-environmental-parameters}}
 
@@
 
+\hypertarget{vertical-structure-and-trophic-roles}{%
 \subsubsection{Vertical Structure and Trophic
-Roles}\label{vertical-structure-and-trophic-roles}
+Roles}\label{vertical-structure-and-trophic-roles}}
 
@@
 Current Ecosystem (Biard and Ohman, 2020). It should be noted however,
-that the secondary abundance peak reported in the present study is
-lower. This is likely due to the more oligotrophic nature of the study
-region, were the euphotic zone penetrates deeper into the water column.
-Most prevalent in the epipelagic were Collodaria. These mixotrophic
-Radiolaria have long been reported to contribute to primary productivity
-in the euphotic zone (Michaels \emph{et al.}, 1995; Dennett \emph{et
-al.}, 2002). Collodaria are thought to be particularly successful
-globally in oligotrophic regions due to their photosymbiotic
+that the secondary abundance peak reported in the present study is lower
+in magnitude. This is likely due to the more oligotrophic nature of the
+study region, were the euphotic zone penetrates deeper into the water
+column. Most prevalent in the epipelagic were Collodaria. These
+mixotrophic Radiolaria have long been reported to contribute to primary
+productivity in the euphotic zone (Michaels \emph{et al.}, 1995; Dennett
+\emph{et al.}, 2002). Collodaria are thought to be particularly
+successful globally in oligotrophic regions due to their photosymbiotic
 relationships (Biard \emph{et al.}, 2016, 2017). We observed the highest
@@
 local protist community as well as global patterns in Rhizaria ecology
-{[}Laget \emph{et al.} (2024){]}. The upper mesopelagic interestingly
-had relatively low total abundance. This low-abundance region likely
+(Laget \emph{et al.}, 2024). The upper mesopelagic interestingly had
+relatively low total abundance. This low-abundance region likely
 reflects the dynamics of productivity and export throughout the water
@@
 California's Coast, much higher abundances of Aulosphaeridae have been
-reported {[}Zasko and Rusanov (2005); Biard and Ohman (2020){]} and they
-have massive potential to impact silica export (Biard \emph{et al.},
-2018). Coelodendridae were also seemingly restricted to the deeper
-section of the mesopelagic. This is interesting given that in the
-California Current, (Biard and Ohman, 2020) found a bimodal distribution
-in Coelodendridae. There are several morphotypes corresponding to
-different taxa of Coelodendridae (Nakamura and Suzuki, 2015; Biard and
-Ohman, 2020). So it may be that only a few types of Coelodendridae were
+reported (Zasko and Rusanov, 2005; Biard and Ohman, 2020) and they have
+massive potential to impact silica export (Biard \emph{et al.}, 2018).
+Coelodendridae were also seemingly restricted to the deeper section of
+the mesopelagic. This is interesting given that in the California
+Current, (Biard and Ohman, 2020) found a bimodal distribution in
+Coelodendridae. There are several morphotypes corresponding to different
+taxa of Coelodendridae (Nakamura and Suzuki, 2015; Biard and Ohman,
+2020). So it may be that only a few types of Coelodendridae were
 observed in this study, while the epipelagic variety was not.
@@
 rapidly return to the photic zone (Decelle \emph{et al.}, 2012, 2013).
-This hypothesis predicts that Acanthareas in the mesopelagic would be
+This hypothesis predicts that Acantharea in the mesopelagic would be
 smaller (Decelle \emph{et al.}, 2013). Mars Brisbin \emph{et al.} (2020)
@@
 due to lower food availability (Mars Brisbin \emph{et al.}, 2020). Since
-food is more scarce in the mesopelagic, nutritional quality lower, yet
-flux feeders would likely grow larger to increase their feeding range
-(Stukel \emph{et al.}, 2018; Biard and Ohman, 2020). In the data
-collecting in this study, Acanthareas in the mesopelagic were
-significantly smaller than the epipelagic, despite the other bimodal
-taxa (Foraminifera and Aulacanthidae) being significantly larger with
-depth. This provides added support for the hypothesis that cyst-forming
-Acantharea may utilize different sections of the water column throughout
-their life cycle. However to further investigate this, more work is
-needed with higher temporal and taxonomic resolution.
-
+food is more scarce in the mesopelagic, nutritional quality lower (Kim
+\emph{et al.}, 2018), yet flux feeders would likely grow larger to
+increase their feeding range (Stukel \emph{et al.}, 2018; Biard and
+Ohman, 2020). In the data collecting in this study, Acantharea in the
+mesopelagic were significantly smaller than the epipelagic, despite the
+other bimodal taxa (Foraminifera and Aulacanthidae) being significantly
+larger with depth. This provides added support for the hypothesis that
+cyst-forming Acantharea may utilize different sections of the water
+column throughout their life cycle. However to further investigate this,
+more work is needed with higher temporal and taxonomic resolution.
+
+\hypertarget{conclusions-and-considerations}{%
 \subsubsection{Conclusions and
-Considerations}\label{conclusions-and-considerations}
+Considerations}\label{conclusions-and-considerations}}
 
@@
 of oligotrophic regions. However, their role in food webs is not well
-understood. While this study represents a step-forward in our
-understanding of Rhizaria ecology, continued research on Rhizaria is
-much needed to better understand their ecology. Particularly extended
-work describing interannual patterns in Rhizaria abundance. Also work
-defining biotic interactions, feeding rates, productivity, and life
-history are all rich fields of interest in Rhizaria.
-
-\section{References:}\label{references}
-
-\phantomsection\label{refs}
+understood. While this study represents a step forward in our
+understanding of Rhizaria, further research is needed to better
+characterize their ecology and role in biogeochemical cycling. To fully
+understand the functioning of the oceans, not only carbon or nitrogen,
+but other elements such as silica or strontium, extended work addressing
+the interannual patterns in Rhizaria, their biotic interactions, and the
+key biological rates that impact ocean biogeochemistry, will be
+required.
+
+\hypertarget{references}{%
+\section{References:}\label{references}}
+
+\hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-anderson2014}
+\leavevmode\vadjust pre{\hypertarget{ref-anderson2014}{}}%
 Anderson, O. R. (2014)
@@
 
-\bibitem[\citeproctext]{ref-anderson1976}
+\leavevmode\vadjust pre{\hypertarget{ref-anderson1976}{}}%
 Anderson, O. R. and Bé, A. W. H. (1976)
@@
 
-\bibitem[\citeproctext]{ref-barth2022}
+\leavevmode\vadjust pre{\hypertarget{ref-barth2022}{}}%
 Barth, A. and Stone, J. (2022)
@@
 
-\bibitem[\citeproctext]{ref-barth2024}
+\leavevmode\vadjust pre{\hypertarget{ref-barth2024}{}}%
 Barth, A. and Stone, J. (2024)
@@
 
-\bibitem[\citeproctext]{ref-benfield1996}
+\leavevmode\vadjust pre{\hypertarget{ref-benfield1996}{}}%
 Benfield, M. C., Davis, C. S., Wiebe, P. H., Gallager, S. M., Gregory
@@
 
-\bibitem[\citeproctext]{ref-biard2022b}
+\leavevmode\vadjust pre{\hypertarget{ref-biard2022b}{}}%
 Biard, T. (2022a)
@@
 
-\bibitem[\citeproctext]{ref-biard2022}
+\leavevmode\vadjust pre{\hypertarget{ref-biard2022}{}}%
 Biard, T. (2022b)
@@
 
-\bibitem[\citeproctext]{ref-biard2017}
+\leavevmode\vadjust pre{\hypertarget{ref-biard2017}{}}%
 Biard, T., Bigeard, E., Audic, S., Poulain, J., Gutierrez-Rodriguez, A.,
@@
 
-\bibitem[\citeproctext]{ref-biard2018}
+\leavevmode\vadjust pre{\hypertarget{ref-biard2018}{}}%
 Biard, T., Krause, J. W., Stukel, M. R., and Ohman, M. D. (2018)
@@
 
-\bibitem[\citeproctext]{ref-biard2020}
+\leavevmode\vadjust pre{\hypertarget{ref-biard2020}{}}%
 Biard, T. and Ohman, M. D. (2020)
@@
 
-\bibitem[\citeproctext]{ref-biard2016}
+\leavevmode\vadjust pre{\hypertarget{ref-biard2016}{}}%
 Biard, T., Stemmann, L., Picheral, M., Mayot, N., Vandromme, P., Hauss,
@@
 
-\bibitem[\citeproctext]{ref-bijma1990}
+\leavevmode\vadjust pre{\hypertarget{ref-bijma1990}{}}%
 Bijma, J., Erez, J., and Hemleben, C. (1990)
@@
 
-\bibitem[\citeproctext]{ref-blanco-bercial2020}
+\leavevmode\vadjust pre{\hypertarget{ref-blanco-bercial2020}{}}%
 Blanco-Bercial, L. (2020)
@@
 
-\bibitem[\citeproctext]{ref-blanco-bercial2022}
+\leavevmode\vadjust pre{\hypertarget{ref-blanco-bercial2022}{}}%
 Blanco-Bercial, L., Parsons, R., Bolaños, L. M., Johnson, R.,
@@
 
-\bibitem[\citeproctext]{ref-boltovskoy2017}
+\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy2017}{}}%
 Boltovskoy, D. (2017)
@@
 
-\bibitem[\citeproctext]{ref-boltovskoy1993}
+\leavevmode\vadjust pre{\hypertarget{ref-boltovskoy1993}{}}%
 Boltovskoy, D., Alder, V. A., and Abelmann, A. (1993)
@@
 
-\bibitem[\citeproctext]{ref-caron2016}
+\leavevmode\vadjust pre{\hypertarget{ref-caron2016}{}}%
 Caron, D. A. (2016)
@@
 
-\bibitem[\citeproctext]{ref-caron1984}
+\leavevmode\vadjust pre{\hypertarget{ref-caron1984}{}}%
 Caron, D. A. and Be, A. W. H. (1984) PREDICTED AND OBSERVED FEEDING
@@
 
-\bibitem[\citeproctext]{ref-caron1995}
+\leavevmode\vadjust pre{\hypertarget{ref-caron1995}{}}%
 Caron, D. A., Michaels, A. F., Swanberg, N. R., and Howse, F. A. (1995)
@@
 
-\bibitem[\citeproctext]{ref-cohen2023}
+\leavevmode\vadjust pre{\hypertarget{ref-cohen2023}{}}%
 Cohen, N. R., Krinos, A. I., Kell, R. M., Chmiel, R. J., Moran, D. M.,
@@
 
-\bibitem[\citeproctext]{ref-decelle2015}
+\leavevmode\vadjust pre{\hypertarget{ref-decelle2015}{}}%
 Decelle, J., Colin, S., and Foster, R. A. (2015)
@@
 
-\bibitem[\citeproctext]{ref-decelle2013}
+\leavevmode\vadjust pre{\hypertarget{ref-decelle2013}{}}%
 Decelle, J., Martin, P., Paborstava, K., Pond, D. W., Tarling, G., Mahé,
@@
 
-\bibitem[\citeproctext]{ref-decelle2012}
+\leavevmode\vadjust pre{\hypertarget{ref-decelle2012}{}}%
 Decelle, J., Siano, R., Probert, I., Poirier, C., and Not, F. (2012)
@@
 
-\bibitem[\citeproctext]{ref-dennett2002}
+\leavevmode\vadjust pre{\hypertarget{ref-dennett2002}{}}%
 Dennett, M. R., Caron, D. A., Michaels, A. F., Gallager, S. M., and
@@
 
-\bibitem[\citeproctext]{ref-drago2022}
+\leavevmode\vadjust pre{\hypertarget{ref-drago2022}{}}%
 Drago, L., Panaïotis, T., Irisson, J.-O., Babin, M., Biard, T.,
@@
 
-\bibitem[\citeproctext]{ref-gaskell2019}
+\leavevmode\vadjust pre{\hypertarget{ref-gaskell2019}{}}%
 Gaskell, D. E., Ohman, M. D., and Hull, P. M. (2019)
@@
 
-\bibitem[\citeproctext]{ref-gorsky2010}
+\leavevmode\vadjust pre{\hypertarget{ref-gorsky2010}{}}%
 Gorsky, G., Ohman, M. D., Picheral, M., Gasparini, S., Stemmann, L.,
@@
 
-\bibitem[\citeproctext]{ref-gowing1989}
+\leavevmode\vadjust pre{\hypertarget{ref-gowing1989}{}}%
 Gowing, M. M. (1989) \href{https://doi.org/10.1007/BF00391069}{Abundance
@@
 
-\bibitem[\citeproctext]{ref-gutierrez-rodriguez2019}
+\leavevmode\vadjust pre{\hypertarget{ref-gutierrez-rodriguez2019}{}}%
 Gutierrez-Rodriguez, A., Stukel, M. R., Lopes dos Santos, A., Biard, T.,
@@
 
-\bibitem[\citeproctext]{ref-gutiuxe9rrez-rodruxedguez2022}
+\leavevmode\vadjust pre{\hypertarget{ref-gutiuxe9rrez-rodruxedguez2022}{}}%
 Gutiérrez-Rodríguez, A., Lopes dos Santos, A., Safi, K., Probert, I.,
@@
 
-\bibitem[\citeproctext]{ref-haekel1887}
+\leavevmode\vadjust pre{\hypertarget{ref-haekel1887}{}}%
 Haekel, E. (1887)
@@
 
-\bibitem[\citeproctext]{ref-hull2011}
+\leavevmode\vadjust pre{\hypertarget{ref-hull2011}{}}%
 Hull, P. M., Osborn, K. J., Norris, R. D., and Robison, B. H. (2011)
@@
 
-\bibitem[\citeproctext]{ref-ikenoue2019}
+\leavevmode\vadjust pre{\hypertarget{ref-ikenoue2019}{}}%
 Ikenoue, T., Kimoto, K., Okazaki, Y., Sato, M., Honda, M. C., Takahashi,
@@
 
-\bibitem[\citeproctext]{ref-kimoto2015}
+\leavevmode\vadjust pre{\hypertarget{ref-kim2018}{}}%
+Kim, B. K., Lee, S., Ha, S.-Y., Jung, J., Kim, T. W., Yang, E. J., Jo,
+N., Lim, Y. J., \emph{et al.} (2018)
+\href{https://doi.org/10.1002/2017JC013457}{Vertical Distributions of
+Macromolecular Composition of Particulate Organic Matter in the Water
+Column of the Amundsen Sea Polynya During the Summer in 2014}.
+\emph{Journal of Geophysical Research: Oceans}, \textbf{123},
+1393--1405.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kimoto2015}{}}%
 Kimoto, K. (2015)
@@
 
-\bibitem[\citeproctext]{ref-knap1997}
+\leavevmode\vadjust pre{\hypertarget{ref-knap1997}{}}%
 Knap, A. H., Michaels, A. F., Steinberg, D. K., Bahr, F., Bates, N. R.,
@@
 
-\bibitem[\citeproctext]{ref-laget2024}
+\leavevmode\vadjust pre{\hypertarget{ref-laget2024}{}}%
 Laget, M., Drago, L., Panaïotis, T., Kiko, R., Stemmann, L., Rogge, A.,
@@
 
-\bibitem[\citeproctext]{ref-lee2006}
+\leavevmode\vadjust pre{\hypertarget{ref-lee2006}{}}%
 Lee, J. J. (2006)
@@
 
-\bibitem[\citeproctext]{ref-llopismonferrer2022}
+\leavevmode\vadjust pre{\hypertarget{ref-llopismonferrer2022}{}}%
 Llopis Monferrer, N., Biard, T., Sandin, M. M., Lombard, F., Picheral,
@@
 
-\bibitem[\citeproctext]{ref-lomas2013}
+\leavevmode\vadjust pre{\hypertarget{ref-lomas2013}{}}%
 Lomas, M. W., Bates, N. R., Johnson, R. J., Knap, A. H., Steinberg, D.
@@
 
-\bibitem[\citeproctext]{ref-lombard2019}
+\leavevmode\vadjust pre{\hypertarget{ref-lombard2019}{}}%
 Lombard, F., Boss, E., Waite, A. M., Vogt, M., Uitz, J., Stemmann, L.,
@@
 
-\bibitem[\citeproctext]{ref-marra2011}
+\leavevmode\vadjust pre{\hypertarget{ref-marra2011}{}}%
 Marra, G. and Wood, S. N. (2011)
@@
 
-\bibitem[\citeproctext]{ref-marsbrisbin2020}
+\leavevmode\vadjust pre{\hypertarget{ref-marsbrisbin2020}{}}%
 Mars Brisbin, M., Brunner, O. D., Grossmann, M. M., and Mitarai, S.
@@
 
-\bibitem[\citeproctext]{ref-mcgillicuddy1998}
+\leavevmode\vadjust pre{\hypertarget{ref-mcgillicuddy1998}{}}%
 McGillicuddy, D. J., Robinson, A. R., Siegel, D. A., Jannasch, H. W.,
@@
 
-\bibitem[\citeproctext]{ref-michaels1988}
+\leavevmode\vadjust pre{\hypertarget{ref-michaels1988}{}}%
 Michaels, A. F. (1988)
@@
 
-\bibitem[\citeproctext]{ref-michaels1995}
+\leavevmode\vadjust pre{\hypertarget{ref-michaels1995}{}}%
 Michaels, A. F., Caron, D. A., Swanberg, N. R., Howse, F. A., and
@@
 
-\bibitem[\citeproctext]{ref-michaels1996}
+\leavevmode\vadjust pre{\hypertarget{ref-michaels1996}{}}%
 Michaels, A. F. and Knap, A. H. (1996)
@@
 
-\bibitem[\citeproctext]{ref-nakamura2023}
+\leavevmode\vadjust pre{\hypertarget{ref-nakamura2023}{}}%
 Nakamura, Y., Itagaki, H., Tuji, A., Shimode, S., Yamaguchi, A., Hidaka,
@@
 
-\bibitem[\citeproctext]{ref-nakamura2015}
+\leavevmode\vadjust pre{\hypertarget{ref-nakamura2015}{}}%
 Nakamura, Y. and Suzuki, N. (2015)
@@
 
-\bibitem[\citeproctext]{ref-ohman2019a}
+\leavevmode\vadjust pre{\hypertarget{ref-ohman2019a}{}}%
 Ohman, M. D. (2019) \href{https://doi.org/10.1093/icesjms/fsz184}{A sea
@@
 
-\bibitem[\citeproctext]{ref-panauxefotis2023}
+\leavevmode\vadjust pre{\hypertarget{ref-panauxefotis2023}{}}%
 Panaïotis, T., Babin, M., Biard, T., Carlotti, F., Coppola, L., Guidi,
@@
 
-\bibitem[\citeproctext]{ref-picheral}
+\leavevmode\vadjust pre{\hypertarget{ref-picheral}{}}%
 Picheral, M., Colin, S. P., and Irisson, J.-O.
@@
 
-\bibitem[\citeproctext]{ref-picheral2010}
+\leavevmode\vadjust pre{\hypertarget{ref-picheral2010}{}}%
 Picheral, M., Guidi, L., Stemmann, L., Karl, D. M., Iddaoud, G., and
@@
 
-\bibitem[\citeproctext]{ref-sogawa2022}
+\leavevmode\vadjust pre{\hypertarget{ref-sogawa2022}{}}%
 Sogawa, S., Nakamura, Y., Nagai, S., Nishi, N., Hidaka, K., Shimizu, Y.,
@@
 
-\bibitem[\citeproctext]{ref-stukel2018}
+\leavevmode\vadjust pre{\hypertarget{ref-stukel2018}{}}%
 Stukel, M. R., Biard, T., Krause, J., and Ohman, M. D. (2018)
@@
 
-\bibitem[\citeproctext]{ref-stukel2019}
+\leavevmode\vadjust pre{\hypertarget{ref-stukel2019}{}}%
 Stukel, M. R., Ohman, M. D., Kelly, T. B., and Biard, T. (2019)
@@
 
-\bibitem[\citeproctext]{ref-suzuki2015}
+\leavevmode\vadjust pre{\hypertarget{ref-suzuki2015}{}}%
 Suzuki, N. and Not, F. (2015)
@@
 
-\bibitem[\citeproctext]{ref-swanberg1981}
+\leavevmode\vadjust pre{\hypertarget{ref-swanberg1981}{}}%
 Swanberg, N. R. and Anderson, O. R. (1981)
@@
 
-\bibitem[\citeproctext]{ref-takahashi1983}
+\leavevmode\vadjust pre{\hypertarget{ref-takahashi1983}{}}%
 Takahashi, K., Hurd, D. C., and Honjo, S. (1983)
@@
 
-\bibitem[\citeproctext]{ref-whitmore2021}
+\leavevmode\vadjust pre{\hypertarget{ref-whitmore2021}{}}%
 Whitmore, B. M. and Ohman, M. D. (2021)
@@
 
-\bibitem[\citeproctext]{ref-wood2017}
+\leavevmode\vadjust pre{\hypertarget{ref-wood2017}{}}%
 Wood, S. N. (2017) \emph{Generalized additive models: an introduction
@@
 
-\bibitem[\citeproctext]{ref-wood2001}
+\leavevmode\vadjust pre{\hypertarget{ref-wood2001}{}}%
 Wood, S. N. (2001) mgcv: GAMs and Generalized Ridge Regression for R.
@@
 
-\bibitem[\citeproctext]{ref-zasko2005}
+\leavevmode\vadjust pre{\hypertarget{ref-zasko2005}{}}%
 Zasko, D. N. and Rusanov, I. I. (2005)
